### 2012-03-23 many.tex
Initial import.

\documentclass[aps,prd,reprint]{revtex4-1}

\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}
\newcommand{\fixme}[1]{\textbf{FIXME: #1}}

\begin{document}

\title{Bayesian Rate Estimation}

\author{Will M. Farr}
\email{w-farr@northwestern.edu}
%\homepage{http://faculty.wcas.northwestern.edu/will-farr/}
% \affiliation{Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA)\\
% Department of Physics and Astronomy\\
% Northwestern University, 2145 Sheridan Road, Evanston, IL 60208}
\affiliation{Northwestern University and CIERA}

\author{Ilya Mandel}
\email{ilya@chgk.info}
%\homepage{http://www.chgk.info/~ilyamandel/}
% \affiliation{School of Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT Birmingham\\United Kingdom}
\affiliation{University of Birmingham}

\author{Jon Gair}

\begin{abstract}
  We show how to obtain a Bayesian estimate of the rate of signal
  events from a set of signal and background events indexed by a
  ranking statistic when the shapes of the signal and background
  distributions are known, can be estimated, or approximated.  We
  focus on the specific application of estimating astrophysical rates
  of the coalescence of compact binary black holes or neutron stars
  from a set of triggers in the LIGO/Virgo gravitational wave
  detectors, but our framework is fully general.  We discuss the
  systematic effects on the rate estimate due to differences between
  the assumed and true shapes of the foreground or background
  distributions, identifying ways these effects can be minimized.
  Similarly, we discuss the effects of various priors on the rate,
  including uninformative priors, weakly-informative priors, and the
  use of priors from previous rate experiments.  In the limit where
  the expected signal rate gives high probability of zero or one
  signal in the data, our technique reduces to the ``loudest event
  statistic,'' but it is generally applicable to arbitrarily large
  signal rates.
\end{abstract}

\maketitle

\section{Introduction}

\fixme{Introduce the necessity of estimating rates, prior work (like
  \cite{Biswas2009}), Bayesian inference.}

\section{Model}

We assume that we are presented with a data set of $N$ events.  Each
event may be due to either a signal of interest or an uninteresting
background.  Each event is associated with a ranking statistic, $x$.
Our data set therefore consists of the ranking statistics for the set
of events:
\begin{equation}
  d = \{ x_i | i = 1, \ldots, N \}.
\end{equation}
We assume that the events are sorted by ranking statistic, so that $i
< j$ implies that $x_i < x_j$.

We assume that both the foreground and background events are samples
from an inhomogeneous Poisson process with rates (per unit ranking
statistic, $x$)
\begin{equation}
  \frac{dN_f}{dx} = f(x)
\end{equation}
and 
\begin{equation}
  \frac{dN_b}{dx} = b(x).
\end{equation}
The cumulative rates of the two processes are therefore
\begin{equation}
  F(x) \equiv \int_{-\infty}^x ds\, f(s)
\end{equation}
and
\begin{equation}
  B(x) \equiv \int_{-\infty}^x ds\, b(s).
\end{equation}
The assumption that the foreground and background events form an
inhomogeneous Poisson process implies
\begin{enumerate}
\item The number of events in any range of ranking statistics, $x \in
  [x_1, x_2]$ is Poisson distributed with rate $F(x_2) - F(x_1)$ or
  $B(x_2) - B(x_1)$.
\item The numbers of events in non-overlapping ranges of ranking
  statistics are independent. 
\item The probability of exactly one foreground event between $x$ and
  $x+h$ is given by
  \begin{equation}
    P(N = 1 \in [x, x+h]) = f(x) h + \order{h^2}.
  \end{equation}
  and similarly for background events.
\item The probability of two or more events in a small range of
  ranking statistic is negligable
  \begin{equation}
    P(N = 2 \in [x, x+h]) = \order{h^2}.
  \end{equation}
\end{enumerate}
The foreground and background rates can in general depend on several
parameters; the goal of our analysis is to determine the posterior
probability distributions for these parameters that are implied by the
data.  At the least, we will want to know the overall amplitude of the
foreground and background rates.  Let
\begin{equation}
  f(x) = R_f \hat{f}(x),
\end{equation}
and 
\begin{equation}
  b(x) = R_b \hat{b}(x),
\end{equation}
where $\hat{F}(\infty) = \hat{B}(\infty) = 1$.  Then $R_f$ and $R_b$
are the total number of foreground and background events expected,
respectively.  Other parameters may describe the shape of the rate
functions, but these will depend on the details of the system being
analyzed.  In \S \ref{sec:GW-example} we give an example of fitting
such shape paremeters.

We do not know a priori which of the events are foreground and which
are background.  For each event, we introduce a flag, $s_i$, which is
either 0 or 1, indicating a background or foreground event,
respectively.  These ``state'' flags are additional parameters in our
model.  We can marginalize over our uncertainty in the state of any
given event by summing all posteriors over $s_i = \{0,1\}$.  Given an
identification of each event as foreground or background, and the
rates of each, the likelihood of our data is
\begin{multline}
  p\left( d | \left\{ f(x), b(x), \left\{ s_i \right\} \right\}
  \right) = \\ \left[ \prod_{\left\{ i | s_i = 1 \right\}} f\left( x_i
    \right) \right]  \left[ \prod_{\left\{ j | s_j = 0 \right\}}
    b\left( x_j \right) \right] \\
  \times \exp\left[ - F\left( \infty \right) \right] 
  \exp\left[ -B\left( \infty \right) \right].
\end{multline}
Written in terms of the rate parameters, this becomes
\begin{multline}
  p\left( d | \left\{ f(x), b(x), \left\{ s_i \right\} \right\}
  \right) = \\ R_f^{N_f} \left[ \prod_{\left\{ i | s_i = 1 \right\}}
    \hat{f}\left( x_i \right) \right] \exp\left[ - R_f \right] \\
  \times R_b^{N_b} \left[ \prod_{\left\{ j | s_j = 0 \right\}}
    \hat{b}\left( x_j \right) \right] \exp\left[ - R_b \right],
\end{multline}
where $N_f$ is the number of the $s_i$ that are 1 (i.e.\ the number of
assumed foreground events), and $N_b$ is the number of the $s_i$ that
are 0 (i.e.\ the assumed number of background events).  Note that $N_f
+ N_b = N$, as each event is considered either foreground or
background in our model.

\subsection{Priors}

\fixme{Discussion of the various priors one might want to impose on
  the rates, $R_f$ and $R_b$, such as a flat prior on rate, the
  Poisson Jeffrey's prior (which is $\propto 1/\sqrt{R}$?), .}

A choice of the foreground and background rates implies a prior on the
state flags.  For fixed foreground and background rates, the
probability that any particular event is foreground is given by 
\begin{equation}
  p\left(s_i = 1 | R_f, R_b \right) = \frac{R_f}{R_f + R_b}.
\end{equation}

The complete prior is then
\begin{multline}
  p\left( R_f, R_b, \left\{ s_i \right\} \right) = \\ p\left( \left\{ s_i
  \right\} | R_f, R_b \right) p\left( R_f, R_b \right) = \\ \frac{R_f^{N_f} R_b^{N_b}}{\left( R_f
    + R_b\right)^N} p\left( R_f, R_b \right).
\end{multline}

\section{Gravitational Waves from Compact Binary Inspirals}
\label{sec:GW-example}

\begin{acknowledgments}
  Richard O'Shaughnessy for discussions.
\end{acknowledgments}

\appendix

\section{Likelihood for Inhomogeneous Poisson Processes}

Consider a set of samples, $\{ x_i \}$, from an

%%%%%%%%%%%%%%%%%%%% Bibliography 
\bibliographystyle{apsrev4-1}
\bibliography{many}

\end{document}

### 2012-03-27 many.tex
Edits to many.tex.

@@
 We do not know a priori which of the events are foreground and which
-are background.  For each event, we introduce a flag, $s_i$, which is
+are background.  For each event, we introduce a flag, $f_i$, which is
 either 0 or 1, indicating a background or foreground event,
@@
 model.  We can marginalize over our uncertainty in the state of any
-given event by summing all posteriors over $s_i = \{0,1\}$.  Given an
+given event by summing all posteriors over $f_i = \{0,1\}$.  Given an
 identification of each event as foreground or background, and the
-rates of each, the likelihood of our data is
+rates of each, the likelihood%
+\footnote{See \S \ref{sec:likelihood-derivation} for a derivation of
+  Eq.~\eqref{eq:likelihood}.} %
+of our data is
 \begin{multline}
-  p\left( d | \left\{ f(x), b(x), \left\{ s_i \right\} \right\}
-  \right) = \\ \left[ \prod_{\left\{ i | s_i = 1 \right\}} f\left( x_i
-    \right) \right]  \left[ \prod_{\left\{ j | s_j = 0 \right\}}
+  \label{eq:likelihood}
+  p\left( d, \left\{ f_i \right\} | f(x), b(x)
+  \right) = \\ \left[ \prod_{\left\{ i | f_i = 1 \right\}} f\left( x_i
+    \right) \right]  \left[ \prod_{\left\{ j | f_j = 0 \right\}}
     b\left( x_j \right) \right] \\
@@
 \begin{multline}
-  p\left( d | \left\{ f(x), b(x), \left\{ s_i \right\} \right\}
-  \right) = \\ R_f^{N_f} \left[ \prod_{\left\{ i | s_i = 1 \right\}}
+  p\left( d, \left\{ f_i \right\} | f(x), b(x)\right) = \\ R_f^{N_f}
+  \left[ \prod_{\left\{ i | f_i = 1 \right\}}
     \hat{f}\left( x_i \right) \right] \exp\left[ - R_f \right] \\
-  \times R_b^{N_b} \left[ \prod_{\left\{ j | s_j = 0 \right\}}
+  \times R_b^{N_b} \left[ \prod_{\left\{ j | f_j = 0 \right\}}
     \hat{b}\left( x_j \right) \right] \exp\left[ - R_b \right],
 \end{multline}
-where $N_f$ is the number of the $s_i$ that are 1 (i.e.\ the number of
-assumed foreground events), and $N_b$ is the number of the $s_i$ that
+where $N_f$ is the number of the $f_i$ that are 1 (i.e.\ the number of
+assumed foreground events), and $N_b$ is the number of the $f_i$ that
 are 0 (i.e.\ the assumed number of background events).  Note that $N_f
@@
 
-A choice of the foreground and background rates implies a prior on the
-state flags.  For fixed foreground and background rates, the
-probability that any particular event is foreground is given by 
+\subsection{Posteriors}
+
+Putting everything together, the posterior probability of each event's
+state and the overall rate is given by
+\begin{multline}
+  p(R_f, R_b, \left\{ f_i \right\} | d) \propto R_f^{N_f}
+  \left[ \prod_{\left\{ i | f_i = 1 \right\}}
+    \hat{f}\left( x_i \right) \right] \exp\left[ - R_f \right] \\
+  \times R_b^{N_b} \left[ \prod_{\left\{ j | f_j = 0 \right\}}
+    \hat{b}\left( x_j \right) \right] \exp\left[ - R_b \right] p(R_f, R_b).
+\end{multline}
+If we marginalize over all the flags, $f_i$, then we obtain
+\begin{multline}
+  p(R_f, R_b | d) \propto \prod_{i} \left[ R_f \hat{f}\left( x_i
+    \right) + R_b \hat{b}\left( x_i \right) \right] \\ \times
+  \exp\left[ - \left( R_f + R_b \right) \right] p\left( R_f, R_b
+  \right).
+\end{multline}
+If, instead, we marginalize over all but one flag, $f_k$, we obtain
+that the posterior ratio of signal to noise probability for event $k$
+given the rates $R_f$ and $R_b$ is 
 \begin{equation}
-  p\left(s_i = 1 | R_f, R_b \right) = \frac{R_f}{R_f + R_b}.
+  \frac{p(R_f, R_b, f_k = 1 | d)}{p(R_f, R_b, f_k = 0 | d)} =
+  \frac{R_f \hat{f}\left( x_k \right)}{R_b \hat{b}\left( x_k \right)}. 
 \end{equation}
-
-The complete prior is then
-\begin{multline}
-  p\left( R_f, R_b, \left\{ s_i \right\} \right) = \\ p\left( \left\{ s_i
-  \right\} | R_f, R_b \right) p\left( R_f, R_b \right) = \\ \frac{R_f^{N_f} R_b^{N_b}}{\left( R_f
-    + R_b\right)^N} p\left( R_f, R_b \right).
-\end{multline}
 
@@
 \section{Likelihood for Inhomogeneous Poisson Processes}
+\label{sec:likelihood-derivation}
 

### 2012-03-31 many.tex
Updates to many.tex.  Added section describing analytic model and recovery of rates.

@@
 \usepackage{amssymb}
+\usepackage{graphicx}
 
@@
 \newcommand{\fixme}[1]{\textbf{FIXME: #1}}
+
+\newcommand{\xmin}{x_\mathrm{min}}
+
+\DeclareMathOperator{\erf}{erf}
 
@@
 \end{equation}
-We assume that the events are sorted by ranking statistic, so that $i
-< j$ implies that $x_i < x_j$.
 
@@
 either 0 or 1, indicating a background or foreground event,
-respectively.  These ``state'' flags are additional parameters in our
-model.  We can marginalize over our uncertainty in the state of any
-given event by summing all posteriors over $f_i = \{0,1\}$.  Given an
-identification of each event as foreground or background, and the
+respectively.  These ``state'' flags are additional (unobserved) data
+in our model.  We can marginalize over our uncertainty in the state of
+any given event by summing all posteriors over $f_i = \{0,1\}$.  Given
+an identification of each event as foreground or background, and the
 rates of each, the likelihood%
@@
 \begin{multline}
+  \label{eq:likelihood-rates}
   p\left( d, \left\{ f_i \right\} | f(x), b(x)\right) = \\ R_f^{N_f}
@@
 
+It will occasionally be convenient to work with the ratio,
+$\hat{f}(x)/\hat{b}(x)$ instead of the values of $\hat{f}$ and
+$\hat{b}$ separately.  This ratio can be interpreted as a likelihood
+ratio between the foreground and background models for an event that
+has occurred with ranking statistic $x$:
+\begin{equation}
+  \label{eq:loudness-likelihood}
+  \frac{\hat{f}(x)}{\hat{b}(x)} = \frac{p(x|\mathrm{foreground})}{p(x|\mathrm{background})}.
+\end{equation}
+In some circumstances, it is possible to compute this ratio
+empirically (for example, in \citet{Cannon2012}, the authors give a
+practical method for estimating the distribution of loudnesses for
+known-background events and known-foreground events in a mock data
+stream from the Advanced LIGO and Advanced Virgo gravitational wave
+detectors).  Dividing Eq.~\eqref{eq:likelihood-rates} by
+$\prod_{\left\{ i | f_i = 0 \right\}} \hat{b}(x_i)$, we obtain
+\begin{multline}
+  \label{eq:likelihood-ratio}
+  p\left(d, \left\{ f_i \right\} | f(x), b(x) \right) \propto  \\ \left[
+    \prod_{\left\{ i | f_i = 1 \right\}}
+    \frac{\hat{f}(x_i)}{\hat{b}(x_i)} \right] R_f^{N_f} R_b^{N_b}
+  \exp\left[ - \left( R_f + R_b \right) \right].
+\end{multline}
+
 \subsection{Priors}
 
-\fixme{Discussion of the various priors one might want to impose on
-  the rates, $R_f$ and $R_b$, such as a flat prior on rate, the
-  Poisson Jeffrey's prior (which is $\propto 1/\sqrt{R}$?), .}
+Because the overall rates enter the likelihood in
+Eq.~\eqref{eq:likelihood-rates} in the same form as in the
+constant-rate Poisson likelihood, a reasonable prior for $R_f$ and
+$R_b$ is the Poisson Jeffrey's prior%
+\footnote{Recall that the Jeffrey's prior for a parameter $\theta$
+  with likeilhood function, $p(d|\theta)$, is
+  \begin{equation}
+    p(\theta) \propto \sqrt{\left\langle \left( \frac{\partial
+          p(d | \theta) }{\partial \theta} \right)^2 \right\rangle},
+  \end{equation}
+  where the expectation is taken over all data, $d$.  The Jeffrey's
+  prior assigns prior weight to locations in parameter space
+  proportional to the Fisher information that is expected to result
+  from a measurement at that parameter value.}, %
+\begin{equation}
+  \label{eq:Jeffrey-prior}
+  p(R) \propto \frac{1}{\sqrt{R}}.
+\end{equation}
+One advantage of the Jeffrey's prior in this circumstance is that it
+has finite probability mass as $R \to 0$, and the likelihood in
+Eq.~\eqref{eq:likelihood-rates} controls its behavior as $R\to \infty$
+to ensure a normalizable posterior.
+
+Of course, other priors could be used; in particular, previous
+experiments may have placed constraints on the rate that can be used
+as priors in subsequent experiments.  Throughout the remainder of this
+paper, however, we use the Jeffrey's prior,
+Eq.~\eqref{eq:Jeffrey-prior}.
 
@@
 \begin{multline}
+  \label{eq:rates-marginalized}
   p(R_f, R_b | d) \propto \prod_{i} \left[ R_f \hat{f}\left( x_i
@@
 \begin{equation}
+  \label{eq:foreground-marginalized}
   \frac{p(R_f, R_b, f_k = 1 | d)}{p(R_f, R_b, f_k = 0 | d)} =
-  \frac{R_f \hat{f}\left( x_k \right)}{R_b \hat{b}\left( x_k \right)}. 
+  \frac{R_f \hat{f}\left( x_k \right)}{R_b \hat{b}\left( x_k
+    \right)}p\left(R_f, R_b\right). 
 \end{equation}
+Note that the \emph{posterior} estimate of the ratio of signal to
+noise probabilities for event $i$ differs from the likelihood ratio,
+Eq.~\eqref{eq:loudness-likelihood}, because it depends on the overall
+rates of foreground and background events.
 
@@
 
+\subsection{Analytic Example}
+\label{sec:analytic-GW-example}
+
+Suppose we attempt to detect gravitational wave signals in a data
+stream by matched filtering against a set of $N$ templates.  The data
+stream consists of stationary Gaussian noise with a power spectral
+density $S(f)$ combined additively with some number of signals.  The
+signal to noise ratio (SNR) of a template, $h(f)$, given data, $d(f)$,
+is given by
+\begin{equation}
+  \rho_h \equiv \frac{\left\langle h, d \right\rangle}{\sqrt{\left
+        \langle h, h \right\rangle}},
+\end{equation}
+where $\left \langle \cdot \right\rangle$ denotes the noise-weighted
+inner product:
+\begin{equation}
+  \left\langle a, b \right\rangle \equiv \int_0^\infty df\,
+  \frac{a^*(f) b(f)}{S(f)}.
+\end{equation}
+We suppose for simplicity that the templates are sufficiently distinct
+that 
+\begin{equation}
+  \left\langle h_1, h_2 \right\rangle \sim 0.
+\end{equation}
+
+For a data stream of pure noise, $d(f) = n(f)$, the SNR of a each
+template follows a $N(0,1)$ distribution.  If our ranking statistic
+for events in each interval of data is their maximum SNR over the $N$
+templates, 
+\begin{equation}
+  x \equiv \max_h \rho_h,
+\end{equation}
+then the background distribution is
+\begin{equation}
+  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{2}
+      \right)}{2} \right)^N
+\end{equation}
+If, in addition, we impose a threshold ranking statistic, $\xmin$, to
+be considered an ``event,'' then $\hat{B}$ becomes
+\begin{equation}
+  \label{eq:analytic-background-rate}
+  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{2} \right)
+    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
+    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
+    \right)^N }
+\end{equation}
+
+The SNR of a gravitational wave signal in an interferometric detector
+scales as $1/d$, where $d$ is the distance to the source.  Ignoring
+possible cosmological effects, the number of sources will scale as
+$d^3$.  Thus, we expect that the foreground cumulative distribution of
+events will follow
+\begin{equation}
+  \label{eq:analytic-foreground-rate}
+  \hat{F}(x) = 1 - \frac{\xmin^3}{x^3}.
+\end{equation}
+
+To demonstrate the effectiveness of our formalism, we applied it to a
+synthetic data set with foreground and background distributions drawn
+from Eqs.~\eqref{eq:analytic-background-rate} and
+\eqref{eq:analytic-foreground-rate} with $R_f^\mathrm{true} = 10.4$
+and $R_b^\mathrm{true} = 95.1$.  The cumulative distribution for the
+ranking statistic of the synthetic data we used appears in Figure
+\ref{fig:analytic-data-cumulative}.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{data}
+  \caption{\label{fig:analytic-data-cumulative} The cumulative
+    distribution of the ranking statistics for the synthetic data used
+    to test the the formalism on the model from \S
+    \ref{sec:analytic-GW-example}.  The solid line gives the
+    cumulative distribution of the synthetic data; the dashed line
+    gives the theoretical cumulative distribution for the models in
+    Eqs.~\eqref{eq:analytic-background-rate} and
+    \eqref{eq:analytic-foreground-rate} combined with $R_f = 10.4$ and
+    $R_b = 95.1$.}
+\end{figure}
+
+In Figure \ref{fig:analytic-rate-recovery}, we show the marginalized
+posterior densities for the foreground and background rates recovered
+from a synthetic data set drawn from the distributions in
+Eqs.~\eqref{eq:analytic-background-rate} and
+\eqref{eq:analytic-foreground-rate} with .  (Refer to
+Eq.~\eqref{eq:rates-marginalized}.)  Figure
+\ref{fig:analytic-rate-foreground-probs} shows the posterior
+foreground probability for each event marginalized over all other
+events' types and the foreground and background rates.  (Refer to
+Eq.~\eqref{eq:foreground-marginalized}.)
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{rates}
+  \caption{\label{fig:analytic-rate-recovery} The posterior densities
+    for $R_f$ (in blue) and $R_b$ (in black) for the analytic model
+    discussed in \S \ref{sec:analytic-GW-example}.  The vertical lines
+    indicate the ``true'' used to generate the synthetic data set.}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{pfore}
+  \caption{\label{fig:analytic-rate-foreground-probs} Foreground
+    probability for each event in the synthetic data set of \S
+    \ref{sec:analytic-GW-example} marginalized over all other
+    parameters.  True foreground events are in blue, background events
+    in red.  The solid line is the likelihood ratio
+    $p(x|\mathrm{foreground})/p(x|\mathrm{background})$ (see
+    Eq.~\eqref{eq:loudness-likelihood}) for this model; this exceeds
+    the marginalized foreground probability for many of the events
+    because in this data set there are approximately nine times as
+    many background events as foreground events (see
+    Eq.~\eqref{eq:foreground-marginalized}). }
+\end{figure}
+
 \begin{acknowledgments}
-  Richard O'Shaughnessy for discussions.
+  Kipp Cannon, Chad Hanna, and Drew Keppel for providing the data set
+  from \citet{Cannon2012} and assisting us in its use. Richard
+  O'Shaughnessy for discussions.
 \end{acknowledgments}
@@
 
-Consider a set of samples, $\{ x_i \}$, from an
+\fixme{To do.}
 

### 2012-03-31 many.tex
Fixed minor typo in III A.

@@
 In Figure \ref{fig:analytic-rate-recovery}, we show the marginalized
-posterior densities for the foreground and background rates recovered
-from a synthetic data set drawn from the distributions in
-Eqs.~\eqref{eq:analytic-background-rate} and
-\eqref{eq:analytic-foreground-rate} with .  (Refer to
-Eq.~\eqref{eq:rates-marginalized}.)  Figure
+posterior densities for the foreground and background rates.  (Refer
+to Eq.~\eqref{eq:rates-marginalized}.)  Figure
 \ref{fig:analytic-rate-foreground-probs} shows the posterior

### 2012-06-08 RateEstimation.tex
.tex of Jon's Rate Estimation document

%\documentclass[12pt]{iopart}
\documentclass[prd,aps,eqsecnum]{revtex4}
\usepackage{epsfig,graphicx,latexsym}
\def\rmd{{\rm d}}
\def\msun{$M_{\odot}$}
\def\MNRAS{{\em Mon. Not. Roy. Astron. Soc. }}
\def\CQG{{\em Class. Quantum Grav. }}
\def\PRD{{\em Phys. Rev. D }}
\def\apj{{\em Astrophys. J.}}
\newcommand{\erf}[1]{(\ref{#1})}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\rmd{{\rm d}}

\begin{document}

\title{Rate Estimation}
%\author{Jonathan R Gair}
%\address{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
%\email{jgair@ast.cam.ac.uk}
\author{Will, Jon and Ilya}

\date{\today}

\begin{abstract}
Some notes on the discussion Will and I had today (March 23rd 2012).
\end{abstract}

\maketitle

\section{Rate Estimation}
\subsection{Likelihood}
The data we observe is a sequence of $N$ events with statistic values $x_i$, above some threshold $x_0$. These events could be drawn from a background described by the rate distribution $\nu_b(x)$, which specifies the rate of events arriving with statistic value greater than $x$, or from a foreground described by the rate distribution $\nu_f(x)$. We use flags $f_i \in \{0,1\}$ to denote if an event if is foreground ($f_i=1$) or background ($f_i=0$). We want to use the data to estimate the $f_i$'s and some property of the rate distributions (we'll come back to this later).

The probability of seeing data $x = \{x_i\}$ {\it and} flags $f_i$ given the rates is the probability that we would draw events $\{x_j : f_j=1\}$ from the foreground distribution and the events $\{x_j : f_j=0\}$ from the background distribution. The probability of getting a set of $M$ events $\{y_i\}$ above threshold $y_0$ and with $y_1 < y_2 < \cdots y_M$ from a Poisson process with loudness distribution $\nu(y)$ is the probability of having $0$ events between $y_0$ and $y_1$, which is ${\rm e}^{-(\nu(y_0)-\nu(y_1))}$, times the probability of having an event at $y_1$, which is $-\nu'(y_1)$, times the probability of having no events between $y_1$ and $y_2$, which is ${\rm e}^{-(\nu(y_0)-\nu(y_1))}$, etc. Overall we find
\begin{equation}
p(\{y_i\}) = {\rm e}^{-\nu(y_0)} \prod_{i=1}^M [-\nu'(y_i)].
\end{equation}
Combining the foreground and background distributions we then have
\begin{equation}
p(x, \{f_i\} | \nu_f, \nu_b) = {\rm e}^{-\nu_f(x_0)} {\rm e}^{-\nu_b(x_0)} (-1)^N \prod_{\{j :   f_j=1\}} \nu_f'(x_j)\prod_{\{k : f_k=0\}} \nu_b'(x_k) .
\end{equation}
We want the posterior distribution for the parameters of the rate distributions and the $f_i$'s, which is
\be
p(\nu_f, \nu_b, \{f_i\} | x) \propto p(x, \{f_i\} | \nu_f, \nu_b) p(\nu_f, \nu_b)
\ee
The last term is just the prior probability of the rate parameters.

To relate this to Ilya's parameterisation, we say that the loudness distribution is determined by an overall scale, $\lambda$, and a probability distribution of events $p(x)$ with loudness greater than $x_0$, which we temporarily assume is known. Then we have 
\be
\frac{\rmd \nu}{\rmd x} = - \lambda p(x)
\ee
and $\nu(x_0) = \lambda$. The posterior probability then becomes
\be
p(\lambda_f, \lambda_b, \{f_i\} | x) \propto  p(\lambda_f, \lambda_b) \, {\rm e}^{-\lambda_f} {\rm e}^{-\lambda_b} \lambda_f^{N_f} \lambda_b^{N_b} \prod_{\{j :   f_j=1\}} p_f(x_j)\prod_{\{k : f_k=0\}} p_b(x_k)
\ee
where $N_f/N_b$ are the number of foreground/background events and $p_f/p_b$ are the corresponding loudness probability distributions. This is similar to what Ilya was writing down, except possible for the $\lambda_b$ and $\lambda_f$ dependent terms.

We can marginalise this probability distribution over the $f_i$'s which gives
\be
p(\lambda_f, \lambda_b | x) \propto p(\lambda_f, \lambda_b) \, {\rm e}^{-\lambda_f} {\rm e}^{-\lambda_b}  \prod_{j=1}^N (\lambda_f p_f(x_j) + \lambda_b p_b(x_j))
\ee
or we could marginalise over the rate parameters. This requires a specification of the prior on the rates. As an illustrative example, we can take a flat prior on the rate parameters and then we have
\be
p(\{f_i\} | x) \propto  N_f! N_b! \prod_{\{j :   f_j=1\}} p_f(x_j)\prod_{\{k : f_k=0\}} p_b(x_k).
\ee 
We can go further and work out the marginal distribution of a single $f_i$, but these distributions are not particularly pleasant. We can also look at 2D probabilities of pairs of $f_i$'s and would expect the see correlations. Thus, everything can be written down exactly, but given the number of possible combinations of $f_i$'s that we would have to evaluate, an MCMC will allow us to find the combined distribution more easily.
% e.g., for $f_1$ we have
%\be
%p(f_1=1|x) 
%\ee

\subsection{A few more thoughts}
The framework above allows us to estimate the form of the rate functions as well, i.e., the $p_f(x)$ and $p_b(x)$, if we assume some model for them. If we parameterize $p_f(x)$ by parameters $\{\mu^i_f\}$ and parameterize $p_b(x)$ by parameters $\{\mu^i_b\}$ then the combined posterior becomes
\be
p(\lambda_f, \lambda_b, \{\mu_f^i\}, \{\mu_b^i\}, \{f_i\} | x) \propto  p(\lambda_f, \lambda_b, \{\mu_f^i\}, \{\mu_b^i\}) \, {\rm e}^{-\lambda_f} {\rm e}^{-\lambda_b} \lambda_f^{N_f} \lambda_b^{N_b} \prod_{\{j :   f_j=1\}} p_f(x_j | \mu_f^i)\prod_{\{k : f_k=0\}} p_b(x_k | \mu_b^i) \label{fullpost}
\ee
so we can jointly recover the shape of the loudness distribution and the rate, if we assume some model. I think using the power-law or exponential models would be a good illustration even if it was not realistic for LIGO. However, we could also mimimc what Drew et al. do. If we divide the statistic space up into bins, we can then model $p_f(x)$ to be flat in each of these bins. The number of free parameters, $\{\mu_f^i\}$, is then the number of bins used. This is essentially what they do when they bin the data and look at the distribution of events from injections and background (timeslides).

Going one step further, we can use the injection and timeslide data directly in this framework. We just assume that we have another set of $N_I$ events, $\{ w_i \}$, known to be drawn from the foreground distribution and a set of $N_T$ events $\{z_i\}$ known to be dtawn from the background distribution. These events were not drawn with the correct rate parameters, but a fixed number of events were computed and so they do not contribute to the $\lambda_f$ and $\lambda_b$ terms, but they do contribute an extra factor
\be
\prod_{l=1}^{N_I} p_f(w_l | | \mu_f^i) \prod_{m=1}^{N_T} p_b(z_i | | \mu_b^i) 
\ee
on the right hand side of Eq.~(\ref{fullpost}). So, we could simultaneously fit for the shape of the background as well as the rates and the identification of individual sources. If we were using the binned model for the probability distributions we would presumably find that, if there were many more events in the timeslide and injection data set, this would reduce to the analysis that was described above which assumes the shape of the rate distributions are given by the distribution of those events. However, this framework would be more flexible in the limit where the amount of timeslide or injection data was limited.

Another thought was that the injection distribution might not reflect the astrophysical distribution of signals, in which case we might want to see what happens if we only use timeslide data and ignore injection data. Similarly, in the detection era we might not be able to estimate background very easily since there will be too any sources in the data set (we wish!). In that case maybe we would assume that foreground was reliable and background wasn't. All of this is easy enough within this framework, but we might want to think a bit more.

\end{document}

### 2012-06-08 many.tex
Incorporate notes from RateEstimation.tex into many.tex

@@
 \affiliation{Northwestern University and CIERA}
+
+%\author{Jonathan R Gair}
+%\email{jgair@ast.cam.ac.uk}
+%\affiliation{University of Cambridge}
+%\affiliation{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
 
@@
 detectors).  Dividing Eq.~\eqref{eq:likelihood-rates} by
-$\prod_{\left\{ i | f_i = 0 \right\}} \hat{b}(x_i)$, we obtain
+$\prod_{i=1}^N \hat{b}(x_i)$, we obtain
 \begin{multline}
@@
 
+\section{Estimating the form of the foreground and background distributions}
+We can extend this framework to estimate the form of the rate functions as well, i.e., the $\hat{f}(x)$ and $\hat{b}(x)$, assuming a model.
+If we parameterize $\hat{f}(x)$ by parameters $\{\mu^i_f\}$ and parameterize $\hat{b}(x)$ by parameters $\{\mu^i_b\}$ then the combined posterior becomes
+\begin{eqnarray}
+&&p(R_f, R_b, \{\mu_f^i\}, \{\mu_b^i\}, \{f_i\} | d)  \propto \nonumber \\   && \hspace{2cm} \prod_{\{j :   f_j=1\}} \hat{f}(x_j ; \mu_f^i)\prod_{\{k : f_k=0\}} \hat{b}(x_k ; \mu_b^i) \nonumber \\ &&\hspace{2.25cm}\times \exp\left[-(R_f+R_b)\right]    R_f^{N_f} R_b^{N_b} \nonumber \\ && \hspace{2.5cm}\times p(R_f, R_b, \{\mu_f^i\}, \{\mu_b^i\}) \label{fullpost}
+\end{eqnarray}
+in which $p(R_f, R_b, \{\mu_f^i\}, \{\mu_b^i\})$ is the combined prior on the rate distribution and rate function parameters.
+The model for the rate functions could take a simple analytic form, for instance a power law
+\begin{eqnarray}
+\hat{f}(x; \alpha_f, x_{\rm min}) = \alpha_f \frac{x_{\rm min}^{\alpha_f}}{x^{\alpha_f+1}} \nonumber \\
+\hat{b}(x; \alpha_b, x_{\rm min}) = \alpha_b \frac{x_{\rm min}^{\alpha_b}}{x^{\alpha_b+1}}
+\end{eqnarray}
+or an exponential distribution.
+If the same model is used for both the foreground and background distributions then care must be taken with the priors otherwise it might not be possible to distinguish the foreground and background events.
+
+An alternative, almost model-free approach, would be to divide the statistic space up into bins and use a model that $\hat{f}(x)$ and $\hat{b}(x)$ are constant in each bin.
+The number of free parameters, $\{\mu_f^i\}$, is then the number of bins used.
+
+\subsection{Incorporating timeslide and injection data}
+Detection rates for LIGO are commonly estimated by injecting signals with known parameters into the detector stream.
+False alarm rates are also commonly estimated by using timeslides, in which data from different detectors is offset in time so that all events passing the specified threshold are known to be noise artefacts.
+Injection and timeslide data can also be readily incorporated into this framework.
+We assume that we have another set of $N_I$ events, labelled as $i=N+1, \cdots N+N_I$, that are known to be drawn from the foreground distribution and a set of $N_T$ events, labelled as $i=N+N_I+1, \cdots N+N_I+N_T$, that are known to be drawn from the background distribution.
+These events will not be drawn with the correct rate parameters and so they do not contribute to the determination of $R_f$ and $R_b$ terms, but they do contribute an extra factor
+\begin{equation}
+\prod_{i=N+1}^{N+N_I} \hat{f}(x_i; \mu_f^i) \prod_{m=N+N_I+1}^{N+N_I+N_T} \hat{b}(x_i; \mu_b^i) 
+\end{equation}
+on the right hand side of Eq.~(\ref{fullpost}).
+Given the timeslides and injections we can therefore simultaneously fit for the shape of the background as well as for the rates.
+If we use the binned model for the probability distributions then if there were many more events in the timeslide and injection data set than in the data set being analysed, the analysis would be equivalent to assuming that the shape of the foreground and background distributions was given by the distribution of the statistics for the timeslide and injection events.
+However, this framework is more flexible when the amount of timeslide or injection data is limited.
+
+In practice, the foreground events will not all be identical and so the distribution of statistics for the foreground population will depend on the astrophysical distribution of parameters of the sources.
+If this is not correctly reproduced by the injections, the assumption that the injections represent the same distribution as the foreground in the data is not valid.
+In that case, it might be safer not to use injection information in the analysis.
+
 \begin{acknowledgments}

### 2012-07-31 many.tex
Major edits.
* Updated the derivation of the likelihood, approaching first from the
  normalized foreground and background distributions, and treating the
  Poisson factor for the rates as a sort of ``prior'' for the rate
  flags.

* Added two examples:

  * Fitting the effective number of independent templates from the a
    Gaussian-noise data set as a shape parameter of the background
    distribution.

  * An example of fitting globular cluster parameters from observed
    stars given a non-uniform background.

--- 2012-03-31 many.tex
@@
-\documentclass[aps,prd,reprint]{revtex4-1}
+\documentclass[aps,prd]{revtex4-1}
 
@@
 \newcommand{\fixme}[1]{\textbf{FIXME: #1}}
+\newcommand{\mathset}[1]{\left\{ #1 \right\}}
 
@@
 \email{w-farr@northwestern.edu}
-%\homepage{http://faculty.wcas.northwestern.edu/will-farr/}
-% \affiliation{Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA)\\
-% Department of Physics and Astronomy\\
-% Northwestern University, 2145 Sheridan Road, Evanston, IL 60208}
-\affiliation{Northwestern University and CIERA}
+\homepage{http://faculty.wcas.northwestern.edu/will-farr/}
+\affiliation{Center for Interdisciplinary Exploration and Research in Astrophysics\\
+Department of Physics and Astronomy\\
+Northwestern University, 2145 Sheridan Road, Evanston, IL 60208}
+%\affiliation{Northwestern University and CIERA}
 
@@
 \email{ilya@chgk.info}
-%\homepage{http://www.chgk.info/~ilyamandel/}
-% \affiliation{School of Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT Birmingham\\United Kingdom}
-\affiliation{University of Birmingham}
+\homepage{http://www.chgk.info/~ilyamandel/}
+\affiliation{School of Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT Birmingham\\United Kingdom}
+%\affiliation{University of Birmingham}
 
 \author{Jon Gair}
+\email{jrg23@cam.ac.uk}
 
@@
 \begin{equation}
-  d = \{ x_i | i = 1, \ldots, N \}.
+  d = \mathset{ x_i | i = 1, \ldots, N }
 \end{equation}
@@
 We assume that both the foreground and background events are samples
-from an inhomogeneous Poisson process with rates (per unit ranking
-statistic, $x$)
+from an inhomogeneous Poisson process with differential rates
 \begin{equation}
-  \frac{dN_f}{dx} = f(x)
+  \frac{dN_f}{dx} = f(x, \theta)
 \end{equation}
@@
 \begin{equation}
-  \frac{dN_b}{dx} = b(x).
+  \frac{dN_b}{dx} = b(x, \theta),
 \end{equation}
-The cumulative rates of the two processes are therefore
+where the $\theta$ argument represents additional ``shape'' parameters
+that may affect the distribution, and for which we will eventually
+fit.  The cumulative rates of the two processes are therefore
 \begin{equation}
-  F(x) \equiv \int_{-\infty}^x ds\, f(s)
+  F(x,\theta) \equiv \int_{-\infty}^x ds\, f(s, \theta)
 \end{equation}
@@
 \begin{equation}
-  B(x) \equiv \int_{-\infty}^x ds\, b(s).
+  B(x,\theta) \equiv \int_{-\infty}^x ds\, b(s,\theta).
 \end{equation}
@@
 \begin{enumerate}
-\item The number of events in any range of ranking statistics, $x \in
-  [x_1, x_2]$ is Poisson distributed with rate $F(x_2) - F(x_1)$ or
-  $B(x_2) - B(x_1)$.
+\item \label{prop:Poisson} The number of events in any range of
+  ranking statistics, $x \in [x_1, x_2]$ is Poisson distributed with
+  rate $F(x_2, \theta) - F(x_1,\theta)$ or $B(x_2,\theta) -
+  B(x_1,\theta)$.
 \item The numbers of events in non-overlapping ranges of ranking
@@
   \begin{equation}
-    P(N = 1 \in [x, x+h]) = f(x) h + \order{h^2}.
+    P(N = 1 \in [x, x+h]) = f(x,\theta) h + \order{h^2}.
   \end{equation}
@@
 \begin{equation}
-  f(x) = R_f \hat{f}(x),
+  f(x,\theta) = R_f \hat{f}(x,\theta'),
 \end{equation}
@@
 \begin{equation}
-  b(x) = R_b \hat{b}(x),
+  b(x, \theta) = R_b \hat{b}(x, \theta'),
 \end{equation}
-where $\hat{F}(\infty) = \hat{B}(\infty) = 1$.  Then $R_f$ and $R_b$
-are the total number of foreground and background events expected,
-respectively.  Other parameters may describe the shape of the rate
-functions, but these will depend on the details of the system being
-analyzed.  In \S \ref{sec:GW-example} we give an example of fitting
-such shape paremeters.
+where $\hat{F}(\infty, \theta') = \hat{B}(\infty, \theta') = 1$, and
+$\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f$ and $R_b$
+are the total number of foreground and background events expected and
+$\hat{f}(x, \theta')$ and $\hat{b}(x, \theta')$ are the likelihood of
+obtaining an event with ranking statistic $x$ under the foreground and
+background distributions.  In what follows, we will drop the prime,
+using $\theta$ to denote all parameters of the rate distributions
+except $R_f$ and $R_b$.
 
@@
 are background.  For each event, we introduce a flag, $f_i$, which is
-either 0 or 1, indicating a background or foreground event,
-respectively.  These ``state'' flags are additional (unobserved) data
-in our model.  We can marginalize over our uncertainty in the state of
-any given event by summing all posteriors over $f_i = \{0,1\}$.  Given
-an identification of each event as foreground or background, and the
-rates of each, the likelihood%
-\footnote{See \S \ref{sec:likelihood-derivation} for a derivation of
-  Eq.~\eqref{eq:likelihood}.} %
-of our data is
+either 0 or 1, indicating a background or foreground event.  These
+``state'' flags are parameters in our model, along with $R_f$, $R_b$,
+and $\theta$.  We can marginalize over our uncertainty in the state of
+any given event by summing posteriors over $f_i = \mathset{0,1}$.
+
+Bayes' theorem relates the posterior probability of the state flags,
+rates, and shape parameters, $p\left(\mathset{f_i}, R_f, R_b, \theta |
+  d \right)$, the likelihood of the data, $p\left( d | \mathset{f_i},
+  R_f, R_b, \theta\right)$, and the prior probability of state flags,
+rates and shape parameters before any data are obtained, $p\left(
+  \mathset{f_i}, R_f, R_b, \theta\right)$:
+\begin{equation}
+  \label{eq:Bayes}
+  p\left(\mathset{f_i}, R_f, R_b, \theta | d \right)  \propto
+  p\left( d | \mathset{f_i}, R_f, R_b, \theta\right)
+  p\left(\mathset{f_i}, R_f, R_b, \theta\right). 
+\end{equation}
+The constant of proportionality, called the evidence, $p(d)$, is a
+normalization constant independent of the state flags, rates, and
+shape parameters.
+
+Each foreground event is drawn from the probability distribution
+$\hat{f}$ and each background event is drawn from the probability
+distribution $\hat{b}$.  The events are independent of each other.
+Therefore, the likelihood of the data is 
+\begin{equation}
+  p\left( d | \mathset{f_i}, R_f, R_b, \theta\right)  = \left[
+    \prod_{\mathset{i | f_i = 1}} \hat{f}\left(x_i, \theta\right) \right]
+  \left[ \prod_{\mathset{i | f_i = 0}} \hat{b}\left( x_i, \theta\right)
+  \right].  
+\end{equation}
+The prior distribution can be factorized as
+\begin{equation}
+  \label{eq:combined-flag-rate-prior}
+  p\left(\mathset{f_i}, R_f, R_b, \theta\right)  = p\left(
+    \mathset{f_i} | R_f, R_b, \theta\right)p(R_f, R_b, \theta).
+\end{equation}
+The first term is not acutally a ``prior'' in the usual sense; the
+conditional probability of the $\mathset{f_i}$ on the rates is given
+by
+\begin{equation}
+  \label{eq:flag-conditional-prior}
+  p\left(\mathset{f_i} | R_f, R_b, \theta\right) = \frac{R_f^{N_f}
+    R_b^{N_b}}{N_f! N_b!} \exp\left[ - \left(R_f + R_b\right) \right]
+  N_f! N_b! = R_f^{N_f}R_b^{N_b} \exp\left[ - \left(R_f + R_b\right) \right],
+\end{equation}
+where $N_f$ and $N_b$ are the number of foreground and background
+events.  This expression follows from Property \ref{prop:Poisson} of
+inhomogenous Poisson processes and the $N_f! N_b!$ differnt ways of
+assigning foreground and background flags to events for fixed $N_f$,
+$N_b$.
+
+The second term in Eq.~\eqref{eq:combined-flag-rate-prior} is a
+traditional prior.  Because the rate parameters enter the posterior in
+the same form as Poisson rates, we choose here the Poisson Jeffrey's
+prior on rates, independent of the shape parameters
+\begin{equation}
+  p\left( R_f, R_b, \theta\right) = \frac{1}{\sqrt{R_f R_b}} p(\theta),
+\end{equation}
+but of course other choices are possible.  This choice has the
+advantage that the prior is normalizable as $R_f, R_b \to 0$, and the
+exponentials in Eq.~\eqref{eq:flag-conditional-prior} regularize the
+posterior as $R_f, R_b \to \infty$.
+
+Putting everything together, the posterior is
+\begin{equation}
+  \label{eq:posterior}
+  p\left( \mathset{f_i}, R_f, R_b, \theta | d \right) \propto \left[
+    \prod_{\mathset{i | f_i = 1}} R_f \hat{f}\left(x_i, \theta\right) \right]
+  \left[ \prod_{\mathset{i | f_i = 0}} R_b \hat{b}\left( x_i, \theta\right)
+  \right] \exp\left[ - \left(R_f + R_b\right) \right]
+  \frac{p(\theta)}{\sqrt{R_f R_b}}
+\end{equation}
+Not surprisingly, this posterior equivalent to one derived by assuming
+that the flags, $\mathset{f_i}$, are un-observed data and treating the
+sets $\mathset{x_i | f_i = 1}$ and $\mathset{x_i | f_i = 0}$ as
+samples from an inhomogeneous Poisson process:
 \begin{multline}
-  \label{eq:likelihood}
-  p\left( d, \left\{ f_i \right\} | f(x), b(x)
-  \right) = \\ \left[ \prod_{\left\{ i | f_i = 1 \right\}} f\left( x_i
-    \right) \right]  \left[ \prod_{\left\{ j | f_j = 0 \right\}}
-    b\left( x_j \right) \right] \\
-  \times \exp\left[ - F\left( \infty \right) \right] 
-  \exp\left[ -B\left( \infty \right) \right].
-\end{multline}
-Written in terms of the rate parameters, this becomes
-\begin{multline}
-  \label{eq:likelihood-rates}
-  p\left( d, \left\{ f_i \right\} | f(x), b(x)\right) = \\ R_f^{N_f}
-  \left[ \prod_{\left\{ i | f_i = 1 \right\}}
-    \hat{f}\left( x_i \right) \right] \exp\left[ - R_f \right] \\
-  \times R_b^{N_b} \left[ \prod_{\left\{ j | f_j = 0 \right\}}
-    \hat{b}\left( x_j \right) \right] \exp\left[ - R_b \right],
-\end{multline}
-where $N_f$ is the number of the $f_i$ that are 1 (i.e.\ the number of
-assumed foreground events), and $N_b$ is the number of the $f_i$ that
-are 0 (i.e.\ the assumed number of background events).  Note that $N_f
-+ N_b = N$, as each event is considered either foreground or
-background in our model.
-
-It will occasionally be convenient to work with the ratio,
-$\hat{f}(x)/\hat{b}(x)$ instead of the values of $\hat{f}$ and
-$\hat{b}$ separately.  This ratio can be interpreted as a likelihood
-ratio between the foreground and background models for an event that
-has occurred with ranking statistic $x$:
-\begin{equation}
-  \label{eq:loudness-likelihood}
-  \frac{\hat{f}(x)}{\hat{b}(x)} = \frac{p(x|\mathrm{foreground})}{p(x|\mathrm{background})}.
-\end{equation}
-In some circumstances, it is possible to compute this ratio
-empirically (for example, in \citet{Cannon2012}, the authors give a
-practical method for estimating the distribution of loudnesses for
-known-background events and known-foreground events in a mock data
-stream from the Advanced LIGO and Advanced Virgo gravitational wave
-detectors).  Dividing Eq.~\eqref{eq:likelihood-rates} by
-$\prod_{\left\{ i | f_i = 0 \right\}} \hat{b}(x_i)$, we obtain
-\begin{multline}
-  \label{eq:likelihood-ratio}
-  p\left(d, \left\{ f_i \right\} | f(x), b(x) \right) \propto  \\ \left[
-    \prod_{\left\{ i | f_i = 1 \right\}}
-    \frac{\hat{f}(x_i)}{\hat{b}(x_i)} \right] R_f^{N_f} R_b^{N_b}
-  \exp\left[ - \left( R_f + R_b \right) \right].
+  p\left( \mathset{f_i}, R_f, R_b, \theta | d \right) \propto p\left(
+    d, \mathset{f_i} | R_f, R_b, \theta \right) p\left( R_f, R_b,
+    \theta \right) \\ = \left[ \prod_{\mathset{i|f_i=1}} f\left( x_i,
+      \theta \right) \right] \left[ \prod_{\mathset{i|f_i=0}} b\left(
+      x_i, \theta\right) \right] \exp\left[ -\left( R_f + R_b\right)
+  \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
 \end{multline}
 
-\subsection{Priors}
+We can marginalize the posterior over the flags, $f_i$, obtaining
+\begin{equation}
+  \label{eq:rate-shape-posterior}
+  p\left( R_f, R_b, \theta | d \right) = \sum_{\mathset{f_i} \in
+    \mathset{0,1}^N} p\left( \mathset{f_i}, R_f, R_b, \theta | d \right)
+  \propto \prod_{i} \left[ R_f \hat{f}\left(x_i, \theta\right) + R_b
+    \hat{b}\left( x_i, \theta\right) \right] \exp\left[-\left( R_f +
+      R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
+\end{equation}
+This expression is useful if we are only interested in rates and not
+the probability that any particular event is foreground or background.
+Unlike the full posterior (Eq.~\eqref{eq:posterior}),
+Eq.~\eqref{eq:rate-shape-posterior} contains only continuous
+parameters.
 
-Because the overall rates enter the likelihood in
-Eq.~\eqref{eq:likelihood-rates} in the same form as in the
-constant-rate Poisson likelihood, a reasonable prior for $R_f$ and
-$R_b$ is the Poisson Jeffrey's prior%
-\footnote{Recall that the Jeffrey's prior for a parameter $\theta$
-  with likeilhood function, $p(d|\theta)$, is
+Eq.~\eqref{eq:posterior} is unchanged if the ranking statistic is
+multi-dimensional; in this case, the rates are 
 \begin{equation}
-    p(\theta) \propto \sqrt{\left\langle \left( \frac{\partial
-          p(d | \theta) }{\partial \theta} \right)^2 \right\rangle},
+  R_f = \int d^n x \, f(x, \theta)
 \end{equation}
-  where the expectation is taken over all data, $d$.  The Jeffrey's
-  prior assigns prior weight to locations in parameter space
-  proportional to the Fisher information that is expected to result
-  from a measurement at that parameter value.}, %
+and
 \begin{equation}
-  \label{eq:Jeffrey-prior}
-  p(R) \propto \frac{1}{\sqrt{R}}.
+  R_b = \int d^n x \, b(x, \theta),  
 \end{equation}
-One advantage of the Jeffrey's prior in this circumstance is that it
-has finite probability mass as $R \to 0$, and the likelihood in
-Eq.~\eqref{eq:likelihood-rates} controls its behavior as $R\to \infty$
-to ensure a normalizable posterior.
+where $f$ and $b$ are rate densities on the $n$-dimensional space of
+ranking statistics.  We give an example of fitting for
+multi-dimensional rate densities in \S~\ref{sec:star-cluster}.
 
-Of course, other priors could be used; in particular, previous
-experiments may have placed constraints on the rate that can be used
-as priors in subsequent experiments.  Throughout the remainder of this
-paper, however, we use the Jeffrey's prior,
-Eq.~\eqref{eq:Jeffrey-prior}.
-
-\subsection{Posteriors}
-
-Putting everything together, the posterior probability of each event's
-state and the overall rate is given by
-\begin{multline}
-  p(R_f, R_b, \left\{ f_i \right\} | d) \propto R_f^{N_f}
-  \left[ \prod_{\left\{ i | f_i = 1 \right\}}
-    \hat{f}\left( x_i \right) \right] \exp\left[ - R_f \right] \\
-  \times R_b^{N_b} \left[ \prod_{\left\{ j | f_j = 0 \right\}}
-    \hat{b}\left( x_j \right) \right] \exp\left[ - R_b \right] p(R_f, R_b).
-\end{multline}
-If we marginalize over all the flags, $f_i$, then we obtain
-\begin{multline}
-  \label{eq:rates-marginalized}
-  p(R_f, R_b | d) \propto \prod_{i} \left[ R_f \hat{f}\left( x_i
-    \right) + R_b \hat{b}\left( x_i \right) \right] \\ \times
-  \exp\left[ - \left( R_f + R_b \right) \right] p\left( R_f, R_b
-  \right).
-\end{multline}
-If, instead, we marginalize over all but one flag, $f_k$, we obtain
-that the posterior ratio of signal to noise probability for event $k$
-given the rates $R_f$ and $R_b$ is 
-\begin{equation}
-  \label{eq:foreground-marginalized}
-  \frac{p(R_f, R_b, f_k = 1 | d)}{p(R_f, R_b, f_k = 0 | d)} =
-  \frac{R_f \hat{f}\left( x_k \right)}{R_b \hat{b}\left( x_k
-    \right)}p\left(R_f, R_b\right). 
-\end{equation}
-Note that the \emph{posterior} estimate of the ratio of signal to
-noise probabilities for event $i$ differs from the likelihood ratio,
-Eq.~\eqref{eq:loudness-likelihood}, because it depends on the overall
-rates of foreground and background events.
-
-\section{Gravitational Waves from Compact Binary Inspirals}
+\section{Examples}
 \label{sec:GW-example}
 
-\subsection{Analytic Example}
+In this section we present several examples of the application of our
+framework to various rate estimation problems in the presence of
+background.
+
+\subsection{Gravitational Waves with Non-Overlapping Templates}
 \label{sec:analytic-GW-example}
@@
 Suppose we attempt to detect gravitational wave signals in a data
-stream by matched filtering against a set of $N$ templates.  The data
-stream consists of stationary Gaussian noise with a power spectral
-density $S(f)$ combined additively with some number of signals.  The
-signal to noise ratio (SNR) of a template, $h(f)$, given data, $d(f)$,
-is given by
+stream by matched filtering in the frequency domain against a set of
+$N$ template waveforms \citep[e.g.][]{LVC2011}.  In our simplistic
+model, we suppose the data stream consists of stationary Gaussian
+noise with a power spectral density $S(f)$ combined additively with
+some number of gravitational wave signals.  We assume that the signals
+are sufficiently rare that they do not overlap in the data stream.
+The signal to noise ratio (SNR) of a template, $h(f)$, given data,
+$d(f)$, is
 \begin{equation}
@@
 \begin{equation}
-  \left\langle h_1, h_2 \right\rangle \sim 0.
+  \left\langle h_i, h_j \right\rangle \simeq \delta_{ij}.
 \end{equation}
+In the following subsection, we will generalize the model to
+overlapping templates.  We rank candidate events by their maxmim SNR
+over the entire template bank,
+\begin{equation}
+  x \equiv \max_{h} \rho_h,
+\end{equation}
+and consider only events that have a maximum SNR above some threshold,
+$x > \xmin$.
 
 For a data stream of pure noise, $d(f) = n(f)$, the SNR of a each
-template follows a $N(0,1)$ distribution.  If our ranking statistic
-for events in each interval of data is their maximum SNR over the $N$
-templates, 
-\begin{equation}
-  x \equiv \max_h \rho_h,
-\end{equation}
-then the background distribution is
+template follows a $N(0,1)$ distribution.  The background ranking
+statistic (i.e.\ the maximum SNR over the template bank) then has a
+cumulative distribution without thresholding of
 \begin{equation}
@@
 \end{equation}
-If, in addition, we impose a threshold ranking statistic, $\xmin$, to
-be considered an ``event,'' then $\hat{B}$ becomes
+Imposing the threshold, $x > \xmin$, the
+cumulative distribution of the background becomes
 \begin{equation}
@@
 The SNR of a gravitational wave signal in an interferometric detector
-scales as $1/d$, where $d$ is the distance to the source.  Ignoring
-possible cosmological effects, the number of sources will scale as
-$d^3$.  Thus, we expect that the foreground cumulative distribution of
-events will follow
+scales as $1/d$ \citep{Finn1992}, where $d$ is the distance to the
+source.  Ignoring cosmological effects, the number of sources scales
+as $d^3$.  Thus, we expect that the foreground cumulative distribution
+of events will follow
 \begin{equation}
@@
 
+Note that this scenario has no shape parameters for the foreground and
+background distributions.
+
 To demonstrate the effectiveness of our formalism, we applied it to a
@@
 \eqref{eq:analytic-foreground-rate} with $R_f^\mathrm{true} = 10.4$
-and $R_b^\mathrm{true} = 95.1$.  The cumulative distribution for the
-ranking statistic of the synthetic data we used appears in Figure
-\ref{fig:analytic-data-cumulative}.
+and $R_b^\mathrm{true} = 95.1$, using 1000 templates.  The synthetic
+data consisted of 13 foreground events and 85 background events; the
+cumulative distribution for the ranking statistic of the synthetic
+data appears in Figure \ref{fig:analytic-data-cumulative}.  We used a
+Markov Chain Monte Carlo simulation to draw samples of state flags and
+rates from the joint posterior (Eq.~\eqref{eq:posterior}).
 
@@
 posterior densities for the foreground and background rates.  (Refer
-to Eq.~\eqref{eq:rates-marginalized}.)  Figure
+to Eq.~\eqref{eq:rate-shape-posterior}.)  Figure
 \ref{fig:analytic-rate-foreground-probs} shows the posterior
 foreground probability for each event marginalized over all other
-events' types and the foreground and background rates.  (Refer to
-Eq.~\eqref{eq:foreground-marginalized}.)
+events' types and the foreground and background rates.
 
@@
   \includegraphics[width=\columnwidth]{rates}
-  \caption{\label{fig:analytic-rate-recovery} The posterior densities
-    for $R_f$ (in blue) and $R_b$ (in black) for the analytic model
-    discussed in \S \ref{sec:analytic-GW-example}.  The vertical lines
-    indicate the ``true'' used to generate the synthetic data set.}
+  \caption{\label{fig:analytic-rate-recovery} The marginalized
+    posterior densities for $R_f$ (solid line) and $R_b$ (dashed line)
+    for the analytic model discussed in \S
+    \ref{sec:analytic-GW-example}.  The vertical lines indicate the
+    ``true'' values used to generate the synthetic data set.}
 \end{figure}
@@
     in red.  The solid line is the likelihood ratio
-    $p(x|\mathrm{foreground})/p(x|\mathrm{background})$ (see
-    Eq.~\eqref{eq:loudness-likelihood}) for this model; this exceeds
-    the marginalized foreground probability for many of the events
-    because in this data set there are approximately nine times as
-    many background events as foreground events (see
-    Eq.~\eqref{eq:foreground-marginalized}). }
+    $p(x|\mathrm{foreground})/p(x|\mathrm{background}) =
+    \hat{f}(x)/\hat{b}(x)$ for this model; this exceeds the
+    marginalized foreground probability for many of the events because
+    in this data set there are approximately nine times as many
+    background events as foreground events. }
 \end{figure}
 
+\subsection{Gravitational Waves With Overlapping Templates}
+\label{sec:gw-overlapping-template}
+
+In \S \ref{sec:analytic-GW-example} we assumed that the overlap
+between templates in the template bank vanished, so the SNR's in
+different templates are independent random variables.  In fact,
+template banks are not constructed in this way
+\citep{Caudill2011,Cannon2011}, because signals can fall in the gaps
+between the non-overlapping templates.  We can model this effect by
+assuming that a template bank of $N$ actual templates will behave as
+if it had $N_\mathrm{eff}$ \emph{independent} templates, and fit for
+the shape parameter $N_\mathrm{eff}$.  That is, we assume that
+$N_\mathrm{eff}$ is a shape parameter for the background cumulative
+distribution:
+\begin{equation}
+  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left( \frac{x}{2} \right)
+    \right)^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
+    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
+    \right)^{N_\mathrm{eff}} }.
+\end{equation}
+
+Results from such an analysis appear in Figures \ref{fig:rates-nt} and
+\ref{fig:ntemplates}.  We use the same parameters and data set as in
+\S \ref{sec:analytic-GW-example}, with $R_f = 10.4$, $R_b = 95.1$, and
+$N_\mathrm{eff} = 1000$, but now allow $N_\mathrm{eff}$ to be a
+parameter of the background distribution, with a flat prior.  Both the
+rates and the number of effective templates are recovered without
+significant loss of accuracy relative to the fixed $N_\mathrm{eff}$
+situation in \S \ref{sec:analytic-GW-example}.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{rates-nt}
+  \caption{\label{fig:rates-nt} The foreground (solid lines) and
+    background (dashed lines) rate posterior, marginalized over all
+    flags and the $N_\mathrm{eff}$ parameter, for the gravitational
+    wave template detection scenario with overlapping templates
+    discussed in \S \ref{sec:gw-overlapping-template}.  The true
+    values of the rates, $R_f = 10.4$ and $R_b=95.1$, are indicated
+    with vertical lines.  The distributions are not significantly
+    wider than those of Figure \ref{fig:analytic-rate-recovery}, in
+    spite of the extra parameter.}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{ntemplates}
+  \caption{\label{fig:ntemplates} The posterior on the number of
+    effective templates, $N_\mathrm{eff}$, for the model and data
+    discussed in \S \ref{sec:gw-overlapping-template}, marginalized
+    over all state flags and rates.  The true value, $N_\mathrm{eff} =
+    1000$, is indicated by the vertical line.}
+\end{figure}
+
+\subsection{Star Cluster Parameters With Background Contamination}
+\label{sec:star-cluster}
+
+Our final example concerns fitting for the location and shape
+parameters of a cluster of stars observed on top of a stellar
+background with a density gradient.  We assume that a star cluster has
+a Plummer surface-density profile \citep{Plummer1911,Aarseth1974},
+\begin{equation}
+  \label{eq:plummer-surface-density}
+  \hat{f}(\vec{x}, \theta) = \frac{1}{\pi r_0^2 \left( 1 + \frac{\left|
+          \vec{x} - \vec{x}_0 \right|^2}{r_0^2} \right)^2},
+\end{equation}
+where $\vec{x}_0$ is the location on the sky of the center of the
+cluster, and $r_0$ is a radial scale parameter.  We assume a square
+observational domain\footnote{The observational domain is not
+  infinite, so the normalization of the cluster density in
+  Eq.~\eqref{eq:plummer-surface-density} is not quite correct.  In our
+  modeling we properly take this into account, but for simplicity here
+  we ignore it.}, $\vec{x} \in [0,1]^2$, and a background that has a
+density gradient at an arbitrary orientation with respect to the
+observational axes:
+\begin{equation}
+  \hat{b}\left(\vec{x}, \theta\right) = 1 + \vec{\gamma} \cdot \left( \vec{x}
+    - \vec{x}_{1/2} \right),
+\end{equation}
+where $\vec{\gamma}$ is the gradient, and $\vec{x}_{1/2} = [1/2, 1/2]$
+is the centroid of the observational domain.  We use assume a gradient
+for the background of $\vec{\gamma} = [-0.5, 0.5]$.  We assume that
+the total number of background stars, $R_b = 10000$, exceeds by a
+factor of ten the total number of cluster stars, $R_f = 1000$.  We
+choose a true cluster center $\vec{x}_0 = \vec{x}_{1/2}$, and a radial
+scale parameter of $r_0 = 0.1$.  In all, the shape parameters for the
+true distributions are
+\begin{equation}
+\label{eq:true-cluster-parameters}
+\theta_0 \equiv \left\{ x_0, y_0, r_0, \gamma_x, \gamma_y \right\} =
+\left\{ \frac{1}{2}, \frac{1}{2}, \frac{1}{10}, -\frac{1}{2},
+  \frac{1}{2} \right\}.
+\end{equation}
+Figure \ref{fig:sky-density} shows the density on the sky due to the
+cluster foreground and the background with these parameters.  We drew
+a synthetic data set from this density according to the corresponding
+inhomogeneous Poisson process.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{sky-density}
+  \caption{\label{fig:sky-density} The density contours on the sky of
+    the forergound cluster and background discussed in \S
+    \ref{sec:star-cluster}, with true parameters given in
+    Eq.~\eqref{eq:true-cluster-parameters}.  There are a factor of 10
+    more stars (10000) in the background than in the cluster (1000),
+    but the peak density for the cluster is much higher.}
+\end{figure}
+
+To analyize our synthetic data set, we analytically marginalized over
+the state flags, using the likelihood in
+Eq.~\eqref{eq:rate-shape-posterior}.  We did this to take advantage of
+the \textit{emcee} sampler of \citet{ForemanMackey2012}, which
+requires all parameters to be in $\mathbb{R}$.  We applied a prior on
+the shape parameters that is flat in $\vec{x}_0$ and $\vec{\gamma}$,
+and the Jeffrey's prior on $r_0$,
+\begin{equation}
+  p\left( r_0 \right) = \frac{\sqrt{R_f}}{r_0}.
+\end{equation}
+(Note that this factor of $\sqrt{R_f}$ cancels with the Jeffrey's
+prior on the rate, $1/\sqrt{R_f}$; we have verified that the priors on
+these parameters are irrelevant to our results, as would be expected
+from the measurement of $\sim 1000$ foreground stars.)
+
+Figure \ref{fig:sky-loc} shows the posterior for the location
+parameters, $\vec{x}_0$; the center of the cluster is localized to
+within a few percent of its scale.  Figure \ref{fig:cluster-number}
+shows the posteriors inferred on the cluster and background numbers,
+$R_f$ and $R_b$, and Figure \ref{fig:cluster-scale} shows the
+posterior for the cluster's scale parameter.  In spite of the
+significant background, the cluster scale and total number are
+accurately recoverd by our analysis.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{sky_location}
+  \caption{\label{fig:sky-loc} Contours of the posterior probability
+    distribution for the center of the cluster, $\vec{x}_0$, in the
+    example from \S~\ref{sec:star-cluster}.  The center $(x,y) =
+    \left(x_0, y_0\right)$ is determined to within a few percent of
+    the structural radius of the cluster, $r_0$ (see
+    Eq.~\eqref{eq:true-cluster-parameters}).}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{numbers}
+  \caption{\label{fig:cluster-number} Posterior densities for the
+    number of stars in the cluster ($R_f$) and in the field ($R_b$) in
+    the example from \S~\ref{sec:star-cluster}.  Vertical lines
+    indicate the true values (see
+    Eq.~\eqref{eq:true-cluster-parameters}). }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{scale}
+  \caption{\label{fig:cluster-scale} Posterior density for the scale
+    parameter for the cluster, $r_0$, from the example in
+    \S~\ref{sec:star-cluster}.  The true value is indicated by the
+    vertical line (see Eq.~\ref{eq:true-cluster-parameters}).}
+\end{figure}
+
+\section{Conclusion}
+We conclude.
+
 \begin{acknowledgments}
-  Kipp Cannon, Chad Hanna, and Drew Keppel for providing the data set
-  from \citet{Cannon2012} and assisting us in its use. Richard
-  O'Shaughnessy for discussions.
+  We thank Kipp Cannon, Chad Hanna, Drew Keppel, and Richard
+  O'Shaughnessy for discussions and suggestions about this manuscript.
 \end{acknowledgments}
-
-\appendix
-
-\section{Likelihood for Inhomogeneous Poisson Processes}
-\label{sec:likelihood-derivation}
-
-\fixme{To do.}
 

### 2012-08-01 many.tex
Abstract improvements; updated to match current paper.

@@
   We show how to obtain a Bayesian estimate of the rate of signal
-  events from a set of signal and background events indexed by a
-  ranking statistic when the shapes of the signal and background
-  distributions are known, can be estimated, or approximated.  We
-  focus on the specific application of estimating astrophysical rates
-  of the coalescence of compact binary black holes or neutron stars
-  from a set of triggers in the LIGO/Virgo gravitational wave
-  detectors, but our framework is fully general.  We discuss the
-  systematic effects on the rate estimate due to differences between
-  the assumed and true shapes of the foreground or background
-  distributions, identifying ways these effects can be minimized.
-  Similarly, we discuss the effects of various priors on the rate,
-  including uninformative priors, weakly-informative priors, and the
-  use of priors from previous rate experiments.  In the limit where
-  the expected signal rate gives high probability of zero or one
-  signal in the data, our technique reduces to the ``loudest event
-  statistic,'' but it is generally applicable to arbitrarily large
-  signal rates.
+  events from a set of signal and background events when the shapes of
+  the signal and background distributions are known, can be estimated,
+  or approximated; our method works well even if the foreground and
+  background event distributions overlap significantly.  We give
+  examples of determining the rates of gravitaitonal wave events in
+  the presence of background triggers from a template bank when noise
+  parameters are known and/or can be fit from the trigger data.  We
+  also give an example of determining globular cluster shape and
+  location parameters from an observation of a stellar field that
+  contains a non-uniform background density of stars superimposed on
+  the cluster stars.
 \end{abstract}

### 2012-08-09 many.tex
Added comments on Section II.

@@
 \usepackage{graphicx}
+\usepackage{color}
 
@@
 \newcommand{\xmin}{x_\mathrm{min}}
+
+\newcommand{\ilya}[1]{{\color{red} \bf #1}}
+\newcommand{\will}[1]{{\color{blue} \bf #1}}
+\newcommand{\jon}[1]{{\color{orange} \bf  #1}}
 
@@
 \author{Ilya Mandel}
-\email{ilya@chgk.info}
-\homepage{http://www.chgk.info/~ilyamandel/}
+\email{imandel@star.sr.bham.ac.uk}
+\homepage{http://www.star.sr.bham.ac.uk/~imandel}
 \affiliation{School of Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT Birmingham\\United Kingdom}
@@
   background event distributions overlap significantly.  We give
-  examples of determining the rates of gravitaitonal wave events in
+  examples of determining the rates of gravitatonal-wave events in
   the presence of background triggers from a template bank when noise
   parameters are known and/or can be fit from the trigger data.  We
-  also give an example of determining globular cluster shape and
+  also give an example of determining globular-cluster shape and
   location parameters from an observation of a stellar field that
@@
 We assume that both the foreground and background events are samples
-from an inhomogeneous Poisson process with differential rates
+from an inhomogeneous \ilya{[pardon my ignorance, what does "inhomogeneous" refer to in this context (i.e., what would a homogeneous Poisson process be?)]} Poisson process with \ilya{respective} differential rates
 \begin{equation}
@@
   \begin{equation}
-    P(N = 1 \in [x, x+h]) = f(x,\theta) h + \order{h^2}.
+    P(\ilya{n} = 1 \in [x, x+h]) = f(x,\theta) h + \order{h^2}.
   \end{equation}
@@
 \item The probability of two or more events in a small range of
-  ranking statistic is negligable
+  ranking statistic is negligible
   \begin{equation}
-    P(N = 2 \in [x, x+h]) = \order{h^2}.
+    P(\ilya{n} = 2 \in [x, x+h]) = \order{h^2}.
   \end{equation}
@@
 where $\hat{F}(\infty, \theta') = \hat{B}(\infty, \theta') = 1$, and
-$\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f$ and $R_b$
+$\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f\ilya{\equiv F(\infty,\theta)}$ and $R_b\ilya{\equiv B(\infty,\theta)}$
 are the total number of foreground and background events expected and
@@
 are background.  For each event, we introduce a flag, $f_i$, which is
-either 0 or 1, indicating a background or foreground event.  These
-``state'' flags are parameters in our model, along with $R_f$, $R_b$,
-and $\theta$.  We can marginalize over our uncertainty in the state of
+either 0 \ilya{(background)} or 1 \ilya{(foreground)}.
+These ``state'' flags are parameters in our model, along with $R_f$, $R_b$, and $\theta$.
+We can marginalize over our uncertainty in the state of
 any given event by summing posteriors over $f_i = \mathset{0,1}$.
@@
   \label{eq:Bayes}
-  p\left(\mathset{f_i}, R_f, R_b, \theta | d \right)  \propto
-  p\left( d | \mathset{f_i}, R_f, R_b, \theta\right)
-  p\left(\mathset{f_i}, R_f, R_b, \theta\right). 
+  p\left(\mathset{f_i}, R_f, R_b, \theta | d \right)  \ilya{=}
+  \frac{p\left( d | \mathset{f_i}, R_f, R_b, \theta\right)
+  p\left(\mathset{f_i}, R_f, R_b, \theta\right)}{\ilya{p(d)}}. 
 \end{equation}
-The constant of proportionality, called the evidence, $p(d)$, is a
-normalization constant independent of the state flags, rates, and
-shape parameters.
+The \ilya{normalization constant}, called the evidence, $p(d)$, is independent of the state flags, rates, and shape parameters.
 
@@
 \end{equation}
-The first term is not acutally a ``prior'' in the usual sense; the
-conditional probability of the $\mathset{f_i}$ on the rates is given
+The first term is not actually a ``prior'' in the usual sense; the
+conditional probability of the flags $\mathset{f_i}$ on the rates is given
 by
@@
 events.  This expression follows from Property \ref{prop:Poisson} of
-inhomogenous Poisson processes and the $N_f! N_b!$ differnt ways of
+inhomogeneous Poisson processes and the $N_f! N_b!$ different ways of
 assigning foreground and background flags to events for fixed $N_f$,
 $N_b$.
+
+\ilya{[I agree with the final result, but I find this way of deriving it to be confusing.
+First of all, the number of ways to assign flags to events for fixed $N_f$ and $N_b$ is $N!/(N_f!
+N_b!)$, not $N_f!
+N_b!$ as written.
+Secondly, you are computing the probability of having a particular ordered list of flags, so there's no reason to consider different permutations.
+Thirdly, you are not properly including the fact that the length of the flags vector must be exactly $N$, the total number of data points.
+(Well, you have the right result, so you are obviously doing these things properly, but the explanation just doesn't seem clear to me.
+:) )  I would make that $N$ explicit, by writing Eq.~(12) as
+
+$$
+  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right)  = 
+  p\left( \mathset{f_i} | N, R_f, R_b\right) p\left(N |R_f, R_b\right)  
+ p\left(R_f, R_b, \theta\right).
+$$
+
+Then 
+$$ p\left( \mathset{f_i} | N, R_f, R_b\right) = 
+\prod_{\mathset{i|f_i=1}} \left(\frac{R_f}{R_f+R_b}\right) 
+\prod_{\mathset{i|f_i=0}} \left(\frac{R_b}{R_f+R_b}\right) =
+\left(\frac{R_f}{R_f+R_b}\right)^{N_f} \left(\frac{R_b}{R_f+R_b}\right)^{N_b}. 
+$$
+Note that there are no permutations involved here -- once we assume that there are precisely $N=N_f+N_b$ events, the probability for the i'th event to have flag $f_i$ is either $R_f/(R_f+R_b)$ or $R_b/(R_f+R_b)$ depending on whether $f_i$ is 1 or 0.
+Meanwhile,
+$$
+p\left(N |R_f, R_b\right) = \frac{\left(R_f+R_b\right)^N}{N!} e^{-(R_f+R_b)},
+$$
+since the distribution of total event number is a Poisson process with rate $R_f+R_b$.
+Multiplying these, we get the same result as in (13), but in a way that even I can understand. :)
+]}
 
@@
 the same form as Poisson rates, we choose here the Poisson Jeffrey's
-prior on rates, independent of the shape parameters
+prior on rates \ilya{[citation?]}, independent of the shape parameters
 \begin{equation}
-  p\left( R_f, R_b, \theta\right) = \frac{1}{\sqrt{R_f R_b}} p(\theta),
+  p\left( R_f, R_b, \theta\right) \ilya{\propto} \frac{1}{\sqrt{R_f R_b}} p(\theta),
 \end{equation}
@@
 \end{equation}
-Not surprisingly, this posterior equivalent to one derived by assuming
+Not surprisingly, this posterior is equivalent to one derived by assuming
 that the flags, $\mathset{f_i}$, are un-observed data and treating the
@@
 \end{multline}
+
+\ilya{[I can't derive the preceding equation in 1 line.
+;)  So I'd suggest either making the alternative derivation more explicit, or, if you don't think it's worth the space, just saying that one could derive the same thing in this alternative way -- otherwise, just repeating Eq.~(15) doesn't seem very useful.]}
 
@@
 \begin{equation}
-  R_f = \int d^n x \, f(x, \theta)
+  R_f = \int d^\ilya{k} \ilya{\vec{x}} \, f(x, \theta)
 \end{equation}
@@
 \begin{equation}
-  R_b = \int d^n x \, b(x, \theta),  
+  R_b = \int d^\ilya{k} \ilya{\vec{x}} \, b(x, \theta),  
 \end{equation}
-where $f$ and $b$ are rate densities on the $n$-dimensional space of
+where $f$ and $b$ are rate densities on the \ilya{$k$}-dimensional space of
 ranking statistics.  We give an example of fitting for

### 2012-08-09 many.tex
Fixed URL typo.

@@
 \email{imandel@star.sr.bham.ac.uk}
-\homepage{http://www.star.sr.bham.ac.uk/~imandel}
+\homepage{http://www.sr.bham.ac.uk/~imandel}
 \affiliation{School of Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT Birmingham\\United Kingdom}
@@
 We assume that both the foreground and background events are samples
-from an inhomogeneous \ilya{[pardon my ignorance, what does "inhomogeneous" refer to in this context (i.e., what would a homogeneous Poisson process be?)]} Poisson process with \ilya{respective} differential rates
+from an inhomogeneous Poisson process \ilya{[pardon my ignorance, but I thought an "inhomogeneous" Poisson process meant that the rate changed as a function of time -- and I don't think that's what we are assuming...]} with \ilya{respective} differential rates
 \begin{equation}

### 2012-08-11 many.tex
Added Jeffreys1946 citation.

@@
 the same form as Poisson rates, we choose here the Poisson Jeffrey's
-prior on rates \ilya{[citation?]}, independent of the shape parameters
+prior on rates \citep{Jeffreys1946}, independent of the shape
+parameters
 \begin{equation}

### 2012-08-11 many.tex
Added derivation of likelihood for inhomogeneous Poisson process.

@@
 \end{equation}
-Not surprisingly, this posterior is equivalent to one derived by assuming
-that the flags, $\mathset{f_i}$, are un-observed data and treating the
-sets $\mathset{x_i | f_i = 1}$ and $\mathset{x_i | f_i = 0}$ as
-samples from an inhomogeneous Poisson process:
+Not surprisingly, this posterior is equivalent to one derived by
+assuming that the flags, $\mathset{f_i}$, are un-observed data and
+treating the sets $\mathset{x_i | f_i = 1}$ and $\mathset{x_i | f_i =
+  0}$ as samples from an inhomogeneous Poisson process.  For an
+inhomogeneous Poisson process with rate function $r(y)$ (cumulative
+rate $R(y)$), the likelihood of a set of samples $\mathset{y_i}$ is
+given by 
 \begin{multline}
-  p\left( \mathset{f_i}, R_f, R_b, \theta | d \right) \propto p\left(
-    d, \mathset{f_i} | R_f, R_b, \theta \right) p\left( R_f, R_b,
-    \theta \right) \\ = \left[ \prod_{\mathset{i|f_i=1}} f\left( x_i,
-      \theta \right) \right] \left[ \prod_{\mathset{i|f_i=0}} b\left(
-      x_i, \theta\right) \right] \exp\left[ -\left( R_f + R_b\right)
-  \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
+  p\left( \mathset{y_i} | r \right) = \lim_{\epsilon \to 0} P\left(
+  \textnormal{zero events below } y_0 - \epsilon \right) \times
+  P\left( \textnormal{one event between } y_0 - \epsilon \textnormal{
+    and } y_0 + \epsilon \right) \\ \times P\left( \textnormal{zero
+    events between } y_0+\epsilon \textnormal{ and } y_1 - \epsilon
+  \right) \ldots \\ = \lim_{\epsilon \to 0} \exp\left[ - R\left(y_0 -
+    \epsilon \right) \right] \left[ r\left( y_0 \right) +
+    \order{\epsilon^2} \right] \times \exp\left[ - \left[ R\left( y_1
+      - \epsilon\right) - R\left( y_0 + \epsilon\right) \right]
+    \right] \times \ldots = \left[\prod_i r\left( y_i
+    \right)\right] \exp\left[ - R\left( \infty \right) \right].
 \end{multline}
-
-\ilya{[I can't derive the preceding equation in 1 line.
-;)  So I'd suggest either making the alternative derivation more explicit, or, if you don't think it's worth the space, just saying that one could derive the same thing in this alternative way -- otherwise, just repeating Eq.~(15) doesn't seem very useful.]}
+Applying this likelihood once to the foreground samples and once to
+the background samples reproduces Eq.~\eqref{eq:posterior}.
 

### 2012-08-12 many.tex
Added comments on rest of paper.

@@
   \cite{Biswas2009}), Bayesian inference.}
+
+Difficulty of measuring rate in presence of background
+
+Want to add more than just the loudest event (cf. loudest event statistic)
+
+Unknown background distribution, $\Lambda$
+
+Cf. number of gold-plated detections in sensitive volume -- frequentist approach OK if tons of events, otherwise need this.
 
@@
 traditional prior.  Because the rate parameters enter the posterior in
-the same form as Poisson rates, we choose here the Poisson Jeffrey's
+the same form as Poisson rates, we choose here the Poisson Jeffreys
 prior on rates \citep{Jeffreys1946}, independent of the shape
@@
 stream by matched filtering in the frequency domain against a set of
-$N$ template waveforms \citep[e.g.][]{LVC2011}.  In our simplistic
+$N$ template waveforms \citep[e.g.,][]{findchirppaper,LVC2011}.  In our simplistic
 model, we suppose the data stream consists of stationary Gaussian
@@
 \begin{equation}
-  \left\langle a, b \right\rangle \equiv \int_0^\infty df\,
+  \left\langle a, b \right\rangle \equiv \ilya{4 \Re} \int_0^\infty df\,
   \frac{a^*(f) b(f)}{S(f)}.
@@
 \end{equation}
+\ilya{for $x>\xmin$, $0$ otherwise.}
 
@@
 
-Note that this scenario has no shape parameters for the foreground and
+Note that this scenario has no shape parameters $\theta$ for the foreground and
 background distributions.
@@
 \eqref{eq:analytic-foreground-rate} with $R_f^\mathrm{true} = 10.4$
-and $R_b^\mathrm{true} = 95.1$, using 1000 templates.  The synthetic
+and $R_b^\mathrm{true} = 95.1$, using 1000 templates.  \ilya{[Specify $\xmin$?]} The synthetic
 data consisted of 13 foreground events and 85 background events; the
@@
     distribution of the ranking statistics for the synthetic data used
-    to test the the formalism on the model from \S
+    to test the formalism on the model from \S
     \ref{sec:analytic-GW-example}.  The solid line gives the
@@
     \eqref{eq:analytic-foreground-rate} combined with $R_f = 10.4$ and
-    $R_b = 95.1$.}
+    $R_b = 95.1$. \ilya{[I think both axis labels should say $x$ rather than $\rho$, since that's the ranking statistic ($\max \rho$).]}}
 \end{figure}
@@
 events' types and the foreground and background rates.
+
+\ilya{[General comment on the text: I think we need to say a bit more about the figures and the interpretation of the results, instead of just relying on the reader to make the best of it.
+For example, even though it may seem obvious to us, we should explicitly say that Fig.~2 shows that we succeed in recovering rates within the expected uncertainty envelope.
+On Fig.~3, could say something obvious like point out that we cannot separate background from foreground for individual events, but that doesn't necessarily matter for rate measurements.
+Now, one interesting question that Fig.~3 brings to mind is the following: if we only used "gold-plated" events (say, those with $p(x|fore)/p(x|back)>0.99$, how much worse / more uncertain would the rate estimation be?]}
 
@@
     \ref{sec:analytic-GW-example} marginalized over all other
-    parameters.  True foreground events are in blue, background events
-    in red.  The solid line is the likelihood ratio
+    parameters.  True foreground events are in blue \ilya{(dark grey)}, background events
+    in red \ilya{(light grey)}.  The solid line is the likelihood ratio
     $p(x|\mathrm{foreground})/p(x|\mathrm{background}) =
-    \hat{f}(x)/\hat{b}(x)$ for this model; this exceeds the
+    \hat{f}(x)/\hat{b}(x)$ for this model \ilya{evaluated at the simulated parameters (?)}; this exceeds the
     marginalized foreground probability for many of the events because
     in this data set there are approximately nine times as many
-    background events as foreground events. }
+    background events as foreground events.
+\ilya{[I'm not sure what the reader is meant to take away from the solid lines  If there's something worth taking away, we should describe the message in the text.
+Otherwise, if it only confuses things, we should drop it.]}}
 \end{figure}
@@
 In \S \ref{sec:analytic-GW-example} we assumed that the overlap
-between templates in the template bank vanished, so the SNR's in
+between \ilya{different} templates in the template bank \ilya{was negligible}, so the SNRs \ilya{recovered by}
 different templates are independent random variables.  In fact,
-template banks are not constructed in this way
-\citep{Caudill2011,Cannon2011}, because signals can fall in the gaps
+template banks are not constructed in this way \citep[e.g.,][]{Owen:1998dk,Ajith:2008},
+%\citep{Caudill2011,Cannon2011}, 
+because signals could fall in the gaps
 between the non-overlapping templates.  We can model this effect by
 assuming that a template bank of $N$ actual templates will behave as
-if it had $N_\mathrm{eff}$ \emph{independent} templates, and fit for
-the shape parameter $N_\mathrm{eff}$.  That is, we assume that
-$N_\mathrm{eff}$ is a shape parameter for the background cumulative
+if it had $N_\mathrm{eff}$ \emph{independent} templates.  \ilya{Rather than pre-computing $N_\mathrm{eff}$,} we can fit for \ilya{it as a} shape parameter.  That is, we assume that
+\ilya{$\theta = \{N_\mathrm{eff}\}$} is a shape parameter for the background cumulative
 distribution:
@@
 
+\ilya{[Again, I think we need to explicitly compare this method with two alternatives.
+Alternative one is the loudest event statistic.
+Alternative two is just counting gold-plated events, and dividing the volume in which we are sensitive to them by their number.
+In principle, alternative one should be suboptimal when the number of events is more than 0 or 1, while alternative two should be OK when the number of events is large, but error-prone otherwise.
+But if we want people to use this technique, we have to defend the point that there is a regime where it is clearly superior (and say what that regime is).]}
+
 \begin{figure}
@@
     wider than those of Figure \ref{fig:analytic-rate-recovery}, in
-    spite of the extra parameter.}
+    spite of the extra parameter.  \ilya{[$R_f$ seems to be more of an outlier than before; why was there a significant shift, given that $N_\mathrm{eff}$ seems fairly spot on?]}}
 \end{figure}
@@
 where $\vec{\gamma}$ is the gradient, and $\vec{x}_{1/2} = [1/2, 1/2]$
-is the centroid of the observational domain.  We use assume a gradient
-for the background of $\vec{\gamma} = [-0.5, 0.5]$.  We assume that
+is the centroid of the observational domain.
+\ilya{[I think that if the domain is finite, the previous equation is only normalized for all $\vec{\gamma}$ if $\vec{x}_{1/2}$ is in the center of a symmetric domain (as it happens to be in this case).]}
+We use assume a gradient
+for the background of $\vec{\gamma} = [-0.5, 0.5]$. 
+\ilya{[I find it a bit confusing to use the word "assume" for both the model and for the specific parameters injected into the simulation (but then being fit for, so their values are not "assumed" in the recovery).
+Maybe "assume" for the model, and "simulate" for the parameter values to be fit?]}
+We assume that
 the total number of background stars, $R_b = 10000$, exceeds by a
-factor of ten the total number of cluster stars, $R_f = 1000$.  We
-choose a true cluster center $\vec{x}_0 = \vec{x}_{1/2}$, and a radial
+factor of ten the total number of cluster stars, $R_f = 1000$.  
+\ilya{[But the cluster has only $\sim 0.01$ of the total area, so it's over-dense by a factor of $10$ or more.
+Fig.~6 shows a very bright spot where the cluster is, which sort of makes the problem look too simple for the chosen parameters...]}
+We choose a true cluster center $\vec{x}_0 = \vec{x}_{1/2}$, and a radial
 scale parameter of $r_0 = 0.1$.  In all, the shape parameters for the
@@
     more stars (10000) in the background than in the cluster (1000),
-    but the peak density for the cluster is much higher.}
+    but the peak density for the cluster is much higher.
+\ilya{[It seems too obvious that where the cluster is by eye, and probably fairly straightforward to compute the number of "excess" stars there...
+Reader might ask why we need something so complicated when the answer is seems so obvious and a "frequentist" approach would seem to do here.]}}
 \end{figure}
@@
 To analyize our synthetic data set, we analytically marginalized over
-the state flags, using the likelihood in
-Eq.~\eqref{eq:rate-shape-posterior}.  We did this to take advantage of
+the state \ilya{[this is the only time we call them "state" flags...]} flags, using the likelihood in
+Eq.~\eqref{eq:rate-shape-posterior}. 
+\ilya{[I think it would be kind to the reader to be more explicit about the connection with the formalism in Section II.
+What are the "events"?
+What is the ranking statistic $x$?
+What are the flags (which aren't in $\mathbb{R}$?)?]}
+ We did this to take advantage of
 the \textit{emcee} sampler of \citet{ForemanMackey2012}, which
@@
 the shape parameters that is flat in $\vec{x}_0$ and $\vec{\gamma}$,
-and the Jeffrey's prior on $r_0$,
+and the Jeffreys prior on $r_0$,
 \begin{equation}
@@
 \end{equation}
-(Note that this factor of $\sqrt{R_f}$ cancels with the Jeffrey's
+\ilya{[Why is this the Jeffreys prior (BTW, it's Jeffreys, or perhaps Jeffreys', but definitely not Jeffrey's ;) )?  It's not obvious why the prior on $r_0$ should depend on $R_f$?]}
+(Note that this factor of $\sqrt{R_f}$ cancels with the Jeffreys
 prior on the rate, $1/\sqrt{R_f}$; we have verified that the priors on
@@
 significant background, the cluster scale and total number are
-accurately recoverd by our analysis.
+accurately recovered by our analysis.
 
@@
     the structural radius of the cluster, $r_0$ (see
-    Eq.~\eqref{eq:true-cluster-parameters}).}
+    Eq.~\eqref{eq:true-cluster-parameters}). \ilya{[Given the cluster size of 0.1, it seems a bit surprising that the center is off by $\sim 0.05$...]}}
 \end{figure}
@@
     \S~\ref{sec:star-cluster}.  The true value is indicated by the
-    vertical line (see Eq.~\ref{eq:true-cluster-parameters}).}
+    vertical line (see Eq.~\ref{eq:true-cluster-parameters}).  \ilya{[Perhaps combine Figs.~8 and 9 (put an extra scale for $r_0$ at the top of 8?]} }
 \end{figure}
@@
 \section{Conclusion}
-We conclude.
+
+Separating classes of events
+
+Comments on background probability estimation --  Cannon et al. in prep.
+
+Relevance (vs. loudest statistic, or frequentist approach using only gold-plated events)
 

### 2012-08-12 many.tex
Linebreaks.

@@
 
-Want to add more than just the loudest event (cf. loudest event statistic)
+Want to add more than just the loudest event (cf. loudest event
+statistic)
 
@@
 
-Cf. number of gold-plated detections in sensitive volume -- frequentist approach OK if tons of events, otherwise need this.
+Cf. number of gold-plated detections in sensitive volume --
+frequentist approach OK if tons of events, otherwise need this.
 
@@
 We assume that both the foreground and background events are samples
-from an inhomogeneous Poisson process \ilya{[pardon my ignorance, but I thought an "inhomogeneous" Poisson process meant that the rate changed as a function of time -- and I don't think that's what we are assuming...]} with \ilya{respective} differential rates
+from an inhomogeneous Poisson process \ilya{[pardon my ignorance, but
+    I thought an "inhomogeneous" Poisson process meant that the rate
+    changed as a function of time -- and I don't think that's what we
+    are assuming...]} with \ilya{respective} differential rates
 \begin{equation}
@@
 where $\hat{F}(\infty, \theta') = \hat{B}(\infty, \theta') = 1$, and
-$\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f\ilya{\equiv F(\infty,\theta)}$ and $R_b\ilya{\equiv B(\infty,\theta)}$
-are the total number of foreground and background events expected and
+$\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f\ilya{\equiv
+  F(\infty,\theta)}$ and $R_b\ilya{\equiv B(\infty,\theta)}$ are the
+total number of foreground and background events expected and
 $\hat{f}(x, \theta')$ and $\hat{b}(x, \theta')$ are the likelihood of
@@
 are background.  For each event, we introduce a flag, $f_i$, which is
-either 0 \ilya{(background)} or 1 \ilya{(foreground)}.
-These ``state'' flags are parameters in our model, along with $R_f$, $R_b$, and $\theta$.
-We can marginalize over our uncertainty in the state of
+either 0 \ilya{(background)} or 1 \ilya{(foreground)}.  These
+``state'' flags are parameters in our model, along with $R_f$, $R_b$,
+and $\theta$.  We can marginalize over our uncertainty in the state of
 any given event by summing posteriors over $f_i = \mathset{0,1}$.
@@
 \end{equation}
-The \ilya{normalization constant}, called the evidence, $p(d)$, is independent of the state flags, rates, and shape parameters.
+The \ilya{normalization constant}, called the evidence, $p(d)$, is
+independent of the state flags, rates, and shape parameters.
 
@@
 
-\ilya{[I agree with the final result, but I find this way of deriving it to be confusing.
-First of all, the number of ways to assign flags to events for fixed $N_f$ and $N_b$ is $N!/(N_f!
-N_b!)$, not $N_f!
-N_b!$ as written.
-Secondly, you are computing the probability of having a particular ordered list of flags, so there's no reason to consider different permutations.
-Thirdly, you are not properly including the fact that the length of the flags vector must be exactly $N$, the total number of data points.
-(Well, you have the right result, so you are obviously doing these things properly, but the explanation just doesn't seem clear to me.
-:) )  I would make that $N$ explicit, by writing Eq.~(12) as
+\ilya{[I agree with the final result, but I find this way of deriving
+    it to be confusing.  First of all, the number of ways to assign
+    flags to events for fixed $N_f$ and $N_b$ is $N!/(N_f! N_b!)$, not
+    $N_f! N_b!$ as written.  Secondly, you are computing the
+    probability of having a particular ordered list of flags, so
+    there's no reason to consider different permutations.  Thirdly,
+    you are not properly including the fact that the length of the
+    flags vector must be exactly $N$, the total number of data points.
+    (Well, you have the right result, so you are obviously doing these
+    things properly, but the explanation just doesn't seem clear to
+    me. :) ) I would make that $N$ explicit, by writing Eq.~(12) as
 
@@
 \left(\frac{R_f}{R_f+R_b}\right)^{N_f} \left(\frac{R_b}{R_f+R_b}\right)^{N_b}. 
-$$
-Note that there are no permutations involved here -- once we assume that there are precisely $N=N_f+N_b$ events, the probability for the i'th event to have flag $f_i$ is either $R_f/(R_f+R_b)$ or $R_b/(R_f+R_b)$ depending on whether $f_i$ is 1 or 0.
-Meanwhile,
+$$ Note that there are no permutations involved here -- once we assume
+that there are precisely $N=N_f+N_b$ events, the probability for the
+i'th event to have flag $f_i$ is either $R_f/(R_f+R_b)$ or
+$R_b/(R_f+R_b)$ depending on whether $f_i$ is 1 or 0.  Meanwhile,
 $$
 p\left(N |R_f, R_b\right) = \frac{\left(R_f+R_b\right)^N}{N!} e^{-(R_f+R_b)},
-$$
-since the distribution of total event number is a Poisson process with rate $R_f+R_b$.
-Multiplying these, we get the same result as in (13), but in a way that even I can understand. :)
-]}
+$$ since the distribution of total event number is a Poisson process
+with rate $R_f+R_b$.  Multiplying these, we get the same result as in
+(13), but in a way that even I can understand. :) ]}
 
@@
 \end{equation}
-where $f$ and $b$ are rate densities on the \ilya{$k$}-dimensional space of
-ranking statistics.  We give an example of fitting for
+where $f$ and $b$ are rate densities on the \ilya{$k$}-dimensional
+space of ranking statistics.  We give an example of fitting for
 multi-dimensional rate densities in \S~\ref{sec:star-cluster}.
@@
 \eqref{eq:analytic-foreground-rate} with $R_f^\mathrm{true} = 10.4$
-and $R_b^\mathrm{true} = 95.1$, using 1000 templates.  \ilya{[Specify $\xmin$?]} The synthetic
-data consisted of 13 foreground events and 85 background events; the
-cumulative distribution for the ranking statistic of the synthetic
-data appears in Figure \ref{fig:analytic-data-cumulative}.  We used a
-Markov Chain Monte Carlo simulation to draw samples of state flags and
-rates from the joint posterior (Eq.~\eqref{eq:posterior}).
+and $R_b^\mathrm{true} = 95.1$, using 1000 templates.  \ilya{[Specify
+    $\xmin$?]} The synthetic data consisted of 13 foreground events
+and 85 background events; the cumulative distribution for the ranking
+statistic of the synthetic data appears in Figure
+\ref{fig:analytic-data-cumulative}.  We used a Markov Chain Monte
+Carlo simulation to draw samples of state flags and rates from the
+joint posterior (Eq.~\eqref{eq:posterior}).
 
@@
     \eqref{eq:analytic-foreground-rate} combined with $R_f = 10.4$ and
-    $R_b = 95.1$. \ilya{[I think both axis labels should say $x$ rather than $\rho$, since that's the ranking statistic ($\max \rho$).]}}
+    $R_b = 95.1$. \ilya{[I think both axis labels should say $x$
+        rather than $\rho$, since that's the ranking statistic ($\max
+        \rho$).]}}
 \end{figure}
@@
 
-\ilya{[General comment on the text: I think we need to say a bit more about the figures and the interpretation of the results, instead of just relying on the reader to make the best of it.
-For example, even though it may seem obvious to us, we should explicitly say that Fig.~2 shows that we succeed in recovering rates within the expected uncertainty envelope.
-On Fig.~3, could say something obvious like point out that we cannot separate background from foreground for individual events, but that doesn't necessarily matter for rate measurements.
-Now, one interesting question that Fig.~3 brings to mind is the following: if we only used "gold-plated" events (say, those with $p(x|fore)/p(x|back)>0.99$, how much worse / more uncertain would the rate estimation be?]}
+\ilya{[General comment on the text: I think we need to say a bit more
+    about the figures and the interpretation of the results, instead
+    of just relying on the reader to make the best of it.  For
+    example, even though it may seem obvious to us, we should
+    explicitly say that Fig.~2 shows that we succeed in recovering
+    rates within the expected uncertainty envelope.  On Fig.~3, could
+    say something obvious like point out that we cannot separate
+    background from foreground for individual events, but that doesn't
+    necessarily matter for rate measurements.  Now, one interesting
+    question that Fig.~3 brings to mind is the following: if we only
+    used "gold-plated" events (say, those with
+    $p(x|fore)/p(x|back)>0.99$, how much worse / more uncertain would
+    the rate estimation be?]}
 
@@
     \ref{sec:analytic-GW-example} marginalized over all other
-    parameters.  True foreground events are in blue \ilya{(dark grey)}, background events
-    in red \ilya{(light grey)}.  The solid line is the likelihood ratio
+    parameters.  True foreground events are in blue \ilya{(dark
+      grey)}, background events in red \ilya{(light grey)}.  The solid
+    line is the likelihood ratio
     $p(x|\mathrm{foreground})/p(x|\mathrm{background}) =
-    \hat{f}(x)/\hat{b}(x)$ for this model \ilya{evaluated at the simulated parameters (?)}; this exceeds the
-    marginalized foreground probability for many of the events because
-    in this data set there are approximately nine times as many
-    background events as foreground events.
-\ilya{[I'm not sure what the reader is meant to take away from the solid lines  If there's something worth taking away, we should describe the message in the text.
+    \hat{f}(x)/\hat{b}(x)$ for this model \ilya{evaluated at the
+      simulated parameters (?)}; this exceeds the marginalized
+    foreground probability for many of the events because in this data
+    set there are approximately nine times as many background events
+    as foreground events. \ilya{[I'm not sure what the reader is meant
+        to take away from the solid lines If there's something worth
+        taking away, we should describe the message in the text.
         Otherwise, if it only confuses things, we should drop it.]}}
@@
 In \S \ref{sec:analytic-GW-example} we assumed that the overlap
-between \ilya{different} templates in the template bank \ilya{was negligible}, so the SNRs \ilya{recovered by}
-different templates are independent random variables.  In fact,
-template banks are not constructed in this way \citep[e.g.,][]{Owen:1998dk,Ajith:2008},
+between \ilya{different} templates in the template bank \ilya{was
+  negligible}, so the SNRs \ilya{recovered by} different templates are
+independent random variables.  In fact, template banks are not
+constructed in this way \citep[e.g.,][]{Owen:1998dk,Ajith:2008},
 %\citep{Caudill2011,Cannon2011}, 
-because signals could fall in the gaps
-between the non-overlapping templates.  We can model this effect by
-assuming that a template bank of $N$ actual templates will behave as
-if it had $N_\mathrm{eff}$ \emph{independent} templates.  \ilya{Rather than pre-computing $N_\mathrm{eff}$,} we can fit for \ilya{it as a} shape parameter.  That is, we assume that
-\ilya{$\theta = \{N_\mathrm{eff}\}$} is a shape parameter for the background cumulative
-distribution:
+because signals could fall in the gaps between the non-overlapping
+templates.  We can model this effect by assuming that a template bank
+of $N$ actual templates will behave as if it had $N_\mathrm{eff}$
+\emph{independent} templates.  \ilya{Rather than pre-computing
+  $N_\mathrm{eff}$,} we can fit for \ilya{it as a} shape parameter.
+That is, we assume that \ilya{$\theta = \{N_\mathrm{eff}\}$} is a
+shape parameter for the background cumulative distribution:
 \begin{equation}
@@
 
-\ilya{[Again, I think we need to explicitly compare this method with two alternatives.
-Alternative one is the loudest event statistic.
-Alternative two is just counting gold-plated events, and dividing the volume in which we are sensitive to them by their number.
-In principle, alternative one should be suboptimal when the number of events is more than 0 or 1, while alternative two should be OK when the number of events is large, but error-prone otherwise.
-But if we want people to use this technique, we have to defend the point that there is a regime where it is clearly superior (and say what that regime is).]}
+\ilya{[Again, I think we need to explicitly compare this method with
+    two alternatives.  Alternative one is the loudest event statistic.
+    Alternative two is just counting gold-plated events, and dividing
+    the volume in which we are sensitive to them by their number.  In
+    principle, alternative one should be suboptimal when the number of
+    events is more than 0 or 1, while alternative two should be OK
+    when the number of events is large, but error-prone otherwise.
+    But if we want people to use this technique, we have to defend the
+    point that there is a regime where it is clearly superior (and say
+    what that regime is).]}
 
@@
     wider than those of Figure \ref{fig:analytic-rate-recovery}, in
-    spite of the extra parameter.  \ilya{[$R_f$ seems to be more of an outlier than before; why was there a significant shift, given that $N_\mathrm{eff}$ seems fairly spot on?]}}
+    spite of the extra parameter.  \ilya{[$R_f$ seems to be more of an
+        outlier than before; why was there a significant shift, given
+        that $N_\mathrm{eff}$ seems fairly spot on?]}}
 \end{figure}
@@
 where $\vec{\gamma}$ is the gradient, and $\vec{x}_{1/2} = [1/2, 1/2]$
-is the centroid of the observational domain.
-\ilya{[I think that if the domain is finite, the previous equation is only normalized for all $\vec{\gamma}$ if $\vec{x}_{1/2}$ is in the center of a symmetric domain (as it happens to be in this case).]}
-We use assume a gradient
-for the background of $\vec{\gamma} = [-0.5, 0.5]$. 
-\ilya{[I find it a bit confusing to use the word "assume" for both the model and for the specific parameters injected into the simulation (but then being fit for, so their values are not "assumed" in the recovery).
-Maybe "assume" for the model, and "simulate" for the parameter values to be fit?]}
-We assume that
-the total number of background stars, $R_b = 10000$, exceeds by a
-factor of ten the total number of cluster stars, $R_f = 1000$.  
-\ilya{[But the cluster has only $\sim 0.01$ of the total area, so it's over-dense by a factor of $10$ or more.
-Fig.~6 shows a very bright spot where the cluster is, which sort of makes the problem look too simple for the chosen parameters...]}
-We choose a true cluster center $\vec{x}_0 = \vec{x}_{1/2}$, and a radial
-scale parameter of $r_0 = 0.1$.  In all, the shape parameters for the
-true distributions are
+is the centroid of the observational domain.  \ilya{[I think that if
+    the domain is finite, the previous equation is only normalized for
+    all $\vec{\gamma}$ if $\vec{x}_{1/2}$ is in the center of a
+    symmetric domain (as it happens to be in this case).]}  We use
+assume a gradient for the background of $\vec{\gamma} = [-0.5, 0.5]$.
+\ilya{[I find it a bit confusing to use the word "assume" for both the
+    model and for the specific parameters injected into the simulation
+    (but then being fit for, so their values are not "assumed" in the
+    recovery).  Maybe "assume" for the model, and "simulate" for the
+    parameter values to be fit?]}  We assume that the total number of
+background stars, $R_b = 10000$, exceeds by a factor of ten the total
+number of cluster stars, $R_f = 1000$.  \ilya{[But the cluster has
+    only $\sim 0.01$ of the total area, so it's over-dense by a factor
+    of $10$ or more.  Fig.~6 shows a very bright spot where the
+    cluster is, which sort of makes the problem look too simple for
+    the chosen parameters...]}  We choose a true cluster center
+$\vec{x}_0 = \vec{x}_{1/2}$, and a radial scale parameter of $r_0 =
+0.1$.  In all, the shape parameters for the true distributions are
 \begin{equation}
@@
     more stars (10000) in the background than in the cluster (1000),
-    but the peak density for the cluster is much higher.
-\ilya{[It seems too obvious that where the cluster is by eye, and probably fairly straightforward to compute the number of "excess" stars there...
-Reader might ask why we need something so complicated when the answer is seems so obvious and a "frequentist" approach would seem to do here.]}}
+    but the peak density for the cluster is much higher.  \ilya{[It
+        seems too obvious that where the cluster is by eye, and
+        probably fairly straightforward to compute the number of
+        "excess" stars there...  Reader might ask why we need
+        something so complicated when the answer is seems so obvious
+        and a "frequentist" approach would seem to do here.]}}
 \end{figure}
@@
 To analyize our synthetic data set, we analytically marginalized over
-the state \ilya{[this is the only time we call them "state" flags...]} flags, using the likelihood in
-Eq.~\eqref{eq:rate-shape-posterior}. 
-\ilya{[I think it would be kind to the reader to be more explicit about the connection with the formalism in Section II.
-What are the "events"?
-What is the ranking statistic $x$?
-What are the flags (which aren't in $\mathbb{R}$?)?]}
- We did this to take advantage of
-the \textit{emcee} sampler of \citet{ForemanMackey2012}, which
-requires all parameters to be in $\mathbb{R}$.  We applied a prior on
-the shape parameters that is flat in $\vec{x}_0$ and $\vec{\gamma}$,
-and the Jeffreys prior on $r_0$,
+the state \ilya{[this is the only time we call them "state" flags...]}
+flags, using the likelihood in Eq.~\eqref{eq:rate-shape-posterior}.
+\ilya{[I think it would be kind to the reader to be more explicit
+    about the connection with the formalism in Section II.  What are
+    the "events"?  What is the ranking statistic $x$?  What are the
+    flags (which aren't in $\mathbb{R}$?)?]}  We did this to take
+advantage of the \textit{emcee} sampler of \citet{ForemanMackey2012},
+which requires all parameters to be in $\mathbb{R}$.  We applied a
+prior on the shape parameters that is flat in $\vec{x}_0$ and
+$\vec{\gamma}$, and the Jeffreys prior on $r_0$,
 \begin{equation}
@@
 \end{equation}
-\ilya{[Why is this the Jeffreys prior (BTW, it's Jeffreys, or perhaps Jeffreys', but definitely not Jeffrey's ;) )?  It's not obvious why the prior on $r_0$ should depend on $R_f$?]}
-(Note that this factor of $\sqrt{R_f}$ cancels with the Jeffreys
-prior on the rate, $1/\sqrt{R_f}$; we have verified that the priors on
-these parameters are irrelevant to our results, as would be expected
-from the measurement of $\sim 1000$ foreground stars.)
+\ilya{[Why is this the Jeffreys prior (BTW, it's Jeffreys, or perhaps
+    Jeffreys', but definitely not Jeffrey's ;) )?  It's not obvious
+    why the prior on $r_0$ should depend on $R_f$?]}  (Note that this
+factor of $\sqrt{R_f}$ cancels with the Jeffreys prior on the rate,
+$1/\sqrt{R_f}$; we have verified that the priors on these parameters
+are irrelevant to our results, as would be expected from the
+measurement of $\sim 1000$ foreground stars.)
 
@@
     the structural radius of the cluster, $r_0$ (see
-    Eq.~\eqref{eq:true-cluster-parameters}). \ilya{[Given the cluster size of 0.1, it seems a bit surprising that the center is off by $\sim 0.05$...]}}
+    Eq.~\eqref{eq:true-cluster-parameters}). \ilya{[Given the cluster
+        size of 0.1, it seems a bit surprising that the center is off
+        by $\sim 0.05$...]}}
 \end{figure}
@@
     \S~\ref{sec:star-cluster}.  The true value is indicated by the
-    vertical line (see Eq.~\ref{eq:true-cluster-parameters}).  \ilya{[Perhaps combine Figs.~8 and 9 (put an extra scale for $r_0$ at the top of 8?]} }
+    vertical line (see Eq.~\ref{eq:true-cluster-parameters}).
+    \ilya{[Perhaps combine Figs.~8 and 9 (put an extra scale for $r_0$
+        at the top of 8?]} }
 \end{figure}
@@
 
-Relevance (vs. loudest statistic, or frequentist approach using only gold-plated events)
+Relevance (vs. loudest statistic, or frequentist approach using only
+gold-plated events)
 

### 2012-08-13 many.tex
Removed red from Sec. II.

@@
 
-\author{Jon Gair}
+\author{Jonathan R. Gair}
 \email{jrg23@cam.ac.uk}
@@
 We assume that both the foreground and background events are samples
-from an inhomogeneous Poisson process \ilya{[pardon my ignorance, but
-    I thought an "inhomogeneous" Poisson process meant that the rate
-    changed as a function of time -- and I don't think that's what we
-    are assuming...]} with \ilya{respective} differential rates
+from an inhomogeneous Poisson process with respective differential rates
 \begin{equation}
@@
   \begin{equation}
-    P(\ilya{n} = 1 \in [x, x+h]) = f(x,\theta) h + \order{h^2}.
+    P(n = 1 \in [x, x+h]) = f(x,\theta) h + \order{h^2}.
   \end{equation}
@@
   \begin{equation}
-    P(\ilya{n} = 2 \in [x, x+h]) = \order{h^2}.
+    P(n = 2 \in [x, x+h]) = \order{h^2}.
   \end{equation}
@@
 where $\hat{F}(\infty, \theta') = \hat{B}(\infty, \theta') = 1$, and
-$\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f\ilya{\equiv
-  F(\infty,\theta)}$ and $R_b\ilya{\equiv B(\infty,\theta)}$ are the
+$\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f \equiv
+  F(\infty,\theta)$ and $R_b \equiv B(\infty,\theta)$ are the
 total number of foreground and background events expected and
@@
 are background.  For each event, we introduce a flag, $f_i$, which is
-either 0 \ilya{(background)} or 1 \ilya{(foreground)}.  These
+either 0 (background) or 1 (foreground).  These
 ``state'' flags are parameters in our model, along with $R_f$, $R_b$,
@@
   \label{eq:Bayes}
-  p\left(\mathset{f_i}, R_f, R_b, \theta | d \right)  \ilya{=}
+  p\left(\mathset{f_i}, R_f, R_b, \theta | d \right)  =
   \frac{p\left( d | \mathset{f_i}, R_f, R_b, \theta\right)
-  p\left(\mathset{f_i}, R_f, R_b, \theta\right)}{\ilya{p(d)}}. 
+  p\left(\mathset{f_i}, R_f, R_b, \theta\right)}{p(d)}. 
 \end{equation}
-The \ilya{normalization constant}, called the evidence, $p(d)$, is
+The normalization constant, called the evidence, $p(d)$, is
 independent of the state flags, rates, and shape parameters.
@@
 \end{equation}
+
+%The prior distribution can be factorized as
+%\begin{equation}
+%  \label{eq:combined-flag-rate-prior}
+%  p\left(\mathset{f_i}, R_f, R_b, \theta\right)  = p\left(
+%   \mathset{f_i} | R_f, R_b, \theta\right)p(R_f, R_b, \theta).
+%\end{equation}
+%The first term is not actually a ``prior'' in the usual sense; the
+%conditional probability of the flags $\mathset{f_i}$ on the rates is given
+%by
+%\begin{equation}
+%  \label{eq:flag-conditional-prior}
+%  p\left(\mathset{f_i} | R_f, R_b, \theta\right) = \frac{R_f^{N_f}
+%    R_b^{N_b}}{N_f! N_b!} \exp\left[ - \left(R_f + R_b\right) \right]
+%  N_f! N_b! = R_f^{N_f}R_b^{N_b} \exp\left[ - \left(R_f + R_b\right) \right],
+%\end{equation}
+%where $N_f$ and $N_b$ are the number of foreground and background
+%events.  This expression follows from Property \ref{prop:Poisson} of
+%inhomogeneous Poisson processes and the $N_f! N_b!$ different ways of
+%assigning foreground and background flags to events for fixed $N_f$,
+%$N_b$.  
+%
+%\ilya{[I agree with the final result, but I find this way of deriving
+%    it to be confusing.  First of all, the number of ways to assign
+%    flags to events for fixed $N_f$ and $N_b$ is $N!/(N_f! N_b!)$, not
+%    $N_f! N_b!$ as written.  Secondly, you are computing the
+%    probability of having a particular ordered list of flags, so
+%    there's no reason to consider different permutations.  Thirdly,
+%    you are not properly including the fact that the length of the
+%    flags vector must be exactly $N$, the total number of data points.
+%    (Well, you have the right result, so you are obviously doing these
+%    things properly, but the explanation just doesn't seem clear to
+%    me. :) ) I would make that $N$ explicit, by writing Eq.~(12) as
+
 The prior distribution can be factorized as
@@
   \label{eq:combined-flag-rate-prior}
-  p\left(\mathset{f_i}, R_f, R_b, \theta\right)  = p\left(
-    \mathset{f_i} | R_f, R_b, \theta\right)p(R_f, R_b, \theta).
-\end{equation}
-The first term is not actually a ``prior'' in the usual sense; the
-conditional probability of the flags $\mathset{f_i}$ on the rates is given
-by
-\begin{equation}
-  \label{eq:flag-conditional-prior}
-  p\left(\mathset{f_i} | R_f, R_b, \theta\right) = \frac{R_f^{N_f}
-    R_b^{N_b}}{N_f! N_b!} \exp\left[ - \left(R_f + R_b\right) \right]
-  N_f! N_b! = R_f^{N_f}R_b^{N_b} \exp\left[ - \left(R_f + R_b\right) \right],
-\end{equation}
-where $N_f$ and $N_b$ are the number of foreground and background
-events.  This expression follows from Property \ref{prop:Poisson} of
-inhomogeneous Poisson processes and the $N_f! N_b!$ different ways of
-assigning foreground and background flags to events for fixed $N_f$,
-$N_b$.
-
-\ilya{[I agree with the final result, but I find this way of deriving
-    it to be confusing.  First of all, the number of ways to assign
-    flags to events for fixed $N_f$ and $N_b$ is $N!/(N_f! N_b!)$, not
-    $N_f! N_b!$ as written.  Secondly, you are computing the
-    probability of having a particular ordered list of flags, so
-    there's no reason to consider different permutations.  Thirdly,
-    you are not properly including the fact that the length of the
-    flags vector must be exactly $N$, the total number of data points.
-    (Well, you have the right result, so you are obviously doing these
-    things properly, but the explanation just doesn't seem clear to
-    me. :) ) I would make that $N$ explicit, by writing Eq.~(12) as
-
-$$
   p\left(\mathset{f_i}, N, R_f, R_b, \theta\right)  = 
@@
  p\left(R_f, R_b, \theta\right).
-$$
+ \end{equation}
 
-Then 
-$$ p\left( \mathset{f_i} | N, R_f, R_b\right) = 
+The {\it a priori} probability that the $i$'th state flag is $f_i=1$ is given by $R_f/(R_f+R_b)$, while the probability that it is zero is $R_b/(R_f+R_b)$.  Then
+\begin{equation}
+p\left( \mathset{f_i} | N, R_f, R_b\right) = 
 \prod_{\mathset{i|f_i=1}} \left(\frac{R_f}{R_f+R_b}\right) 
 \prod_{\mathset{i|f_i=0}} \left(\frac{R_b}{R_f+R_b}\right) =
-\left(\frac{R_f}{R_f+R_b}\right)^{N_f} \left(\frac{R_b}{R_f+R_b}\right)^{N_b}. 
-$$ Note that there are no permutations involved here -- once we assume
-that there are precisely $N=N_f+N_b$ events, the probability for the
-i'th event to have flag $f_i$ is either $R_f/(R_f+R_b)$ or
-$R_b/(R_f+R_b)$ depending on whether $f_i$ is 1 or 0.  Meanwhile,
-$$
+\left(\frac{R_f}{R_f+R_b}\right)^{N_f} \left(\frac{R_b}{R_f+R_b}\right)^{N_b},
+\end{equation}
+where $N_f$ and $N_b$ are the numbers of foreground and background flags, $N_f+N_b=N$.
+Meanwhile,
+\begin{equation}
 p\left(N |R_f, R_b\right) = \frac{\left(R_f+R_b\right)^N}{N!} e^{-(R_f+R_b)},
-$$ since the distribution of total event number is a Poisson process
-with rate $R_f+R_b$.  Multiplying these, we get the same result as in
-(13), but in a way that even I can understand. :) ]}
+\end{equation}
+since the distribution of total event number is a Poisson process
+with rate $R_f+R_b$.  Combining these yields the conditional probability of the flags on the rates:
+\begin{equation}
+  \label{eq:flag-conditional-prior}
+  p\left(\mathset{f_i} | R_f, R_b\right) = \frac{R_f^{N_f}
+    R_b^{N_b}}{N_f! N_b!} \exp\left[ - \left(R_f + R_b\right) \right]
+  N_f! N_b! = R_f^{N_f}R_b^{N_b} \exp\left[ - \left(R_f + R_b\right) \right].
+\end{equation}
 
@@
 \begin{equation}
-  p\left( R_f, R_b, \theta\right) \ilya{\propto} \frac{1}{\sqrt{R_f R_b}} p(\theta),
+  p\left( R_f, R_b, \theta\right) \propto \frac{1}{\sqrt{R_f R_b}} p(\theta),
 \end{equation}
@@
 \begin{equation}
-  R_f = \int d^\ilya{k} \ilya{\vec{x}} \, f(x, \theta)
+  R_f = \int d^k \vec{x} \, f(x, \theta)
 \end{equation}
@@
 \begin{equation}
-  R_b = \int d^\ilya{k} \ilya{\vec{x}} \, b(x, \theta),  
+  R_b = \int d^k \vec{x} \, b(x, \theta),  
 \end{equation}
-where $f$ and $b$ are rate densities on the \ilya{$k$}-dimensional
+where $f$ and $b$ are rate densities on the $k$-dimensional
 space of ranking statistics.  We give an example of fitting for
@@
 To analyize our synthetic data set, we analytically marginalized over
-the state \ilya{[this is the only time we call them "state" flags...]}
-flags, using the likelihood in Eq.~\eqref{eq:rate-shape-posterior}.
+the state flags, using the likelihood in Eq.~\eqref{eq:rate-shape-posterior}.
 \ilya{[I think it would be kind to the reader to be more explicit

### 2012-08-13 many.tex
Spotted two more typos.

@@
 \end{multline}
+\ilya{[Note: I don't think this is right -- the expression you've written down should be zero in the limit $\epsilon \to 0$, and it would be if the $\epsilon$s were kept correctly (e.g., $ P\left( \textnormal{one event between } y_0 - \epsilon \textnormal{ and } y_0 + \epsilon \right) \sim  r\left( y_0 \right) \epsilon +
+    \order{\epsilon^2}$).]}
 Applying this likelihood once to the foreground samples and once to
@@
 \begin{equation}
-  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{2}
+  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}}
       \right)}{2} \right)^N
@@
   \label{eq:analytic-background-rate}
-  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{2} \right)
-    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
-    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
+  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}} \right)
+    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
+    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
     \right)^N }
@@
 
-To analyize our synthetic data set, we analytically marginalized over
+To analyze our synthetic data set, we analytically marginalized over
 the state flags, using the likelihood in Eq.~\eqref{eq:rate-shape-posterior}.

### 2012-08-13 many.tex
Comment on normalization.

@@
 \end{equation}
+\ilya{[Note: some of the expressions here aren't really properly normalized -- should check carefully which ones are, and replace $=$ by $\propto$ for the others.]}
 This expression is useful if we are only interested in rates and not

### 2012-08-14 many.tex
Started to add in theory relating our work to loudest event etc.

@@
 
+\author{Jonathan R. Gair}
+\email{jrg23@cam.ac.uk}
+\affiliation{Institute of Astronomy\\University of Cambridge\\Madingley Road, Cambridge CB3 0HA\\United Kingdom}
+
 \author{Ilya Mandel}
@@
 %\affiliation{University of Birmingham}
-
-\author{Jonathan R. Gair}
-\email{jrg23@cam.ac.uk}
 
@@
 \begin{multline}
-  p\left( \mathset{y_i} | r \right) = \lim_{\epsilon \to 0} P\left(
-  \textnormal{zero events below } y_0 - \epsilon \right) \times
-  P\left( \textnormal{one event between } y_0 - \epsilon \textnormal{
-    and } y_0 + \epsilon \right) \\ \times P\left( \textnormal{zero
-    events between } y_0+\epsilon \textnormal{ and } y_1 - \epsilon
-  \right) \ldots \\ = \lim_{\epsilon \to 0} \exp\left[ - R\left(y_0 -
-    \epsilon \right) \right] \left[ r\left( y_0 \right) +
-    \order{\epsilon^2} \right] \times \exp\left[ - \left[ R\left( y_1
-      - \epsilon\right) - R\left( y_0 + \epsilon\right) \right]
-    \right] \times \ldots = \left[\prod_i r\left( y_i
+  p\left( \mathset{y_i} | r \right) \,{\rm d}^Ny_i= P\left(
+  \textnormal{zero events below } y_0  \right) \times
+  P\left( \textnormal{one event between } y_0  \textnormal{
+    and } y_0 + {\rm d}y_0 \right) \\ \times P\left( \textnormal{zero
+    events between } y_0+{\rm d}y_0 \textnormal{ and } y_1
+  \right) \ldots \\ p\left( \mathset{y_i} | r \right) = \lim_{\delta y_i \to 0} \exp\left[ - R\left(y_0\right) \right] \left[ r\left( y_0 \right) +
+    \order{\delta y_0} \right] \times \exp\left[ - \left[ R\left( y_1
+      \right) - R\left( y_0 + \delta y_0\right) \right]
+    \right] \times \ldots \\ = \left[\prod_i r\left( y_i
     \right)\right] \exp\left[ - R\left( \infty \right) \right].
 \end{multline}
-\ilya{[Note: I don't think this is right -- the expression you've written down should be zero in the limit $\epsilon \to 0$, and it would be if the $\epsilon$s were kept correctly (e.g., $ P\left( \textnormal{one event between } y_0 - \epsilon \textnormal{ and } y_0 + \epsilon \right) \sim  r\left( y_0 \right) \epsilon +
-    \order{\epsilon^2}$).]}
 Applying this likelihood once to the foreground samples and once to
@@
 parameters.
+We note that the terms that depend on the overall rate parameters, $R_b$ or $R_f$, are of the form $R_b^{n-1/2} \exp(-R_b)$ and so marginalisation over either $R_b$ or $R_f$ can be achieved analytically using
+\begin{equation}
+I_n = \int_0^{\infty} x^{n-\frac{1}{2}} \, {\rm e}^{-x} {\rm d} x = \frac{(2n-1)!!}{2^n} \sqrt{\pi}
+\end{equation}
+using the usual notation $(2n-1)!! \equiv (2n-1)(2n-3)\cdots1$.
 
@@
 multi-dimensional rate densities in \S~\ref{sec:star-cluster}.
+
+It is informative to relate these results to two other ways to estimate the foreground rate parameter --- the loudest event statistic and the foreground-dominated statistic.
+
+\subsubsection{Loudest event statistic}
+If we were to include only the $k$-loudest events in the posterior distribution, rather than all observed events, the posterior (Eq.~\eqref{eq:posterior}) would be modified by an additional factor of $\exp[-(R_f \hat{F}(x_{N-k+1}) + R_b \hat{B}(x_{N-k+1}))]$, where we have assumed events are ordered by loudness, so that $x_{N-k+1}$ is the $k$-th loudest event.
+This term accounts for the data-dependent threshold that a loudest event statistic employs.
+
+For the usual $k=1$ case, the marginalised posterior (Eq.~\eqref{eq:rate-shape-posterior}) becomes
+\begin{equation}
+p_{\rm LE}\left( R_f, R_b, \theta | d \right) \propto \left(R_f \hat{f}(x_N,\theta) + R_b \hat{b}(x_N,\theta)\right)\exp\left[-\left( R_f (1-\hat{F}(x_N,\theta)) +
+      R_b (1-\hat{B}(x_N,\theta)) \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
+\end{equation}
+where $x_N$ denotes the loudness of the loudest event. Marginalising over $R_b$ we obtain
+\begin{equation}
+p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f \hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}} \exp(-R_f) .
+\end{equation}
+This posterior has a maximum in $R_f$ at
+\begin{equation}
+R_f = \frac{\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta) \pm \sqrt{(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta))^2 - 4\tilde{b}(x_N,\theta)(1-\hat{F}(x_N,\theta))\hat{f}(x_N,\theta)}}{4 \hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta)}.
+\end{equation}
+where $\tilde{b}(x_N,\theta) =\hat{b}(x_N,\theta)/(1-\hat{B}(x_N,\theta))$.
+If we assume that $\tilde{b}(x_N,\theta) \ll 1$, i.e., that the loudest event is very likely to be foreground we obtain the expected result $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$.
 

### 2012-08-15 many.tex
Changes to section II for clarity, and foreground-dominate result included. Not yet finished.

@@
 
-We assume that we are presented with a data set of $N$ events.  Each
+We assume that we are presented with a data set of $N$ events that exceed a pre-specified threshold in ranking statistic, $x_{min}$.  Each
 event may be due to either a signal of interest or an uninteresting
@@
 \begin{equation}
-  d = \mathset{ x_i | i = 1, \ldots, N }
+  d = \mathset{ x_i | i = 1, \ldots, N } .
 \end{equation}
+The number of events $N$ is also part of the observed data, but we separate out $N$ and the observed ranking statistics, $d$, for convenience.
+We can choose how to label our events.
+Ultimately we will label the events in order of ranking statistic, i.e., $x_1 < x_2 < \cdots < x_N$, but some of the derivations that follow are simpler if the events are ordered by time of arrival.
+We will use $d$ to denote ranking statistic-ordered events, and $d_{\rm to}$ to denote time-ordered events.
 
@@
 
-Bayes' theorem relates the posterior probability of the state flags,
+Assuming time-ordered data, $d_{\rm to}$, in the following, Bayes' theorem relates the posterior probability of the state flags,
 rates, and shape parameters, $p\left(\mathset{f_i}, R_f, R_b, \theta |
-  d \right)$, the likelihood of the data, $p\left( d | \mathset{f_i},
+  d_{\rm to}, N \right)$, the likelihood of the data, $p\left( d_{\rm to} | \mathset{f_i}, N, 
   R_f, R_b, \theta\right)$, and the prior probability of state flags,
 rates and shape parameters before any data are obtained, $p\left(
-  \mathset{f_i}, R_f, R_b, \theta\right)$:
+  \mathset{f_i}, N, R_f, R_b, \theta\right)$:
 \begin{equation}
   \label{eq:Bayes}
-  p\left(\mathset{f_i}, R_f, R_b, \theta | d \right)  =
-  \frac{p\left( d | \mathset{f_i}, R_f, R_b, \theta\right)
-  p\left(\mathset{f_i}, R_f, R_b, \theta\right)}{p(d)}. 
+  p\left(\mathset{f_i}, R_f, R_b, \theta | d_{\rm to}, N \right)  =
+  \frac{p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right)
+  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right)}{p(d_{\rm to}, N)}. 
 \end{equation}
-The normalization constant, called the evidence, $p(d)$, is
+The normalization constant, called the evidence, $p(d_{\rm to}, N)$, is
 independent of the state flags, rates, and shape parameters.
@@
 \begin{equation}
-  p\left( d | \mathset{f_i}, R_f, R_b, \theta\right)  = \left[
+  p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right)  = \left[
     \prod_{\mathset{i | f_i = 1}} \hat{f}\left(x_i, \theta\right) \right]
@@
 \end{equation}
+This is the probability that the first observed event is a fore/background event (if $f_i=1,0$) with ranking statistic $x_1$ and the second observed event is a fore/background event (if $f_i=1,0$) with ranking statistic $x_2$ etc.
+If the events are ordered by ranking statistic the corresponding expression is more complicated, since $x_1$ is now the event  from foreground or background, with the smallest ranking statistic etc.
+We will return to the statistic-ordered case later.
 
@@
   \label{eq:combined-flag-rate-prior}
-  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right)  = 
-  p\left( \mathset{f_i} | N, R_f, R_b\right) p\left(N |R_f, R_b\right)  
- p\left(R_f, R_b, \theta\right).
+  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right)  = p\left( \mathset{f_i} | N, R_f, R_b\right) p\left(N |R_f, R_b\right)   p\left(R_f, R_b, \theta\right) =  p\left( \mathset{f_i},N | R_f, R_b\right)  p\left(R_f, R_b, \theta\right).
  \end{equation}
 
-The {\it a priori} probability that the $i$'th state flag is $f_i=1$ is given by $R_f/(R_f+R_b)$, while the probability that it is zero is $R_b/(R_f+R_b)$.  Then
+The probability that the $i$'th state flag is $f_i=1$ is given by $R_f/(R_f+R_b)$, while the probability that it is zero is $R_b/(R_f+R_b)$, provided the data are time-ordered as we have assumed.
+Then
 \begin{equation}
@@
   \label{eq:flag-conditional-prior}
-  p\left(\mathset{f_i} | R_f, R_b\right) = \frac{R_f^{N_f}
-    R_b^{N_b}}{N_f! N_b!} \exp\left[ - \left(R_f + R_b\right) \right]
-  N_f! N_b! = R_f^{N_f}R_b^{N_b} \exp\left[ - \left(R_f + R_b\right) \right].
+  p\left(\mathset{f_i},N | R_f, R_b\right) = \frac{R_f^{N_f}
+    R_b^{N_b}}{N!} \exp\left[ - \left(R_f + R_b\right) \right].
 \end{equation}
@@
 \begin{equation}
-  p\left( R_f, R_b, \theta\right) \propto \frac{1}{\sqrt{R_f R_b}} p(\theta),
+  p\left( R_f, R_b, \theta\right) = \alpha \frac{1}{\sqrt{R_f R_b}} p(\theta),
 \end{equation}
-but of course other choices are possible.  This choice has the
-advantage that the prior is normalizable as $R_f, R_b \to 0$, and the
+where $\alpha$ is a normalisation constant, but of course other choices are possible.  This choice has the advantage that the prior is normalizable as $R_f, R_b \to 0$, and the
 exponentials in Eq.~\eqref{eq:flag-conditional-prior} regularize the
@@
   \label{eq:posterior}
-  p\left( \mathset{f_i}, R_f, R_b, \theta | d \right) \propto \left[
+  p\left( \mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right) = \frac{\alpha}{p(d_{\rm to},N)\,N!} \left[
     \prod_{\mathset{i | f_i = 1}} R_f \hat{f}\left(x_i, \theta\right) \right]
@@
 \end{equation}
-Not surprisingly, this posterior is equivalent to one derived by
+When sampling the posterior, the first term can be omitted and the equals sign replaced by proportionality, since it is independent of the parameters of interest, but we have kept this term explicitly so that we can see the equivalence to ranking-statistic ordered data.
+Once data have been observed, there is a unique loudness ordering and time ordering of those events, and so there is a one to one correspondence between a time-ordered posterior $p\left( \mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right)$ and the corresponding statistic-ordered posterior $ p\left( \mathset{f_i}, R_f, R_b, \theta | d,N \right)$, which means $p\left( \mathset{f_i}, R_f, R_b, \theta | d,N \right) = p\left( \mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right)$.
+However, the evidence $p(d,N) = N!\,p(d_{\rm to},N)$, since there are $N!$ ways in which $N$ events can be ordered in time and would have the same set of ranking statistics.
+
+Not surprisingly, this ranking-statistic ordered posterior can be computed directly by
 assuming that the flags, $\mathset{f_i}$, are un-observed data and
@@
 \end{multline}
-Applying this likelihood once to the foreground samples and once to
-the background samples reproduces Eq.~\eqref{eq:posterior}.
+Applying this once to the foreground samples, once to
+the background samples and taking the product, we obtain $p(d,\mathset{f_i}, N | R_f, R_b, \theta)$ and thence $p(\mathset{f_i}, R_f, R_b, \theta | d,N) = p(d,\mathset{f_i}, N | R_f, R_b, \theta) p*R_f, R_b, \theta)/p(d,N)$.
+With the identification $p(d,N) = N!\,p(d_{\rm to},N)$, as justified above, we reproduce Eq.~\eqref{eq:posterior}.
 
@@
   \label{eq:rate-shape-posterior}
-  p\left( R_f, R_b, \theta | d \right) = \sum_{\mathset{f_i} \in
-    \mathset{0,1}^N} p\left( \mathset{f_i}, R_f, R_b, \theta | d \right)
+  p\left( R_f, R_b, \theta | d, N \right) = \sum_{\mathset{f_i} \in
+    \mathset{0,1}^N} p\left( \mathset{f_i}, R_f, R_b, \theta | d, N \right)
   \propto \prod_{i} \left[ R_f \hat{f}\left(x_i, \theta\right) + R_b
@@
 \end{equation}
-\ilya{[Note: some of the expressions here aren't really properly normalized -- should check carefully which ones are, and replace $=$ by $\propto$ for the others.]}
 This expression is useful if we are only interested in rates and not
@@
 \subsubsection{Loudest event statistic}
-If we were to include only the $k$-loudest events in the posterior distribution, rather than all observed events, the posterior (Eq.~\eqref{eq:posterior}) would be modified by an additional factor of $\exp[-(R_f \hat{F}(x_{N-k+1}) + R_b \hat{B}(x_{N-k+1}))]$, where we have assumed events are ordered by loudness, so that $x_{N-k+1}$ is the $k$-th loudest event.
+If we were to include only the $k$-loudest events in the posterior distribution, rather than all observed events, the posterior (Eq.~\eqref{eq:posterior}) would be modified by an additional factor of $\exp[R_f \hat{F}(x_{N-k+1}) + R_b \hat{B}(x_{N-k+1})]$, where we have assumed events are ordered by loudness, so that $x_{N-k+1}$ is the $k$-th loudest event.
 This term accounts for the data-dependent threshold that a loudest event statistic employs.
@@
 \begin{equation}
-p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f \hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}} \exp(-R_f) .
+p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f \hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}} \exp(-R_f(1-\hat{F}(x_N,\theta))) .
 \end{equation}
@@
 where $\tilde{b}(x_N,\theta) =\hat{b}(x_N,\theta)/(1-\hat{B}(x_N,\theta))$.
-If we assume that $\tilde{b}(x_N,\theta) \ll 1$, i.e., that the loudest event is very likely to be foreground we obtain the expected result $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$.
+If $\tilde{b}(x_N,\theta) \ll 1$, i.e., the loudest event is very likely to be foreground, we obtain the result $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$.
+This is can be understood as the statement that the rate of foreground events with ranking statistic greater than $x_N$,  $(1- \hat{F}(x_N,\theta))\,R_f$, is of order $1$, as expected.
+If $\tilde{b}(x_N,\theta) >> 1$, however, the posterior on $R_f$ is peaked at $0$.
+
+\subsubsection{Foreground dominated statistic}
+If we set the threshold for including an event, $x_{\rm min}$, sufficiently high, we can ensure that $\hat{f}(x_i,\theta) \gg \hat{b}(x_i,\theta)$ for all ranking statistics $x_i$ in the data set.
+The posterior can then be approximated by
+\begin{equation}
+p_{\rm FD}\left( R_f, R_b, \theta | d \right) \propto  \prod_{i} \left[ \hat{f}\left(x_i, \theta\right)\right] R_f^N \exp\left[-\left( R_f + R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
+\end{equation}
+Normalisation over $R_b$ gives a constant factor and the posterior on the foreground rate becomes
+\begin{equation}
+p_{\rm FD}\left( R_f, \theta | d \right) \propto  \prod_{i} \left[ \hat{f}\left(x_i, \theta\right)\right] R_f^{N-\frac{1}{2}} \exp\left[-R_f  \right] p(\theta).
+\end{equation}
+Ignoring the dependence on $\theta$, this is peaked at a rate $R_f = N-1/2$, so we have the expected result that, in the foreground dominated regime, the rate is approximately equal to the number of events observed.
+We note that this is the rate of events occurring above threshold and so if we were to use this statistic with different choices of threshold we would have to renormalise the rates to a fixed threshold in order to compare the results.
 

### 2012-08-16 many.tex
Indentation updates.

@@
 
-\usepackage{hyperref}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{color}
+\usepackage{hyperref} \usepackage{amsmath} \usepackage{amssymb}
+\usepackage{graphicx} \usepackage{color}
 
@@
 
-\author{Will M. Farr}
-\email{w-farr@northwestern.edu}
+\author{Will M. Farr} \email{w-farr@northwestern.edu}
 \homepage{http://faculty.wcas.northwestern.edu/will-farr/}
-\affiliation{Center for Interdisciplinary Exploration and Research in Astrophysics\\
-Department of Physics and Astronomy\\
-Northwestern University, 2145 Sheridan Road, Evanston, IL 60208}
+\affiliation{Center for Interdisciplinary Exploration and Research in
+  Astrophysics\\ Department of Physics and Astronomy\\ Northwestern
+  University, 2145 Sheridan Road, Evanston, IL 60208}
 %\affiliation{Northwestern University and CIERA}
 
-\author{Jonathan R. Gair}
-\email{jrg23@cam.ac.uk}
-\affiliation{Institute of Astronomy\\University of Cambridge\\Madingley Road, Cambridge CB3 0HA\\United Kingdom}
+\author{Jonathan R. Gair} \email{jrg23@cam.ac.uk}
+\affiliation{Institute of Astronomy\\University of
+  Cambridge\\Madingley Road, Cambridge CB3 0HA\\United Kingdom}
 
-\author{Ilya Mandel}
-\email{imandel@star.sr.bham.ac.uk}
-\homepage{http://www.sr.bham.ac.uk/~imandel}
-\affiliation{School of Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT Birmingham\\United Kingdom}
+\author{Ilya Mandel} \email{imandel@star.sr.bham.ac.uk}
+\homepage{http://www.sr.bham.ac.uk/~imandel} \affiliation{School of
+  Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT
+  Birmingham\\United Kingdom}
 %\affiliation{University of Birmingham}
@@
   background event distributions overlap significantly.  We give
-  examples of determining the rates of gravitatonal-wave events in
-  the presence of background triggers from a template bank when noise
+  examples of determining the rates of gravitatonal-wave events in the
+  presence of background triggers from a template bank when noise
   parameters are known and/or can be fit from the trigger data.  We
@@
 
-We assume that we are presented with a data set of $N$ events that exceed a pre-specified threshold in ranking statistic, $x_{min}$.  Each
-event may be due to either a signal of interest or an uninteresting
-background.  Each event is associated with a ranking statistic, $x$.
-Our data set therefore consists of the ranking statistics for the set
-of events:
+We assume that we are presented with a data set of $N$ events that
+exceed a pre-specified threshold in ranking statistic, $x_{min}$.
+Each event may be due to either a signal of interest or an
+uninteresting background.  Each event is associated with a ranking
+statistic, $x$.  Our data set therefore consists of the ranking
+statistics for the set of events:
 \begin{equation}
@@
 \end{equation}
-The number of events $N$ is also part of the observed data, but we separate out $N$ and the observed ranking statistics, $d$, for convenience.
-We can choose how to label our events.
-Ultimately we will label the events in order of ranking statistic, i.e., $x_1 < x_2 < \cdots < x_N$, but some of the derivations that follow are simpler if the events are ordered by time of arrival.
-We will use $d$ to denote ranking statistic-ordered events, and $d_{\rm to}$ to denote time-ordered events.
+The number of events $N$ is also part of the observed data, but we
+separate out $N$ and the observed ranking statistics, $d$, for
+convenience. We can choose how to label our events. Ultimately we will
+label the events in order of ranking statistic, i.e., $x_1 < x_2 <
+\cdots < x_N$, but some of the derivations that follow are simpler if
+the events are ordered by time of arrival. We will use $d$ to denote
+ranking statistic-ordered events, and $d_{\rm to}$ to denote
+time-ordered events.
 
 We assume that both the foreground and background events are samples
-from an inhomogeneous Poisson process with respective differential rates
+from an inhomogeneous Poisson process with respective differential
+rates
 \begin{equation}
@@
 $\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f \equiv
-  F(\infty,\theta)$ and $R_b \equiv B(\infty,\theta)$ are the
-total number of foreground and background events expected and
-$\hat{f}(x, \theta')$ and $\hat{b}(x, \theta')$ are the likelihood of
-obtaining an event with ranking statistic $x$ under the foreground and
-background distributions.  In what follows, we will drop the prime,
-using $\theta$ to denote all parameters of the rate distributions
-except $R_f$ and $R_b$.
+F(\infty,\theta)$ and $R_b \equiv B(\infty,\theta)$ are the total
+number of foreground and background events expected and $\hat{f}(x,
+\theta')$ and $\hat{b}(x, \theta')$ are the likelihood of obtaining an
+event with ranking statistic $x$ under the foreground and background
+distributions.  In what follows, we will drop the prime, using
+$\theta$ to denote all parameters of the rate distributions except
+$R_f$ and $R_b$.
 
@@
 are background.  For each event, we introduce a flag, $f_i$, which is
-either 0 (background) or 1 (foreground).  These
-``state'' flags are parameters in our model, along with $R_f$, $R_b$,
-and $\theta$.  We can marginalize over our uncertainty in the state of
-any given event by summing posteriors over $f_i = \mathset{0,1}$.
+either 0 (background) or 1 (foreground).  These ``state'' flags are
+parameters in our model, along with $R_f$, $R_b$, and $\theta$.  We
+can marginalize over our uncertainty in the state of any given event
+by summing posteriors over $f_i = \mathset{0,1}$.
 
-Assuming time-ordered data, $d_{\rm to}$, in the following, Bayes' theorem relates the posterior probability of the state flags,
-rates, and shape parameters, $p\left(\mathset{f_i}, R_f, R_b, \theta |
-  d_{\rm to}, N \right)$, the likelihood of the data, $p\left( d_{\rm to} | \mathset{f_i}, N, 
-  R_f, R_b, \theta\right)$, and the prior probability of state flags,
-rates and shape parameters before any data are obtained, $p\left(
-  \mathset{f_i}, N, R_f, R_b, \theta\right)$:
+Assuming time-ordered data, $d_{\rm to}$, in the following, Bayes'
+theorem relates the posterior probability of the state flags, rates,
+and shape parameters, $p\left(\mathset{f_i}, R_f, R_b, \theta | d_{\rm
+  to}, N \right)$, the likelihood of the data, $p\left( d_{\rm to} |
+\mathset{f_i}, N, R_f, R_b, \theta\right)$, and the prior probability
+of state flags, rates and shape parameters before any data are
+obtained, $p\left( \mathset{f_i}, N, R_f, R_b, \theta\right)$:
 \begin{equation}
@@
   \frac{p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right)
-  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right)}{p(d_{\rm to}, N)}. 
+    p\left(\mathset{f_i}, N, R_f, R_b, \theta\right)}{p(d_{\rm to},
+    N)}.
 \end{equation}
-The normalization constant, called the evidence, $p(d_{\rm to}, N)$, is
-independent of the state flags, rates, and shape parameters.
+The normalization constant, called the evidence, $p(d_{\rm to}, N)$,
+is independent of the state flags, rates, and shape parameters.
 
@@
 \begin{equation}
-  p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right)  = \left[
-    \prod_{\mathset{i | f_i = 1}} \hat{f}\left(x_i, \theta\right) \right]
-  \left[ \prod_{\mathset{i | f_i = 0}} \hat{b}\left( x_i, \theta\right)
-  \right].  
+  p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right) =
+  \left[ \prod_{\mathset{i | f_i = 1}} \hat{f}\left(x_i, \theta\right)
+    \right] \left[ \prod_{\mathset{i | f_i = 0}} \hat{b}\left( x_i,
+    \theta\right) \right].
 \end{equation}
-This is the probability that the first observed event is a fore/background event (if $f_i=1,0$) with ranking statistic $x_1$ and the second observed event is a fore/background event (if $f_i=1,0$) with ranking statistic $x_2$ etc.
-If the events are ordered by ranking statistic the corresponding expression is more complicated, since $x_1$ is now the event  from foreground or background, with the smallest ranking statistic etc.
-We will return to the statistic-ordered case later.
+This is the probability that the first observed event is a
+fore/background event (if $f_i=1,0$) with ranking statistic $x_1$ and
+the second observed event is a fore/background event (if $f_i=1,0$)
+with ranking statistic $x_2$ etc. If the events are ordered by ranking
+statistic the corresponding expression is more complicated, since
+$x_1$ is now the event from foreground or background, with the
+smallest ranking statistic etc. We will return to the
+statistic-ordered case later.
 
@@
   \label{eq:combined-flag-rate-prior}
-  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right)  = p\left( \mathset{f_i} | N, R_f, R_b\right) p\left(N |R_f, R_b\right)   p\left(R_f, R_b, \theta\right) =  p\left( \mathset{f_i},N | R_f, R_b\right)  p\left(R_f, R_b, \theta\right).
+  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right) = p\left(
+  \mathset{f_i} | N, R_f, R_b\right) p\left(N |R_f, R_b\right)
+  p\left(R_f, R_b, \theta\right) = p\left( \mathset{f_i},N | R_f,
+  R_b\right) p\left(R_f, R_b, \theta\right).
  \end{equation}
 
-The probability that the $i$'th state flag is $f_i=1$ is given by $R_f/(R_f+R_b)$, while the probability that it is zero is $R_b/(R_f+R_b)$, provided the data are time-ordered as we have assumed.
-Then
+The probability that the $i$'th state flag is $f_i=1$ is given by
+$R_f/(R_f+R_b)$, while the probability that it is zero is
+$R_b/(R_f+R_b)$, provided the data are time-ordered as we have
+assumed.  Then
 \begin{equation}
-p\left( \mathset{f_i} | N, R_f, R_b\right) = 
-\prod_{\mathset{i|f_i=1}} \left(\frac{R_f}{R_f+R_b}\right) 
-\prod_{\mathset{i|f_i=0}} \left(\frac{R_b}{R_f+R_b}\right) =
-\left(\frac{R_f}{R_f+R_b}\right)^{N_f} \left(\frac{R_b}{R_f+R_b}\right)^{N_b},
+p\left( \mathset{f_i} | N, R_f, R_b\right) = \prod_{\mathset{i|f_i=1}}
+\left(\frac{R_f}{R_f+R_b}\right) \prod_{\mathset{i|f_i=0}}
+\left(\frac{R_b}{R_f+R_b}\right) =
+\left(\frac{R_f}{R_f+R_b}\right)^{N_f}
+\left(\frac{R_b}{R_f+R_b}\right)^{N_b},
 \end{equation}
-where $N_f$ and $N_b$ are the numbers of foreground and background flags, $N_f+N_b=N$.
-Meanwhile,
+where $N_f$ and $N_b$ are the numbers of foreground and background
+flags, $N_f+N_b=N$.  Meanwhile,
 \begin{equation}
-p\left(N |R_f, R_b\right) = \frac{\left(R_f+R_b\right)^N}{N!} e^{-(R_f+R_b)},
+p\left(N |R_f, R_b\right) = \frac{\left(R_f+R_b\right)^N}{N!}
+e^{-(R_f+R_b)},
 \end{equation}
-since the distribution of total event number is a Poisson process
-with rate $R_f+R_b$.  Combining these yields the conditional probability of the flags on the rates:
+since the distribution of total event number is a Poisson process with
+rate $R_f+R_b$.  Combining these yields the conditional probability of
+the flags on the rates:
 \begin{equation}
@@
 \begin{equation}
-  p\left( R_f, R_b, \theta\right) = \alpha \frac{1}{\sqrt{R_f R_b}} p(\theta),
+  p\left( R_f, R_b, \theta\right) = \alpha \frac{1}{\sqrt{R_f R_b}}
+  p(\theta),
 \end{equation}
-where $\alpha$ is a normalisation constant, but of course other choices are possible.  This choice has the advantage that the prior is normalizable as $R_f, R_b \to 0$, and the
-exponentials in Eq.~\eqref{eq:flag-conditional-prior} regularize the
-posterior as $R_f, R_b \to \infty$.
+where $\alpha$ is a normalisation constant, but of course other
+choices are possible.  This choice has the advantage that the prior is
+normalizable as $R_f, R_b \to 0$, and the exponentials in
+Eq.~\eqref{eq:flag-conditional-prior} regularize the posterior as
+$R_f, R_b \to \infty$.
 
@@
   \label{eq:posterior}
-  p\left( \mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right) = \frac{\alpha}{p(d_{\rm to},N)\,N!} \left[
-    \prod_{\mathset{i | f_i = 1}} R_f \hat{f}\left(x_i, \theta\right) \right]
-  \left[ \prod_{\mathset{i | f_i = 0}} R_b \hat{b}\left( x_i, \theta\right)
+  p\left( \mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right) =
+  \frac{\alpha}{p(d_{\rm to},N)\,N!} \left[ \prod_{\mathset{i | f_i =
+        1}} R_f \hat{f}\left(x_i, \theta\right) \right] \left[
+    \prod_{\mathset{i | f_i = 0}} R_b \hat{b}\left( x_i, \theta\right)
     \right] \exp\left[ - \left(R_f + R_b\right) \right]
@@
 \end{equation}
-When sampling the posterior, the first term can be omitted and the equals sign replaced by proportionality, since it is independent of the parameters of interest, but we have kept this term explicitly so that we can see the equivalence to ranking-statistic ordered data.
-Once data have been observed, there is a unique loudness ordering and time ordering of those events, and so there is a one to one correspondence between a time-ordered posterior $p\left( \mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right)$ and the corresponding statistic-ordered posterior $ p\left( \mathset{f_i}, R_f, R_b, \theta | d,N \right)$, which means $p\left( \mathset{f_i}, R_f, R_b, \theta | d,N \right) = p\left( \mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right)$.
-However, the evidence $p(d,N) = N!\,p(d_{\rm to},N)$, since there are $N!$ ways in which $N$ events can be ordered in time and would have the same set of ranking statistics.
+When sampling the posterior, the first term can be omitted and the
+equals sign replaced by proportionality, since it is independent of
+the parameters of interest, but we have kept this term explicitly so
+that we can see the equivalence to ranking-statistic ordered
+data. Once data have been observed, there is a unique loudness
+ordering and time ordering of those events, and so there is a one to
+one correspondence between a time-ordered posterior $p\left(
+\mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right)$ and the
+corresponding statistic-ordered posterior $ p\left( \mathset{f_i},
+R_f, R_b, \theta | d,N \right)$, which means $p\left( \mathset{f_i},
+R_f, R_b, \theta | d,N \right) = p\left( \mathset{f_i}, R_f, R_b,
+\theta | d_{\rm to},N \right)$. However, the evidence $p(d,N) =
+N!\,p(d_{\rm to},N)$, since there are $N!$ ways in which $N$ events
+can be ordered in time and would have the same set of ranking
+statistics.
 
-Not surprisingly, this ranking-statistic ordered posterior can be computed directly by
-assuming that the flags, $\mathset{f_i}$, are un-observed data and
-treating the sets $\mathset{x_i | f_i = 1}$ and $\mathset{x_i | f_i =
-  0}$ as samples from an inhomogeneous Poisson process.  For an
-inhomogeneous Poisson process with rate function $r(y)$ (cumulative
-rate $R(y)$), the likelihood of a set of samples $\mathset{y_i}$ is
-given by 
+Not surprisingly, this ranking-statistic ordered posterior can be
+computed directly by assuming that the flags, $\mathset{f_i}$, are
+un-observed data and treating the sets $\mathset{x_i | f_i = 1}$ and
+$\mathset{x_i | f_i = 0}$ as samples from an inhomogeneous Poisson
+process.  For an inhomogeneous Poisson process with rate function
+$r(y)$ (cumulative rate $R(y)$), the likelihood of a set of samples
+$\mathset{y_i}$ is given by
 \begin{multline}
   p\left( \mathset{y_i} | r \right) \,{\rm d}^Ny_i= P\left(
-  \textnormal{zero events below } y_0  \right) \times
-  P\left( \textnormal{one event between } y_0  \textnormal{
-    and } y_0 + {\rm d}y_0 \right) \\ \times P\left( \textnormal{zero
-    events between } y_0+{\rm d}y_0 \textnormal{ and } y_1
-  \right) \ldots \\ p\left( \mathset{y_i} | r \right) = \lim_{\delta y_i \to 0} \exp\left[ - R\left(y_0\right) \right] \left[ r\left( y_0 \right) +
+  \textnormal{zero events below } y_0 \right) \times P\left(
+  \textnormal{one event between } y_0 \textnormal{ and } y_0 + {\rm
+    d}y_0 \right) \\ \times P\left( \textnormal{zero events between }
+  y_0+{\rm d}y_0 \textnormal{ and } y_1 \right) \ldots \\ p\left(
+  \mathset{y_i} | r \right) = \lim_{\delta y_i \to 0} \exp\left[ -
+    R\left(y_0\right) \right] \left[ r\left( y_0 \right) +
     \order{\delta y_0} \right] \times \exp\left[ - \left[ R\left( y_1
-      \right) - R\left( y_0 + \delta y_0\right) \right]
-    \right] \times \ldots \\ = \left[\prod_i r\left( y_i
-    \right)\right] \exp\left[ - R\left( \infty \right) \right].
+      \right) - R\left( y_0 + \delta y_0\right) \right] \right] \times
+  \ldots \\ = \left[\prod_i r\left( y_i \right)\right] \exp\left[ -
+    R\left( \infty \right) \right].
 \end{multline}
-Applying this once to the foreground samples, once to
-the background samples and taking the product, we obtain $p(d,\mathset{f_i}, N | R_f, R_b, \theta)$ and thence $p(\mathset{f_i}, R_f, R_b, \theta | d,N) = p(d,\mathset{f_i}, N | R_f, R_b, \theta) p*R_f, R_b, \theta)/p(d,N)$.
-With the identification $p(d,N) = N!\,p(d_{\rm to},N)$, as justified above, we reproduce Eq.~\eqref{eq:posterior}.
+Applying this once to the foreground samples, once to the background
+samples and taking the product, we obtain $p(d,\mathset{f_i}, N | R_f,
+R_b, \theta)$ and thence $p(\mathset{f_i}, R_f, R_b, \theta | d,N) =
+p(d,\mathset{f_i}, N | R_f, R_b, \theta) p*R_f, R_b,
+\theta)/p(d,N)$. With the identification $p(d,N) = N!\,p(d_{\rm
+  to},N)$, as justified above, we reproduce Eq.~\eqref{eq:posterior}.
 
@@
   p\left( R_f, R_b, \theta | d, N \right) = \sum_{\mathset{f_i} \in
-    \mathset{0,1}^N} p\left( \mathset{f_i}, R_f, R_b, \theta | d, N \right)
-  \propto \prod_{i} \left[ R_f \hat{f}\left(x_i, \theta\right) + R_b
-    \hat{b}\left( x_i, \theta\right) \right] \exp\left[-\left( R_f +
-      R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
+    \mathset{0,1}^N} p\left( \mathset{f_i}, R_f, R_b, \theta | d, N
+  \right) \propto \prod_{i} \left[ R_f \hat{f}\left(x_i, \theta\right)
+    + R_b \hat{b}\left( x_i, \theta\right) \right] \exp\left[-\left(
+    R_f + R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
 \end{equation}
@@
 Eq.~\eqref{eq:rate-shape-posterior} contains only continuous
-parameters.
-We note that the terms that depend on the overall rate parameters, $R_b$ or $R_f$, are of the form $R_b^{n-1/2} \exp(-R_b)$ and so marginalisation over either $R_b$ or $R_f$ can be achieved analytically using
+parameters. We note that the terms that depend on the overall rate
+parameters, $R_b$ or $R_f$, are of the form $R_b^{n-1/2} \exp(-R_b)$
+and so marginalisation over either $R_b$ or $R_f$ can be achieved
+analytically using
 \begin{equation}
-I_n = \int_0^{\infty} x^{n-\frac{1}{2}} \, {\rm e}^{-x} {\rm d} x = \frac{(2n-1)!!}{2^n} \sqrt{\pi}
+I_n = \int_0^{\infty} x^{n-\frac{1}{2}} \, {\rm e}^{-x} {\rm d} x =
+\frac{(2n-1)!!}{2^n} \sqrt{\pi}
 \end{equation}
@@
 \end{equation}
-where $f$ and $b$ are rate densities on the $k$-dimensional
-space of ranking statistics.  We give an example of fitting for
+where $f$ and $b$ are rate densities on the $k$-dimensional space of
+ranking statistics.  We give an example of fitting for
 multi-dimensional rate densities in \S~\ref{sec:star-cluster}.
 
-It is informative to relate these results to two other ways to estimate the foreground rate parameter --- the loudest event statistic and the foreground-dominated statistic.
+It is informative to relate these results to two other ways to
+estimate the foreground rate parameter --- the loudest event statistic
+and the foreground-dominated statistic.
 
 \subsubsection{Loudest event statistic}
-If we were to include only the $k$-loudest events in the posterior distribution, rather than all observed events, the posterior (Eq.~\eqref{eq:posterior}) would be modified by an additional factor of $\exp[R_f \hat{F}(x_{N-k+1}) + R_b \hat{B}(x_{N-k+1})]$, where we have assumed events are ordered by loudness, so that $x_{N-k+1}$ is the $k$-th loudest event.
-This term accounts for the data-dependent threshold that a loudest event statistic employs.
+If we were to include only the $k$-loudest events in the posterior
+distribution, rather than all observed events, the posterior
+(Eq.~\eqref{eq:posterior}) would be modified by an additional factor
+of $\exp[R_f \hat{F}(x_{N-k+1}) + R_b \hat{B}(x_{N-k+1})]$, where we
+have assumed events are ordered by loudness, so that $x_{N-k+1}$ is
+the $k$-th loudest event. This term accounts for the data-dependent
+threshold that a loudest event statistic employs.
 
-For the usual $k=1$ case, the marginalised posterior (Eq.~\eqref{eq:rate-shape-posterior}) becomes
+For the usual $k=1$ case, the marginalised posterior
+(Eq.~\eqref{eq:rate-shape-posterior}) becomes
 \begin{equation}
-p_{\rm LE}\left( R_f, R_b, \theta | d \right) \propto \left(R_f \hat{f}(x_N,\theta) + R_b \hat{b}(x_N,\theta)\right)\exp\left[-\left( R_f (1-\hat{F}(x_N,\theta)) +
-      R_b (1-\hat{B}(x_N,\theta)) \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
+p_{\rm LE}\left( R_f, R_b, \theta | d \right) \propto \left(R_f
+\hat{f}(x_N,\theta) + R_b \hat{b}(x_N,\theta)\right)\exp\left[-\left(
+  R_f (1-\hat{F}(x_N,\theta)) + R_b (1-\hat{B}(x_N,\theta)) \right)
+  \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
 \end{equation}
-where $x_N$ denotes the loudness of the loudest event. Marginalising over $R_b$ we obtain
+where $x_N$ denotes the loudness of the loudest event. Marginalising
+over $R_b$ we obtain
 \begin{equation}
-p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f \hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}} \exp(-R_f(1-\hat{F}(x_N,\theta))) .
+p_{\rm LE}\left( R_f, \theta | d \right) \propto \left(
+\frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f
+\hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}}
+\exp(-R_f(1-\hat{F}(x_N,\theta))) .
 \end{equation}
@@
 \begin{equation}
-R_f = \frac{\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta) \pm \sqrt{(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta))^2 - 4\tilde{b}(x_N,\theta)(1-\hat{F}(x_N,\theta))\hat{f}(x_N,\theta)}}{4 \hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta)}.
+R_f =
+\frac{\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta)
+  \pm
+  \sqrt{(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta))^2
+    -
+    4\tilde{b}(x_N,\theta)(1-\hat{F}(x_N,\theta))\hat{f}(x_N,\theta)}}{4
+  \hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta)}.
 \end{equation}
-where $\tilde{b}(x_N,\theta) =\hat{b}(x_N,\theta)/(1-\hat{B}(x_N,\theta))$.
-If $\tilde{b}(x_N,\theta) \ll 1$, i.e., the loudest event is very likely to be foreground, we obtain the result $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$.
-This is can be understood as the statement that the rate of foreground events with ranking statistic greater than $x_N$,  $(1- \hat{F}(x_N,\theta))\,R_f$, is of order $1$, as expected.
-If $\tilde{b}(x_N,\theta) >> 1$, however, the posterior on $R_f$ is peaked at $0$.
+where $\tilde{b}(x_N,\theta)
+=\hat{b}(x_N,\theta)/(1-\hat{B}(x_N,\theta))$. If
+$\tilde{b}(x_N,\theta) \ll 1$, i.e., the loudest event is very likely
+to be foreground, we obtain the result $(1- \hat{F}(x_N,\theta))\,R_f
+\approx 1/2$. This is can be understood as the statement that the rate
+of foreground events with ranking statistic greater than $x_N$, $(1-
+\hat{F}(x_N,\theta))\,R_f$, is of order $1$, as expected. If
+$\tilde{b}(x_N,\theta) >> 1$, however, the posterior on $R_f$ is
+peaked at $0$.
 
 \subsubsection{Foreground dominated statistic}
-If we set the threshold for including an event, $x_{\rm min}$, sufficiently high, we can ensure that $\hat{f}(x_i,\theta) \gg \hat{b}(x_i,\theta)$ for all ranking statistics $x_i$ in the data set.
-The posterior can then be approximated by
+If we set the threshold for including an event, $x_{\rm min}$,
+sufficiently high, we can ensure that $\hat{f}(x_i,\theta) \gg
+\hat{b}(x_i,\theta)$ for all ranking statistics $x_i$ in the data
+set. The posterior can then be approximated by
 \begin{equation}
-p_{\rm FD}\left( R_f, R_b, \theta | d \right) \propto  \prod_{i} \left[ \hat{f}\left(x_i, \theta\right)\right] R_f^N \exp\left[-\left( R_f + R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
+p_{\rm FD}\left( R_f, R_b, \theta | d \right) \propto \prod_{i} \left[
+  \hat{f}\left(x_i, \theta\right)\right] R_f^N \exp\left[-\left( R_f +
+  R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
 \end{equation}
-Normalisation over $R_b$ gives a constant factor and the posterior on the foreground rate becomes
+Normalisation over $R_b$ gives a constant factor and the posterior on
+the foreground rate becomes
 \begin{equation}
-p_{\rm FD}\left( R_f, \theta | d \right) \propto  \prod_{i} \left[ \hat{f}\left(x_i, \theta\right)\right] R_f^{N-\frac{1}{2}} \exp\left[-R_f  \right] p(\theta).
+p_{\rm FD}\left( R_f, \theta | d \right) \propto \prod_{i} \left[
+  \hat{f}\left(x_i, \theta\right)\right] R_f^{N-\frac{1}{2}}
+\exp\left[-R_f \right] p(\theta).
 \end{equation}
-Ignoring the dependence on $\theta$, this is peaked at a rate $R_f = N-1/2$, so we have the expected result that, in the foreground dominated regime, the rate is approximately equal to the number of events observed.
-We note that this is the rate of events occurring above threshold and so if we were to use this statistic with different choices of threshold we would have to renormalise the rates to a fixed threshold in order to compare the results.
+Ignoring the dependence on $\theta$, this is peaked at a rate $R_f =
+N-1/2$, so we have the expected result that, in the foreground
+dominated regime, the rate is approximately equal to the number of
+events observed. We note that this is the rate of events occurring
+above threshold and so if we were to use this statistic with different
+choices of threshold we would have to renormalise the rates to a fixed
+threshold in order to compare the results.
 
@@
 stream by matched filtering in the frequency domain against a set of
-$N$ template waveforms \citep[e.g.,][]{findchirppaper,LVC2011}.  In our simplistic
-model, we suppose the data stream consists of stationary Gaussian
-noise with a power spectral density $S(f)$ combined additively with
-some number of gravitational wave signals.  We assume that the signals
-are sufficiently rare that they do not overlap in the data stream.
-The signal to noise ratio (SNR) of a template, $h(f)$, given data,
-$d(f)$, is
+$N$ template waveforms \citep[e.g.,][]{findchirppaper,LVC2011}.  In
+our simplistic model, we suppose the data stream consists of
+stationary Gaussian noise with a power spectral density $S(f)$
+combined additively with some number of gravitational wave signals.
+We assume that the signals are sufficiently rare that they do not
+overlap in the data stream.  The signal to noise ratio (SNR) of a
+template, $h(f)$, given data, $d(f)$, is
 \begin{equation}
@@
 \begin{equation}
-  \left\langle a, b \right\rangle \equiv \ilya{4 \Re} \int_0^\infty df\,
-  \frac{a^*(f) b(f)}{S(f)}.
+  \left\langle a, b \right\rangle \equiv \ilya{4 \Re} \int_0^\infty
+  df\, \frac{a^*(f) b(f)}{S(f)}.
 \end{equation}
@@
 \end{equation}
-Imposing the threshold, $x > \xmin$, the
-cumulative distribution of the background becomes
+Imposing the threshold, $x > \xmin$, the cumulative distribution of
+the background becomes
 \begin{equation}
   \label{eq:analytic-background-rate}
-  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}} \right)
-    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
-    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
-    \right)^N }
+  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}}
+    \right) \right)^N - \left( 1 + \erf\left(
+    \frac{\xmin}{\ilya{\sqrt{2}}} \right) \right)^N}{2^N - \left( 1 +
+    \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right) \right)^N }
 \end{equation}
@@
 
-Note that this scenario has no shape parameters $\theta$ for the foreground and
-background distributions.
+Note that this scenario has no shape parameters $\theta$ for the
+foreground and background distributions.
 
@@
 \begin{equation}
-  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left( \frac{x}{2} \right)
-    \right)^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
-    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
-    \right)^{N_\mathrm{eff}} }.
+  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left(
+    \frac{x}{2} \right) \right)^{N_\mathrm{eff}} - \left( 1 +
+    \erf\left( \frac{\xmin}{2} \right)
+    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 +
+    \erf\left( \frac{\xmin}{2} \right) \right)^{N_\mathrm{eff}} }.
 \end{equation}
@@
   \label{eq:plummer-surface-density}
-  \hat{f}(\vec{x}, \theta) = \frac{1}{\pi r_0^2 \left( 1 + \frac{\left|
-          \vec{x} - \vec{x}_0 \right|^2}{r_0^2} \right)^2},
+  \hat{f}(\vec{x}, \theta) = \frac{1}{\pi r_0^2 \left( 1 +
+    \frac{\left| \vec{x} - \vec{x}_0 \right|^2}{r_0^2} \right)^2},
 \end{equation}
@@
 \begin{equation}
-  \hat{b}\left(\vec{x}, \theta\right) = 1 + \vec{\gamma} \cdot \left( \vec{x}
-    - \vec{x}_{1/2} \right),
+  \hat{b}\left(\vec{x}, \theta\right) = 1 + \vec{\gamma} \cdot \left(
+  \vec{x} - \vec{x}_{1/2} \right),
 \end{equation}
@@
 To analyze our synthetic data set, we analytically marginalized over
-the state flags, using the likelihood in Eq.~\eqref{eq:rate-shape-posterior}.
-\ilya{[I think it would be kind to the reader to be more explicit
-    about the connection with the formalism in Section II.  What are
-    the "events"?  What is the ranking statistic $x$?  What are the
-    flags (which aren't in $\mathbb{R}$?)?]}  We did this to take
-advantage of the \textit{emcee} sampler of \citet{ForemanMackey2012},
-which requires all parameters to be in $\mathbb{R}$.  We applied a
-prior on the shape parameters that is flat in $\vec{x}_0$ and
-$\vec{\gamma}$, and the Jeffreys prior on $r_0$,
+the state flags, using the likelihood in
+Eq.~\eqref{eq:rate-shape-posterior}.  \ilya{[I think it would be kind
+    to the reader to be more explicit about the connection with the
+    formalism in Section II.  What are the "events"?  What is the
+    ranking statistic $x$?  What are the flags (which aren't in
+    $\mathbb{R}$?)?]}  We did this to take advantage of the
+\textit{emcee} sampler of \citet{ForemanMackey2012}, which requires
+all parameters to be in $\mathbb{R}$.  We applied a prior on the shape
+parameters that is flat in $\vec{x}_0$ and $\vec{\gamma}$, and the
+Jeffreys prior on $r_0$,
 \begin{equation}
@@
 
-Comments on background probability estimation --  Cannon et al. in prep.
+Comments on background probability estimation -- Cannon et al. in
+prep.
 
@@
 %%%%%%%%%%%%%%%%%%%% Bibliography 
-\bibliographystyle{apsrev4-1}
-\bibliography{many}
+\bibliographystyle{apsrev4-1} \bibliography{many}
 

### 2012-08-16 many.tex
Some edits to cluster example.

@@
 
+\will{Look up von Hippel paper about Bayes 9, the cluster isochrone
+  fitting software---we should cite, since our approach to cluster
+  membership.}
+
 Our final example concerns fitting for the location and shape
 parameters of a cluster of stars observed on top of a stellar
-background with a density gradient.  We assume that a star cluster has
-a Plummer surface-density profile \citep{Plummer1911,Aarseth1974},
+background with a density gradient.  In this example, stars are either
+members of the cluster (i.e.~foreground) or background contamination,
+with a spatially varying density (i.e.~our rate functions are
+two-dimensional).  We assume that a star cluster has a Plummer
+surface-density profile \citep{Plummer1911,Aarseth1974},
 \begin{equation}
@@
 where $\vec{x}_0$ is the location on the sky of the center of the
-cluster, and $r_0$ is a radial scale parameter.  We assume a square
-observational domain\footnote{The observational domain is not
-  infinite, so the normalization of the cluster density in
+cluster, $r_0$ is a radial scale parameter, and $\vec{x} = \left( x, y
+\right)$ is the position on the sky.  We assume a square observational
+domain\footnote{The observational domain is not infinite, so the
+  normalization of the cluster density in
   Eq.~\eqref{eq:plummer-surface-density} is not quite correct.  In our
@@
     all $\vec{\gamma}$ if $\vec{x}_{1/2}$ is in the center of a
-    symmetric domain (as it happens to be in this case).]}  We use
-assume a gradient for the background of $\vec{\gamma} = [-0.5, 0.5]$.
-\ilya{[I find it a bit confusing to use the word "assume" for both the
-    model and for the specific parameters injected into the simulation
-    (but then being fit for, so their values are not "assumed" in the
-    recovery).  Maybe "assume" for the model, and "simulate" for the
-    parameter values to be fit?]}  We assume that the total number of
-background stars, $R_b = 10000$, exceeds by a factor of ten the total
-number of cluster stars, $R_f = 1000$.  \ilya{[But the cluster has
-    only $\sim 0.01$ of the total area, so it's over-dense by a factor
-    of $10$ or more.  Fig.~6 shows a very bright spot where the
-    cluster is, which sort of makes the problem look too simple for
-    the chosen parameters...]}  We choose a true cluster center
-$\vec{x}_0 = \vec{x}_{1/2}$, and a radial scale parameter of $r_0 =
-0.1$.  In all, the shape parameters for the true distributions are
+    symmetric domain (as it happens to be in this case).]}  \will{Yes,
+  as you say, this function is only normalized for $\vec{x}_{1/2}$ is
+  the center, and also for small enough $\vec{\gamma}$; both of these
+  conditions obtain here.}
+
+We use simulated data drawn from our model with parameters
 \begin{equation}
@@
 \theta_0 \equiv \left\{ x_0, y_0, r_0, \gamma_x, \gamma_y \right\} =
-\left\{ \frac{1}{2}, \frac{1}{2}, \frac{1}{10}, -\frac{1}{2},
-\frac{1}{2} \right\}.
+\left\{ \frac{1}{2}, \frac{1}{2}, 0.18, -\frac{1}{2},
+\frac{1}{2} \right\},
 \end{equation}
-Figure \ref{fig:sky-density} shows the density on the sky due to the
-cluster foreground and the background with these parameters.  We drew
-a synthetic data set from this density according to the corresponding
-inhomogeneous Poisson process.
+with $R_f = 1000$ and $R_b = 10000$.  For this set of parameters, the
+average density of the background and the peak density of the cluster
+are comparable; there are an order of magnitude more background stars
+than cluster stars in the field.  Figure \ref{fig:sky-density} shows
+the density of stars (cluster and background) on the sky for our
+particular data set.
 
@@
   \caption{\label{fig:sky-density} The density contours on the sky of
-    the forergound cluster and background discussed in \S
-    \ref{sec:star-cluster}, with true parameters given in
-    Eq.~\eqref{eq:true-cluster-parameters}.  There are a factor of 10
-    more stars (10000) in the background than in the cluster (1000),
-    but the peak density for the cluster is much higher.  \ilya{[It
-        seems too obvious that where the cluster is by eye, and
-        probably fairly straightforward to compute the number of
-        "excess" stars there...  Reader might ask why we need
-        something so complicated when the answer is seems so obvious
-        and a "frequentist" approach would seem to do here.]}}
+    our synthetic data set of cluster and background stars for the
+    parametrs in Eq.~\eqref{eq:true-cluster-parameters}.  There are a
+    factor of 10 more stars (10000) in the background than in the
+    cluster (1000), but the average density of cluster and background
+    stars is similar.}
 \end{figure}
@@
 To analyze our synthetic data set, we analytically marginalized over
-the state flags, using the likelihood in
-Eq.~\eqref{eq:rate-shape-posterior}.  \ilya{[I think it would be kind
-    to the reader to be more explicit about the connection with the
-    formalism in Section II.  What are the "events"?  What is the
-    ranking statistic $x$?  What are the flags (which aren't in
-    $\mathbb{R}$?)?]}  We did this to take advantage of the
-\textit{emcee} sampler of \citet{ForemanMackey2012}, which requires
-all parameters to be in $\mathbb{R}$.  We applied a prior on the shape
-parameters that is flat in $\vec{x}_0$ and $\vec{\gamma}$, and the
-Jeffreys prior on $r_0$,
+the state flags (i.e.~cluster membership), using the likelihood in
+Eq.~\eqref{eq:rate-shape-posterior}.  We did this to take advantage of
+the \textit{emcee} sampler of \citet{ForemanMackey2012}, which
+requires all parameters to be in $\mathbb{R}$.  We applied a prior on
+the shape parameters that is flat in $\vec{x}_0$ and $\vec{\gamma}$,
+and an (approximately) Jeffreys prior on $r_0$,
 \begin{equation}
@@
     Jeffreys', but definitely not Jeffrey's ;) )?  It's not obvious
-    why the prior on $r_0$ should depend on $R_f$?]}  (Note that this
-factor of $\sqrt{R_f}$ cancels with the Jeffreys prior on the rate,
-$1/\sqrt{R_f}$; we have verified that the priors on these parameters
-are irrelevant to our results, as would be expected from the
-measurement of $\sim 1000$ foreground stars.)
+    why the prior on $r_0$ should depend on $R_f$?]}  \will{Here is
+  the calculation:
+\begin{equation}
+  p\left( r_0 \right) = \sqrt{\left\langle \left(\frac{\partial \log
+      \mathcal{L}}{\partial r_0 } \right)^2 \right\rangle} = \sqrt{
+    R_f \left\langle \left(\frac{\partial \log\hat{f}}{\partial r_0
+    }\right)^2 \right\rangle } = \sqrt{R_f \left\langle \left(
+    \frac{2}{r_0} - \frac{4 r_0}{r^2 + r_0^2} \right)^2 \right\rangle}
+  = \sqrt{\frac{ 4 R_f}{3 r_0^2}} \propto \frac{\sqrt{R_f}}{r_0}.
+\end{equation}
+Of course, we are ignoring the fact that $R_f$ and $r_0$ are
+correlated.  In truth, the Jeffreys prior would be the determinant of
+the matrix built out of terms like $\partial \log\mathcal{L} /
+\partial \theta_i \times \partial \log\mathcal{L} /\partial \theta_j$,
+but let's not get carried away\ldots.  As another aside, note that we
+could have predicted the $1/r_0$ dependence because $r_0$ has units,
+so the only unit-invariant prior is to be flat in $\log r_0$; I'm
+don't have a similar argument for the appearance of the $\sqrt{R_f}$
+factor.}  (Note that this factor of $\sqrt{R_f}$ cancels with the
+Jeffreys prior on the rate, $1/\sqrt{R_f}$; we have verified that the
+priors on these parameters are irrelevant to our results, as would be
+expected from the measurement of $\sim 1000$ foreground stars.)
 
@@
 parameters, $\vec{x}_0$; the center of the cluster is localized to
-within a few percent of its scale.  Figure \ref{fig:cluster-number}
-shows the posteriors inferred on the cluster and background numbers,
-$R_f$ and $R_b$, and Figure \ref{fig:cluster-scale} shows the
-posterior for the cluster's scale parameter.  In spite of the
-significant background, the cluster scale and total number are
-accurately recovered by our analysis.
+within about 10\% of the cluster scale.  Figure
+\ref{fig:cluster-number} shows the posteriors inferred on the cluster
+and background numbers, $R_f$ and $R_b$, and Figure
+\ref{fig:cluster-scale} shows the posterior for the cluster's scale
+parameter.  In spite of the significant background, the cluster scale
+and total number are accurately recovered by our analysis.
 

### 2012-08-16 many.tex
Finished section on loudest event and foreground dominated statistics. Added figure for GW example.

--- 2012-08-15 many.tex
@@
 \newcommand{\will}[1]{{\color{blue} \bf #1}}
-\newcommand{\jon}[1]{{\color{orange} \bf  #1}}
+\newcommand{\jon}[1]{{\color{green} \bf  #1}}
 
@@
 
-We assume that we are presented with a data set of $N$ events that exceed a pre-specified threshold in ranking statistic, $x_{min}$.  Each
+We assume that we are presented with a data set of $N$ events that exceed a pre-specified threshold in ranking statistic, $x_{\rm min}$.  Each
 event may be due to either a signal of interest or an uninteresting
@@
 \begin{equation}
-p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f \hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}} \exp(-R_f(1-\hat{F}(x_N,\theta))) .
+p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f \hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}\, p(\theta)}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}} \exp\left(-R_f(1-\hat{F}(x_N,\theta))\right) .
 \end{equation}
@@
 where $\tilde{b}(x_N,\theta) =\hat{b}(x_N,\theta)/(1-\hat{B}(x_N,\theta))$.
-If $\tilde{b}(x_N,\theta) \ll 1$, i.e., the loudest event is very likely to be foreground, we obtain the result $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$.
+If $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$.
 This is can be understood as the statement that the rate of foreground events with ranking statistic greater than $x_N$,  $(1- \hat{F}(x_N,\theta))\,R_f$, is of order $1$, as expected.
-If $\tilde{b}(x_N,\theta) >> 1$, however, the posterior on $R_f$ is peaked at $0$.
+However, $\tilde{b}(x_N,\theta) = -{\rm d} [\ln(1-\hat{B}(x,\theta))]/{\rm d}x$ and $(1-\hat{B}(x,\theta)) \rightarrow 0$ as $x \rightarrow \infty$, so this term will be divergent and for many reasonable examples, we will find $\tilde{b}(x_N,\theta) >> \hat{f}(x_N,\theta)$, in which case the posterior on $R_f$ is peaked at $0$.
+This issue highlights the problem with using a loudest-event statistic with an improper prior on the background rate $R_b$.
+No matter how improbable an event with $x=x_N$ is under the background distribution, it can become likely that the event at $x_N$ is from the background distribution by taking the background rate to be sufficiently large.
+Although this predicts many more events with $x < x_N$, by using only the loudest event we do not incorporate the information that no such events are seen.
+This problem is avoided in the new framework described here, since we use all events detected above threshold and combined rates, $R_f+R_b$, significantly greater than the total number of observed events are strongly disfavoured.
+
+This problem can be avoided in the loudest event framework, by including an upper limit on the rate, $R_{\rm max}$, in the prior for $R_b$.
+The marginalised distribution for the foreground rate then becomes
+\begin{eqnarray}
+p_{\rm LE}\left( R_f, \theta | d \right) &\propto& \left( \frac{\hat{b}(x_N,\theta)}{(1-\hat{B}(x_N,\theta))^{\frac{3}{2}}} \left[\frac{\sqrt{\pi}}{2}{\rm erf}\left(\sqrt{(1-\hat{B}(x_N,\theta))\,R_{\rm max}}\right) - \sqrt{(1-\hat{B}(x_N,\theta))\,R_{\rm max}} \, {\rm e}^{-(1-\hat{B}(x_N,\theta)) \, R_{\rm max}}\right] \right.
+\nonumber \\
+&& \left.+ R_f \hat{f}(x_N,\theta)\frac{\sqrt{\pi}\,{\rm erf}\left(\sqrt{(1-\hat{B}(x_N,\theta))\,R_{\rm max}}\right)}{\sqrt{(1-\hat{B}(x_N,\theta))}}\right)\frac{p(\theta)}{\sqrt{R_f}} \exp\left(-R_f(1-\hat{F}(x_N,\theta))\right) \label{LEwithRmax},
+\end{eqnarray}
+where ${\rm erf}(x)$ is the error function, defined in the usual way ${\rm erf}(x) = (2/\sqrt{\pi}) \int_0^x \exp(-u^2){\rm d}u$.
+If $(1-\hat{B}(x_N,\theta)) \, R_{\rm max} \ll 1$, Eq.~(\ref{LEwithRmax}) can be approximated by
+\begin{equation}
+p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{R_{\rm max}}{3} \hat{b}(x_N,\theta) + R_f \hat{f}(x_N,\theta)\right)\frac{p(\theta)}{\sqrt{R_f}} \exp\left(-R_f(1-\hat{F}(x_N,\theta))\right)
+\end{equation}
+and if $\hat{f}(x_N,\theta) \gg R_{\rm max} \hat{b}(x_N.\theta)$ we find the same result as before $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$.
 
@@
 Ignoring the dependence on $\theta$, this is peaked at a rate $R_f = N-1/2$, so we have the expected result that, in the foreground dominated regime, the rate is approximately equal to the number of events observed.
-We note that this is the rate of events occurring above threshold and so if we were to use this statistic with different choices of threshold we would have to renormalise the rates to a fixed threshold in order to compare the results.
+We note that the rate $R_f$ is defined as the rate of events occurring above the specified threshold and therefore if the foreground dominated statistic were used to estimated rates $R_f^1$ and $R_f^2$ for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$ respectively, these rates should be compared by equating $R_f^1 (1-\hat{F}(x_2,\theta))$ and $R_f^2 (1-\hat{F}(x_1,\theta))$.
 
@@
 \begin{equation}
-  \left\langle a, b \right\rangle \equiv \ilya{4 \Re} \int_0^\infty df\,
+  \left\langle a, b \right\rangle \equiv 4 \Re \int_0^\infty df\,
   \frac{a^*(f) b(f)}{S(f)}.
@@
 \begin{equation}
-  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}}
+  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\sqrt{2}}
       \right)}{2} \right)^N
 \end{equation}
-Imposing the threshold, $x > \xmin$, the
+where ${\rm erf}(x)$ is the error function as before. Imposing the threshold, $x > \xmin$, the
 cumulative distribution of the background becomes
@@
   \label{eq:analytic-background-rate}
-  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}} \right)
-    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
-    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
+  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{{\sqrt{2}}} \right)
+    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{{\sqrt{2}}} \right)
+    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{{\sqrt{2}}} \right)
     \right)^N }
 \end{equation}
-\ilya{for $x>\xmin$, $0$ otherwise.}
+for $x>\xmin$, $0$ otherwise.
 
@@
 
+We can compare these results to results obtained using the two approximations described earlier, the loudest-event statistic and the foreground-dominated statistic.
+The marginalised distribution for the foreground rate using these alternatives are shown in Figure~\ref{fig:altmeth}.
+In this case, the loudest event had $x_N=9.5$\jon{(correct this)}.
+The loudest-event statistic depends on a specification of the maximum, $R_{\rm max}$, for the background rate.
+We show results for $R_{\rm max}=\infty$, i.e.m, the improper prior, and $R_{\rm max}=10000$.
+The results for other reasonable  choices of $R_{\rm max} = 100, 1000, 100000$ etc.
+gave exactly the same posterior, since $\hat{b}(x_N,) R_{\rm max} \ll \hat{f}(x_N)$ for all these choices and we are therefore in the regime where the posterior is insensitive to $R_{\rm max}$.
+To apply the foreground-dominated statistic we must specify a threshold above which we assume all events are foreground.
+It is reasonable to do this based on a specification for the relative probability of an event being fore/background, $\hat{f}(x)/\hat{b}(x)=p_{\rm thresh}$.
+Setting $p_{\rm thresh}=0.99$ gives $x_{\rm min}=4.07$ and there are $N=13$ events exceeding that threshold.
+Setting $p_{\rm thresh}=0.5$ gives $x_{\rm min}=3.82$ and there are $N=15$ events exceeding that threshold \jon{(correct these numbers using Will's data)}.
+We show results for both of these choices of $x_{\rm min}$ in Figure~\ref{fig:altmeth}.
+We also show similar results for two extreme cases --- setting $x_{\rm min}=3.5$ so that we assume all $98$ observed events are foreground; and setting $x_{\rm min}=9$ so that only the loudest event is classified as foreground.
+
+The loudest event statistic with the improper-prior gives, as expected, a poor approximation to the foreground rate.
+The peak is more accurately located when a prior maximum rate is defined, but the distribution is much wider than using the full analysis described here.
+This is to be expected as much of the information is being thrown away.
+The foreground-dominated statistic gives a reasonable approximation to the true foreground rate, and a distribution that is not much wider than the full analysis, for the non-extreme choices of $x_{\rm min}$.
+For $x_{\rm min}=9$, the foreground dominated statistic has equivalent performance to the loudest event statistic, as we might expect, while when $x_{\rm min}=3.5$, it performs poorly since we are approximating the foreground rate by the total foreground plus background rate.
+This indicates that, provided the threshold is chosen appropriately, the foreground dominated statistic can perform quite well at estimating the rate.
+The fact that it reproduces the posterior from the full analysis so well is indicative of the fact that most of the information about the foreground comes from the loudest events.
+Indeed, for $x_{\rm min} \approx 4$, the set of events consists mostly of foreground sources and that set contains a decent number of events ($13$), so it is not surprising that this statistic can estimate the foreground rate well.
+The choice of $x_{\rm min}$ was motivated by knowledge of $\hat{b}(x,\theta)$ and $\hat{f}(x,\theta)$, which we have in this case, but this might not be possible if the parameters $\theta$ were simultaneously being estimated.
+The full analysis naturally incorporates inference about the rate parameters $\theta$ and the background rate $R_b$ along with the foreground rate and incorporates maximum information from the data set and should therefore lead to narrower posteriors in general.
+
 \ilya{[General comment on the text: I think we need to say a bit more
@@
     background from foreground for individual events, but that doesn't
-    necessarily matter for rate measurements.  Now, one interesting
-    question that Fig.~3 brings to mind is the following: if we only
-    used "gold-plated" events (say, those with
-    $p(x|fore)/p(x|back)>0.99$, how much worse / more uncertain would
-    the rate estimation be?]}
+    necessarily matter for rate measurements.}
 
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{AltMethodComp}
+  \caption{\label{fig:altmeth}Posteriors on foreground rate obtained using alternative methods --- the loudest event statistic and the foreground dominated statistic.
+Results are shown for two different choices of the maximum a priori background rate, $R_{\rm max}$, in the case of the loudest event (``LE'') statistic, and for several different choices of the threshold, $x_{\rm min}$, in the case of the foreground dominated (``FD'') statistic.}
+\end{figure}
+
 \subsection{Gravitational Waves With Overlapping Templates}
@@
 In \S \ref{sec:analytic-GW-example} we assumed that the overlap
-between \ilya{different} templates in the template bank \ilya{was
-  negligible}, so the SNRs \ilya{recovered by} different templates are
+between different templates in the template bank was
+  negligible, so the SNRs recovered by different templates are
 independent random variables.  In fact, template banks are not
@@
 of $N$ actual templates will behave as if it had $N_\mathrm{eff}$
-\emph{independent} templates.  \ilya{Rather than pre-computing
-  $N_\mathrm{eff}$,} we can fit for \ilya{it as a} shape parameter.
-That is, we assume that \ilya{$\theta = \{N_\mathrm{eff}\}$} is a
+\emph{independent} templates.   Rather than pre-computing
+  $N_\mathrm{eff}$, we can fit for it as a shape parameter.
+That is, we assume that $\theta = \{N_\mathrm{eff}\}$ is a
 shape parameter for the background cumulative distribution:
 \begin{equation}
-  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left( \frac{x}{2} \right)
-    \right)^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
-    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
+  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left( \frac{x}{\sqrt{2}} \right)
+    \right)^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{\sqrt{2}} \right)
+    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{\sqrt{2}} \right)
     \right)^{N_\mathrm{eff}} }.
@@
 
-\ilya{[Again, I think we need to explicitly compare this method with
-    two alternatives.  Alternative one is the loudest event statistic.
-    Alternative two is just counting gold-plated events, and dividing
-    the volume in which we are sensitive to them by their number.  In
-    principle, alternative one should be suboptimal when the number of
-    events is more than 0 or 1, while alternative two should be OK
-    when the number of events is large, but error-prone otherwise.
-    But if we want people to use this technique, we have to defend the
-    point that there is a regime where it is clearly superior (and say
-    what that regime is).]}
+If we consider the two alternative methods, the loudest event and foreground dominated statistics, we will recover the same foreground distributions as in the case with $N_{\rm eff}$ fixed, which are shown in Figure~\ref{fig:altmeth}.
+This is because the parameter $N_{\rm eff}$ affects only the background distribution, to which the foreground-dominated statistic is insensitive, and in the loudest event case, after marginalisation over $N_{\rm eff}$ we find $\int_0^{N_{\rm max}} \hat{b}(x_N,N_{\rm eff}) {\rm d}N_{\rm eff} \ll 3 N_{\rm max}/R_{\rm max}  \hat{f}(x_N, N_{\rm eff})$ and so we are still in the foreground-dominated regime in which the loudest event tells us nothing about the background.
+Similarly, neither of these alternative methods can inform us about the value of $N_{\rm eff}$, a property of the background.
+\jon{Technically, if we allowed really large values of $N_{\rm eff}$ we would get into a regime where the loudest event could be background and we'd get something here.
+But, we would need $N_{\rm max} \sim 10^{20}$ and so I don't think we should consider these alternatives further in this case.}
+
+%\ilya{[Again, I think we need to explicitly compare this method with
+    %two alternatives.  Alternative one is the loudest event statistic.
+    %Alternative two is just counting gold-plated events, and dividing
+    %the volume in which we are sensitive to them by their number.  In
+    %principle, alternative one should be suboptimal when the number of
+    %events is more than 0 or 1, while alternative two should be OK
+    %when the number of events is large, but error-prone otherwise.
+    %But if we want people to use this technique, we have to defend the
+    %point that there is a regime where it is clearly superior (and say
+    %what that regime is).]}
 

### 2012-08-16 many.tex
Better note about r_0 Jeffreys prior.

--- 2012-08-16 many.tex
@@
 \end{equation}
-Of course, we are ignoring the fact that $R_f$ and $r_0$ are
-correlated.  In truth, the Jeffreys prior would be the determinant of
-the matrix built out of terms like $\partial \log\mathcal{L} /
-\partial \theta_i \times \partial \log\mathcal{L} /\partial \theta_j$,
-but let's not get carried away\ldots.  As another aside, note that we
-could have predicted the $1/r_0$ dependence because $r_0$ has units,
-so the only unit-invariant prior is to be flat in $\log r_0$; I'm
-don't have a similar argument for the appearance of the $\sqrt{R_f}$
-factor.}  (Note that this factor of $\sqrt{R_f}$ cancels with the
-Jeffreys prior on the rate, $1/\sqrt{R_f}$; we have verified that the
-priors on these parameters are irrelevant to our results, as would be
-expected from the measurement of $\sim 1000$ foreground stars.)
+Note that the calculation of the prior uses the likelihood before
+analytical marginalization over the flags, $f_i$.  We are ignoring the
+fact that $R_f$ and $r_0$ are correlated.  In truth, the Jeffreys
+prior would be the determinant of the matrix built out of terms like
+$\partial \log\mathcal{L} / \partial \theta_i \times \partial
+\log\mathcal{L} /\partial \theta_j$, but let's not get carried
+away\ldots.  As another aside, note that we could have predicted the
+$1/r_0$ dependence because $r_0$ has units, so the only unit-invariant
+prior is to be flat in $\log r_0$; I'm don't have a similar argument
+for the appearance of the $\sqrt{R_f}$ factor.}  (Note that this
+factor of $\sqrt{R_f}$ cancels with the Jeffreys prior on the rate,
+$1/\sqrt{R_f}$; we have verified that the priors on these parameters
+are irrelevant to our results, as would be expected from the
+measurement of $\sim 1000$ foreground stars.)
 

### 2012-08-16 many.tex
erf(x/2) => erf(x/sqrt(2)) typo fixed.
Changed labels in data cumulative distribution.
Updated N-templates fitting.

@@
 \begin{equation}
-  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}}
+  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\sqrt{2}}
     \right)}{2} \right)^N
@@
   \label{eq:analytic-background-rate}
-  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}}
+  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\sqrt{2}}
     \right) \right)^N - \left( 1 + \erf\left(
-    \frac{\xmin}{\ilya{\sqrt{2}}} \right) \right)^N}{2^N - \left( 1 +
-    \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right) \right)^N }
+    \frac{\xmin}{\sqrt{2}} \right) \right)^N}{2^N - \left( 1 +
+    \erf\left( \frac{\xmin}{\sqrt{2}} \right) \right)^N }
 \end{equation}
@@
     \eqref{eq:analytic-foreground-rate} combined with $R_f = 10.4$ and
-    $R_b = 95.1$. \ilya{[I think both axis labels should say $x$
-        rather than $\rho$, since that's the ranking statistic ($\max
-        \rho$).]}}
+    $R_b = 95.1$.}
 \end{figure}
@@
   \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left(
-    \frac{x}{2} \right) \right)^{N_\mathrm{eff}} - \left( 1 +
-    \erf\left( \frac{\xmin}{2} \right)
+    \frac{x}{\sqrt{2}} \right) \right)^{N_\mathrm{eff}} - \left( 1 +
+    \erf\left( \frac{\xmin}{\sqrt{2}} \right)
     \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 +
-    \erf\left( \frac{\xmin}{2} \right) \right)^{N_\mathrm{eff}} }.
+    \erf\left( \frac{\xmin}{\sqrt{2}} \right) \right)^{N_\mathrm{eff}} }.
 \end{equation}
@@
     wider than those of Figure \ref{fig:analytic-rate-recovery}, in
-    spite of the extra parameter.  \ilya{[$R_f$ seems to be more of an
-        outlier than before; why was there a significant shift, given
-        that $N_\mathrm{eff}$ seems fairly spot on?]}}
+    spite of the extra parameter.}
 \end{figure}

### 2012-08-16 many.tex
Updates to analytic example per Ilya's comments; better figures, more captions.

@@
 
-For the usual $k=1$ case, the marginalised posterior
-(Eq.~\eqref{eq:rate-shape-posterior}) becomes
+For the usual $k=1$ case \citep{Biswas2009}, the marginalised
+posterior (Eq.~\eqref{eq:rate-shape-posterior}) becomes
 \begin{equation}
@@
 
-\ilya{[General comment on the text: I think we need to say a bit more
-    about the figures and the interpretation of the results, instead
-    of just relying on the reader to make the best of it.  For
-    example, even though it may seem obvious to us, we should
-    explicitly say that Fig.~2 shows that we succeed in recovering
-    rates within the expected uncertainty envelope.  On Fig.~3, could
-    say something obvious like point out that we cannot separate
-    background from foreground for individual events, but that doesn't
-    necessarily matter for rate measurements.  Now, one interesting
-    question that Fig.~3 brings to mind is the following: if we only
-    used "gold-plated" events (say, those with
+\ilya{Now, one interesting question that Fig.~3 brings to mind is the
+  following: if we only used "gold-plated" events (say, those with
   $p(x|fore)/p(x|back)>0.99$, how much worse / more uncertain would
@@
     \ref{sec:analytic-GW-example}.  The vertical lines indicate the
-    ``true'' values used to generate the synthetic data set.}
+    ``true'' values used to generate the synthetic data set.  Both the
+    true foreground and background rates lie well within the
+    probability envelope for $R_f$ and $R_b$.}
 \end{figure}
@@
     \ref{sec:analytic-GW-example} marginalized over all other
-    parameters.  True foreground events are in blue \ilya{(dark
-      grey)}, background events in red \ilya{(light grey)}.  The solid
-    line is the likelihood ratio
-    $p(x|\mathrm{foreground})/p(x|\mathrm{background}) =
-    \hat{f}(x)/\hat{b}(x)$ for this model \ilya{evaluated at the
-      simulated parameters (?)}; this exceeds the marginalized
-    foreground probability for many of the events because in this data
-    set there are approximately nine times as many background events
-    as foreground events. \ilya{[I'm not sure what the reader is meant
-        to take away from the solid lines If there's something worth
-        taking away, we should describe the message in the text.
-        Otherwise, if it only confuses things, we should drop it.]}}
+    parameters (errorbars are one standard deviation).  True
+    foreground events are in dark grey, background events in light
+    grey.  Even though our method cannot identify the status of most
+    events with confidence, it can still correctly estimate the rates
+    (Figure \ref{fig:analytic-rate-recovery}).}
 \end{figure}
@@
 
-In \S \ref{sec:analytic-GW-example} we assumed that the overlap
-between \ilya{different} templates in the template bank \ilya{was
-  negligible}, so the SNRs \ilya{recovered by} different templates are
-independent random variables.  In fact, template banks are not
-constructed in this way \citep[e.g.,][]{Owen:1998dk,Ajith:2008},
-%\citep{Caudill2011,Cannon2011}, 
-because signals could fall in the gaps between the non-overlapping
-templates.  We can model this effect by assuming that a template bank
-of $N$ actual templates will behave as if it had $N_\mathrm{eff}$
-\emph{independent} templates.  \ilya{Rather than pre-computing
-  $N_\mathrm{eff}$,} we can fit for \ilya{it as a} shape parameter.
-That is, we assume that \ilya{$\theta = \{N_\mathrm{eff}\}$} is a
-shape parameter for the background cumulative distribution:
+In \S~\ref{sec:analytic-GW-example} we assumed that the overlap
+between different templates in the template bank was negligible, so
+the SNRs recovered by different templates are independent random
+variables.  In fact, template banks are not constructed in this way
+\citep[e.g.,][]{Owen:1998dk,Ajith:2008}, because signals could fall in
+the gaps between the non-overlapping templates.  We can model this
+effect by assuming that a template bank of $N$ actual templates will
+behave as if it had $N_\mathrm{eff}$ \emph{independent} templates.
+Rather than pre-computing $N_\mathrm{eff}$, we can fit for it as a
+shape parameter.  That is, we assume that $\theta =
+\{N_\mathrm{eff}\}$ is a shape parameter for the background cumulative
+distribution:
 \begin{equation}
@@
 
-Results from such an analysis appear in Figures \ref{fig:rates-nt} and
-\ref{fig:ntemplates}.  We use the same parameters and data set as in
-\S \ref{sec:analytic-GW-example}, with $R_f = 10.4$, $R_b = 95.1$, and
-$N_\mathrm{eff} = 1000$, but now allow $N_\mathrm{eff}$ to be a
-parameter of the background distribution, with a flat prior.  Both the
-rates and the number of effective templates are recovered without
-significant loss of accuracy relative to the fixed $N_\mathrm{eff}$
-situation in \S \ref{sec:analytic-GW-example}.
+Results from such an analysis on the data in
+\S~\ref{sec:analytic-GW-example} appear in Figures \ref{fig:rates-nt}
+and \ref{fig:ntemplates}.  Recall that the parameters used to generate
+this data set are $R_f = 10.4$, $R_b = 95.1$, and $N_\mathrm{eff} =
+1000$.  We use a flat prior on $N_\mathrm{eff}$.  Both the rates and
+the number of effective templates are recovered without significant
+loss of accuracy relative to the fixed $N_\mathrm{eff}$ situation in
+\S~\ref{sec:analytic-GW-example}.
 

### 2012-08-16 many.tex
Typo in k-loudest events fixed.

@@
 (Eq.~\eqref{eq:posterior}) would be modified by an additional factor
-of $\exp[R_f \hat{F}(x_{N-k+1}) + R_b \hat{B}(x_{N-k+1})]$, where we
-have assumed events are ordered by loudness, so that $x_{N-k+1}$ is
-the $k$-th loudest event. This term accounts for the data-dependent
-threshold that a loudest event statistic employs.
+of $\exp[R_f \hat{F}(x_{N-k+1}, \theta) + R_b
+  \hat{B}(x_{N-k+1},\theta)]$, where we have assumed events are
+ordered by loudness, so that $x_{N-k+1}$ is the $k$-th loudest
+event. This term accounts for the data-dependent threshold that a
+loudest event statistic employs.
 

### 2012-08-16 many.tex

@@
 \newcommand{\will}[1]{{\color{blue} \bf #1}}
-\newcommand{\jon}[1]{{\color{orange} \bf #1}}
+\newcommand{\jon}[1]{{\color{green} \bf  #1}}
 
@@
 
-We assume that we are presented with a data set of $N$ events that
-exceed a pre-specified threshold in ranking statistic, $x_{min}$.
-Each event may be due to either a signal of interest or an
-uninteresting background.  Each event is associated with a ranking
-statistic, $x$.  Our data set therefore consists of the ranking
-statistics for the set of events:
+We assume that we are presented with a data set of $N$ events that exceed a pre-specified threshold in ranking statistic, $x_{\rm min}$.  Each
+event may be due to either a signal of interest or an uninteresting
+background.  Each event is associated with a ranking statistic, $x$.
+Our data set therefore consists of the ranking statistics for the set
+of events:
 \begin{equation}
@@
 (Eq.~\eqref{eq:posterior}) would be modified by an additional factor
-of $\exp[R_f \hat{F}(x_{N-k+1}, \theta) + R_b
+of $\exp[R_f \hat{F}(x_{N-k+1},\theta) + R_b
   \hat{B}(x_{N-k+1},\theta)]$, where we have assumed events are
@@
 \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f
-\hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}}
-\exp(-R_f(1-\hat{F}(x_N,\theta))) .
+\hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}\,
+  p(\theta)}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}}
+\exp\left(-R_f(1-\hat{F}(x_N,\theta))\right) .
 \end{equation}
@@
 =\hat{b}(x_N,\theta)/(1-\hat{B}(x_N,\theta))$. If
-$\tilde{b}(x_N,\theta) \ll 1$, i.e., the loudest event is very likely
-to be foreground, we obtain the result $(1- \hat{F}(x_N,\theta))\,R_f
-\approx 1/2$. This is can be understood as the statement that the rate
-of foreground events with ranking statistic greater than $x_N$, $(1-
-\hat{F}(x_N,\theta))\,R_f$, is of order $1$, as expected. If
-$\tilde{b}(x_N,\theta) >> 1$, however, the posterior on $R_f$ is
-peaked at $0$.
+$\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-
+\hat{F}(x_N,\theta))\,R_f \approx 1/2$. This is can be understood as
+the statement that the rate of foreground events with ranking
+statistic greater than $x_N$, $(1- \hat{F}(x_N,\theta))\,R_f$, is of
+order $1$, as expected. However, $\tilde{b}(x_N,\theta) = -{\rm d}
+[\ln(1-\hat{B}(x,\theta))]/{\rm d}x$ and $(1-\hat{B}(x,\theta))
+\rightarrow 0$ as $x \rightarrow \infty$, so this term will be
+divergent and for many reasonable examples, we will find
+$\tilde{b}(x_N,\theta) >> \hat{f}(x_N,\theta)$, in which case the
+posterior on $R_f$ is peaked at $0$. This issue highlights the problem
+with using a loudest-event statistic with an improper prior on the
+background rate $R_b$. No matter how improbable an event with $x=x_N$
+is under the background distribution, it can become likely that the
+event at $x_N$ is from the background distribution by taking the
+background rate to be sufficiently large. Although this predicts many
+more events with $x < x_N$, by using only the loudest event we do not
+incorporate the information that no such events are seen. This problem
+is avoided in the new framework described here, since we use all
+events detected above threshold and combined rates, $R_f+R_b$,
+significantly greater than the total number of observed events are
+strongly disfavoured.
+
+This problem can be avoided in the loudest event framework, by
+including an upper limit on the rate, $R_{\rm max}$, in the prior for
+$R_b$. The marginalised distribution for the foreground rate then
+becomes
+\begin{eqnarray}
+p_{\rm LE}\left( R_f, \theta | d \right) &\propto& \left(
+\frac{\hat{b}(x_N,\theta)}{(1-\hat{B}(x_N,\theta))^{\frac{3}{2}}}
+\left[\frac{\sqrt{\pi}}{2}{\rm
+    erf}\left(\sqrt{(1-\hat{B}(x_N,\theta))\,R_{\rm max}}\right) -
+  \sqrt{(1-\hat{B}(x_N,\theta))\,R_{\rm max}} \, {\rm
+    e}^{-(1-\hat{B}(x_N,\theta)) \, R_{\rm max}}\right]
+\right. \nonumber \\ && \left.+ R_f
+\hat{f}(x_N,\theta)\frac{\sqrt{\pi}\,{\rm
+    erf}\left(\sqrt{(1-\hat{B}(x_N,\theta))\,R_{\rm
+      max}}\right)}{\sqrt{(1-\hat{B}(x_N,\theta))}}\right)\frac{p(\theta)}{\sqrt{R_f}}
+\exp\left(-R_f(1-\hat{F}(x_N,\theta))\right) \label{LEwithRmax},
+\end{eqnarray}
+where ${\rm erf}(x)$ is the error function, defined in the usual way
+${\rm erf}(x) = (2/\sqrt{\pi}) \int_0^x \exp(-u^2){\rm d}u$. If
+$(1-\hat{B}(x_N,\theta)) \, R_{\rm max} \ll 1$, Eq.~(\ref{LEwithRmax})
+can be approximated by
+\begin{equation}
+p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{R_{\rm
+    max}}{3} \hat{b}(x_N,\theta) + R_f
+\hat{f}(x_N,\theta)\right)\frac{p(\theta)}{\sqrt{R_f}}
+\exp\left(-R_f(1-\hat{F}(x_N,\theta))\right)
+\end{equation}
+and if $\hat{f}(x_N,\theta) \gg R_{\rm max} \hat{b}(x_N.\theta)$ we
+find the same result as before $(1- \hat{F}(x_N,\theta))\,R_f \approx
+1/2$.
 
@@
 dominated regime, the rate is approximately equal to the number of
-events observed. We note that this is the rate of events occurring
-above threshold and so if we were to use this statistic with different
-choices of threshold we would have to renormalise the rates to a fixed
-threshold in order to compare the results.
+events observed. We note that the rate $R_f$ is defined as the rate of
+events occurring above the specified threshold and therefore if the
+foreground dominated statistic were used to estimated rates $R_f^1$
+and $R_f^2$ for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$
+respectively, these rates should be compared by equating $R_f^1
+(1-\hat{F}(x_2,\theta))$ and $R_f^2 (1-\hat{F}(x_1,\theta))$.
 
@@
 \begin{equation}
-  \left\langle a, b \right\rangle \equiv \ilya{4 \Re} \int_0^\infty
-  df\, \frac{a^*(f) b(f)}{S(f)}.
+  \left\langle a, b \right\rangle \equiv 4 \Re \int_0^\infty df\,
+  \frac{a^*(f) b(f)}{S(f)}.
 \end{equation}
@@
 \end{equation}
-Imposing the threshold, $x > \xmin$, the cumulative distribution of
-the background becomes
+where ${\rm erf}(x)$ is the error function as before. Imposing the
+threshold, $x > \xmin$, the cumulative distribution of the background
+becomes
 \begin{equation}
   \label{eq:analytic-background-rate}
-  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\sqrt{2}}
-    \right) \right)^N - \left( 1 + \erf\left(
-    \frac{\xmin}{\sqrt{2}} \right) \right)^N}{2^N - \left( 1 +
-    \erf\left( \frac{\xmin}{\sqrt{2}} \right) \right)^N }
+  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{{\sqrt{2}}} \right)
+    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{{\sqrt{2}}} \right)
+    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{{\sqrt{2}}} \right)
+    \right)^N }
 \end{equation}
-\ilya{for $x>\xmin$, $0$ otherwise.}
+for $x>\xmin$, $0$ otherwise.
 
@@
 
-\ilya{Now, one interesting question that Fig.~3 brings to mind is the
-  following: if we only used "gold-plated" events (say, those with
-  $p(x|fore)/p(x|back)>0.99$, how much worse / more uncertain would
-  the rate estimation be?]}
+We can compare these results to results obtained using the two
+approximations described earlier, the loudest-event statistic and the
+foreground-dominated statistic. The marginalised distribution for the
+foreground rate using these alternatives are shown in
+Figure~\ref{fig:altmeth}. In this case, the loudest event had
+$x_N=9.5$\jon{(correct this)}. The loudest-event statistic depends on
+a specification of the maximum, $R_{\rm max}$, for the background
+rate. We show results for $R_{\rm max}=\infty$, i.e.m, the improper
+prior, and $R_{\rm max}=10000$. The results for other reasonable
+choices of $R_{\rm max} = 100, 1000, 100000$ etc. gave exactly the
+same posterior, since $\hat{b}(x_N,) R_{\rm max} \ll \hat{f}(x_N)$ for
+all these choices and we are therefore in the regime where the
+posterior is insensitive to $R_{\rm max}$. To apply the
+foreground-dominated statistic we must specify a threshold above which
+we assume all events are foreground. It is reasonable to do this based
+on a specification for the relative probability of an event being
+fore/background, $\hat{f}(x)/\hat{b}(x)=p_{\rm thresh}$. Setting
+$p_{\rm thresh}=0.99$ gives $x_{\rm min}=4.07$ and there are $N=13$
+events exceeding that threshold. Setting $p_{\rm thresh}=0.5$ gives
+$x_{\rm min}=3.82$ and there are $N=15$ events exceeding that
+threshold \jon{(correct these numbers using Will's data)}. We show
+results for both of these choices of $x_{\rm min}$ in
+Figure~\ref{fig:altmeth}. We also show similar results for two extreme
+cases --- setting $x_{\rm min}=3.5$ so that we assume all $98$
+observed events are foreground; and setting $x_{\rm min}=9$ so that
+only the loudest event is classified as foreground.
+
+The loudest event statistic with the improper-prior gives, as
+expected, a poor approximation to the foreground rate. The peak is
+more accurately located when a prior maximum rate is defined, but the
+distribution is much wider than using the full analysis described
+here. This is to be expected as much of the information is being
+thrown away. The foreground-dominated statistic gives a reasonable
+approximation to the true foreground rate, and a distribution that is
+not much wider than the full analysis, for the non-extreme choices of
+$x_{\rm min}$. For $x_{\rm min}=9$, the foreground dominated statistic
+has equivalent performance to the loudest event statistic, as we might
+expect, while when $x_{\rm min}=3.5$, it performs poorly since we are
+approximating the foreground rate by the total foreground plus
+background rate. This indicates that, provided the threshold is chosen
+appropriately, the foreground dominated statistic can perform quite
+well at estimating the rate. The fact that it reproduces the posterior
+from the full analysis so well is indicative of the fact that most of
+the information about the foreground comes from the loudest
+events. Indeed, for $x_{\rm min} \approx 4$, the set of events
+consists mostly of foreground sources and that set contains a decent
+number of events ($13$), so it is not surprising that this statistic
+can estimate the foreground rate well. The choice of $x_{\rm min}$ was
+motivated by knowledge of $\hat{b}(x,\theta)$ and $\hat{f}(x,\theta)$,
+which we have in this case, but this might not be possible if the
+parameters $\theta$ were simultaneously being estimated. The full
+analysis naturally incorporates inference about the rate parameters
+$\theta$ and the background rate $R_b$ along with the foreground rate
+and incorporates maximum information from the data set and should
+therefore lead to narrower posteriors in general.
 
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{AltMethodComp}
+  \caption{\label{fig:altmeth}Posteriors on foreground rate obtained
+    using alternative methods --- the loudest event statistic and the
+    foreground dominated statistic. Results are shown for two
+    different choices of the maximum a priori background rate, $R_{\rm
+      max}$, in the case of the loudest event (``LE'') statistic, and
+    for several different choices of the threshold, $x_{\rm min}$, in
+    the case of the foreground dominated (``FD'') statistic.}
+\end{figure}
+
 \subsection{Gravitational Waves With Overlapping Templates}
@@
 
-In \S~\ref{sec:analytic-GW-example} we assumed that the overlap
+In \S \ref{sec:analytic-GW-example} we assumed that the overlap
 between different templates in the template bank was negligible, so
@@
 variables.  In fact, template banks are not constructed in this way
-\citep[e.g.,][]{Owen:1998dk,Ajith:2008}, because signals could fall in
-the gaps between the non-overlapping templates.  We can model this
-effect by assuming that a template bank of $N$ actual templates will
-behave as if it had $N_\mathrm{eff}$ \emph{independent} templates.
-Rather than pre-computing $N_\mathrm{eff}$, we can fit for it as a
-shape parameter.  That is, we assume that $\theta =
-\{N_\mathrm{eff}\}$ is a shape parameter for the background cumulative
-distribution:
+\citep[e.g.,][]{Owen:1998dk,Ajith:2008},
+%\citep{Caudill2011,Cannon2011}, 
+because signals could fall in the gaps between the non-overlapping
+templates.  We can model this effect by assuming that a template bank
+of $N$ actual templates will behave as if it had $N_\mathrm{eff}$
+\emph{independent} templates.  Rather than pre-computing
+$N_\mathrm{eff}$, we can fit for it as a shape parameter.  That is, we
+assume that $\theta = \{N_\mathrm{eff}\}$ is a shape parameter for the
+background cumulative distribution:
 \begin{equation}
-  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left(
-    \frac{x}{\sqrt{2}} \right) \right)^{N_\mathrm{eff}} - \left( 1 +
-    \erf\left( \frac{\xmin}{\sqrt{2}} \right)
-    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 +
-    \erf\left( \frac{\xmin}{\sqrt{2}} \right) \right)^{N_\mathrm{eff}} }.
+  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left( \frac{x}{\sqrt{2}} \right)
+    \right)^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{\sqrt{2}} \right)
+    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{\sqrt{2}} \right)
+    \right)^{N_\mathrm{eff}} }.
 \end{equation}
 
-Results from such an analysis on the data in
-\S~\ref{sec:analytic-GW-example} appear in Figures \ref{fig:rates-nt}
-and \ref{fig:ntemplates}.  Recall that the parameters used to generate
-this data set are $R_f = 10.4$, $R_b = 95.1$, and $N_\mathrm{eff} =
-1000$.  We use a flat prior on $N_\mathrm{eff}$.  Both the rates and
-the number of effective templates are recovered without significant
-loss of accuracy relative to the fixed $N_\mathrm{eff}$ situation in
-\S~\ref{sec:analytic-GW-example}.
+Results from such an analysis appear in Figures \ref{fig:rates-nt} and
+\ref{fig:ntemplates}.  We use the same parameters and data set as in
+\S \ref{sec:analytic-GW-example}, with $R_f = 10.4$, $R_b = 95.1$, and
+$N_\mathrm{eff} = 1000$, but now allow $N_\mathrm{eff}$ to be a
+parameter of the background distribution, with a flat prior.  Both the
+rates and the number of effective templates are recovered without
+significant loss of accuracy relative to the fixed $N_\mathrm{eff}$
+situation in \S \ref{sec:analytic-GW-example}.
 
-\ilya{[Again, I think we need to explicitly compare this method with
-    two alternatives.  Alternative one is the loudest event statistic.
-    Alternative two is just counting gold-plated events, and dividing
-    the volume in which we are sensitive to them by their number.  In
-    principle, alternative one should be suboptimal when the number of
-    events is more than 0 or 1, while alternative two should be OK
-    when the number of events is large, but error-prone otherwise.
-    But if we want people to use this technique, we have to defend the
-    point that there is a regime where it is clearly superior (and say
-    what that regime is).]}
+If we consider the two alternative methods, the loudest event and
+foreground dominated statistics, we will recover the same foreground
+distributions as in the case with $N_{\rm eff}$ fixed, which are shown
+in Figure~\ref{fig:altmeth}. This is because the parameter $N_{\rm
+  eff}$ affects only the background distribution, to which the
+foreground-dominated statistic is insensitive, and in the loudest
+event case, after marginalisation over $N_{\rm eff}$ we find
+$\int_0^{N_{\rm max}} \hat{b}(x_N,N_{\rm eff}) {\rm d}N_{\rm eff} \ll
+3 N_{\rm max}/R_{\rm max} \hat{f}(x_N, N_{\rm eff})$ and so we are
+still in the foreground-dominated regime in which the loudest event
+tells us nothing about the background. Similarly, neither of these
+alternative methods can inform us about the value of $N_{\rm eff}$, a
+property of the background. \jon{Technically, if we allowed really
+  large values of $N_{\rm eff}$ we would get into a regime where the
+  loudest event could be background and we'd get something here. But,
+  we would need $N_{\rm max} \sim 10^{20}$ and so I don't think we
+  should consider these alternatives further in this case.}
 
@@
     example from \S~\ref{sec:star-cluster}.  The center $(x,y) =
-    \left(x_0, y_0\right)$ is determined to within a few percent of
-    the structural radius of the cluster, $r_0$ (see
+    \left(x_0, y_0\right)$ is determined to within about 5\%
+    of the structural radius of the cluster, $r_0$ (see
     Eq.~\eqref{eq:true-cluster-parameters}). \ilya{[Given the cluster
         size of 0.1, it seems a bit surprising that the center is off
-        by $\sim 0.05$...]}}
+        by $\sim 0.05$...]} \will{That is $0.1*r_0$, not $0.1$!}}
 \end{figure}

### 2012-09-03 many.tex
Minor edits to Section II.

@@
   University, 2145 Sheridan Road, Evanston, IL 60208}
-%\affiliation{Northwestern University and CIERA}
 
@@
   Birmingham\\United Kingdom}
-%\affiliation{University of Birmingham}
+
+\author{Curt Cutler} \email{Curt.J.Cutler@jpl.nasa.gov}
+\affiliation{Jet Propulsion Lab, 4800 Oak Grove Dr., Pasadena, CA 91109}
+\affiliation{Theoretical Astrophysics, California Institute of Technology, Pasadena, California 91125}
 
@@
 This is the probability that the first observed event is a
-fore/background event (if $f_i=1,0$) with ranking statistic $x_1$ and
-the second observed event is a fore/background event (if $f_i=1,0$)
+fore/background event (if $f_1=1,0$) with ranking statistic $x_1$ and
+the second observed event is a fore/background event (if $f_2=1,0$)
 with ranking statistic $x_2$ etc. If the events are ordered by ranking
 statistic the corresponding expression is more complicated, since
-$x_1$ is now the event from foreground or background, with the
-smallest ranking statistic etc. We will return to the
+$x_1$ is now the event from foreground or background with the
+smallest ranking statistic, etc. We will return to the
 statistic-ordered case later.
@@
 \end{equation}
-When sampling the posterior, the first term can be omitted and the
-equals sign replaced by proportionality, since it is independent of
-the parameters of interest, but we have kept this term explicitly so
+When sampling the posterior, the first term, which is independent of
+the parameters of interest, can be omitted and the
+equals sign replaced by proportionality; however, we have kept this term explicitly so
 that we can see the equivalence to ranking-statistic ordered
@@
   p\left( \mathset{y_i} | r \right) \,{\rm d}^Ny_i= P\left(
-  \textnormal{zero events below } y_0 \right) \times P\left(
-  \textnormal{one event between } y_0 \textnormal{ and } y_0 + {\rm
-    d}y_0 \right) \\ \times P\left( \textnormal{zero events between }
-  y_0+{\rm d}y_0 \textnormal{ and } y_1 \right) \ldots \\ p\left(
+  \textnormal{zero events below } y_1 \right) \times P\left(
+  \textnormal{one event between } y_1 \textnormal{ and } y_1 + {\rm
+    d}y_1 \right) \\ \times P\left( \textnormal{zero events between }
+  y_1+{\rm d}y_1 \textnormal{ and } y_2 \right) \ldots \\ p\left(
   \mathset{y_i} | r \right) = \lim_{\delta y_i \to 0} \exp\left[ -
-    R\left(y_0\right) \right] \left[ r\left( y_0 \right) +
-    \order{\delta y_0} \right] \times \exp\left[ - \left[ R\left( y_1
-      \right) - R\left( y_0 + \delta y_0\right) \right] \right] \times
+    R\left(y_1\right) \right] \left[ r\left( y_1 \right) +
+    \order{\delta y_1} \right] \times \exp\left[ - \left[ R\left( y_2
+      \right) - R\left( y_1 + \delta y_1\right) \right] \right] \times
   \ldots \\ = \left[\prod_i r\left( y_i \right)\right] \exp\left[ -
@@
 R_b, \theta)$ and thence $p(\mathset{f_i}, R_f, R_b, \theta | d,N) =
-p(d,\mathset{f_i}, N | R_f, R_b, \theta) p*R_f, R_b,
-\theta)/p(d,N)$. With the identification $p(d,N) = N!\,p(d_{\rm
+p(d,\mathset{f_i}, N | R_f, R_b, \theta)\ p(R_f, R_b,
+\theta)\, /\, p(d,N)$. With the identification $p(d,N) = N!\,p(d_{\rm
   to},N)$, as justified above, we reproduce Eq.~\eqref{eq:posterior}.
@@
 
-It is informative to relate these results to two other ways to
-estimate the foreground rate parameter --- the loudest event statistic
+It is informative to relate these results to two other methods for
+estimating the foreground rate parameter --- the loudest event statistic
 and the foreground-dominated statistic.
@@
 \subsubsection{Loudest event statistic}
-If we were to include only the $k$-loudest events in the posterior
+If we were to include only the $k$ loudest events in the posterior
 distribution, rather than all observed events, the posterior
@@
 events occurring above the specified threshold and therefore if the
-foreground dominated statistic were used to estimated rates $R_f^1$
-and $R_f^2$ for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$
-respectively, these rates should be compared by equating $R_f^1
-(1-\hat{F}(x_2,\theta))$ and $R_f^2 (1-\hat{F}(x_1,\theta))$.
+foreground dominated statistic were used to estimate rates $R_{f,1}$
+and $R_{f,2}$ for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$
+respectively, these rates should be compared by equating $R_{f,1}
+(1-\hat{F}(x_2,\theta))$ and $R_{f,2} (1-\hat{F}(x_1,\theta))$.
 

### 2012-09-04 many.tex
More minor edits to 2.

@@
 R_f =
-\frac{\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta)
-  \pm
-  \sqrt{(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta))^2
+\frac{\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta))\tilde{b}(x_N,\theta)
+  +
+  \sqrt{\left(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta))\tilde{b}(x_N,\theta)\right)^2
     -
     4\tilde{b}(x_N,\theta)(1-\hat{F}(x_N,\theta))\hat{f}(x_N,\theta)}}{4
-  \hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta)}.
+  \hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta))}.
 \end{equation}
@@
 $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-
-\hat{F}(x_N,\theta))\,R_f \approx 1/2$. This is can be understood as
+\hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be understood as
 the statement that the rate of foreground events with ranking
@@
 and if $\hat{f}(x_N,\theta) \gg R_{\rm max} \hat{b}(x_N.\theta)$ we
-find the same result as before $(1- \hat{F}(x_N,\theta))\,R_f \approx
+find the same result as before, $(1- \hat{F}(x_N,\theta))\,R_f \approx
 1/2$.
@@
 a specification of the maximum, $R_{\rm max}$, for the background
-rate. We show results for $R_{\rm max}=\infty$, i.e.m, the improper
+rate. We show results for $R_{\rm max}=\infty$, i.e., the improper
 prior, and $R_{\rm max}=10000$. The results for other reasonable
 choices of $R_{\rm max} = 100, 1000, 100000$ etc. gave exactly the
-same posterior, since $\hat{b}(x_N,) R_{\rm max} \ll \hat{f}(x_N)$ for
+same posterior, since $\hat{b}(x_N) R_{\rm max} \ll \hat{f}(x_N)$ for
 all these choices and we are therefore in the regime where the
@@
 
-The loudest event statistic with the improper-prior gives, as
+The loudest event statistic with the improper prior gives, as
 expected, a poor approximation to the foreground rate. The peak is
@@
     \ref{sec:analytic-GW-example} marginalized over all other
-    parameters (errorbars are one standard deviation).  True
+    parameters (error bars are one standard deviation).  True
     foreground events are in dark grey, background events in light
@@
     for several different choices of the threshold, $x_{\rm min}$, in
-    the case of the foreground dominated (``FD'') statistic.}
+    the case of the foreground dominated (``FD'') statistic. \ilya{I can see the figure on the screen, but can't print it...}}
 \end{figure}
@@
   we would need $N_{\rm max} \sim 10^{20}$ and so I don't think we
-  should consider these alternatives further in this case.}
+  should consider these alternatives further in this case.} \ilya{On the other hand, this ties in with the comment at the end of the previous subsection about the difficulty of choosing $x_{min}$ for the foreground-dominated statistic when there are other parameters $\theta$ to be estimated simultaneously ($N_{eff}$ in this case), so perhaps it's worthwhile to add some comment about this?}
 
@@
     of the structural radius of the cluster, $r_0$ (see
-    Eq.~\eqref{eq:true-cluster-parameters}). \ilya{[Given the cluster
-        size of 0.1, it seems a bit surprising that the center is off
-        by $\sim 0.05$...]} \will{That is $0.1*r_0$, not $0.1$!}}
+    Eq.~\eqref{eq:true-cluster-parameters}).}
 \end{figure}

### 2012-09-04 many.tex
Done with first draft of intro.

@@
 
-\fixme{Introduce the necessity of estimating rates, prior work (like
-  \cite{Biswas2009}), Bayesian inference.}
+The task of estimating a rate of foreground events when a mixture of foreground and background events is present in the data is a common one in physical and astrophysical applications.
+This problem comes up, among others, in gravitational-wave data analysis \cite[e.g.,][]{s5-highmass,s6-lowmass} and in astronomical observations of a field of objects of mixed provenance \cite[GC cluster paper]{}..
+In this paper, we introduce a robust formalism for estimating event rates from the data when the shape of foreground and background distributions are known (but may have additional free parameters), but the provenance of individual events as either background or foreground is unknown.
 
-Difficulty of measuring rate in presence of background
+We use a Bayesian approach and consider all available data to ensure that the inferred rates are both unbiased and maximally constrained in the presence of limited observations.
+Bayes' theorem yields the \emph{posterior} probability density function on a set of parameters, $\vec{\theta}$, given the observed data, $d$, under a model $M$:
+\begin{equation}
+p(\vec{\theta} | d, M) = \frac{p(\vec{\theta}|M) p(d|\vec{\theta},M)}{p(d|M)},
+\end{equation}
+where $p(\vec{\theta}|M)$ are the \emph{prior} probabilities of the model parameters, $p(d|\vec{\theta},M)$ is the \emph{likelihood} of obtaining the data given a particular choice of parameters, and the normalizing factor $p(d|M)$ is known as the \emph{evidence}.
 
-Want to add more than just the loudest event (cf. loudest event
-statistic)
+Two alternative approaches have been suggested in the past.
+One, known as the \emph{loudest-event statistic} \cite{Biswas2009}, uses only the information from the highest-ranked event in the data to infer the rate distribution.
+This approach has been used successfully \cite{s5-highmass,s6-lowmass} when the number of loud foreground events is small (typically zero or one) to obtain upper limits on foreground rates.
+However, the loudest-event statistic ignores all events except the loudest one, and so suffers from an unnecessary loss of information; therefore, we expect it to yield a much larger variance than strictly necessary when multiple events are present in the data.
 
-Unknown background distribution, $\Lambda$
+Another possible approach is based on the use of only loud, ``gold-plated'' events, ones which are certain (or nearly certain) to come from the foreground, to derive rates.
+We refer to this approach as the \emph{foreground-dominated statistic}.
+The foreground-dominated statistic may yield accurate results when the foreground and background are cleanly separated, at least for the loudest events, and the number of such loud events is sufficiently large.
+However, it cannot properly account for marginal events.
+While either the loudest-event statistic or the foreground-dominated statistic can approach the accuracy of our proposed method in specific regimes, both are suboptimal in a general case.
 
-Cf. number of gold-plated detections in sensitive volume --
-frequentist approach OK if tons of events, otherwise need this.
+In order to demonstrate our method, we consider three different examples.
+The first two come from the field of gravitational-wave data analysis, but could equally arise in any application that employs matched filtering \cite{FINDCHIRP} to extract weak signals with known shapes from the data.
+The last example considers the case of a globular cluster on a background of field stars.
+Throughout, we compare the results obtained with our technique to the loudest-event and foreground-dominated statistics, which make use of a limited subset of the available information.
 
-\section{Model}
+ This paper is organized as follows...
+
+%\fixme{Introduce the necessity of estimating rates, prior work (like
+%  \cite{Biswas2009}), Bayesian inference.}
+
+%Difficulty of measuring rate in presence of background
+
+%Want to add more than just the loudest event (cf. loudest event statistic)
+
+%Unknown background distribution, $\Lambda$
+
+%Cf. number of gold-plated detections in sensitive volume --
+%frequentist approach OK if tons of events, otherwise need this.
+
+\section{Model}\label{sec:model}
 

### 2012-09-04 many.tex
Added conclusion, a couple of references.

@@
 In order to demonstrate our method, we consider three different examples.
-The first two come from the field of gravitational-wave data analysis, but could equally arise in any application that employs matched filtering \cite{FINDCHIRP} to extract weak signals with known shapes from the data.
+The first two come from the field of gravitational-wave data analysis, but could equally arise in any application that employs matched filtering \cite{findchirppaper} to extract weak signals with known shapes from the data.
 The last example considers the case of a globular cluster on a background of field stars.
@@
 
- This paper is organized as follows...
-
-%\fixme{Introduce the necessity of estimating rates, prior work (like
-%  \cite{Biswas2009}), Bayesian inference.}
-
-%Difficulty of measuring rate in presence of background
-
-%Want to add more than just the loudest event (cf. loudest event statistic)
-
-%Unknown background distribution, $\Lambda$
-
-%Cf. number of gold-plated detections in sensitive volume --
-%frequentist approach OK if tons of events, otherwise need this.
+This paper is organized as follows.
+In section \ref{sec:model} we introduce our method for Bayesian rate estimation, and describe loudest-event and foreground-dominated statistics for comparison.
+Section \ref{sec:GW-example} includes three examples of the application of these techniques.
+We conclude with a discussion of important remaining issues in section \ref{sec:discussion}.
 
@@
 
-\section{Conclusion}
+\section{Discussion}\label{sec:discussion}
 
-Separating classes of events
+In this paper, we have developed a Bayesian framework for rate estimation when the data consists of a mixture of foreground and background events.
+We demonstrated the application of this framework using several examples from gravitational-wave data analysis in the presence of signatures of binary mergers and noise triggers, and astronomical image analysis in the presence of several populations of stars.
+We showed that this framework is generally superior to both the loudest-event statistic and and the foreground-dominated statistic.
 
-Comments on background probability estimation -- Cannon et al. in
-prep.
+Throughout this paper, we have assumed that the shape of the foreground and background distributions is known, or at least can be modeled with several additional parameters.
+This is not necessarily easy to do.
+For example, in the case of gravitational-wave data analysis, the shape of the foreground distribution of events may depend on the details of a complex data-analysis pipeline as well as the astrophysical source distribution, while the background event distribution depends on data quality and may deviate significantly from the simple Gaussian-noise behavior modeled in section \ref{sec:GW-example}.
+Several approaches have been developed to accurately model both distributions, e.g., through the use of injected signals to model the background \cite{ihope,CannonHannaKeppel:2012}.
+However, this is a difficult problem (e.g., because of the need to estimate the background at the very tails of the distribution), and will require significant future work.
 
-Relevance (vs. loudest statistic, or frequentist approach using only
-gold-plated events)
+A further complication is that we have considered the rate of events in the data as products of some analysis pipeline.
+This rate may be different from the physical rate of interest, such as the rate of compact-binary mergers per unit time per unit volume which generate gravitational waves, or the physical numbers of stars in the cluster and field populations which produce the observed luminosities.
+Again, the conversion between the two will depend on the details of the data-analysis algorithm and ranking statistic, including any selection effects, and would need to be determined on a case-by-case basis.
+
+Furthermore, in a practical application there could be multiple classes of events, not just foreground and background.
+For example, we are not necessarily interested in the rate of gravitational-wave signals per se, but separately in the rate of signals from mergers of binary neutron stars and binary black holes -- populations that may sometimes be difficult to distinguish.
+Our approach is readily extendable to this particular complication, however.
+Note that it is symmetric with respect to foreground and background events (as expected, since one physicist's background is another physicist's foreground).
+We could relabel foreground and background events into other competing event classes, and further classes could be added in a straightforward way.
 
@@
   We thank Kipp Cannon, Chad Hanna, Drew Keppel, and Richard
-  O'Shaughnessy for discussions and suggestions about this manuscript.
+  O'Shaughnessy for discussions and suggestions about this manuscript.  IM thanks KITP.
 \end{acknowledgments}

### 2012-09-04 many.tex
Added a few lines to discussion.

@@
 We could relabel foreground and background events into other competing event classes, and further classes could be added in a straightforward way.
+However, the ability to distinguish classes relies on different distributions of their statistics.
+In general, rankings may need to be extended to include other statistics in addition to the signal ``loudness'' statistic in order to indicate both event significance and the probability of event attribution to a particular class.
 

### 2012-09-04 many.tex
Formatting adjustments of LaTeX document; no content changes.

@@
 \author{Curt Cutler} \email{Curt.J.Cutler@jpl.nasa.gov}
-\affiliation{Jet Propulsion Lab, 4800 Oak Grove Dr., Pasadena, CA 91109}
-\affiliation{Theoretical Astrophysics, California Institute of Technology, Pasadena, California 91125}
+\affiliation{Jet Propulsion Lab, 4800 Oak Grove Dr., Pasadena, CA
+  91109} \affiliation{Theoretical Astrophysics, California Institute
+  of Technology, Pasadena, California 91125}
 
@@
 
-The task of estimating a rate of foreground events when a mixture of foreground and background events is present in the data is a common one in physical and astrophysical applications.
-This problem comes up, among others, in gravitational-wave data analysis \cite[e.g.,][]{s5-highmass,s6-lowmass} and in astronomical observations of a field of objects of mixed provenance \cite[GC cluster paper]{}..
-In this paper, we introduce a robust formalism for estimating event rates from the data when the shape of foreground and background distributions are known (but may have additional free parameters), but the provenance of individual events as either background or foreground is unknown.
+The task of estimating a rate of foreground events when a mixture of
+foreground and background events is present in the data is a common
+one in physical and astrophysical applications.  This problem comes
+up, among others, in gravitational-wave data analysis
+\cite[e.g.,][]{s5-highmass,s6-lowmass} and in astronomical
+observations of a field of objects of mixed provenance \cite[GC
+  cluster paper]{}..  In this paper, we introduce a robust formalism
+for estimating event rates from the data when the shape of foreground
+and background distributions are known (but may have additional free
+parameters), but the provenance of individual events as either
+background or foreground is unknown.
 
-We use a Bayesian approach and consider all available data to ensure that the inferred rates are both unbiased and maximally constrained in the presence of limited observations.
-Bayes' theorem yields the \emph{posterior} probability density function on a set of parameters, $\vec{\theta}$, given the observed data, $d$, under a model $M$:
+We use a Bayesian approach and consider all available data to ensure
+that the inferred rates are both unbiased and maximally constrained in
+the presence of limited observations.  Bayes' theorem yields the
+\emph{posterior} probability density function on a set of parameters,
+$\vec{\theta}$, given the observed data, $d$, under a model $M$:
 \begin{equation}
-p(\vec{\theta} | d, M) = \frac{p(\vec{\theta}|M) p(d|\vec{\theta},M)}{p(d|M)},
+p(\vec{\theta} | d, M) = \frac{p(\vec{\theta}|M)
+  p(d|\vec{\theta},M)}{p(d|M)},
 \end{equation}
-where $p(\vec{\theta}|M)$ are the \emph{prior} probabilities of the model parameters, $p(d|\vec{\theta},M)$ is the \emph{likelihood} of obtaining the data given a particular choice of parameters, and the normalizing factor $p(d|M)$ is known as the \emph{evidence}.
+where $p(\vec{\theta}|M)$ are the \emph{prior} probabilities of the
+model parameters, $p(d|\vec{\theta},M)$ is the \emph{likelihood} of
+obtaining the data given a particular choice of parameters, and the
+normalizing factor $p(d|M)$ is known as the \emph{evidence}.
 
-Two alternative approaches have been suggested in the past.
-One, known as the \emph{loudest-event statistic} \cite{Biswas2009}, uses only the information from the highest-ranked event in the data to infer the rate distribution.
-This approach has been used successfully \cite{s5-highmass,s6-lowmass} when the number of loud foreground events is small (typically zero or one) to obtain upper limits on foreground rates.
-However, the loudest-event statistic ignores all events except the loudest one, and so suffers from an unnecessary loss of information; therefore, we expect it to yield a much larger variance than strictly necessary when multiple events are present in the data.
+Two alternative approaches have been suggested in the past.  One,
+known as the \emph{loudest-event statistic} \cite{Biswas2009}, uses
+only the information from the highest-ranked event in the data to
+infer the rate distribution.  This approach has been used successfully
+\cite{s5-highmass,s6-lowmass} when the number of loud foreground
+events is small (typically zero or one) to obtain upper limits on
+foreground rates.  However, the loudest-event statistic ignores all
+events except the loudest one, and so suffers from an unnecessary loss
+of information; therefore, we expect it to yield a much larger
+variance than strictly necessary when multiple events are present in
+the data.
 
-Another possible approach is based on the use of only loud, ``gold-plated'' events, ones which are certain (or nearly certain) to come from the foreground, to derive rates.
-We refer to this approach as the \emph{foreground-dominated statistic}.
-The foreground-dominated statistic may yield accurate results when the foreground and background are cleanly separated, at least for the loudest events, and the number of such loud events is sufficiently large.
-However, it cannot properly account for marginal events.
-While either the loudest-event statistic or the foreground-dominated statistic can approach the accuracy of our proposed method in specific regimes, both are suboptimal in a general case.
+Another possible approach is based on the use of only loud,
+``gold-plated'' events, ones which are certain (or nearly certain) to
+come from the foreground, to derive rates.  We refer to this approach
+as the \emph{foreground-dominated statistic}.  The
+foreground-dominated statistic may yield accurate results when the
+foreground and background are cleanly separated, at least for the
+loudest events, and the number of such loud events is sufficiently
+large.  However, it cannot properly account for marginal events.
+While either the loudest-event statistic or the foreground-dominated
+statistic can approach the accuracy of our proposed method in specific
+regimes, both are suboptimal in a general case.
 
-In order to demonstrate our method, we consider three different examples.
-The first two come from the field of gravitational-wave data analysis, but could equally arise in any application that employs matched filtering \cite{findchirppaper} to extract weak signals with known shapes from the data.
-The last example considers the case of a globular cluster on a background of field stars.
-Throughout, we compare the results obtained with our technique to the loudest-event and foreground-dominated statistics, which make use of a limited subset of the available information.
+In order to demonstrate our method, we consider three different
+examples.  The first two come from the field of gravitational-wave
+data analysis, but could equally arise in any application that employs
+matched filtering \cite{findchirppaper} to extract weak signals with
+known shapes from the data.  The last example considers the case of a
+globular cluster on a background of field stars.  Throughout, we
+compare the results obtained with our technique to the loudest-event
+and foreground-dominated statistics, which make use of a limited
+subset of the available information.
 
-This paper is organized as follows.
-In section \ref{sec:model} we introduce our method for Bayesian rate estimation, and describe loudest-event and foreground-dominated statistics for comparison.
-Section \ref{sec:GW-example} includes three examples of the application of these techniques.
-We conclude with a discussion of important remaining issues in section \ref{sec:discussion}.
+This paper is organized as follows.  In section \ref{sec:model} we
+introduce our method for Bayesian rate estimation, and describe
+loudest-event and foreground-dominated statistics for comparison.
+Section \ref{sec:GW-example} includes three examples of the
+application of these techniques.  We conclude with a discussion of
+important remaining issues in section \ref{sec:discussion}.
 
@@
 
-We assume that we are presented with a data set of $N$ events that exceed a pre-specified threshold in ranking statistic, $x_{\rm min}$.  Each
-event may be due to either a signal of interest or an uninteresting
-background.  Each event is associated with a ranking statistic, $x$.
-Our data set therefore consists of the ranking statistics for the set
-of events:
+We assume that we are presented with a data set of $N$ events that
+exceed a pre-specified threshold in ranking statistic, $x_{\rm min}$.
+Each event may be due to either a signal of interest or an
+uninteresting background.  Each event is associated with a ranking
+statistic, $x$.  Our data set therefore consists of the ranking
+statistics for the set of events:
 \begin{equation}
@@
 statistic the corresponding expression is more complicated, since
-$x_1$ is now the event from foreground or background with the
-smallest ranking statistic, etc. We will return to the
-statistic-ordered case later.
+$x_1$ is now the event from foreground or background with the smallest
+ranking statistic, etc. We will return to the statistic-ordered case
+later.
 
@@
 When sampling the posterior, the first term, which is independent of
-the parameters of interest, can be omitted and the
-equals sign replaced by proportionality; however, we have kept this term explicitly so
-that we can see the equivalence to ranking-statistic ordered
-data. Once data have been observed, there is a unique loudness
+the parameters of interest, can be omitted and the equals sign
+replaced by proportionality; however, we have kept this term
+explicitly so that we can see the equivalence to ranking-statistic
+ordered data. Once data have been observed, there is a unique loudness
 ordering and time ordering of those events, and so there is a one to
@@
 R_b, \theta)$ and thence $p(\mathset{f_i}, R_f, R_b, \theta | d,N) =
-p(d,\mathset{f_i}, N | R_f, R_b, \theta)\ p(R_f, R_b,
-\theta)\, /\, p(d,N)$. With the identification $p(d,N) = N!\,p(d_{\rm
-  to},N)$, as justified above, we reproduce Eq.~\eqref{eq:posterior}.
+p(d,\mathset{f_i}, N | R_f, R_b, \theta)\ p(R_f, R_b, \theta)\, /\,
+p(d,N)$. With the identification $p(d,N) = N!\,p(d_{\rm to},N)$, as
+justified above, we reproduce Eq.~\eqref{eq:posterior}.
 
@@
 It is informative to relate these results to two other methods for
-estimating the foreground rate parameter --- the loudest event statistic
-and the foreground-dominated statistic.
+estimating the foreground rate parameter --- the loudest event
+statistic and the foreground-dominated statistic.
 
@@
 $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-
-\hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be understood as
-the statement that the rate of foreground events with ranking
-statistic greater than $x_N$, $(1- \hat{F}(x_N,\theta))\,R_f$, is of
-order $1$, as expected. However, $\tilde{b}(x_N,\theta) = -{\rm d}
+\hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be understood as the
+statement that the rate of foreground events with ranking statistic
+greater than $x_N$, $(1- \hat{F}(x_N,\theta))\,R_f$, is of order $1$,
+as expected. However, $\tilde{b}(x_N,\theta) = -{\rm d}
 [\ln(1-\hat{B}(x,\theta))]/{\rm d}x$ and $(1-\hat{B}(x,\theta))
@@
   \label{eq:analytic-background-rate}
-  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{{\sqrt{2}}} \right)
-    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{{\sqrt{2}}} \right)
-    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{{\sqrt{2}}} \right)
-    \right)^N }
+  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{{\sqrt{2}}}
+    \right) \right)^N - \left( 1 + \erf\left( \frac{\xmin}{{\sqrt{2}}}
+    \right) \right)^N}{2^N - \left( 1 + \erf\left(
+    \frac{\xmin}{{\sqrt{2}}} \right) \right)^N }
 \end{equation}
@@
     for several different choices of the threshold, $x_{\rm min}$, in
-    the case of the foreground dominated (``FD'') statistic. \ilya{I can see the figure on the screen, but can't print it...}}
+    the case of the foreground dominated (``FD'') statistic. \ilya{I
+      can see the figure on the screen, but can't print it...}}
 \end{figure}
@@
 \begin{equation}
-  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left( \frac{x}{\sqrt{2}} \right)
-    \right)^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{\sqrt{2}} \right)
-    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{\sqrt{2}} \right)
-    \right)^{N_\mathrm{eff}} }.
+  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left(
+    \frac{x}{\sqrt{2}} \right) \right)^{N_\mathrm{eff}} - \left( 1 +
+    \erf\left( \frac{\xmin}{\sqrt{2}} \right)
+    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 +
+    \erf\left( \frac{\xmin}{\sqrt{2}} \right) \right)^{N_\mathrm{eff}}
+  }.
 \end{equation}
@@
   we would need $N_{\rm max} \sim 10^{20}$ and so I don't think we
-  should consider these alternatives further in this case.} \ilya{On the other hand, this ties in with the comment at the end of the previous subsection about the difficulty of choosing $x_{min}$ for the foreground-dominated statistic when there are other parameters $\theta$ to be estimated simultaneously ($N_{eff}$ in this case), so perhaps it's worthwhile to add some comment about this?}
+  should consider these alternatives further in this case.} \ilya{On
+  the other hand, this ties in with the comment at the end of the
+  previous subsection about the difficulty of choosing $x_{min}$ for
+  the foreground-dominated statistic when there are other parameters
+  $\theta$ to be estimated simultaneously ($N_{eff}$ in this case), so
+  perhaps it's worthwhile to add some comment about this?}
 
@@
 \theta_0 \equiv \left\{ x_0, y_0, r_0, \gamma_x, \gamma_y \right\} =
-\left\{ \frac{1}{2}, \frac{1}{2}, 0.18, -\frac{1}{2},
-\frac{1}{2} \right\},
+\left\{ \frac{1}{2}, \frac{1}{2}, 0.18, -\frac{1}{2}, \frac{1}{2}
+\right\},
 \end{equation}
@@
     example from \S~\ref{sec:star-cluster}.  The center $(x,y) =
-    \left(x_0, y_0\right)$ is determined to within about 5\%
-    of the structural radius of the cluster, $r_0$ (see
+    \left(x_0, y_0\right)$ is determined to within about 5\% of the
+    structural radius of the cluster, $r_0$ (see
     Eq.~\eqref{eq:true-cluster-parameters}).}
@@
 
-In this paper, we have developed a Bayesian framework for rate estimation when the data consists of a mixture of foreground and background events.
-We demonstrated the application of this framework using several examples from gravitational-wave data analysis in the presence of signatures of binary mergers and noise triggers, and astronomical image analysis in the presence of several populations of stars.
-We showed that this framework is generally superior to both the loudest-event statistic and and the foreground-dominated statistic.
+In this paper, we have developed a Bayesian framework for rate
+estimation when the data consists of a mixture of foreground and
+background events.  We demonstrated the application of this framework
+using several examples from gravitational-wave data analysis in the
+presence of signatures of binary mergers and noise triggers, and
+astronomical image analysis in the presence of several populations of
+stars.  We showed that this framework is generally superior to both
+the loudest-event statistic and and the foreground-dominated
+statistic.
 
-Throughout this paper, we have assumed that the shape of the foreground and background distributions is known, or at least can be modeled with several additional parameters.
-This is not necessarily easy to do.
-For example, in the case of gravitational-wave data analysis, the shape of the foreground distribution of events may depend on the details of a complex data-analysis pipeline as well as the astrophysical source distribution, while the background event distribution depends on data quality and may deviate significantly from the simple Gaussian-noise behavior modeled in section \ref{sec:GW-example}.
-Several approaches have been developed to accurately model both distributions, e.g., through the use of injected signals to model the background \cite{ihope,CannonHannaKeppel:2012}.
-However, this is a difficult problem (e.g., because of the need to estimate the background at the very tails of the distribution), and will require significant future work.
+Throughout this paper, we have assumed that the shape of the
+foreground and background distributions is known, or at least can be
+modeled with several additional parameters.  This is not necessarily
+easy to do.  For example, in the case of gravitational-wave data
+analysis, the shape of the foreground distribution of events may
+depend on the details of a complex data-analysis pipeline as well as
+the astrophysical source distribution, while the background event
+distribution depends on data quality and may deviate significantly
+from the simple Gaussian-noise behavior modeled in section
+\ref{sec:GW-example}.  Several approaches have been developed to
+accurately model both distributions, e.g., through the use of injected
+signals to model the background \cite{ihope,CannonHannaKeppel:2012}.
+However, this is a difficult problem (e.g., because of the need to
+estimate the background at the very tails of the distribution), and
+will require significant future work.
 
-A further complication is that we have considered the rate of events in the data as products of some analysis pipeline.
-This rate may be different from the physical rate of interest, such as the rate of compact-binary mergers per unit time per unit volume which generate gravitational waves, or the physical numbers of stars in the cluster and field populations which produce the observed luminosities.
-Again, the conversion between the two will depend on the details of the data-analysis algorithm and ranking statistic, including any selection effects, and would need to be determined on a case-by-case basis.
+A further complication is that we have considered the rate of events
+in the data as products of some analysis pipeline.  This rate may be
+different from the physical rate of interest, such as the rate of
+compact-binary mergers per unit time per unit volume which generate
+gravitational waves, or the physical numbers of stars in the cluster
+and field populations which produce the observed luminosities.  Again,
+the conversion between the two will depend on the details of the
+data-analysis algorithm and ranking statistic, including any selection
+effects, and would need to be determined on a case-by-case basis.
 
-Furthermore, in a practical application there could be multiple classes of events, not just foreground and background.
-For example, we are not necessarily interested in the rate of gravitational-wave signals per se, but separately in the rate of signals from mergers of binary neutron stars and binary black holes -- populations that may sometimes be difficult to distinguish.
-Our approach is readily extendable to this particular complication, however.
-Note that it is symmetric with respect to foreground and background events (as expected, since one physicist's background is another physicist's foreground).
-We could relabel foreground and background events into other competing event classes, and further classes could be added in a straightforward way.
-However, the ability to distinguish classes relies on different distributions of their statistics.
-In general, rankings may need to be extended to include other statistics in addition to the signal ``loudness'' statistic in order to indicate both event significance and the probability of event attribution to a particular class.
+Furthermore, in a practical application there could be multiple
+classes of events, not just foreground and background.  For example,
+we are not necessarily interested in the rate of gravitational-wave
+signals per se, but separately in the rate of signals from mergers of
+binary neutron stars and binary black holes -- populations that may
+sometimes be difficult to distinguish.  Our approach is readily
+extendable to this particular complication, however.  Note that it is
+symmetric with respect to foreground and background events (as
+expected, since one physicist's background is another physicist's
+foreground).  We could relabel foreground and background events into
+other competing event classes, and further classes could be added in a
+straightforward way.  However, the ability to distinguish classes
+relies on different distributions of their statistics.  In general,
+rankings may need to be extended to include other statistics in
+addition to the signal ``loudness'' statistic in order to indicate
+both event significance and the probability of event attribution to a
+particular class.
 
@@
   We thank Kipp Cannon, Chad Hanna, Drew Keppel, and Richard
-  O'Shaughnessy for discussions and suggestions about this manuscript.  IM thanks KITP.
+  O'Shaughnessy for discussions and suggestions about this manuscript.
+  IM and WMF thank KITP.
 \end{acknowledgments}

### 2012-09-10 many.tex
Fixed typo in conclusion.  Added query on Jeffries prior.

@@
 \end{equation}
-Note that the calculation of the prior uses the likelihood before
+}\ilya{Sorry, I still don't get it.  Isn't $\mathcal{L} = R_f \hat{f} + R_b \hat{b}$?  If so, how do you get the second equality above?}
+\will{Note that the calculation of the prior uses the likelihood before
 analytical marginalization over the flags, $f_i$.  We are ignoring the
@@
 accurately model both distributions, e.g., through the use of injected
-signals to model the background \cite{ihope,CannonHannaKeppel:2012}.
+signals to model the foreground distribution \cite{ihope,CannonHannaKeppel:2012}.
 However, this is a difficult problem (e.g., because of the need to

### 2012-09-19 many.tex
Adding Curt's changes.

@@
 \newcommand{\jon}[1]{{\color{green} \bf #1}}
+\newcommand{\curt}[1]{{\color{magenta} \bf #1}}
+
+\newcommand{\be}{\begin{equation}}
+\newcommand{\ee}{\end{equation}}
+\newcommand{\ba}{\begin{eqnarray}}
+\newcommand{\ea}{\end{eqnarray}}
 
@@
 
+\subsection{Dependence of the Bayesian Rate Estimate on the Value of the Threshold}
+This paper is concerned with Bayesian rate estimates based on a  lists of events (which are a certain kind of summary of the complete data set): either the loudest event, or a list of all events whose SNR exceeds a certain, pre-set threshold.
+In this subsection we address the question of how the estimate depends on the threshold value.
+
+To begin with, we recall the well-known fact that the Bayesian estimator is
+unbiased, in the following sense.  For simplicity, assume that the model consists of a single
+rate parameter $R$, with prior distribution $p(R)$.  Consider an ensemble of data sets 
+whose distribution is consistent with that prior; i.e., such that $p(data)$ is given by
+%Construct and ensemble of synthetic data sets consistent with $p(R)$; i.e., draw data from the %distribution
+\be
+p(data) = \int{ p(data|R)\,p(R) dR}\, .
+\ee
+For each data set in the ensemble, compute the Bayesian estimator $R_{BE} = \int{ R p(R|data) dR}$.
+Then it is immediate that 
+\be 
+\int{R_{BE}(data)} d(data) = \int{ R p(R) dR}\, .
+\ee
+\noindent I.e, the data-weighted average of the Bayesian estimate $R_{BE}$ equals the prior-weighted average $R$. \curt{Guys:  would you like me to supply the argument?}
+%
+Therefore (for any fixed prior) all threshold values will yield, on average, the same estimate of the rate.
+However this equality of averages does {\it not} imply that all threshold values yield the same information.
+In general, as the threshold is lowered to include more events, the error bar on the estimate shrink.
+In this subsection we give quantitative illustrations of how the error bar shrinks when the threshold is lowered.
+
+Consider the following model problem.
+% in which the distribution of noise events is Gaussian.
+Let $P(\rho)\Delta\rho$ be the probability that data yields a putative event with SNR in the
+range $[\rho - \Delta\rho/2, \rho + \Delta\rho/2]$  We write $P = P_n + P_t$, where $P_n$ is
+the distribution of noise events and $P_t$ is the distribution of events from "true" astrophysical sources.
+Both $P_n$ and $P_t$ are the product of a rate (of noise or true events) and the observation time.
+Here we will assume that the noise distribution is Gaussian, so that $P_n$
+has the general form
+\be 
+P_n(\rho) = \Lambda_n {\rm exp}\big[-\rho^2/2\big] \, .
+\ee
+\noindent
+We find it useful to define $\rho_1$ as the SNR such that a data set will have on average
+a single noise event louder than $\rho_1$; i.e., such that
+\be
+\int_{\rho_1}^{\infty} P_n(\rho)\, d\rho  = 1 \, .
+\ee
+For this example we will assume that the spatial distribution of true events can be approximated as uniform in Euclidean space. Then we can write the full $P$ as
+%and that the spatial distribution of true events can be approximated as uniform in Euclidean %space. Then $P$ has the general form  
+\be
+P(\rho) = \big[\sqrt{\pi/2} {\rm erfc}(\rho_1/\sqrt{2})\big]^{-1} \ exp[-\rho^2/2] + 3 R_t \frac{\rho_1^3}{\rho^4} \, .
+\ee
+\noindent where $R_t$ is the mean number of true events with $\rho > \rho_1$.
+%What matters for the following analysis is the rate times the observation time.
+
+For any pair $(\rho_1, R_t)$, it is straightforward to construct random event lists drawn from the corresponding $P(\rho)$, and straightforward to apply a threshold by "throwing away" all events with $\rho$ less than some threshold value $\rho_{th}$.
+For any thresholded event list, we use
+Eqs. ..... to construct the probability density $p(R)$ (except that in this example, instead of the
+Jeffrey's prior, we assume a "hat-shaped" prior; see below).
+% that is uniform in some interval and zero elsewhere; see below).  
+For that event list, we define $\Delta R_t$ by
+\be
+(\Delta R_t)^2 = \int{(R - R_t)^2 p(R)\, dR }\, .
+\ee
+We initially set a rather low threshold, and see how $\Delta R_t$ varies as $rho_{th}$ is increased.
+To find the average behavior, we do many draws of $R_t$ and corresponding data sets.
+
+Illustrative results are displayed in Fig.~\ref{fig:delR}, where the mean value of $\Delta R_t$ is plotted as a function of $\rho_{th}$.
+%$x_{th} \equiv \rho_{th}/\rho_1$.  
+To create this figure, we chose $\rho_1 = 8$,
+and took our prior on $R_t$ to be a uniform distribution between $1$ and $10$.
+Fig.~\ref{fig:delR} also shows how the mean number of noise events on the list varies with $\rho_{th}$.
+Note the relative insensitivity of  $\Delta R_t$ to $\rho_{th}$:  as the mean
+number of noise events on the event list shrinks from $8$ to $\sim 10^{-6}$, 
+the average $\Delta R_t$ increases only from 2.4 to 2.8.
+\curt{Guys: I can produce more figs, and will produce cleaner ones.}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{delR}
+  \caption{\label{fig:delR} The meanu uncertainty in the Bayesian rate estimate
+as function of the threshold value $\rho_{th}$.  Also shows how the mean number of
+noise events varies with $\rho_{th}$.}
+\end{figure}
+
 \subsection{Star Cluster Parameters With Background Contamination}

### 2012-10-25 many.tex
curt changed his first fig and discussion around that fig

--- 2012-09-10 many.tex
@@
 
+\subsection{Dependence of the Bayesian Rate Estimate on the Value of the Threshold}
+This paper is concerned with Bayesian rate estimates based on a  lists of events (which are a certain kind of summary of the complete data set): either the loudest event, or a list of all events whose SNR exceeds a certain, pre-set threshold.
+In this subsection we address the question of how the estimate depends on the threshold value.
+
+To begin with, we recall the well-known fact that the Bayesian estimator is
+unbiased, in the following sense.  For simplicity, assume that the model consists of a single
+rate parameter $R$, with prior distribution $p(R)$.  Consider an ensemble of data sets 
+whose distribution is consistent with that prior; i.e., such that $p(data)$ is given by
+%Construct and ensemble of synthetic data sets consistent with $p(R)$; i.e., draw data from the %distribution
+\be
+p(data) = \int{ p(data|R)\,p(R)\, dR}\, .
+\ee
+For each data set in the ensemble, compute the Bayesian estimator $R_{B} = \int{ R p(R|data) dR}$.
+Then it is immediate that 
+\be 
+\int{R_{B}(data)\, p(data)\,d(data) }= \int{ R\, p(R) \,dR}\, .
+\ee
+\noindent I.e, the data-weighted average of the Bayesian estimate $R_{B}$ equals the prior-weighted average $R$. 
+%\curt{Guys:  would you like me to supply the argument?}
+%
+Therefore (for any fixed prior) all threshold values will yield, on average, the same estimate of the rate.
+However this equality of averages does {\it not} imply that all threshold values yield the same information.
+In general, as the threshold is lowered to include more events, the error bar on the estimate shrink.
+In this subsection we give quantitative illustrations of how the error bar shrinks when the threshold is lowered.
+
+Consider the following model problem.
+% in which the distribution of noise events is Gaussian.
+Let $p(\rho)\Delta\rho$ be the probability that data yields a putative event with SNR in the
+range $[\rho - \Delta\rho/2, \rho + \Delta\rho/2]$  We write $p = p_n + p_s$, where $p_n$ is
+the distribution of noise events and $p_s$ is the distribution of events from actual astrophysical sources.
+Both $p_n$ and $p_s$ are the product of a rate (of noise or true events) and the observation time.
+Here we will assume that the noise distribution is Gaussian, so that $p_n$
+has the general form
+\be 
+p_n(\rho) = \Lambda_n {\rm exp}\big[-\rho^2/2\big] \, .
+\ee
+\noindent
+We find it useful to define $\rho_1$ as the SNR such that a data set will have on average
+a single noise event louder than $\rho_1$; i.e., such that
+\be
+\int_{\rho_1}^{\infty} P_n(\rho)\, d\rho  = 1 \, .
+\ee
+For this example we will assume that the spatial distribution of true events can be approximated as uniform in Euclidean space. Then we can write the full $p(\rho)$ as
+%and that the spatial distribution of true events can be approximated as uniform in Euclidean %space. Then $p(\rho)$ has the general form  
+\be\label{eq:ptot}
+p(\rho) = \big[\sqrt{\pi/2} {\rm erfc}(\rho_1/\sqrt{2})\big]^{-1} \ exp[-\rho^2/2] + 3 R_t \frac{\rho_1^3}{\rho^4} \, .
+\ee
+\noindent where $R_t$ is the mean number of true events with $\rho > \rho_1$.
+%What matters for the following analysis is the rate times the observation time.
+
+For any pair $(\rho_1, R_t)$, it is straightforward to construct random event lists drawn from the corresponding $p(\rho)$, and straightforward to apply a threshold by "throwing away" all events with $\rho$ less than some threshold value $\rho_{th}$.
+For any thresholded event list, we use
+Eqs. ..... to construct the probability density $p(R)$ (except that in this example, instead of the
+Jeffrey's prior, we assume a "hat-shaped" prior; see below).
+% that is uniform in some interval and zero elsewhere; see below).  
+For that event list, we define $\Delta R$ by
+\be
+(\Delta R_B)^2 = \int{(R - R_B)^2 \, p(R)\, dR }\, .
+\ee
+%To find the average behavior, we do many draws of $R_t$ and corresponding data sets.
+
+Figs.~\ref{fig:delRb} --\ref{fig:Rt10}  illustrate how $\Delta R$ varies with the threshold value $\rho_{th}$.  In all cases we assumed that $\rho_1 = 8$.
+%
+For Fig.~\ref{fig:delR1}, we adopted a flat prior on $R$ between $1$ and $10$ (i.e, we took
+$p(R) = 1/9$  for $1 < R <10$, and  zero outside that interval).  
+% To create Fig.~\ref{fig:delR1}, 
+We drew $10^3$ 
+%\curt{Curt needs to check this} 
+random realizations $(R_t, data)$, where $R_t$ was drawn from that flat prior and the data--a list of $\rho$ values-- was drawn from the distribution $p(data|R_t)$ given by Eq.~\ref{eq:ptot} .
+For any given threshold value, $\rho_{th}$, we created an event list by taking the subset of the $\rho$ values
+% in the data set that are 
+greater than $\rho_{th}$  For each data set, we used a sequence of increasing threshold  values $\rho_{th}$ to create a corresponding sequence of (ever shorter)  event lists.
+For each event list, we calculated $\Delta R$ and $R_B - R_t$.
+%and take the mean over all realizations.  
+As a code check, we verified that the mean of $R_B - R_t$ (over all realizations) was consistent with zero for all $\rho_{th}$.
+Fig.~\ref{fig:delR1} displays both the mean value of $\Delta R$ and the mean number of noise events (as opposed to actual signals) as a function of $\rho_{th}$.
+Note the relative insensitivity of  $\Delta R_t$ to $\rho_{th}$:  as the mean
+number of noise events on the event list shrinks from $8$ to $\sim 10^{-6}$, 
+the average $\Delta R_B$ increases only $\sim 14\%$, from 0.42to 0.48.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{delRb}
+\caption{\label{fig:delRb} Displays the mean "error bar" $\Delta R_B$ as function of the threshold value $\rho_{th}$.
+In this case, the values of $R_t$ were randomly drawn from a flat prior on the interval $[1,10]$.
+We also show how the (mean) size of the event list varies with $\rho_{th}$.}
+\end{figure}
+
+Figs.~\ref{fig:Rt1} --\ref{fig:Rt10} were constructed similarly, but with the following differences.  First, we took the prior 
+$p(R)$ to be $p(R) \propto 1/R$ for  $0.03 < R < 30$, and zero outside that interval (i.e., $p(R)$ is the Jeffries prior, but restricted to some range).
+Second, we regarded $R_t$ as fixed (as of course it would be in Nature), and only varied the data--randomly drawing $10^3$ realizations from $p(data|R_t)$.
+For each realization we calculated $\Delta R_B$ and $(R_B - R_t)$, and then we calculated the mean over all realizations.
+The mean of $(R_B - R_t)$ was generally {\it not} zero, since the
+prior biases the results.  (Of course, if one {\it updated} the prior after each realization, the 
+bias would decrease towards zero, but that is not what we did: we evaluated each realization with the same Jeffries prior.)
+
+The number of noise events is the same as for the first case, shown in Fig.~\ref{fig:delR}.
+Note that again, for all three values of $R_t$, the mean uncertainty $\Delta R_B$ increases by only $\sim 20\%$  as the mean number of noise events on the event list shrinks from $8$ to $\sim 10^{-6}$.
+And as one would expect, the relative uncertainty shrinks as $R_t$ increases.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{Rt1}
+\caption{\label{fig:Rt1} Displays the mean rms error $\Delta R_B$ and mean $(R_B - R_t)$ (i.e., the bias) as functions of the threshold value $\rho_{th}$.
+In this case, the actual value of $R_t$ was $1.0$, and $p(R|data)$ was computed using the Jeffries prior on the interval $[0.03,30]$.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{Rt3}
+\caption{\label{fig:Rt3} Same as the previous figure, but displaying the size of $\Delta R_B$  relative to $R_t$, with the actual $R_t$ equal to $3.0$.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{Rt10}
+\caption{\label{fig:Rt10} Same as the previous figure, but with the actual $R_t$ equal to $10.0$.}
+\end{figure}
+
 \subsection{Star Cluster Parameters With Background Contamination}

### 2012-11-01 many.tex
curt added subsection on bad behavior of Bayesian rate estimate when there is one very loud outlier

@@
 \ee
-For this example we will assume that the spatial distribution of true events can be approximated as uniform in Euclidean space. Then we can write the full $p(\rho)$ as
-%and that the spatial distribution of true events can be approximated as uniform in Euclidean %space. Then $p(\rho)$ has the general form  
-\be\label{eq:ptot}
-p(\rho) = \big[\sqrt{\pi/2} {\rm erfc}(\rho_1/\sqrt{2})\big]^{-1} \ exp[-\rho^2/2] + 3 R_t \frac{\rho_1^3}{\rho^4} \, .
+For this example we will assume that the spatial distribution of true events can be approximated as uniform in Euclidean space, so 
+\be\label{eq:ps}
+p_s(\rho) = 3 R_t \frac{\rho_1^3}{\rho^4} \, ,
 \ee
 \noindent where $R_t$ is the mean number of true events with $\rho > \rho_1$.
+Then we can write the full $p(\rho)$ as
+%and that the spatial distribution of true events can be approximated as uniform in %Euclidean %space. Then $p(\rho)$ has the general form  
+\be\label{eq:ptot}
+p(\rho) = \big[\sqrt{\pi/2} {\rm erfc}(\rho_1/\sqrt{2})\big]^{-1} \, {\rm exp}[-\rho^2/2]\  + \ 3 R_t \frac{\rho_1^3}{\rho^4} \, .
+\ee
+
 %What matters for the following analysis is the rate times the observation time.
@@
 For any thresholded event list, we use
-Eqs. ..... to construct the probability density $p(R)$ (except that in this example, instead of the
+Eqs.~(\curt{need to supply, from earlier in paper} ) to construct the probability density $p(R|data)$ (except that in this example, instead of the
 Jeffrey's prior, we assume a "hat-shaped" prior; see below).
 % that is uniform in some interval and zero elsewhere; see below).  
-For that event list, we define $\Delta R$ by
+For that event list, we define $\Delta R_B$ by
 \be
-(\Delta R_B)^2 = \int{(R - R_B)^2 \, p(R)\, dR }\, .
+(\Delta R_B)^2 = \int{(R - R_B)^2 \, p(R|data)\, dR }\, .
 \ee
 %To find the average behavior, we do many draws of $R_t$ and corresponding data sets.
-
+%
 Figs.~\ref{fig:delRb} --\ref{fig:Rt10}  illustrate how $\Delta R$ varies with the threshold value $\rho_{th}$.  In all cases we assumed that $\rho_1 = 8$.
 %
-For Fig.~\ref{fig:delR1}, we adopted a flat prior on $R$ between $1$ and $10$ (i.e, we took
+
+For Fig.~\ref{fig:delRb}, we adopted a flat prior on $R$ between $1$ and $10$ (i.e, we took
 $p(R) = 1/9$  for $1 < R <10$, and  zero outside that interval).  
 % To create Fig.~\ref{fig:delR1}, 
-We drew $10^3$ 
-%\curt{Curt needs to check this} 
+We drew $10^4$ 
 random realizations $(R_t, data)$, where $R_t$ was drawn from that flat prior and the data--a list of $\rho$ values-- was drawn from the distribution $p(data|R_t)$ given by Eq.~\ref{eq:ptot} .
-For any given threshold value, $\rho_{th}$, we created an event list by taking the subset of the $\rho$ values
+For any particular threshold value, $\rho_{th}$, we created an  event list by taking the subset of the $\rho$ values
 % in the data set that are 
-greater than $\rho_{th}$  For each data set, we used a sequence of increasing threshold  values $\rho_{th}$ to create a corresponding sequence of (ever shorter)  event lists.
+that were greater than $\rho_{th}$.
+For each data set, we considered a sequence of increasing threshold  values $\rho_{th}$, using them to create a corresponding sequence of (ever shorter)  event lists.
 For each event list, we calculated $\Delta R$ and $R_B - R_t$.
@@
 number of noise events on the event list shrinks from $8$ to $\sim 10^{-6}$, 
-the average $\Delta R_B$ increases only $\sim 14\%$, from 0.42to 0.48.
+the average $\Delta R_B$ increases only $\sim 14\%$, from $0.42$ to $0.48$.
 
@@
 $p(R)$ to be $p(R) \propto 1/R$ for  $0.03 < R < 30$, and zero outside that interval (i.e., $p(R)$ is the Jeffries prior, but restricted to some range).
-Second, we regarded $R_t$ as fixed (as of course it would be in Nature), and only varied the data--randomly drawing $10^3$ realizations from $p(data|R_t)$.
-For each realization we calculated $\Delta R_B$ and $(R_B - R_t)$, and then we calculated the mean over all realizations.
+Second, we regarded $R_t$ as fixed (as of course it would be in Nature), and only varied the data--randomly drawing $10^4$ realizations from $p(data|R_t)$.
+For each realization we calculated $R_t^{-1}\,\Delta R_B$ and $(R_B - R_t)$, and then we calculated the means of these quantities over all realizations.
 The mean of $(R_B - R_t)$ was generally {\it not} zero, since the
@@
 
-The number of noise events is the same as for the first case, shown in Fig.~\ref{fig:delR}.
-Note that again, for all three values of $R_t$, the mean uncertainty $\Delta R_B$ increases by only $\sim 20\%$  as the mean number of noise events on the event list shrinks from $8$ to $\sim 10^{-6}$.
-And as one would expect, the relative uncertainty shrinks as $R_t$ increases.
+For Figs.~\ref{fig:Rt1} --\ref{fig:Rt10}, the mean number of noise events as a function of $\rho_{th}$ was the same as for the case shown in Fig.~\ref{fig:delRb}.
+Note that for all three values of $R_t$, the mean uncertainty $\Delta R_B$ increases by only $\sim 20\%$  as the mean number of noise events on the event list shrinks from $8$ to $\sim 10^{-6}$.
+Thus the inference that the average $\Delta R_B$ is rather insensitive to  $\rho_{th}$ appears to be robust.
+Finally, we note that for all three cases the bias is much smaller than $\Delta R_B$.
 
@@
 \end{figure}
+
+\subsection{Extreme Sensitivity of the LE Rate Estimate to a Single, Unusually Loud Event}
+Here we discuss a very unattractive feature of the Bayesian loudest event estimate of $R$:
+a small percentage of the time it will yield a very large over-estimate.
+
+To explain this, we will use the same model as described in the previous subsection, and we will begin with a very specific example.
+Let the true rate (times the observation time) be $R_t = 1$, meaning (again) that the expected number of actual events with $\rho > \rho_1$ is one.
+Then there is a $1/64$ chance ($\approx 1.6\%$) that the loudest event will have $\rho_{LE} > 4 \rho_1$.
+Consider this case, and let us also assume that there are no events (noise or actual) with $\rho_1 < \rho < \rho_{LE}$.
+The loudest event estimate basically "throws away" the information that there
+are no events in this interval, which (we show below) leads an estimate $R_{LE} \approx 64$,
+  which is $64$ times the true rate!  Now, if the true rate really were $R_t = 64$, then the expected number of 
+events with $\rho > \rho_1$ would be $64$.  So in this case, the loudest event estimate 
+ignores the fact that there are $\sim 56 -72$ "missing" events.  However a Bayesian estimate
+with $\rho_{th}$ set to $\rho_1$ incorporates this information quite naturally, and so (correctly) yields an estimated $R$ of order one.
+
+We now essentially repeat the above discussion, but in greater generality and with the derivations provided.   First recall that $R_B$ is given by
+\be\label{eq:rb1}
+R_B = \frac{  \int{ R\, p(\{\rho_i\}|R)  \, p(R)\, dR}}{  \int{ p(\{\rho_i\}|R)\, p(R)\, dR }}\, ,
+\ee
+\noindent
+where $p(R)$ is the prior on $R$ and (and up to an overall normalization)
+\be\label{eq:rb2}
+p(\{\rho_i\}|R) = {\rm exp}\bigg[\sum_{i \in  list} {\rm ln}\, p(\rho_i|R) - \int_{\rho_{th}}^{\infty} p(\rho)\, d\,\rho  \bigg] \,  ,
+\ee
+\noindent
+with $p(\rho)$ given by Eq.~().
+
+The loudest event estimate is obtained by setting the lower integration limit in the last term on the rhs of Eq.~(\ref{eq:rb2}) to $\rho_{th} = \rho_{LE}$, and replacing the sum 
+$\sum_i {\rm ln}\, p(\rho_i|R)$ by the single term ${\rm ln}\, p(\rho_{LE}|R)$.
+For the case of interest here, where $\rho_{LE}$ is (at least) a few times larger than $\rho_1$ and $R_t$ is of order $1$, we can neglect the noise contribution $p_n(\rho)$.
+Then the integrals in Eq.~(\ref{eq:rb1}) become especially simple, and yield the estimate:
+\be\label{eq:rb3}
+R_B \approx 
+\begin{cases}
+(\rho_{LE}/\rho_1)^3 & \text{if $p(R) \propto 1/R$} \\
+2\, (\rho_{LE}/\rho_1)^3 & \text{if $p(R) =$ const \, .} 
+\end{cases}
+\ee
+
+If instead we set the lower integration limit to $\rho_{th} = \rho_1$, then the $p_n(\rho)$
+contribution to $p(\rho)$ can no longer be neglected.  But the integrals are still 
+straightforward, and we find:
+ \be\label{eq:rb3}
+R_B \approx 
+\begin{cases}
+1 \ \ \  & \text{if $p(R) \propto 1/R$} \\
+2  \ \ \ & \text{if $p(R) =$ const \, .} 
+\end{cases}
+\ee
+\noindent Thus (as emphasized above) simply incorporating the information that there are no events {\it between} $\rho_1$ and  $\rho_{LE}$ results in a far beter estimate.
 

### 2012-11-09 many.tex
Edits complete through SIIIB.

@@
 \cite[e.g.,][]{s5-highmass,s6-lowmass} and in astronomical
-observations of a field of objects of mixed provenance \cite[GC
-  cluster paper]{}..  In this paper, we introduce a robust formalism
+observations of a field of objects of mixed provenance
+\cite{DeGennaro2009}.  In this paper, we introduce a robust formalism
 for estimating event rates from the data when the shape of foreground
@@
 
-We assume that we are presented with a data set of $N$ events that
-exceed a pre-specified threshold in ranking statistic, $x_{\rm min}$.
-Each event may be due to either a signal of interest or an
-uninteresting background.  Each event is associated with a ranking
-statistic, $x$.  Our data set therefore consists of the ranking
-statistics for the set of events:
+We first consider one-dimensional data, but will generalize to the
+multidimensional case below.  We assume that we are presented with a
+data set of $N$ events that exceed a pre-specified threshold in
+ranking statistic, $x_{\rm min}$.  Each event may be due to either a
+signal of interest or an uninteresting background.  Each event is
+associated with a ranking statistic, $x$.  Our data set therefore
+consists of the ranking statistics for the set of events:
 \begin{equation}
@@
 \cdots < x_N$, but some of the derivations that follow are simpler if
-the events are ordered by time of arrival. We will use $d$ to denote
-ranking statistic-ordered events, and $d_{\rm to}$ to denote
-time-ordered events.
+the events are ordered by time of arrival (i.e.\ randomly with respect
+to the $x_i$). We will use $d$ to denote ranking statistic-ordered
+events, and $d_{\rm to}$ to denote time-ordered events.
 
@@
 
-Not surprisingly, this ranking-statistic ordered posterior can be
-computed directly by assuming that the flags, $\mathset{f_i}$, are
-un-observed data and treating the sets $\mathset{x_i | f_i = 1}$ and
-$\mathset{x_i | f_i = 0}$ as samples from an inhomogeneous Poisson
-process.  For an inhomogeneous Poisson process with rate function
-$r(y)$ (cumulative rate $R(y)$), the likelihood of a set of samples
-$\mathset{y_i}$ is given by
+The ranking-statistic ordered posterior can be computed directly by
+assuming that the flags, $\mathset{f_i}$, are un-observed data and
+treating the sets $\mathset{x_i | f_i = 1}$ and $\mathset{x_i | f_i =
+  0}$ as samples from an inhomogeneous Poisson process.  For an
+inhomogeneous Poisson process with rate function $r(y)$ (cumulative
+rate $R(y)$), the likelihood of a set of samples $\mathset{y_i}$ is
+given by
 \begin{multline}
@@
 samples and taking the product, we obtain $p(d,\mathset{f_i}, N | R_f,
-R_b, \theta)$ and thence $p(\mathset{f_i}, R_f, R_b, \theta | d,N) =
+R_b, \theta)$ and thus $p(\mathset{f_i}, R_f, R_b, \theta | d,N) =
 p(d,\mathset{f_i}, N | R_f, R_b, \theta)\ p(R_f, R_b, \theta)\, /\,
@@
 \end{equation}
-where $\tilde{b}(x_N,\theta)
-=\hat{b}(x_N,\theta)/(1-\hat{B}(x_N,\theta))$. If
-$\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-
+where $\tilde{b}(x_N,\theta) = \hat{b}(x_N,\theta) /
+(1-\hat{B}(x_N,\theta))$\footnote{The quantity $\Lambda$ from
+  \citet{Biswas2009} is related to $\hat{b}$ and $\tilde{f}$ by
+  \begin{equation}
+    \Lambda(x, \theta) = \frac{\tilde{f}(x,\theta)}{R_b \hat{b}(x,\theta)}.
+\end{equation}}.  If $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-
 \hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be understood as the
@@
 dominated regime, the rate is approximately equal to the number of
-events observed. We note that the rate $R_f$ is defined as the rate of
+events observed (the $1/2$ comes from our use of the Jeffreys prior on
+the rate).  We note that the rate $R_f$ is defined as the rate of
 events occurring above the specified threshold and therefore if the
@@
 
-For a data stream of pure noise, $d(f) = n(f)$, the SNR of a each
-template follows a $N(0,1)$ distribution.  The background ranking
-statistic (i.e.\ the maximum SNR over the template bank) then has a
-cumulative distribution without thresholding of
+For a data stream of pure noise, $d(f) = n(f)$, the SNRs of the
+templates are independent $N(0,1)$ random variables.  The background
+ranking statistic (i.e.\ the maximum SNR over the template bank) then
+has a cumulative distribution without thresholding of
 \begin{equation}
@@
 from Eqs.~\eqref{eq:analytic-background-rate} and
-\eqref{eq:analytic-foreground-rate} with $R_f^\mathrm{true} = 10.4$
-and $R_b^\mathrm{true} = 95.1$, using 1000 templates.  \ilya{[Specify
-    $\xmin$?]} The synthetic data consisted of 13 foreground events
-and 85 background events; the cumulative distribution for the ranking
+\eqref{eq:analytic-foreground-rate} using $x_\mathrm{min} = 3.5$, with
+$R_f^\mathrm{true} = 10.4$ and $R_b^\mathrm{true} = 95.1$ and 1000
+templates.  The synthetic data consisted of 13 foreground events and
+85 background events; the cumulative distribution for the ranking
 statistic of the synthetic data appears in Figure
@@
 Figure~\ref{fig:altmeth}. In this case, the loudest event had
-$x_N=9.5$\jon{(correct this)}. The loudest-event statistic depends on
-a specification of the maximum, $R_{\rm max}$, for the background
+$x_N\simeq 9.47$. The loudest-event statistic depends on a
+specification of the maximum, $R_{\rm max}$, for the background
 rate. We show results for $R_{\rm max}=\infty$, i.e., the improper
@@
 fore/background, $\hat{f}(x)/\hat{b}(x)=p_{\rm thresh}$. Setting
-$p_{\rm thresh}=0.99$ gives $x_{\rm min}=4.07$ and there are $N=13$
-events exceeding that threshold. Setting $p_{\rm thresh}=0.5$ gives
-$x_{\rm min}=3.82$ and there are $N=15$ events exceeding that
-threshold \jon{(correct these numbers using Will's data)}. We show
-results for both of these choices of $x_{\rm min}$ in
-Figure~\ref{fig:altmeth}. We also show similar results for two extreme
-cases --- setting $x_{\rm min}=3.5$ so that we assume all $98$
-observed events are foreground; and setting $x_{\rm min}=9$ so that
-only the loudest event is classified as foreground.
+$p_{\rm thresh}=0.99$ gives $x_{\rm min}=4.07$ and there are $N=18$
+(11 foreground and 7 background) events exceeding that
+threshold. Setting $p_{\rm thresh}=0.5$ gives $x_{\rm min}=3.82$ and
+there are $N=30$ (11 foreground and 19 background) events exceeding
+that threshold.  Each of these thresholds gives a biased estimate of
+the rate because there are background events still above threshold.
+The ``omniscient'' threshold of $x_\mathrm{min} = 4.38$ produces $N=7$
+(7 foreground and 0 background) events in this data set, and therefore
+an unbiased estimate, but of course this threshold can only be
+determined because we can examine the synthetic foreground and
+background data samples.  We show results for the first two choices of
+$x_{\rm min}$ in Figure~\ref{fig:altmeth}; the omniscient choice
+produces essentially the same posterior as our full analysis.
 
@@
 more accurately located when a prior maximum rate is defined, but the
-distribution is much wider than using the full analysis described
-here. This is to be expected as much of the information is being
-thrown away. The foreground-dominated statistic gives a reasonable
-approximation to the true foreground rate, and a distribution that is
-not much wider than the full analysis, for the non-extreme choices of
-$x_{\rm min}$. For $x_{\rm min}=9$, the foreground dominated statistic
-has equivalent performance to the loudest event statistic, as we might
-expect, while when $x_{\rm min}=3.5$, it performs poorly since we are
+distribution is much wider than using the full analysis described here
+in any case. This is to be expected as much of the information is
+being thrown away. The foreground-dominated statistic gives a
+reasonable approximation to the true foreground rate, and a
+distribution that is essentially equal to the full analysis, for the
+``omniscent'' choice of threshold value that excludes all background
+data. For lower thresholds, even for a threshold where
+$p_\mathrm{thresh} = 0.99$, it performs poorly since we are
 approximating the foreground rate by the total foreground plus
-background rate. This indicates that, provided the threshold is chosen
-appropriately, the foreground dominated statistic can perform quite
-well at estimating the rate. The fact that it reproduces the posterior
-from the full analysis so well is indicative of the fact that most of
-the information about the foreground comes from the loudest
-events. Indeed, for $x_{\rm min} \approx 4$, the set of events
-consists mostly of foreground sources and that set contains a decent
-number of events ($13$), so it is not surprising that this statistic
-can estimate the foreground rate well. The choice of $x_{\rm min}$ was
-motivated by knowledge of $\hat{b}(x,\theta)$ and $\hat{f}(x,\theta)$,
-which we have in this case, but this might not be possible if the
-parameters $\theta$ were simultaneously being estimated. The full
-analysis naturally incorporates inference about the rate parameters
-$\theta$ and the background rate $R_b$ along with the foreground rate
-and incorporates maximum information from the data set and should
-therefore lead to narrower posteriors in general.
+background rate.  This indicates that, provided the threshold is
+chosen appropriately, the foreground dominated statistic can perform
+quite well at estimating the rate---but choosing this threshold
+correctly is difficult. The fact that it reproduces the posterior from
+the full analysis so well is indicative of the fact that most of the
+information about the foreground comes from the loudest events. The
+full analysis naturally incorporates inference about the background
+rate $R_b$ along with the foreground rate and incorporates maximum
+information from the data set and should therefore lead to narrower
+posteriors in general.
 
@@
     \ref{sec:analytic-GW-example} marginalized over all other
-    parameters (error bars are one standard deviation).  True
-    foreground events are in dark grey, background events in light
-    grey.  Even though our method cannot identify the status of most
-    events with confidence, it can still correctly estimate the rates
-    (Figure \ref{fig:analytic-rate-recovery}).}
+    parameters.  True foreground events are in dark grey, background
+    events in light grey.  Even though our method cannot identify the
+    status of most events with confidence, it can still correctly
+    estimate the rates (Figure \ref{fig:analytic-rate-recovery}).}
 \end{figure}
@@
   \caption{\label{fig:altmeth}Posteriors on foreground rate obtained
-    using alternative methods --- the loudest event statistic and the
-    foreground dominated statistic. Results are shown for two
-    different choices of the maximum a priori background rate, $R_{\rm
-      max}$, in the case of the loudest event (``LE'') statistic, and
-    for several different choices of the threshold, $x_{\rm min}$, in
-    the case of the foreground dominated (``FD'') statistic. \ilya{I
-      can see the figure on the screen, but can't print it...}}
+    using the method described in this paper, the loudest event
+    statistic and the foreground dominated analysis for the data set
+    from \S \ref{sec:analytic-GW-example}.  For the loudest event
+    statistic, we present the posterior with and without an upper
+    limit on the background rate, $R_b$; in both cases the rate
+    posterior is significantly wider than the one obtained with the
+    method described in this paper.  For the foreground dominated
+    statistic, the limits $x_\mathrm{min} = 3.82$ and $x_\mathrm{min}
+    = 4.07$ give likelihood ratios of $\hat{f}/\hat{b} = 0.5$ and
+    $0.99$.  For this data set, the thresholds in fact include 19 and
+    7 background events, respectively, so the corresponding rate
+    estimates are significantly biased.  An ``omniscient'' threshold
+    of $x_\mathrm{min} = 4.38$ would produce exactly 7 foreground and
+    zero background events, and the resulting posterior is essentially
+    indistinguishable from the curve for the full analysis.}
 \end{figure}
@@
 \ref{fig:ntemplates}.  We use the same parameters and data set as in
-\S \ref{sec:analytic-GW-example}, with $R_f = 10.4$, $R_b = 95.1$, and
-$N_\mathrm{eff} = 1000$, but now allow $N_\mathrm{eff}$ to be a
-parameter of the background distribution, with a flat prior.  Both the
-rates and the number of effective templates are recovered without
-significant loss of accuracy relative to the fixed $N_\mathrm{eff}$
-situation in \S \ref{sec:analytic-GW-example}.
+\S \ref{sec:analytic-GW-example}, with $x_\mathrm{min} = 3.5$, $R_f =
+10.4$, $R_b = 95.1$, and $N_\mathrm{eff} = 1000$, but now allow
+$N_\mathrm{eff}$ to be a parameter of the background distribution,
+with a flat prior.  Both the rates and the number of effective
+templates are recovered without significant loss of accuracy relative
+to the fixed $N_\mathrm{eff}$ situation in \S
+\ref{sec:analytic-GW-example}.
 
 If we consider the two alternative methods, the loudest event and
-foreground dominated statistics, we will recover the same foreground
-distributions as in the case with $N_{\rm eff}$ fixed, which are shown
-in Figure~\ref{fig:altmeth}. This is because the parameter $N_{\rm
-  eff}$ affects only the background distribution, to which the
-foreground-dominated statistic is insensitive, and in the loudest
-event case, after marginalisation over $N_{\rm eff}$ we find
-$\int_0^{N_{\rm max}} \hat{b}(x_N,N_{\rm eff}) {\rm d}N_{\rm eff} \ll
-3 N_{\rm max}/R_{\rm max} \hat{f}(x_N, N_{\rm eff})$ and so we are
-still in the foreground-dominated regime in which the loudest event
-tells us nothing about the background. Similarly, neither of these
-alternative methods can inform us about the value of $N_{\rm eff}$, a
-property of the background. \jon{Technically, if we allowed really
-  large values of $N_{\rm eff}$ we would get into a regime where the
-  loudest event could be background and we'd get something here. But,
-  we would need $N_{\rm max} \sim 10^{20}$ and so I don't think we
-  should consider these alternatives further in this case.} \ilya{On
-  the other hand, this ties in with the comment at the end of the
-  previous subsection about the difficulty of choosing $x_{min}$ for
-  the foreground-dominated statistic when there are other parameters
-  $\theta$ to be estimated simultaneously ($N_{eff}$ in this case), so
-  perhaps it's worthwhile to add some comment about this?}
+foreground dominated statistics, and apply the same
+foreground-dominated thresholds as before, we will recover the same
+foreground distributions as are shown in Figure~\ref{fig:altmeth}.
+This is because the parameter $N_{\rm eff}$ affects only the
+background distribution, to which the foreground-dominated statistic
+is insensitive, and in the loudest event case, after marginalisation
+over $N_{\rm eff}$ we find $\int_0^{N_{\rm max}} \hat{b}(x_N,N_{\rm
+  eff}) {\rm d}N_{\rm eff} \ll 3 N_{\rm max}/R_{\rm max} \hat{f}(x_N,
+N_{\rm eff})$ and so we are still in the foreground-dominated regime
+in which the loudest event tells us nothing about the background.
+Neither of these alternative methods can inform us about the value of
+$N_{\rm eff}$, a property of the background.  Moreover, the choice of
+threshold value for the foreground-dominated statistic becomes
+significantly more complicated in this case, since $p_\mathrm{thresh}$
+now depends on $N_\mathrm{eff}$.
 

### 2012-11-09 many.tex
Re-formatted IIIC.

@@
 \subsection{Dependence of the Bayesian Rate Estimate on the Value of the Threshold}
-This paper is concerned with Bayesian rate estimates based on a  lists of events (which are a certain kind of summary of the complete data set): either the loudest event, or a list of all events whose SNR exceeds a certain, pre-set threshold.
-In this subsection we address the question of how the estimate depends on the threshold value.
-
-To begin with, we recall the well-known fact that the Bayesian estimator is
-unbiased, in the following sense.  For simplicity, assume that the model consists of a single
-rate parameter $R$, with prior distribution $p(R)$.  Consider an ensemble of data sets 
-whose distribution is consistent with that prior; i.e., such that $p(data)$ is given by
+This paper is concerned with Bayesian rate estimates based on a lists
+of events (which are a certain kind of summary of the complete data
+set): either the loudest event, or a list of all events whose SNR
+exceeds a certain, pre-set threshold.  In this subsection we address
+the question of how the estimate depends on the threshold value.
+
+To begin with, we recall the well-known fact that the Bayesian
+estimator is unbiased, in the following sense.  For simplicity, assume
+that the model consists of a single rate parameter $R$, with prior
+distribution $p(R)$.  Consider an ensemble of data sets whose
+distribution is consistent with that prior; i.e., such that $p(data)$
+is given by
 %Construct and ensemble of synthetic data sets consistent with $p(R)$; i.e., draw data from the %distribution
-\be
-p(data) = \int{ p(data|R)\,p(R)\, dR}\, .
-\ee
-For each data set in the ensemble, compute the Bayesian estimator $R_{B} = \int{ R p(R|data) dR}$.
-Then it is immediate that 
-\be 
-\int{R_{B}(data)\, p(data)\,d(data) }= \int{ R\, p(R) \,dR}\, .
-\ee
-\noindent I.e, the data-weighted average of the Bayesian estimate $R_{B}$ equals the prior-weighted average $R$. 
+\be p(data) = \int{ p(data|R)\,p(R)\, dR}\, .  \ee For each data set
+in the ensemble, compute the Bayesian estimator $R_{B} = \int{ R
+  p(R|data) dR}$.  Then it is immediate that \be \int{R_{B}(data)\,
+  p(data)\,d(data) }= \int{ R\, p(R) \,dR}\, .  \ee
+\noindent I.e, the data-weighted average of the Bayesian estimate
+$R_{B}$ equals the prior-weighted average $R$.
 %\curt{Guys:  would you like me to supply the argument?}
 %
-Therefore (for any fixed prior) all threshold values will yield, on average, the same estimate of the rate.
-However this equality of averages does {\it not} imply that all threshold values yield the same information.
-In general, as the threshold is lowered to include more events, the error bar on the estimate shrink.
-In this subsection we give quantitative illustrations of how the error bar shrinks when the threshold is lowered.
+Therefore (for any fixed prior) all threshold values will yield, on
+average, the same estimate of the rate.  However this equality of
+averages does {\it not} imply that all threshold values yield the same
+information. In general, as the threshold is lowered to include more
+events, the error bar on the estimate shrink.  In this subsection we
+give quantitative illustrations of how the error bar shrinks when the
+threshold is lowered.
 
@@
 % in which the distribution of noise events is Gaussian.
-Let $p(\rho)\Delta\rho$ be the probability that data yields a putative event with SNR in the
-range $[\rho - \Delta\rho/2, \rho + \Delta\rho/2]$  We write $p = p_n + p_s$, where $p_n$ is
-the distribution of noise events and $p_s$ is the distribution of events from actual astrophysical sources.
-Both $p_n$ and $p_s$ are the product of a rate (of noise or true events) and the observation time.
-Here we will assume that the noise distribution is Gaussian, so that $p_n$
-has the general form
-\be 
-p_n(\rho) = \Lambda_n {\rm exp}\big[-\rho^2/2\big] \, .
-\ee
+Let $p(\rho)\Delta\rho$ be the probability that data yields a putative
+event with SNR in the range $[\rho - \Delta\rho/2, \rho +
+  \Delta\rho/2]$ We write $p = p_n + p_s$, where $p_n$ is the
+distribution of noise events and $p_s$ is the distribution of events
+from actual astrophysical sources.  Both $p_n$ and $p_s$ are the
+product of a rate (of noise or true events) and the observation time.
+Here we will assume that the noise distribution is Gaussian, so that
+$p_n$ has the general form \be p_n(\rho) = \Lambda_n {\rm
+  exp}\big[-\rho^2/2\big] \, .  \ee
 \noindent
-We find it useful to define $\rho_1$ as the SNR such that a data set will have on average
-a single noise event louder than $\rho_1$; i.e., such that
-\be
-\int_{\rho_1}^{\infty} P_n(\rho)\, d\rho  = 1 \, .
-\ee
-For this example we will assume that the spatial distribution of true events can be approximated as uniform in Euclidean space, so 
-\be\label{eq:ps}
-p_s(\rho) = 3 R_t \frac{\rho_1^3}{\rho^4} \, ,
-\ee
-\noindent where $R_t$ is the mean number of true events with $\rho > \rho_1$.
-Then we can write the full $p(\rho)$ as
+We find it useful to define $\rho_1$ as the SNR such that a data set
+will have on average a single noise event louder than $\rho_1$; i.e.,
+such that \be \int_{\rho_1}^{\infty} P_n(\rho)\, d\rho = 1 \, .  \ee
+For this example we will assume that the spatial distribution of true
+events can be approximated as uniform in Euclidean space, so
+\be\label{eq:ps} p_s(\rho) = 3 R_t \frac{\rho_1^3}{\rho^4} \, , \ee
+\noindent where $R_t$ is the mean number of true events with $\rho >
+\rho_1$.  Then we can write the full $p(\rho)$ as
 %and that the spatial distribution of true events can be approximated as uniform in %Euclidean %space. Then $p(\rho)$ has the general form  
-\be\label{eq:ptot}
-p(\rho) = \big[\sqrt{\pi/2} {\rm erfc}(\rho_1/\sqrt{2})\big]^{-1} \, {\rm exp}[-\rho^2/2]\  + \ 3 R_t \frac{\rho_1^3}{\rho^4} \, .
-\ee
+\be\label{eq:ptot} p(\rho) = \big[\sqrt{\pi/2} {\rm
+    erfc}(\rho_1/\sqrt{2})\big]^{-1} \, {\rm exp}[-\rho^2/2]\ + \ 3
+R_t \frac{\rho_1^3}{\rho^4} \, .  \ee
 
@@
 
-For any pair $(\rho_1, R_t)$, it is straightforward to construct random event lists drawn from the corresponding $p(\rho)$, and straightforward to apply a threshold by "throwing away" all events with $\rho$ less than some threshold value $\rho_{th}$.
-For any thresholded event list, we use
-Eqs.~(\curt{need to supply, from earlier in paper} ) to construct the probability density $p(R|data)$ (except that in this example, instead of the
-Jeffrey's prior, we assume a "hat-shaped" prior; see below).
+For any pair $(\rho_1, R_t)$, it is straightforward to construct
+random event lists drawn from the corresponding $p(\rho)$, and
+straightforward to apply a threshold by "throwing away" all events
+with $\rho$ less than some threshold value $\rho_{th}$.  For any
+thresholded event list, we use Eqs.~(\curt{need to supply, from
+  earlier in paper} ) to construct the probability density $p(R|data)$
+(except that in this example, instead of the Jeffrey's prior, we
+assume a "hat-shaped" prior; see below).
 % that is uniform in some interval and zero elsewhere; see below).  
-For that event list, we define $\Delta R_B$ by
-\be
-(\Delta R_B)^2 = \int{(R - R_B)^2 \, p(R|data)\, dR }\, .
-\ee
+For that event list, we define $\Delta R_B$ by \be (\Delta R_B)^2 =
+\int{(R - R_B)^2 \, p(R|data)\, dR }\, .  \ee
 %To find the average behavior, we do many draws of $R_t$ and corresponding data sets.
 %
-Figs.~\ref{fig:delRb} --\ref{fig:Rt10}  illustrate how $\Delta R$ varies with the threshold value $\rho_{th}$.  In all cases we assumed that $\rho_1 = 8$.
+Figs.~\ref{fig:delRb} --\ref{fig:Rt10} illustrate how $\Delta R$
+varies with the threshold value $\rho_{th}$.  In all cases we assumed
+that $\rho_1 = 8$.
 %
 
-For Fig.~\ref{fig:delRb}, we adopted a flat prior on $R$ between $1$ and $10$ (i.e, we took
-$p(R) = 1/9$  for $1 < R <10$, and  zero outside that interval).  
+For Fig.~\ref{fig:delRb}, we adopted a flat prior on $R$ between $1$
+and $10$ (i.e, we took $p(R) = 1/9$ for $1 < R <10$, and zero outside
+that interval).
 % To create Fig.~\ref{fig:delR1}, 
-We drew $10^4$ 
-random realizations $(R_t, data)$, where $R_t$ was drawn from that flat prior and the data--a list of $\rho$ values-- was drawn from the distribution $p(data|R_t)$ given by Eq.~\ref{eq:ptot} .
-For any particular threshold value, $\rho_{th}$, we created an  event list by taking the subset of the $\rho$ values
+We drew $10^4$ random realizations $(R_t, data)$, where $R_t$ was
+drawn from that flat prior and the data--a list of $\rho$ values-- was
+drawn from the distribution $p(data|R_t)$ given by Eq.~\ref{eq:ptot} .
+For any particular threshold value, $\rho_{th}$, we created an event
+list by taking the subset of the $\rho$ values
 % in the data set that are 
-that were greater than $\rho_{th}$.
-For each data set, we considered a sequence of increasing threshold  values $\rho_{th}$, using them to create a corresponding sequence of (ever shorter)  event lists.
-For each event list, we calculated $\Delta R$ and $R_B - R_t$.
+that were greater than $\rho_{th}$.  For each data set, we considered
+a sequence of increasing threshold values $\rho_{th}$, using them to
+create a corresponding sequence of (ever shorter) event lists.  For
+each event list, we calculated $\Delta R$ and $R_B - R_t$.
 %and take the mean over all realizations.  
-As a code check, we verified that the mean of $R_B - R_t$ (over all realizations) was consistent with zero for all $\rho_{th}$.
-Fig.~\ref{fig:delR1} displays both the mean value of $\Delta R$ and the mean number of noise events (as opposed to actual signals) as a function of $\rho_{th}$.
-Note the relative insensitivity of  $\Delta R_t$ to $\rho_{th}$:  as the mean
-number of noise events on the event list shrinks from $8$ to $\sim 10^{-6}$, 
-the average $\Delta R_B$ increases only $\sim 14\%$, from $0.42$ to $0.48$.
+As a code check, we verified that the mean of $R_B - R_t$ (over all
+realizations) was consistent with zero for all $\rho_{th}$.
+Fig.~\ref{fig:delR1} displays both the mean value of $\Delta R$ and
+the mean number of noise events (as opposed to actual signals) as a
+function of $\rho_{th}$.  Note the relative insensitivity of $\Delta
+R_t$ to $\rho_{th}$: as the mean number of noise events on the event
+list shrinks from $8$ to $\sim 10^{-6}$, the average $\Delta R_B$
+increases only $\sim 14\%$, from $0.42$ to $0.48$.
 
@@
 \includegraphics[width=\columnwidth]{delRb}
-\caption{\label{fig:delRb} Displays the mean "error bar" $\Delta R_B$ as function of the threshold value $\rho_{th}$.
-In this case, the values of $R_t$ were randomly drawn from a flat prior on the interval $[1,10]$.
-We also show how the (mean) size of the event list varies with $\rho_{th}$.}
-\end{figure}
-
-Figs.~\ref{fig:Rt1} --\ref{fig:Rt10} were constructed similarly, but with the following differences.  First, we took the prior 
-$p(R)$ to be $p(R) \propto 1/R$ for  $0.03 < R < 30$, and zero outside that interval (i.e., $p(R)$ is the Jeffries prior, but restricted to some range).
-Second, we regarded $R_t$ as fixed (as of course it would be in Nature), and only varied the data--randomly drawing $10^4$ realizations from $p(data|R_t)$.
-For each realization we calculated $R_t^{-1}\,\Delta R_B$ and $(R_B - R_t)$, and then we calculated the means of these quantities over all realizations.
-The mean of $(R_B - R_t)$ was generally {\it not} zero, since the
-prior biases the results.  (Of course, if one {\it updated} the prior after each realization, the 
-bias would decrease towards zero, but that is not what we did: we evaluated each realization with the same Jeffries prior.)
-
-For Figs.~\ref{fig:Rt1} --\ref{fig:Rt10}, the mean number of noise events as a function of $\rho_{th}$ was the same as for the case shown in Fig.~\ref{fig:delRb}.
-Note that for all three values of $R_t$, the mean uncertainty $\Delta R_B$ increases by only $\sim 20\%$  as the mean number of noise events on the event list shrinks from $8$ to $\sim 10^{-6}$.
-Thus the inference that the average $\Delta R_B$ is rather insensitive to  $\rho_{th}$ appears to be robust.
-Finally, we note that for all three cases the bias is much smaller than $\Delta R_B$.
+\caption{\label{fig:delRb} Displays the mean "error bar" $\Delta R_B$
+  as function of the threshold value $\rho_{th}$.  In this case, the
+  values of $R_t$ were randomly drawn from a flat prior on the
+  interval $[1,10]$.  We also show how the (mean) size of the event
+  list varies with $\rho_{th}$.}
+\end{figure}
+
+Figs.~\ref{fig:Rt1} --\ref{fig:Rt10} were constructed similarly, but
+with the following differences.  First, we took the prior $p(R)$ to be
+$p(R) \propto 1/R$ for $0.03 < R < 30$, and zero outside that interval
+(i.e., $p(R)$ is the Jeffries prior, but restricted to some range).
+Second, we regarded $R_t$ as fixed (as of course it would be in
+Nature), and only varied the data--randomly drawing $10^4$
+realizations from $p(data|R_t)$.  For each realization we calculated
+$R_t^{-1}\,\Delta R_B$ and $(R_B - R_t)$, and then we calculated the
+means of these quantities over all realizations.  The mean of $(R_B -
+R_t)$ was generally {\it not} zero, since the prior biases the
+results.  (Of course, if one {\it updated} the prior after each
+realization, the bias would decrease towards zero, but that is not
+what we did: we evaluated each realization with the same Jeffries
+prior.)
+
+For Figs.~\ref{fig:Rt1} --\ref{fig:Rt10}, the mean number of noise
+events as a function of $\rho_{th}$ was the same as for the case shown
+in Fig.~\ref{fig:delRb}.  Note that for all three values of $R_t$, the
+mean uncertainty $\Delta R_B$ increases by only $\sim 20\%$ as the
+mean number of noise events on the event list shrinks from $8$ to
+$\sim 10^{-6}$.  Thus the inference that the average $\Delta R_B$ is
+rather insensitive to $\rho_{th}$ appears to be robust.  Finally, we
+note that for all three cases the bias is much smaller than $\Delta
+R_B$.
 
@@
 \includegraphics[width=\columnwidth]{Rt1}
-\caption{\label{fig:Rt1} Displays the mean rms error $\Delta R_B$ and mean $(R_B - R_t)$ (i.e., the bias) as functions of the threshold value $\rho_{th}$.
-In this case, the actual value of $R_t$ was $1.0$, and $p(R|data)$ was computed using the Jeffries prior on the interval $[0.03,30]$.}
+\caption{\label{fig:Rt1} Displays the mean rms error $\Delta R_B$ and
+  mean $(R_B - R_t)$ (i.e., the bias) as functions of the threshold
+  value $\rho_{th}$.  In this case, the actual value of $R_t$ was
+  $1.0$, and $p(R|data)$ was computed using the Jeffries prior on the
+  interval $[0.03,30]$.}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{Rt3}
-\caption{\label{fig:Rt3} Same as the previous figure, but displaying the size of $\Delta R_B$  relative to $R_t$, with the actual $R_t$ equal to $3.0$.}
+\caption{\label{fig:Rt3} Same as the previous figure, but displaying
+  the size of $\Delta R_B$ relative to $R_t$, with the actual $R_t$
+  equal to $3.0$.}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{Rt10}
-\caption{\label{fig:Rt10} Same as the previous figure, but with the actual $R_t$ equal to $10.0$.}
+\caption{\label{fig:Rt10} Same as the previous figure, but with the
+  actual $R_t$ equal to $10.0$.}
 \end{figure}
@@
 \subsection{Extreme Sensitivity of the LE Rate Estimate to a Single, Unusually Loud Event}
-Here we discuss a very unattractive feature of the Bayesian loudest event estimate of $R$:
-a small percentage of the time it will yield a very large over-estimate.
-
-To explain this, we will use the same model as described in the previous subsection, and we will begin with a very specific example.
-Let the true rate (times the observation time) be $R_t = 1$, meaning (again) that the expected number of actual events with $\rho > \rho_1$ is one.
-Then there is a $1/64$ chance ($\approx 1.6\%$) that the loudest event will have $\rho_{LE} > 4 \rho_1$.
-Consider this case, and let us also assume that there are no events (noise or actual) with $\rho_1 < \rho < \rho_{LE}$.
-The loudest event estimate basically "throws away" the information that there
-are no events in this interval, which (we show below) leads an estimate $R_{LE} \approx 64$,
-  which is $64$ times the true rate!  Now, if the true rate really were $R_t = 64$, then the expected number of 
-events with $\rho > \rho_1$ would be $64$.  So in this case, the loudest event estimate 
-ignores the fact that there are $\sim 56 -72$ "missing" events.  However a Bayesian estimate
-with $\rho_{th}$ set to $\rho_1$ incorporates this information quite naturally, and so (correctly) yields an estimated $R$ of order one.
-
-We now essentially repeat the above discussion, but in greater generality and with the derivations provided.   First recall that $R_B$ is given by
-\be\label{eq:rb1}
-R_B = \frac{  \int{ R\, p(\{\rho_i\}|R)  \, p(R)\, dR}}{  \int{ p(\{\rho_i\}|R)\, p(R)\, dR }}\, ,
-\ee
+Here we discuss a very unattractive feature of the Bayesian loudest
+event estimate of $R$: a small percentage of the time it will yield a
+very large over-estimate.
+
+To explain this, we will use the same model as described in the
+previous subsection, and we will begin with a very specific example.
+Let the true rate (times the observation time) be $R_t = 1$, meaning
+(again) that the expected number of actual events with $\rho > \rho_1$
+is one.  Then there is a $1/64$ chance ($\approx 1.6\%$) that the
+loudest event will have $\rho_{LE} > 4 \rho_1$.  Consider this case,
+and let us also assume that there are no events (noise or actual) with
+$\rho_1 < \rho < \rho_{LE}$.  The loudest event estimate basically
+"throws away" the information that there are no events in this
+interval, which (we show below) leads an estimate $R_{LE} \approx 64$,
+which is $64$ times the true rate!  Now, if the true rate really were
+$R_t = 64$, then the expected number of events with $\rho > \rho_1$
+would be $64$.  So in this case, the loudest event estimate ignores
+the fact that there are $\sim 56 -72$ "missing" events.  However a
+Bayesian estimate with $\rho_{th}$ set to $\rho_1$ incorporates this
+information quite naturally, and so (correctly) yields an estimated
+$R$ of order one.
+
+We now essentially repeat the above discussion, but in greater
+generality and with the derivations provided.  First recall that $R_B$
+is given by \be\label{eq:rb1} R_B = \frac{ \int{ R\, p(\{\rho_i\}|R)
+    \, p(R)\, dR}}{ \int{ p(\{\rho_i\}|R)\, p(R)\, dR }}\, , \ee
 \noindent
-where $p(R)$ is the prior on $R$ and (and up to an overall normalization)
-\be\label{eq:rb2}
-p(\{\rho_i\}|R) = {\rm exp}\bigg[\sum_{i \in  list} {\rm ln}\, p(\rho_i|R) - \int_{\rho_{th}}^{\infty} p(\rho)\, d\,\rho  \bigg] \,  ,
-\ee
+where $p(R)$ is the prior on $R$ and (and up to an overall
+normalization) \be\label{eq:rb2} p(\{\rho_i\}|R) = {\rm
+  exp}\bigg[\sum_{i \in list} {\rm ln}\, p(\rho_i|R) -
+  \int_{\rho_{th}}^{\infty} p(\rho)\, d\,\rho \bigg] \, , \ee
 \noindent
@@
 
-The loudest event estimate is obtained by setting the lower integration limit in the last term on the rhs of Eq.~(\ref{eq:rb2}) to $\rho_{th} = \rho_{LE}$, and replacing the sum 
-$\sum_i {\rm ln}\, p(\rho_i|R)$ by the single term ${\rm ln}\, p(\rho_{LE}|R)$.
-For the case of interest here, where $\rho_{LE}$ is (at least) a few times larger than $\rho_1$ and $R_t$ is of order $1$, we can neglect the noise contribution $p_n(\rho)$.
-Then the integrals in Eq.~(\ref{eq:rb1}) become especially simple, and yield the estimate:
-\be\label{eq:rb3}
-R_B \approx 
+The loudest event estimate is obtained by setting the lower
+integration limit in the last term on the rhs of Eq.~(\ref{eq:rb2}) to
+$\rho_{th} = \rho_{LE}$, and replacing the sum $\sum_i {\rm ln}\,
+p(\rho_i|R)$ by the single term ${\rm ln}\, p(\rho_{LE}|R)$.  For the
+case of interest here, where $\rho_{LE}$ is (at least) a few times
+larger than $\rho_1$ and $R_t$ is of order $1$, we can neglect the
+noise contribution $p_n(\rho)$. Then the integrals in
+Eq.~(\ref{eq:rb1}) become especially simple, and yield the estimate:
+\be\label{eq:rb3} R_B \approx
 \begin{cases}
-(\rho_{LE}/\rho_1)^3 & \text{if $p(R) \propto 1/R$} \\
-2\, (\rho_{LE}/\rho_1)^3 & \text{if $p(R) =$ const \, .} 
+(\rho_{LE}/\rho_1)^3 & \text{if $p(R) \propto 1/R$} \\ 2\,
+  (\rho_{LE}/\rho_1)^3 & \text{if $p(R) =$ const \, .}
 \end{cases}
@@
 
-If instead we set the lower integration limit to $\rho_{th} = \rho_1$, then the $p_n(\rho)$
-contribution to $p(\rho)$ can no longer be neglected.  But the integrals are still 
-straightforward, and we find:
- \be\label{eq:rb3}
-R_B \approx 
+If instead we set the lower integration limit to $\rho_{th} = \rho_1$,
+then the $p_n(\rho)$ contribution to $p(\rho)$ can no longer be
+neglected.  But the integrals are still straightforward, and we find:
+\be\label{eq:rb3} R_B \approx
 \begin{cases}
-1 \ \ \  & \text{if $p(R) \propto 1/R$} \\
-2  \ \ \ & \text{if $p(R) =$ const \, .} 
+1 \ \ \ & \text{if $p(R) \propto 1/R$} \\ 2 \ \ \ & \text{if $p(R) =$
+  const \, .}
 \end{cases}
 \ee
-\noindent Thus (as emphasized above) simply incorporating the information that there are no events {\it between} $\rho_1$ and  $\rho_{LE}$ results in a far beter estimate.
+\noindent Thus (as emphasized above) simply incorporating the
+information that there are no events {\it between} $\rho_1$ and
+$\rho_{LE}$ results in a far beter estimate.
 

### 2012-11-10 many.tex
Re-worked cluster example plots.

@@
 
-\will{Look up von Hippel paper about Bayes 9, the cluster isochrone
-  fitting software---we should cite, since our approach to cluster
-  membership.}
-
 Our final example concerns fitting for the location and shape
@@
 with a spatially varying density (i.e.~our rate functions are
-two-dimensional).  We assume that a star cluster has a Plummer
-surface-density profile \citep{Plummer1911,Aarseth1974},
+two-dimensional).  Our method of analysis here is similar to that of
+\citet{DeGennaro2009}, but here we marginalize over membership flags
+and are simultaneously fitting foreground and background densities
+(i.e.\ rates) \emph{and} cluster properties.
+
+We assume that a star cluster has a Plummer surface-density profile
+\citep{Plummer1911,Aarseth1974},
 \begin{equation}
@@
 where $\vec{\gamma}$ is the gradient, and $\vec{x}_{1/2} = [1/2, 1/2]$
-is the centroid of the observational domain.  \ilya{[I think that if
-    the domain is finite, the previous equation is only normalized for
-    all $\vec{\gamma}$ if $\vec{x}_{1/2}$ is in the center of a
-    symmetric domain (as it happens to be in this case).]}  \will{Yes,
-  as you say, this function is only normalized for $\vec{x}_{1/2}$ is
-  the center, and also for small enough $\vec{\gamma}$; both of these
-  conditions obtain here.}
+is the centroid of the observational domain.
 
@@
 than cluster stars in the field.  Figure \ref{fig:sky-density} shows
-the density of stars (cluster and background) on the sky for our
-particular data set.
+the density of stars on the sky and the particular synthetic data set
+used for this analysis.  Because the peak density of the cluster is
+equal to the background density at the center of the domain, there is
+no single star in the domain that is more likely to be a cluster
+member than a background star (i.e.\ $\left \langle f_i \right \rangle
+\lesssim 0.5$ for all stars); nevertheless, we will see that our
+method provides good constraints on the cluster parameters.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{sky-density}
-  \caption{\label{fig:sky-density} The density contours on the sky of
-    our synthetic data set of cluster and background stars for the
-    parametrs in Eq.~\eqref{eq:true-cluster-parameters}.  There are a
-    factor of 10 more stars (10000) in the background than in the
-    cluster (1000), but the average density of cluster and background
-    stars is similar.}
+  \includegraphics[width=\columnwidth]{cluster-density}
+  \caption{\label{fig:sky-density} Density contours and synthetic data
+    for the example in \S~\ref{sec:star-cluster}.  The contours
+    describe the true density profile with the parameters in
+    Eq.~\eqref{eq:true-cluster-parameters}.  The points are the
+    realization of this density profile used as sythetic data in
+    \S~\ref{sec:star-cluster}; the dashed line encloses one Plummer
+    scale radius about the true cluster center.  Because the peak
+    cluster density is equal to the background density at the cluster
+    center, the cluster is barely apparent to the eye.}
 \end{figure}
@@
 \end{equation}
-\ilya{[Why is this the Jeffreys prior (BTW, it's Jeffreys, or perhaps
-    Jeffreys', but definitely not Jeffrey's ;) )?  It's not obvious
-    why the prior on $r_0$ should depend on $R_f$?]}  \will{Here is
-  the calculation:
-\begin{equation}
-  p\left( r_0 \right) = \sqrt{\left\langle \left(\frac{\partial \log
-      \mathcal{L}}{\partial r_0 } \right)^2 \right\rangle} = \sqrt{
-    R_f \left\langle \left(\frac{\partial \log\hat{f}}{\partial r_0
-    }\right)^2 \right\rangle } = \sqrt{R_f \left\langle \left(
-    \frac{2}{r_0} - \frac{4 r_0}{r^2 + r_0^2} \right)^2 \right\rangle}
-  = \sqrt{\frac{ 4 R_f}{3 r_0^2}} \propto \frac{\sqrt{R_f}}{r_0}.
-\end{equation}
-}\ilya{Sorry, I still don't get it.  Isn't $\mathcal{L} = R_f \hat{f} + R_b \hat{b}$?  If so, how do you get the second equality above?}
-\will{Note that the calculation of the prior uses the likelihood before
-analytical marginalization over the flags, $f_i$.  We are ignoring the
-fact that $R_f$ and $r_0$ are correlated.  In truth, the Jeffreys
-prior would be the determinant of the matrix built out of terms like
-$\partial \log\mathcal{L} / \partial \theta_i \times \partial
-\log\mathcal{L} /\partial \theta_j$, but let's not get carried
-away\ldots.  As another aside, note that we could have predicted the
-$1/r_0$ dependence because $r_0$ has units, so the only unit-invariant
-prior is to be flat in $\log r_0$; I'm don't have a similar argument
-for the appearance of the $\sqrt{R_f}$ factor.}  (Note that this
-factor of $\sqrt{R_f}$ cancels with the Jeffreys prior on the rate,
-$1/\sqrt{R_f}$; we have verified that the priors on these parameters
-are irrelevant to our results, as would be expected from the
-measurement of $\sim 1000$ foreground stars.)
-
-Figure \ref{fig:sky-loc} shows the posterior for the location
-parameters, $\vec{x}_0$; the center of the cluster is localized to
-within about 10\% of the cluster scale.  Figure
-\ref{fig:cluster-number} shows the posteriors inferred on the cluster
-and background numbers, $R_f$ and $R_b$, and Figure
-\ref{fig:cluster-scale} shows the posterior for the cluster's scale
-parameter.  In spite of the significant background, the cluster scale
-and total number are accurately recovered by our analysis.
+(Note that this factor of $\sqrt{R_f}$ cancels with the Jeffreys prior
+on the rate, $1/\sqrt{R_f}$; we have verified that the priors on these
+parameters are irrelevant to our results, as would be expected from
+the measurement of $\sim 1000$ foreground stars.)
+
+Figure \ref{fig:cluster-params} shows the posteriors for the cluster
+parameters.  The center of the cluster, $\vec{x}_0$, is localized to
+within about 5\% of the cluster scale, and the cluster radius with a
+relative error of about 10\%.  In spite of the significant background,
+the cluster parameters are recovered to a relative accuracy consistent
+with the expected uncertainty from $N_\mathrm{eff} \simeq R_f = 1000$
+measurements.  Figure \ref{fig:cluster-number} shows the posteriors
+inferred on the cluster and background numbers, $R_f$ and $R_b$.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{sky_location}
-  \caption{\label{fig:sky-loc} Contours of the posterior probability
-    distribution for the center of the cluster, $\vec{x}_0$, in the
-    example from \S~\ref{sec:star-cluster}.  The center $(x,y) =
+  \includegraphics[width=0.4\columnwidth]{sky_location}\includegraphics[width=0.4\columnwidth]{scale}
+  \caption{\label{fig:cluster-params} Posterior density for the
+    cluster parameters for the example from \S~\ref{sec:star-cluster}.
+    (Left) Contours of the posterior probability distribution for the
+    center of the cluster, $\vec{x}_0$.  The center $(x,y) =
     \left(x_0, y_0\right)$ is determined to within about 5\% of the
     structural radius of the cluster, $r_0$ (see
-    Eq.~\eqref{eq:true-cluster-parameters}).}
+    Eq.~\eqref{eq:true-cluster-parameters}).  (Right) Posterior
+    density for the scale parameter for the cluster, $r_0$.  The true
+    value is indicated by the vertical line (see
+    Eq.~\ref{eq:true-cluster-parameters}).}
 \end{figure}
@@
     Eq.~\eqref{eq:true-cluster-parameters}). }
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{scale}
-  \caption{\label{fig:cluster-scale} Posterior density for the scale
-    parameter for the cluster, $r_0$, from the example in
-    \S~\ref{sec:star-cluster}.  The true value is indicated by the
-    vertical line (see Eq.~\ref{eq:true-cluster-parameters}).
-    \ilya{[Perhaps combine Figs.~8 and 9 (put an extra scale for $r_0$
-        at the top of 8?]} }
 \end{figure}

### 2012-11-10 many.tex
Section re-org.

@@
 subset of the available information.
-
-This paper is organized as follows.  In section \ref{sec:model} we
-introduce our method for Bayesian rate estimation, and describe
-loudest-event and foreground-dominated statistics for comparison.
-Section \ref{sec:GW-example} includes three examples of the
-application of these techniques.  We conclude with a discussion of
-important remaining issues in section \ref{sec:discussion}.
 
@@
 
+\section{Comparison to Other Rate Estimation Methods}
+
 It is informative to relate these results to two other methods for
@@
 
-\subsubsection{Loudest event statistic}
+\subsection{Loudest event statistic}
 If we were to include only the $k$ loudest events in the posterior
@@
 
-\subsubsection{Foreground dominated statistic}
+\subsection{Foreground dominated statistic}
 If we set the threshold for including an event, $x_{\rm min}$,
@@
 (1-\hat{F}(x_2,\theta))$ and $R_{f,2} (1-\hat{F}(x_1,\theta))$.
+
+\section{Dependence of the Bayesian Rate Estimate on the Value of the Threshold}
+This paper is concerned with Bayesian rate estimates based on a lists
+of events (which are a certain kind of summary of the complete data
+set): either the loudest event, or a list of all events whose SNR
+exceeds a certain, pre-set threshold.  In this subsection we address
+the question of how the estimate depends on the threshold value.
+
+To begin with, we recall the well-known fact that the Bayesian
+estimator is unbiased, in the following sense.  For simplicity, assume
+that the model consists of a single rate parameter $R$, with prior
+distribution $p(R)$.  Consider an ensemble of data sets whose
+distribution is consistent with that prior; i.e., such that $p(data)$
+is given by
+%Construct and ensemble of synthetic data sets consistent with $p(R)$; i.e., draw data from the %distribution
+\be p(data) = \int{ p(data|R)\,p(R)\, dR}\, .  \ee For each data set
+in the ensemble, compute the Bayesian estimator $R_{B} = \int{ R
+  p(R|data) dR}$.  Then it is immediate that \be \int{R_{B}(data)\,
+  p(data)\,d(data) }= \int{ R\, p(R) \,dR}\, .  \ee
+\noindent I.e, the data-weighted average of the Bayesian estimate
+$R_{B}$ equals the prior-weighted average $R$.
+%\curt{Guys:  would you like me to supply the argument?}
+%
+Therefore (for any fixed prior) all threshold values will yield, on
+average, the same estimate of the rate.  However this equality of
+averages does {\it not} imply that all threshold values yield the same
+information. In general, as the threshold is lowered to include more
+events, the error bar on the estimate shrink.  In this subsection we
+give quantitative illustrations of how the error bar shrinks when the
+threshold is lowered.
+
+Consider the following model problem.
+% in which the distribution of noise events is Gaussian.
+Let $p(\rho)\Delta\rho$ be the probability that data yields a putative
+event with SNR in the range $[\rho - \Delta\rho/2, \rho +
+  \Delta\rho/2]$ We write $p = p_n + p_s$, where $p_n$ is the
+distribution of noise events and $p_s$ is the distribution of events
+from actual astrophysical sources.  Both $p_n$ and $p_s$ are the
+product of a rate (of noise or true events) and the observation time.
+Here we will assume that the noise distribution is Gaussian, so that
+$p_n$ has the general form \be p_n(\rho) = \Lambda_n {\rm
+  exp}\big[-\rho^2/2\big] \, .  \ee
+\noindent
+We find it useful to define $\rho_1$ as the SNR such that a data set
+will have on average a single noise event louder than $\rho_1$; i.e.,
+such that \be \int_{\rho_1}^{\infty} P_n(\rho)\, d\rho = 1 \, .  \ee
+For this example we will assume that the spatial distribution of true
+events can be approximated as uniform in Euclidean space, so
+\be\label{eq:ps} p_s(\rho) = 3 R_t \frac{\rho_1^3}{\rho^4} \, , \ee
+\noindent where $R_t$ is the mean number of true events with $\rho >
+\rho_1$.  Then we can write the full $p(\rho)$ as
+%and that the spatial distribution of true events can be approximated as uniform in %Euclidean %space. Then $p(\rho)$ has the general form  
+\be\label{eq:ptot} p(\rho) = \big[\sqrt{\pi/2} {\rm
+    erfc}(\rho_1/\sqrt{2})\big]^{-1} \, {\rm exp}[-\rho^2/2]\ + \ 3
+R_t \frac{\rho_1^3}{\rho^4} \, .  \ee
+
+%What matters for the following analysis is the rate times the observation time.
+
+For any pair $(\rho_1, R_t)$, it is straightforward to construct
+random event lists drawn from the corresponding $p(\rho)$, and
+straightforward to apply a threshold by "throwing away" all events
+with $\rho$ less than some threshold value $\rho_{th}$.  For any
+thresholded event list, we use Eqs.~(\curt{need to supply, from
+  earlier in paper} ) to construct the probability density $p(R|data)$
+(except that in this example, instead of the Jeffrey's prior, we
+assume a "hat-shaped" prior; see below).
+% that is uniform in some interval and zero elsewhere; see below).  
+For that event list, we define $\Delta R_B$ by \be (\Delta R_B)^2 =
+\int{(R - R_B)^2 \, p(R|data)\, dR }\, .  \ee
+%To find the average behavior, we do many draws of $R_t$ and corresponding data sets.
+%
+Figs.~\ref{fig:delRb} --\ref{fig:Rt10} illustrate how $\Delta R$
+varies with the threshold value $\rho_{th}$.  In all cases we assumed
+that $\rho_1 = 8$.
+%
+
+For Fig.~\ref{fig:delRb}, we adopted a flat prior on $R$ between $1$
+and $10$ (i.e, we took $p(R) = 1/9$ for $1 < R <10$, and zero outside
+that interval).
+% To create Fig.~\ref{fig:delR1}, 
+We drew $10^4$ random realizations $(R_t, data)$, where $R_t$ was
+drawn from that flat prior and the data--a list of $\rho$ values-- was
+drawn from the distribution $p(data|R_t)$ given by Eq.~\ref{eq:ptot} .
+For any particular threshold value, $\rho_{th}$, we created an event
+list by taking the subset of the $\rho$ values
+% in the data set that are 
+that were greater than $\rho_{th}$.  For each data set, we considered
+a sequence of increasing threshold values $\rho_{th}$, using them to
+create a corresponding sequence of (ever shorter) event lists.  For
+each event list, we calculated $\Delta R$ and $R_B - R_t$.
+%and take the mean over all realizations.  
+As a code check, we verified that the mean of $R_B - R_t$ (over all
+realizations) was consistent with zero for all $\rho_{th}$.
+Fig.~\ref{fig:delRb} displays both the mean value of $\Delta R$ and
+the mean number of noise events (as opposed to actual signals) as a
+function of $\rho_{th}$.  Note the relative insensitivity of $\Delta
+R_t$ to $\rho_{th}$: as the mean number of noise events on the event
+list shrinks from $8$ to $\sim 10^{-6}$, the average $\Delta R_B$
+increases only $\sim 14\%$, from $0.42$ to $0.48$.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{delRb}
+\caption{\label{fig:delRb} Displays the mean "error bar" $\Delta R_B$
+  as function of the threshold value $\rho_{th}$.  In this case, the
+  values of $R_t$ were randomly drawn from a flat prior on the
+  interval $[1,10]$.  We also show how the (mean) size of the event
+  list varies with $\rho_{th}$.}
+\end{figure}
+
+Figs.~\ref{fig:Rt1} --\ref{fig:Rt10} were constructed similarly, but
+with the following differences.  First, we took the prior $p(R)$ to be
+$p(R) \propto 1/R$ for $0.03 < R < 30$, and zero outside that interval
+(i.e., $p(R)$ is the Jeffries prior, but restricted to some range).
+Second, we regarded $R_t$ as fixed (as of course it would be in
+Nature), and only varied the data--randomly drawing $10^4$
+realizations from $p(data|R_t)$.  For each realization we calculated
+$R_t^{-1}\,\Delta R_B$ and $(R_B - R_t)$, and then we calculated the
+means of these quantities over all realizations.  The mean of $(R_B -
+R_t)$ was generally {\it not} zero, since the prior biases the
+results.  (Of course, if one {\it updated} the prior after each
+realization, the bias would decrease towards zero, but that is not
+what we did: we evaluated each realization with the same Jeffries
+prior.)
+
+For Figs.~\ref{fig:Rt1} --\ref{fig:Rt10}, the mean number of noise
+events as a function of $\rho_{th}$ was the same as for the case shown
+in Fig.~\ref{fig:delRb}.  Note that for all three values of $R_t$, the
+mean uncertainty $\Delta R_B$ increases by only $\sim 20\%$ as the
+mean number of noise events on the event list shrinks from $8$ to
+$\sim 10^{-6}$.  Thus the inference that the average $\Delta R_B$ is
+rather insensitive to $\rho_{th}$ appears to be robust.  Finally, we
+note that for all three cases the bias is much smaller than $\Delta
+R_B$.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{Rt1}
+\caption{\label{fig:Rt1} Displays the mean rms error $\Delta R_B$ and
+  mean $(R_B - R_t)$ (i.e., the bias) as functions of the threshold
+  value $\rho_{th}$.  In this case, the actual value of $R_t$ was
+  $1.0$, and $p(R|data)$ was computed using the Jeffries prior on the
+  interval $[0.03,30]$.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{Rt3}
+\caption{\label{fig:Rt3} Same as the previous figure, but displaying
+  the size of $\Delta R_B$ relative to $R_t$, with the actual $R_t$
+  equal to $3.0$.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{Rt10}
+\caption{\label{fig:Rt10} Same as the previous figure, but with the
+  actual $R_t$ equal to $10.0$.}
+\end{figure}
+
+\subsection{Extreme Sensitivity of the LE Rate Estimate to a Single, Unusually Loud Event}
+Here we discuss a very unattractive feature of the Bayesian loudest
+event estimate of $R$: a small percentage of the time it will yield a
+very large over-estimate.
+
+To explain this, we will use the same model as described in the
+previous subsection, and we will begin with a very specific example.
+Let the true rate (times the observation time) be $R_t = 1$, meaning
+(again) that the expected number of actual events with $\rho > \rho_1$
+is one.  Then there is a $1/64$ chance ($\approx 1.6\%$) that the
+loudest event will have $\rho_{LE} > 4 \rho_1$.  Consider this case,
+and let us also assume that there are no events (noise or actual) with
+$\rho_1 < \rho < \rho_{LE}$.  The loudest event estimate basically
+"throws away" the information that there are no events in this
+interval, which (we show below) leads an estimate $R_{LE} \approx 64$,
+which is $64$ times the true rate!  Now, if the true rate really were
+$R_t = 64$, then the expected number of events with $\rho > \rho_1$
+would be $64$.  So in this case, the loudest event estimate ignores
+the fact that there are $\sim 56 -72$ "missing" events.  However a
+Bayesian estimate with $\rho_{th}$ set to $\rho_1$ incorporates this
+information quite naturally, and so (correctly) yields an estimated
+$R$ of order one.
+
+We now essentially repeat the above discussion, but in greater
+generality and with the derivations provided.  First recall that $R_B$
+is given by \be\label{eq:rb1} R_B = \frac{ \int{ R\, p(\{\rho_i\}|R)
+    \, p(R)\, dR}}{ \int{ p(\{\rho_i\}|R)\, p(R)\, dR }}\, , \ee
+\noindent
+where $p(R)$ is the prior on $R$ and (and up to an overall
+normalization) \be\label{eq:rb2} p(\{\rho_i\}|R) = {\rm
+  exp}\bigg[\sum_{i \in list} {\rm ln}\, p(\rho_i|R) -
+  \int_{\rho_{th}}^{\infty} p(\rho)\, d\,\rho \bigg] \, , \ee
+\noindent
+with $p(\rho)$ given by Eq.~().
+
+The loudest event estimate is obtained by setting the lower
+integration limit in the last term on the rhs of Eq.~(\ref{eq:rb2}) to
+$\rho_{th} = \rho_{LE}$, and replacing the sum $\sum_i {\rm ln}\,
+p(\rho_i|R)$ by the single term ${\rm ln}\, p(\rho_{LE}|R)$.  For the
+case of interest here, where $\rho_{LE}$ is (at least) a few times
+larger than $\rho_1$ and $R_t$ is of order $1$, we can neglect the
+noise contribution $p_n(\rho)$. Then the integrals in
+Eq.~(\ref{eq:rb1}) become especially simple, and yield the estimate:
+\be\label{eq:rb3} R_B \approx
+\begin{cases}
+(\rho_{LE}/\rho_1)^3 & \text{if $p(R) \propto 1/R$} \\ 2\,
+  (\rho_{LE}/\rho_1)^3 & \text{if $p(R) =$ const \, .}
+\end{cases}
+\ee
+
+If instead we set the lower integration limit to $\rho_{th} = \rho_1$,
+then the $p_n(\rho)$ contribution to $p(\rho)$ can no longer be
+neglected.  But the integrals are still straightforward, and we find:
+\be\label{eq:rb3} R_B \approx
+\begin{cases}
+1 \ \ \ & \text{if $p(R) \propto 1/R$} \\ 2 \ \ \ & \text{if $p(R) =$
+  const \, .}
+\end{cases}
+\ee
+\noindent Thus (as emphasized above) simply incorporating the
+information that there are no events {\it between} $\rho_1$ and
+$\rho_{LE}$ results in a far beter estimate.
 
@@
 
-\subsection{Dependence of the Bayesian Rate Estimate on the Value of the Threshold}
-This paper is concerned with Bayesian rate estimates based on a lists
-of events (which are a certain kind of summary of the complete data
-set): either the loudest event, or a list of all events whose SNR
-exceeds a certain, pre-set threshold.  In this subsection we address
-the question of how the estimate depends on the threshold value.
-
-To begin with, we recall the well-known fact that the Bayesian
-estimator is unbiased, in the following sense.  For simplicity, assume
-that the model consists of a single rate parameter $R$, with prior
-distribution $p(R)$.  Consider an ensemble of data sets whose
-distribution is consistent with that prior; i.e., such that $p(data)$
-is given by
-%Construct and ensemble of synthetic data sets consistent with $p(R)$; i.e., draw data from the %distribution
-\be p(data) = \int{ p(data|R)\,p(R)\, dR}\, .  \ee For each data set
-in the ensemble, compute the Bayesian estimator $R_{B} = \int{ R
-  p(R|data) dR}$.  Then it is immediate that \be \int{R_{B}(data)\,
-  p(data)\,d(data) }= \int{ R\, p(R) \,dR}\, .  \ee
-\noindent I.e, the data-weighted average of the Bayesian estimate
-$R_{B}$ equals the prior-weighted average $R$.
-%\curt{Guys:  would you like me to supply the argument?}
-%
-Therefore (for any fixed prior) all threshold values will yield, on
-average, the same estimate of the rate.  However this equality of
-averages does {\it not} imply that all threshold values yield the same
-information. In general, as the threshold is lowered to include more
-events, the error bar on the estimate shrink.  In this subsection we
-give quantitative illustrations of how the error bar shrinks when the
-threshold is lowered.
-
-Consider the following model problem.
-% in which the distribution of noise events is Gaussian.
-Let $p(\rho)\Delta\rho$ be the probability that data yields a putative
-event with SNR in the range $[\rho - \Delta\rho/2, \rho +
-  \Delta\rho/2]$ We write $p = p_n + p_s$, where $p_n$ is the
-distribution of noise events and $p_s$ is the distribution of events
-from actual astrophysical sources.  Both $p_n$ and $p_s$ are the
-product of a rate (of noise or true events) and the observation time.
-Here we will assume that the noise distribution is Gaussian, so that
-$p_n$ has the general form \be p_n(\rho) = \Lambda_n {\rm
-  exp}\big[-\rho^2/2\big] \, .  \ee
-\noindent
-We find it useful to define $\rho_1$ as the SNR such that a data set
-will have on average a single noise event louder than $\rho_1$; i.e.,
-such that \be \int_{\rho_1}^{\infty} P_n(\rho)\, d\rho = 1 \, .  \ee
-For this example we will assume that the spatial distribution of true
-events can be approximated as uniform in Euclidean space, so
-\be\label{eq:ps} p_s(\rho) = 3 R_t \frac{\rho_1^3}{\rho^4} \, , \ee
-\noindent where $R_t$ is the mean number of true events with $\rho >
-\rho_1$.  Then we can write the full $p(\rho)$ as
-%and that the spatial distribution of true events can be approximated as uniform in %Euclidean %space. Then $p(\rho)$ has the general form  
-\be\label{eq:ptot} p(\rho) = \big[\sqrt{\pi/2} {\rm
-    erfc}(\rho_1/\sqrt{2})\big]^{-1} \, {\rm exp}[-\rho^2/2]\ + \ 3
-R_t \frac{\rho_1^3}{\rho^4} \, .  \ee
-
-%What matters for the following analysis is the rate times the observation time.
-
-For any pair $(\rho_1, R_t)$, it is straightforward to construct
-random event lists drawn from the corresponding $p(\rho)$, and
-straightforward to apply a threshold by "throwing away" all events
-with $\rho$ less than some threshold value $\rho_{th}$.  For any
-thresholded event list, we use Eqs.~(\curt{need to supply, from
-  earlier in paper} ) to construct the probability density $p(R|data)$
-(except that in this example, instead of the Jeffrey's prior, we
-assume a "hat-shaped" prior; see below).
-% that is uniform in some interval and zero elsewhere; see below).  
-For that event list, we define $\Delta R_B$ by \be (\Delta R_B)^2 =
-\int{(R - R_B)^2 \, p(R|data)\, dR }\, .  \ee
-%To find the average behavior, we do many draws of $R_t$ and corresponding data sets.
-%
-Figs.~\ref{fig:delRb} --\ref{fig:Rt10} illustrate how $\Delta R$
-varies with the threshold value $\rho_{th}$.  In all cases we assumed
-that $\rho_1 = 8$.
-%
-
-For Fig.~\ref{fig:delRb}, we adopted a flat prior on $R$ between $1$
-and $10$ (i.e, we took $p(R) = 1/9$ for $1 < R <10$, and zero outside
-that interval).
-% To create Fig.~\ref{fig:delR1}, 
-We drew $10^4$ random realizations $(R_t, data)$, where $R_t$ was
-drawn from that flat prior and the data--a list of $\rho$ values-- was
-drawn from the distribution $p(data|R_t)$ given by Eq.~\ref{eq:ptot} .
-For any particular threshold value, $\rho_{th}$, we created an event
-list by taking the subset of the $\rho$ values
-% in the data set that are 
-that were greater than $\rho_{th}$.  For each data set, we considered
-a sequence of increasing threshold values $\rho_{th}$, using them to
-create a corresponding sequence of (ever shorter) event lists.  For
-each event list, we calculated $\Delta R$ and $R_B - R_t$.
-%and take the mean over all realizations.  
-As a code check, we verified that the mean of $R_B - R_t$ (over all
-realizations) was consistent with zero for all $\rho_{th}$.
-Fig.~\ref{fig:delR1} displays both the mean value of $\Delta R$ and
-the mean number of noise events (as opposed to actual signals) as a
-function of $\rho_{th}$.  Note the relative insensitivity of $\Delta
-R_t$ to $\rho_{th}$: as the mean number of noise events on the event
-list shrinks from $8$ to $\sim 10^{-6}$, the average $\Delta R_B$
-increases only $\sim 14\%$, from $0.42$ to $0.48$.
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{delRb}
-\caption{\label{fig:delRb} Displays the mean "error bar" $\Delta R_B$
-  as function of the threshold value $\rho_{th}$.  In this case, the
-  values of $R_t$ were randomly drawn from a flat prior on the
-  interval $[1,10]$.  We also show how the (mean) size of the event
-  list varies with $\rho_{th}$.}
-\end{figure}
-
-Figs.~\ref{fig:Rt1} --\ref{fig:Rt10} were constructed similarly, but
-with the following differences.  First, we took the prior $p(R)$ to be
-$p(R) \propto 1/R$ for $0.03 < R < 30$, and zero outside that interval
-(i.e., $p(R)$ is the Jeffries prior, but restricted to some range).
-Second, we regarded $R_t$ as fixed (as of course it would be in
-Nature), and only varied the data--randomly drawing $10^4$
-realizations from $p(data|R_t)$.  For each realization we calculated
-$R_t^{-1}\,\Delta R_B$ and $(R_B - R_t)$, and then we calculated the
-means of these quantities over all realizations.  The mean of $(R_B -
-R_t)$ was generally {\it not} zero, since the prior biases the
-results.  (Of course, if one {\it updated} the prior after each
-realization, the bias would decrease towards zero, but that is not
-what we did: we evaluated each realization with the same Jeffries
-prior.)
-
-For Figs.~\ref{fig:Rt1} --\ref{fig:Rt10}, the mean number of noise
-events as a function of $\rho_{th}$ was the same as for the case shown
-in Fig.~\ref{fig:delRb}.  Note that for all three values of $R_t$, the
-mean uncertainty $\Delta R_B$ increases by only $\sim 20\%$ as the
-mean number of noise events on the event list shrinks from $8$ to
-$\sim 10^{-6}$.  Thus the inference that the average $\Delta R_B$ is
-rather insensitive to $\rho_{th}$ appears to be robust.  Finally, we
-note that for all three cases the bias is much smaller than $\Delta
-R_B$.
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{Rt1}
-\caption{\label{fig:Rt1} Displays the mean rms error $\Delta R_B$ and
-  mean $(R_B - R_t)$ (i.e., the bias) as functions of the threshold
-  value $\rho_{th}$.  In this case, the actual value of $R_t$ was
-  $1.0$, and $p(R|data)$ was computed using the Jeffries prior on the
-  interval $[0.03,30]$.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{Rt3}
-\caption{\label{fig:Rt3} Same as the previous figure, but displaying
-  the size of $\Delta R_B$ relative to $R_t$, with the actual $R_t$
-  equal to $3.0$.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{Rt10}
-\caption{\label{fig:Rt10} Same as the previous figure, but with the
-  actual $R_t$ equal to $10.0$.}
-\end{figure}
-
-\subsection{Extreme Sensitivity of the LE Rate Estimate to a Single, Unusually Loud Event}
-Here we discuss a very unattractive feature of the Bayesian loudest
-event estimate of $R$: a small percentage of the time it will yield a
-very large over-estimate.
-
-To explain this, we will use the same model as described in the
-previous subsection, and we will begin with a very specific example.
-Let the true rate (times the observation time) be $R_t = 1$, meaning
-(again) that the expected number of actual events with $\rho > \rho_1$
-is one.  Then there is a $1/64$ chance ($\approx 1.6\%$) that the
-loudest event will have $\rho_{LE} > 4 \rho_1$.  Consider this case,
-and let us also assume that there are no events (noise or actual) with
-$\rho_1 < \rho < \rho_{LE}$.  The loudest event estimate basically
-"throws away" the information that there are no events in this
-interval, which (we show below) leads an estimate $R_{LE} \approx 64$,
-which is $64$ times the true rate!  Now, if the true rate really were
-$R_t = 64$, then the expected number of events with $\rho > \rho_1$
-would be $64$.  So in this case, the loudest event estimate ignores
-the fact that there are $\sim 56 -72$ "missing" events.  However a
-Bayesian estimate with $\rho_{th}$ set to $\rho_1$ incorporates this
-information quite naturally, and so (correctly) yields an estimated
-$R$ of order one.
-
-We now essentially repeat the above discussion, but in greater
-generality and with the derivations provided.  First recall that $R_B$
-is given by \be\label{eq:rb1} R_B = \frac{ \int{ R\, p(\{\rho_i\}|R)
-    \, p(R)\, dR}}{ \int{ p(\{\rho_i\}|R)\, p(R)\, dR }}\, , \ee
-\noindent
-where $p(R)$ is the prior on $R$ and (and up to an overall
-normalization) \be\label{eq:rb2} p(\{\rho_i\}|R) = {\rm
-  exp}\bigg[\sum_{i \in list} {\rm ln}\, p(\rho_i|R) -
-  \int_{\rho_{th}}^{\infty} p(\rho)\, d\,\rho \bigg] \, , \ee
-\noindent
-with $p(\rho)$ given by Eq.~().
-
-The loudest event estimate is obtained by setting the lower
-integration limit in the last term on the rhs of Eq.~(\ref{eq:rb2}) to
-$\rho_{th} = \rho_{LE}$, and replacing the sum $\sum_i {\rm ln}\,
-p(\rho_i|R)$ by the single term ${\rm ln}\, p(\rho_{LE}|R)$.  For the
-case of interest here, where $\rho_{LE}$ is (at least) a few times
-larger than $\rho_1$ and $R_t$ is of order $1$, we can neglect the
-noise contribution $p_n(\rho)$. Then the integrals in
-Eq.~(\ref{eq:rb1}) become especially simple, and yield the estimate:
-\be\label{eq:rb3} R_B \approx
-\begin{cases}
-(\rho_{LE}/\rho_1)^3 & \text{if $p(R) \propto 1/R$} \\ 2\,
-  (\rho_{LE}/\rho_1)^3 & \text{if $p(R) =$ const \, .}
-\end{cases}
-\ee
-
-If instead we set the lower integration limit to $\rho_{th} = \rho_1$,
-then the $p_n(\rho)$ contribution to $p(\rho)$ can no longer be
-neglected.  But the integrals are still straightforward, and we find:
-\be\label{eq:rb3} R_B \approx
-\begin{cases}
-1 \ \ \ & \text{if $p(R) \propto 1/R$} \\ 2 \ \ \ & \text{if $p(R) =$
-  const \, .}
-\end{cases}
-\ee
-\noindent Thus (as emphasized above) simply incorporating the
-information that there are no events {\it between} $\rho_1$ and
-$\rho_{LE}$ results in a far beter estimate.
-
 \subsection{Star Cluster Parameters With Background Contamination}

### 2012-11-23 many.tex
Added section VC on dealing with uncertain foreground and background distributions

@@
 with $\rho$ less than some threshold value $\rho_{th}$.  For any
-thresholded event list, we use Eqs.~(\curt{need to supply, from
-  earlier in paper} ) to construct the probability density $p(R|data)$
+thresholded event list, we use Eq.~(\ref{eq:rate-shape-posterior}) to construct the probability density $p(R|data)$
 (except that in this example, instead of the Jeffrey's prior, we
-assume a "hat-shaped" prior; see below).
+assume a ``hat-shaped'' prior; see below).
 % that is uniform in some interval and zero elsewhere; see below).  
@@
 \noindent
-with $p(\rho)$ given by Eq.~().
+with $p(\rho)$ given by Eq.~(\ref{eq:ptot}).
 
@@
 
+\subsection{Uncertainty in the Foreground and Background Distributions}
+\label{sec:uncertain-dist}
+The framework outlined above relies on the existence of models for the foreground, $\hat{f}(x,\theta)$, and background, $\hat{b}(x,\theta)$, distributions parameterised by a small number of model parameters, $\theta$.
+While in many situations simple analytic functions such as power laws will provide an adequate description, this will not always be the case.
+In the absence of a good analytic model, the space of the ranking statistic, $x$ could be divided into bins and $\hat{f}(x)$/$\hat{b}(x)$ taken to be flat in each of these bins.
+The number of free parameters characterising each of $\hat{f}$ and $\hat{b}$ is then the number of bins used.
+While such a framework is model free, the increase in model parameters will mean that more observed events will typically be required to achieve the same precision on the rates and foreground/background distributions.
+
+In the context of gravitational wave experiments, additional information on the ranking statistic distributions for the foreground and background can be obtained using injections into and timeslides of the data respectively.
+This information can be readily incorporated in the current framework by assuming there is another set of $N_I$ events with ranking statistics $\{ w_i \}$, known to be drawn from the foreground distribution ($f_i=1$) and a set of $N_T$ events with ranking statistics $\{z_i\}$ known to be drawn from the background distribution ($f_i=0$).
+These events will typically not be drawn with the correct rate parameters, so they do not contribute to the estimates of $R_f$ and $R_b$, but they do contribute an extra factor
+\be
+\prod_{l=1}^{N_I} \hat{f}(w_l, \theta ) \prod_{m=1}^{N_T} \hat{b}(z_i, \theta) 
+\ee
+to the right hand sides of Eqs~(\ref{eq:posterior}) and (\ref{eq:rate-shape-posterior}).
+This approach provides a way to incorporate extra information into the analysis in order to simultaneously fit for the shape of the background and foreground as well as the rates.
+In the limit that there are many more events in the timeslide and injection data set, this will reduce to the analysis that was described above with fixed ranking statistic distributions $\hat{f}(x)$ and $\hat{b}(x)$ given by the injection and time slide data.
+We note that this analysis makes the assumption that the background distribution is the same in the time slide and real data and that the foreground distribution is the same between the injection and real data.
+The former assumption is probably reasonable, modulo non-stationarities in the detector data, but the latter relies on knowledge of the astrophysical rate of events which is more uncertain.
+These astrophysical uncertainties could be handled with a hybrid approach, in which injections are used to characterise the statistic distribution for sources of a particular type, while additional rate or shape parameters are introduced to characterise the variation in the astrophysical rate of mergers as a function of source type.
+
 \subsection{Star Cluster Parameters With Background Contamination}

### 2012-12-14 many.tex
Proposed for submission to P&P.
Changes:

* Re-write of ``Thresholding'' section (Curt's notes); condensed figures, too.

* Title changed (more punchy?)

* Updates to abstract.

* Re-format to ``reprint'' style to more closely match journal.

* Standardized on US English spellings (s --> z in normalisation, etc).

@@
-\documentclass[aps,prd]{revtex4-1}
+\documentclass[aps,prd,reprint]{revtex4-1}
 
@@
 \DeclareMathOperator{\erf}{erf}
+\DeclareMathOperator{\erfc}{erfc}
 
@@
 
-\title{Bayesian Rate Estimation}
+\title{Counting With Confusion: Bayesian Rate Estimation With Multiple
+  Populations}
 
@@
 \begin{abstract}
-  We show how to obtain a Bayesian estimate of the rate of signal
-  events from a set of signal and background events when the shapes of
+  We show how to obtain a Bayesian estimate of the rates or numbers of
+  signal and background events from a set of events when the shapes of
   the signal and background distributions are known, can be estimated,
   or approximated; our method works well even if the foreground and
-  background event distributions overlap significantly.  We give
-  examples of determining the rates of gravitatonal-wave events in the
-  presence of background triggers from a template bank when noise
-  parameters are known and/or can be fit from the trigger data.  We
-  also give an example of determining globular-cluster shape and
-  location parameters from an observation of a stellar field that
+  background event distributions overlap significantly and the status
+  of any individual event cannot be determined with any certainty.  We
+  give examples of determining the rates of gravitational-wave events
+  in the presence of background triggers from a template bank when
+  noise parameters are known and/or can be fit from the trigger data.
+  We also give an example of determining globular-cluster shape,
+  location, and density from an observation of a stellar field that
   contains a non-uniform background density of stars superimposed on
@@
 
-The task of estimating a rate of foreground events when a mixture of
-foreground and background events is present in the data is a common
-one in physical and astrophysical applications.  This problem comes
-up, among others, in gravitational-wave data analysis
+The task of estimating rates of events when a mixture of foreground
+and background events is present in data is a common one in physical
+and astrophysical applications.  This problem comes up, among others,
+in gravitational-wave data analysis
 \cite[e.g.,][]{s5-highmass,s6-lowmass} and in astronomical
@@
 for estimating event rates from the data when the shape of foreground
-and background distributions are known (but may have additional free
-parameters), but the provenance of individual events as either
-background or foreground is unknown.
+and background distributions are known (or parameterized), but the
+provenance of individual events as either background or foreground is
+unknown.
 
@@
 
-Two alternative approaches have been suggested in the past.  One,
-known as the \emph{loudest-event statistic} \cite{Biswas2009}, uses
-only the information from the highest-ranked event in the data to
-infer the rate distribution.  This approach has been used successfully
-\cite{s5-highmass,s6-lowmass} when the number of loud foreground
-events is small (typically zero or one) to obtain upper limits on
-foreground rates.  However, the loudest-event statistic ignores all
-events except the loudest one, and so suffers from an unnecessary loss
-of information; therefore, we expect it to yield a much larger
-variance than strictly necessary when multiple events are present in
-the data.
+Two alternative approaches to rate estimation have been suggested and
+are commonly used.  One, known as the \emph{loudest-event statistic}
+\cite{Biswas2009}, uses only the information from the highest-ranked
+event in the data to infer the rate distribution.  This approach has
+been used successfully \cite{s5-highmass,s6-lowmass} when the number
+of loud foreground events is small (typically zero or one) to obtain
+upper limits on foreground rates.  However, the loudest-event
+statistic ignores all events except the loudest one, and so suffers
+from an unnecessary loss of information; therefore, we expect it to
+yield a much larger variance than strictly necessary when multiple
+events are present in the data.
 
@@
 loudest events, and the number of such loud events is sufficiently
-large.  However, it cannot properly account for marginal events.
-While either the loudest-event statistic or the foreground-dominated
-statistic can approach the accuracy of our proposed method in specific
-regimes, both are suboptimal in a general case.
+large.  However, it cannot properly account for marginal events.  In
+addition, the results of the method are very sensitive to
+contamination by the background events, and therefore the method
+requires a careful choice of \emph{threshold} or reliable
+\emph{membership} information to distinguish foregrounds and
+backgrounds for individual events.  While either the loudest-event
+statistic or the foreground-dominated statistic can approach the
+accuracy of our proposed method in specific regimes, both are
+suboptimal in a general case.
 
@@
 obtained, $p\left( \mathset{f_i}, N, R_f, R_b, \theta\right)$:
-\begin{equation}
+\begin{multline}
   \label{eq:Bayes}
-  p\left(\mathset{f_i}, R_f, R_b, \theta | d_{\rm to}, N \right) =
+  p\left(\mathset{f_i}, R_f, R_b, \theta | d_{\rm to}, N \right) \\ =
   \frac{p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right)
@@
     N)}.
-\end{equation}
+\end{multline}
 The normalization constant, called the evidence, $p(d_{\rm to}, N)$,
@@
 Therefore, the likelihood of the data is
-\begin{equation}
-  p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right) =
+\begin{multline}
+  p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right) \\ =
   \left[ \prod_{\mathset{i | f_i = 1}} \hat{f}\left(x_i, \theta\right)
@@
     \theta\right) \right].
-\end{equation}
+\end{multline}
 This is the probability that the first observed event is a
@@
 
-%The prior distribution can be factorized as
-%\begin{equation}
-%  \label{eq:combined-flag-rate-prior}
-%  p\left(\mathset{f_i}, R_f, R_b, \theta\right)  = p\left(
-%   \mathset{f_i} | R_f, R_b, \theta\right)p(R_f, R_b, \theta).
-%\end{equation}
-%The first term is not actually a ``prior'' in the usual sense; the
-%conditional probability of the flags $\mathset{f_i}$ on the rates is given
-%by
-%\begin{equation}
-%  \label{eq:flag-conditional-prior}
-%  p\left(\mathset{f_i} | R_f, R_b, \theta\right) = \frac{R_f^{N_f}
-%    R_b^{N_b}}{N_f! N_b!} \exp\left[ - \left(R_f + R_b\right) \right]
-%  N_f! N_b! = R_f^{N_f}R_b^{N_b} \exp\left[ - \left(R_f + R_b\right) \right],
-%\end{equation}
-%where $N_f$ and $N_b$ are the number of foreground and background
-%events.  This expression follows from Property \ref{prop:Poisson} of
-%inhomogeneous Poisson processes and the $N_f! N_b!$ different ways of
-%assigning foreground and background flags to events for fixed $N_f$,
-%$N_b$.  
-%
-%\ilya{[I agree with the final result, but I find this way of deriving
-%    it to be confusing.  First of all, the number of ways to assign
-%    flags to events for fixed $N_f$ and $N_b$ is $N!/(N_f! N_b!)$, not
-%    $N_f! N_b!$ as written.  Secondly, you are computing the
-%    probability of having a particular ordered list of flags, so
-%    there's no reason to consider different permutations.  Thirdly,
-%    you are not properly including the fact that the length of the
-%    flags vector must be exactly $N$, the total number of data points.
-%    (Well, you have the right result, so you are obviously doing these
-%    things properly, but the explanation just doesn't seem clear to
-%    me. :) ) I would make that $N$ explicit, by writing Eq.~(12) as
-
 The prior distribution can be factorized as
-\begin{equation}
+\begin{multline}
   \label{eq:combined-flag-rate-prior}
-  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right) = p\left(
+  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right) \\ = p\left(
   \mathset{f_i} | N, R_f, R_b\right) p\left(N |R_f, R_b\right)
-  p\left(R_f, R_b, \theta\right) = p\left( \mathset{f_i},N | R_f,
+  p\left(R_f, R_b, \theta\right) \\ = p\left( \mathset{f_i},N | R_f,
   R_b\right) p\left(R_f, R_b, \theta\right).
- \end{equation}
+ \end{multline}
 
@@
 assumed.  Then
-\begin{equation}
-p\left( \mathset{f_i} | N, R_f, R_b\right) = \prod_{\mathset{i|f_i=1}}
+\begin{multline}
+p\left( \mathset{f_i} | N, R_f, R_b\right) \\ = \prod_{\mathset{i|f_i=1}}
 \left(\frac{R_f}{R_f+R_b}\right) \prod_{\mathset{i|f_i=0}}
-\left(\frac{R_b}{R_f+R_b}\right) =
+\left(\frac{R_b}{R_f+R_b}\right) \\ =
 \left(\frac{R_f}{R_f+R_b}\right)^{N_f}
 \left(\frac{R_b}{R_f+R_b}\right)^{N_b},
-\end{equation}
+\end{multline}
 where $N_f$ and $N_b$ are the numbers of foreground and background
@@
 \end{equation}
-where $\alpha$ is a normalisation constant, but of course other
+where $\alpha$ is a normalization constant, but of course other
 choices are possible.  This choice has the advantage that the prior is
@@
 Putting everything together, the posterior is
-\begin{equation}
+\begin{multline}
   \label{eq:posterior}
   p\left( \mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right) =
-  \frac{\alpha}{p(d_{\rm to},N)\,N!} \left[ \prod_{\mathset{i | f_i =
+  \frac{\alpha}{p(d_{\rm to},N)\,N!} \\ \times \left[ \prod_{\mathset{i | f_i =
         1}} R_f \hat{f}\left(x_i, \theta\right) \right] \left[
     \prod_{\mathset{i | f_i = 0}} R_b \hat{b}\left( x_i, \theta\right)
-    \right] \exp\left[ - \left(R_f + R_b\right) \right]
-  \frac{p(\theta)}{\sqrt{R_f R_b}}
-\end{equation}
+    \right] \\ \times \exp\left[ - \left(R_f + R_b\right) \right]
+  \frac{p(\theta)}{\sqrt{R_f R_b}}.
+\end{multline}
 When sampling the posterior, the first term, which is independent of
@@
 \begin{multline}
-  p\left( \mathset{y_i} | r \right) \,{\rm d}^Ny_i= P\left(
-  \textnormal{zero events below } y_1 \right) \times P\left(
+  p\left( \mathset{y_i} | r \right) \,{\rm d}^Ny_i = P\left(
+  \textnormal{zero events below } y_1 \right) \\ \times P\left(
   \textnormal{one event between } y_1 \textnormal{ and } y_1 + {\rm
     d}y_1 \right) \\ \times P\left( \textnormal{zero events between }
-  y_1+{\rm d}y_1 \textnormal{ and } y_2 \right) \ldots \\ p\left(
-  \mathset{y_i} | r \right) = \lim_{\delta y_i \to 0} \exp\left[ -
+  y_1+{\rm d}y_1 \textnormal{ and } y_2 \right) \ldots,
+\end{multline}
+so 
+\begin{multline}
+p\left(\mathset{y_i} | r \right) = \lim_{\delta y_i \to 0} \exp\left[ -
     R\left(y_1\right) \right] \left[ r\left( y_1 \right) +
-    \order{\delta y_1} \right] \times \exp\left[ - \left[ R\left( y_2
+    \order{\delta y_1} \right] \\ \times \exp\left[ - \left[ R\left( y_2
       \right) - R\left( y_1 + \delta y_1\right) \right] \right] \times
@@
 We can marginalize the posterior over the flags, $f_i$, obtaining
-\begin{equation}
+\begin{multline}
   \label{eq:rate-shape-posterior}
@@
     \mathset{0,1}^N} p\left( \mathset{f_i}, R_f, R_b, \theta | d, N
-  \right) \propto \prod_{i} \left[ R_f \hat{f}\left(x_i, \theta\right)
-    + R_b \hat{b}\left( x_i, \theta\right) \right] \exp\left[-\left(
+  \right) \\ \propto \prod_{i} \left[ R_f \hat{f}\left(x_i, \theta\right)
+    + R_b \hat{b}\left( x_i, \theta\right) \right] \\ \times \exp\left[-\left(
     R_f + R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
-\end{equation}
+\end{multline}
 This expression is useful if we are only interested in rates and not
@@
 parameters, $R_b$ or $R_f$, are of the form $R_b^{n-1/2} \exp(-R_b)$
-and so marginalisation over either $R_b$ or $R_f$ can be achieved
+and so marginalization over either $R_b$ or $R_f$ can be achieved
 analytically using
@@
 
-For the usual $k=1$ case \citep{Biswas2009}, the marginalised
+For the usual $k=1$ case \citep{Biswas2009}, the marginalized
 posterior (Eq.~\eqref{eq:rate-shape-posterior}) becomes
-\begin{equation}
+\begin{multline}
 p_{\rm LE}\left( R_f, R_b, \theta | d \right) \propto \left(R_f
-\hat{f}(x_N,\theta) + R_b \hat{b}(x_N,\theta)\right)\exp\left[-\left(
+\hat{f}(x_N,\theta) + R_b \hat{b}(x_N,\theta)\right) \\ \times \exp\left[-\left(
   R_f (1-\hat{F}(x_N,\theta)) + R_b (1-\hat{B}(x_N,\theta)) \right)
-  \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
-\end{equation}
-where $x_N$ denotes the loudness of the loudest event. Marginalising
+  \right] \\ \times \frac{p(\theta)}{\sqrt{R_f R_b}}.
+\end{multline}
+where $x_N$ denotes the loudness of the loudest event. Marginalizing
 over $R_b$ we obtain
-\begin{equation}
+\begin{multline}
 p_{\rm LE}\left( R_f, \theta | d \right) \propto \left(
 \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f
-\hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}\,
+\hat{f}(x_N,\theta)\right) \\ \times \frac{\sqrt{\pi}\,
   p(\theta)}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}}
 \exp\left(-R_f(1-\hat{F}(x_N,\theta))\right) .
-\end{equation}
+\end{multline}
+\begin{widetext}
 This posterior has a maximum in $R_f$ at
@@
     \Lambda(x, \theta) = \frac{\tilde{f}(x,\theta)}{R_b \hat{b}(x,\theta)}.
-\end{equation}}.  If $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-
+\end{equation}}.  
+\end{widetext}
+If $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-
 \hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be understood as the
@@
 significantly greater than the total number of observed events are
-strongly disfavoured.
+strongly disfavored.
 
@@
 including an upper limit on the rate, $R_{\rm max}$, in the prior for
-$R_b$. The marginalised distribution for the foreground rate then
-becomes
+$R_b$. 
+\begin{widetext}
+The marginalized distribution for the foreground rate then becomes
 \begin{eqnarray}
@@
 1/2$.
+\end{widetext}
 
@@
 set. The posterior can then be approximated by
-\begin{equation}
-p_{\rm FD}\left( R_f, R_b, \theta | d \right) \propto \prod_{i} \left[
+\begin{multline}
+p_{\rm FD}\left( R_f, R_b, \theta | d \right) \\ \propto \prod_{i} \left[
   \hat{f}\left(x_i, \theta\right)\right] R_f^N \exp\left[-\left( R_f +
   R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
-\end{equation}
-Normalisation over $R_b$ gives a constant factor and the posterior on
+\end{multline}
+Normalization over $R_b$ gives a constant factor and the posterior on
 the foreground rate becomes
-\begin{equation}
-p_{\rm FD}\left( R_f, \theta | d \right) \propto \prod_{i} \left[
+\begin{multline}
+p_{\rm FD}\left( R_f, \theta | d \right) \\ \propto \prod_{i} \left[
   \hat{f}\left(x_i, \theta\right)\right] R_f^{N-\frac{1}{2}}
 \exp\left[-R_f \right] p(\theta).
-\end{equation}
+\end{multline}
 Ignoring the dependence on $\theta$, this is peaked at a rate $R_f =
@@
 
-\section{Dependence of the Bayesian Rate Estimate on the Value of the Threshold}
-This paper is concerned with Bayesian rate estimates based on a lists
-of events (which are a certain kind of summary of the complete data
-set): either the loudest event, or a list of all events whose SNR
-exceeds a certain, pre-set threshold.  In this subsection we address
-the question of how the estimate depends on the threshold value.
+\section{Thresholding}
+This paper is concerned with Bayesian rate estimates based on lists of
+events.  Ideally, the lists should contain all events in the data set.
+However, for experimental or computational reasons one may wish to
+restrict the events to only those above some loudness threshold; in
+some cases the rate of foreground or background events, or both, is
+even expected to diverge at certain loudnesses.  In this subsection we
+address the question of how the rate estimate depends on the threshold
+value.
 
@@
 distribution $p(R)$.  Consider an ensemble of data sets whose
-distribution is consistent with that prior; i.e., such that $p(data)$
-is given by
+distribution is consistent with that prior; i.e., such that $p(d)$ is
+given by
 %Construct and ensemble of synthetic data sets consistent with $p(R)$; i.e., draw data from the %distribution
-\be p(data) = \int{ p(data|R)\,p(R)\, dR}\, .  \ee For each data set
-in the ensemble, compute the Bayesian estimator $R_{B} = \int{ R
-  p(R|data) dR}$.  Then it is immediate that \be \int{R_{B}(data)\,
-  p(data)\,d(data) }= \int{ R\, p(R) \,dR}\, .  \ee
-\noindent I.e, the data-weighted average of the Bayesian estimate
-$R_{B}$ equals the prior-weighted average $R$.
-%\curt{Guys:  would you like me to supply the argument?}
-%
-Therefore (for any fixed prior) all threshold values will yield, on
-average, the same estimate of the rate.  However this equality of
-averages does {\it not} imply that all threshold values yield the same
-information. In general, as the threshold is lowered to include more
-events, the error bar on the estimate shrink.  In this subsection we
-give quantitative illustrations of how the error bar shrinks when the
-threshold is lowered.
-
-Consider the following model problem.
-% in which the distribution of noise events is Gaussian.
-Let $p(\rho)\Delta\rho$ be the probability that data yields a putative
-event with SNR in the range $[\rho - \Delta\rho/2, \rho +
-  \Delta\rho/2]$ We write $p = p_n + p_s$, where $p_n$ is the
-distribution of noise events and $p_s$ is the distribution of events
-from actual astrophysical sources.  Both $p_n$ and $p_s$ are the
-product of a rate (of noise or true events) and the observation time.
-Here we will assume that the noise distribution is Gaussian, so that
-$p_n$ has the general form \be p_n(\rho) = \Lambda_n {\rm
-  exp}\big[-\rho^2/2\big] \, .  \ee
-\noindent
-We find it useful to define $\rho_1$ as the SNR such that a data set
-will have on average a single noise event louder than $\rho_1$; i.e.,
-such that \be \int_{\rho_1}^{\infty} P_n(\rho)\, d\rho = 1 \, .  \ee
-For this example we will assume that the spatial distribution of true
-events can be approximated as uniform in Euclidean space, so
-\be\label{eq:ps} p_s(\rho) = 3 R_t \frac{\rho_1^3}{\rho^4} \, , \ee
-\noindent where $R_t$ is the mean number of true events with $\rho >
-\rho_1$.  Then we can write the full $p(\rho)$ as
-%and that the spatial distribution of true events can be approximated as uniform in %Euclidean %space. Then $p(\rho)$ has the general form  
-\be\label{eq:ptot} p(\rho) = \big[\sqrt{\pi/2} {\rm
-    erfc}(\rho_1/\sqrt{2})\big]^{-1} \, {\rm exp}[-\rho^2/2]\ + \ 3
-R_t \frac{\rho_1^3}{\rho^4} \, .  \ee
-
-%What matters for the following analysis is the rate times the observation time.
-
-For any pair $(\rho_1, R_t)$, it is straightforward to construct
-random event lists drawn from the corresponding $p(\rho)$, and
-straightforward to apply a threshold by "throwing away" all events
-with $\rho$ less than some threshold value $\rho_{th}$.  For any
-thresholded event list, we use Eq.~(\ref{eq:rate-shape-posterior}) to construct the probability density $p(R|data)$
-(except that in this example, instead of the Jeffrey's prior, we
-assume a ``hat-shaped'' prior; see below).
-% that is uniform in some interval and zero elsewhere; see below).  
-For that event list, we define $\Delta R_B$ by \be (\Delta R_B)^2 =
-\int{(R - R_B)^2 \, p(R|data)\, dR }\, .  \ee
-%To find the average behavior, we do many draws of $R_t$ and corresponding data sets.
-%
-Figs.~\ref{fig:delRb} --\ref{fig:Rt10} illustrate how $\Delta R$
-varies with the threshold value $\rho_{th}$.  In all cases we assumed
-that $\rho_1 = 8$.
-%
-
-For Fig.~\ref{fig:delRb}, we adopted a flat prior on $R$ between $1$
-and $10$ (i.e, we took $p(R) = 1/9$ for $1 < R <10$, and zero outside
-that interval).
-% To create Fig.~\ref{fig:delR1}, 
-We drew $10^4$ random realizations $(R_t, data)$, where $R_t$ was
-drawn from that flat prior and the data--a list of $\rho$ values-- was
-drawn from the distribution $p(data|R_t)$ given by Eq.~\ref{eq:ptot} .
-For any particular threshold value, $\rho_{th}$, we created an event
-list by taking the subset of the $\rho$ values
-% in the data set that are 
-that were greater than $\rho_{th}$.  For each data set, we considered
-a sequence of increasing threshold values $\rho_{th}$, using them to
-create a corresponding sequence of (ever shorter) event lists.  For
-each event list, we calculated $\Delta R$ and $R_B - R_t$.
-%and take the mean over all realizations.  
-As a code check, we verified that the mean of $R_B - R_t$ (over all
-realizations) was consistent with zero for all $\rho_{th}$.
-Fig.~\ref{fig:delRb} displays both the mean value of $\Delta R$ and
-the mean number of noise events (as opposed to actual signals) as a
-function of $\rho_{th}$.  Note the relative insensitivity of $\Delta
-R_t$ to $\rho_{th}$: as the mean number of noise events on the event
-list shrinks from $8$ to $\sim 10^{-6}$, the average $\Delta R_B$
-increases only $\sim 14\%$, from $0.42$ to $0.48$.
+\begin{equation}
+ p(d) = \int{ p(d|R)\,p(R)\, dR}\, .  
+\end{equation}
+For each data set in the ensemble, compute the Bayesian estimator
+$R_{B} = \int{ R p(R|d) dR}$.  Then it is immediate that
+\begin{equation}
+\int{R_{B}(d)\, p(d)\,d(d) }= \int{ R\, p(R) \,dR},
+\end{equation}
+i.e.\ the data-weighted average of the Bayesian estimate $R_{B}$
+equals the prior-weighted average $R$.  Therefore all threshold values
+will yield, on average, the same estimate of the rate.  However this
+equality of averages does {\it not} imply that all threshold values
+yield the same information. In general, as the threshold is lowered to
+include more events, the error bar on the estimate shrinks.  In this
+subsection we give quantitative illustrations of how the error bar
+shrinks when the threshold is lowered.
+
+Consider the following model problem.  Let $p(x) = b(x) + f(x) = R_b
+\hat{b}(x) + R_f \hat{f}(x)$ be the rate density of events (of both
+foreground and background type) per unit loudness.  Here we will
+assume that the background is normally-distributed in loudness, so
+that $b$ has the form
+\begin{equation}
+b(x) = \Lambda_b \exp \left(- \frac{x^2}{2} \right).
+\end{equation}
+We find it useful to define $x_1$ as the loudness such that a data set
+will have on average a single noise event louder than $x_1$; i.e.,
+such that 
+\begin{equation}
+ \int_{x_1}^{\infty} b(x) dx = R_b - B\left(x_1\right) = 1.
+\end{equation}
+This condition fixes 
+\begin{equation}
+  \Lambda_b = \left[ \sqrt{\frac{\pi}{2}} \erfc\left(
+    \frac{x_1}{\sqrt{2}} \right) \right]^{-1},
+\end{equation}
+while $R_b$ will depend on the threshold, $x_\mathrm{th}$, as 
+\begin{equation}
+  R_b = \frac{\erfc\left( \frac{x_\mathrm{th}}{\sqrt{2}}
+    \right)}{\erfc\left( \frac{x_1}{\sqrt{2}} \right)}.
+\end{equation}
+
+Let the foreground distribution follow a power law in loudness (this
+is, for example, the distribution of SNR for gravitational wave events
+from uniformly-distributed sources in a single detector)
+\begin{equation}
+\label{eq:ps}
+f(x) = 3 \Lambda_f \frac{x_1^3}{x^4},
+\end{equation}
+where $\Lambda_f = R_f - F\left(x_1\right)$ is the mean number of
+foreground events with $x > x_1$.  The overall foreground rate is
+given by 
+\begin{equation}
+  \label{eq:foreground-from-lambda}
+  R_f = \Lambda_f \frac{x_1^3}{x_\mathrm{th}^3}.
+\end{equation}
+
+We can write the full $p(x)$ as
+\begin{equation}
+\label{eq:ptot} 
+p(x) = \left[\sqrt{\frac{\pi}{2}}
+  \erfc\left(\frac{x_1}{\sqrt{2}}\right)\right]^{-1}\exp\left[-\frac{x^2}{2}\right]
++ \Lambda_f \frac{x_1^3}{x^4}
+\end{equation}
+
+For any pair $(x_1, \Lambda_f)$, it is straightforward to construct
+random event lists drawn from the corresponding $p(x)$, and
+straightforward to apply a threshold by ``throwing away'' all events
+with $x$ less than the threshold value $x_\mathrm{th}$.  If $\Lambda_f
+\gg 1$, then we are in the foreground-dominated regime at $x = x_1$,
+if $\Lambda_f \ll 1$ we are in the background-dominated regime, and if
+$\Lambda_f \sim 1$ the foreground and background counts above $x_1$
+are about equal.  For any thresholded event list, we use
+Eq.~(\ref{eq:rate-shape-posterior}) to construct the probability
+density $p(R_f|d)$.  For that event list, we define the foreground
+rate uncertainty, $\Delta R_f$, by
+\begin{equation}
+\label{eq:delta-R-definition}
+(\Delta R_f)^2 \equiv \int{(R_f - R_f^\mathrm{true})^2 \, p(R_f|d)\, dR_f },
+\end{equation}
+where $R_f^\mathrm{true}$ is given by
+Eq.~\eqref{eq:foreground-from-lambda}.
 
 \begin{figure}
-\includegraphics[width=\columnwidth]{delRb}
-\caption{\label{fig:delRb} Displays the mean "error bar" $\Delta R_B$
-  as function of the threshold value $\rho_{th}$.  In this case, the
-  values of $R_t$ were randomly drawn from a flat prior on the
-  interval $[1,10]$.  We also show how the (mean) size of the event
-  list varies with $\rho_{th}$.}
+  \includegraphics[width=\columnwidth]{threshold.pdf}
+  \caption{\label{fig:threshold-rate-dependence} The mean foreground
+    rate uncertainty, Eq.~\eqref{eq:delta-R-definition}, as a function
+    of threshold for data sets with $\Lambda_f = 100$ (blue),
+    $\Lambda_f = 10$ (green), and $\Lambda_f = 1$ (red).  Recall that
+    $\Lambda_f$ is the mean number of foreground events above
+    $x_\mathrm{th} = 8$.  The total background rate, $R_b\left(
+    x_\mathrm{th} \right)$, is shown in black; we fix $R_b \left(
+    x_\mathrm{th} = x_1 = 8\right) = 1$, so on average there is one
+    background event above $x = 8$.  For $x_\mathrm{th} \gtrsim x_1$,
+    increasing the threshold tends to increase the foreground rate
+    uncertainty because the rate is foreground-dominated and fewer
+    events are included in the data set.  For $x_\mathrm{th} \lesssim
+    x_1$, the background rate dominates at small loudness, and the
+    foreground rate uncertainty asymptotes to the counting error on
+    the events that stand out from the background, $\Delta R_f / R_f
+    \simeq 1/\sqrt{\Lambda_f}$.  Note that this uncertainty applies
+    even when the total number of background events is orders of
+    magnitude larger than the number of foreground events.  When a
+    threshold must be chosen, it is safest---in the sense of producing
+    the minimal foreground rate uncertainty---to choose the threshold
+    \emph{well into the background-dominated loudness regime}; the
+    extra background events in the data set do not affect the estimate
+    of the foreground rate, and, when the background distribution is
+    parameterized, can help to better determine these parameters (see
+    \S~\ref{sec:gw-overlapping-template}). }
 \end{figure}
 
-Figs.~\ref{fig:Rt1} --\ref{fig:Rt10} were constructed similarly, but
-with the following differences.  First, we took the prior $p(R)$ to be
-$p(R) \propto 1/R$ for $0.03 < R < 30$, and zero outside that interval
-(i.e., $p(R)$ is the Jeffries prior, but restricted to some range).
-Second, we regarded $R_t$ as fixed (as of course it would be in
-Nature), and only varied the data--randomly drawing $10^4$
-realizations from $p(data|R_t)$.  For each realization we calculated
-$R_t^{-1}\,\Delta R_B$ and $(R_B - R_t)$, and then we calculated the
-means of these quantities over all realizations.  The mean of $(R_B -
-R_t)$ was generally {\it not} zero, since the prior biases the
-results.  (Of course, if one {\it updated} the prior after each
-realization, the bias would decrease towards zero, but that is not
-what we did: we evaluated each realization with the same Jeffries
-prior.)
-
-For Figs.~\ref{fig:Rt1} --\ref{fig:Rt10}, the mean number of noise
-events as a function of $\rho_{th}$ was the same as for the case shown
-in Fig.~\ref{fig:delRb}.  Note that for all three values of $R_t$, the
-mean uncertainty $\Delta R_B$ increases by only $\sim 20\%$ as the
-mean number of noise events on the event list shrinks from $8$ to
-$\sim 10^{-6}$.  Thus the inference that the average $\Delta R_B$ is
-rather insensitive to $\rho_{th}$ appears to be robust.  Finally, we
-note that for all three cases the bias is much smaller than $\Delta
-R_B$.
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{Rt1}
-\caption{\label{fig:Rt1} Displays the mean rms error $\Delta R_B$ and
-  mean $(R_B - R_t)$ (i.e., the bias) as functions of the threshold
-  value $\rho_{th}$.  In this case, the actual value of $R_t$ was
-  $1.0$, and $p(R|data)$ was computed using the Jeffries prior on the
-  interval $[0.03,30]$.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{Rt3}
-\caption{\label{fig:Rt3} Same as the previous figure, but displaying
-  the size of $\Delta R_B$ relative to $R_t$, with the actual $R_t$
-  equal to $3.0$.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{Rt10}
-\caption{\label{fig:Rt10} Same as the previous figure, but with the
-  actual $R_t$ equal to $10.0$.}
-\end{figure}
+Figure \ref{fig:threshold-rate-dependence} illustrates how the mean
+fractional foreground uncertainty, $\left\langle \Delta R_f
+\right\rangle / R_f$, varies with the threshold value $x_\mathrm{th}$
+for the foreground-dominated and comparable-rate regime.  In all cases
+we assumed that $x_1 = 8$.  For large thresholds, where $R_b \ll 1$,
+increasing the threshold tends to increase the fractional uncertainty
+on the foreground rate, since fewer foreground events are included in
+the sample.  However, as the threshold passes into the
+background-dominated regime, the uncertainty in the foreground rate
+asymptotes to
+\begin{equation}
+  \frac{\Delta R_f}{R_f} \simeq \frac{1}{\sqrt{\Lambda_f}},
+\end{equation}
+which is the usual Poisson counting uncertainty on the events that
+stand out from the background (those with $x \gtrsim x_1$).  Note that
+this uncertainty applies even when the total number of background
+events is orders of magnitude larger than the number of foreground
+events.  When a threshold must be chosen, it is safest---in the sense
+of producing the minimal foreground rate uncertainty---to choose the
+threshold \emph{well into the background-dominated loudness regime};
+the extra background events in the data set do not affect the estimate
+of the foreground rate, and, when the background distribution is
+parameterized, can help to better determine these parameters (see
+\S~\ref{sec:gw-overlapping-template}).
+
+Though we have only illustrated the behavior of the rate estimate
+quantitatively for this specific example of foreground and background
+rates, the conclusions hold in general.  Consider the Fisher
+information matrix for the posterior distribution in
+Eq.~\eqref{eq:rate-shape-posterior}.  For a model with parameters
+$\mathset{\theta_i}$, the Fisher information matrix has components
+\begin{equation}
+  F_{ij} \equiv \left\langle \frac{\partial \log p\left( \theta | d
+    \right)}{\partial \theta_i} \frac{\partial \log p\left( \theta | d
+    \right)}{\partial \theta_j} \right\rangle,
+\end{equation}
+where the average is taken over the data distribution at fixed
+$\theta$, $p(d | \theta)$.  The components of the Fisher information
+matrix describe the maximum amount of information about the
+corresponding parameters available in a given data set; the inverse of
+the Fisher information matrix gives the Cramer-Rao bound on the
+covariance matrix of unbiased estimators of $\theta$.  For the
+likelihood that enters Eq.~\eqref{eq:rate-shape-posterior}, the Fisher
+information matrix is 
+\begin{multline}
+  \mathbf{F} = \left( R_f + R_b \right) \\ \times \begin{pmatrix}
+    \left \langle \left(\frac{\hat{f}}{R_f \hat{f} + R_b \hat{b}}\right)^2 \right
+    \rangle & \left \langle \frac{\hat{f} \hat{b}}{\left( R_f \hat{f}
+      + R_b \hat{b} \right)^2} \right\rangle \\
+    \left \langle \frac{\hat{f} \hat{b}}{\left( R_f \hat{f}
+      + R_b \hat{b} \right)^2} \right\rangle & \left \langle
+    \left(\frac{\hat{b}}{R_f \hat{f} + R_b \hat{b}}\right)^2 \right 
+    \rangle
+  \end{pmatrix},
+\end{multline}
+where the expectation values are taken over the distributions
+$\hat{f}$ and $\hat{b}$ (i.e.\ they are expectations for one event
+from the combined rate distribution).  If the cross-terms are small,
+then the Cramer-Rao bound on the uncertainty of $R_f$ will be given by
+\begin{equation}
+  \sigma_{R_f} \simeq \frac{1}{\sqrt{R_f + R_b}} \left[ \left\langle \left(
+    \frac{\hat{f}}{R_f \hat{f} + R_b \hat{b}} \right)^2 \right\rangle\right]^{-1/2}
+\end{equation}
+Extending a threshold into regions where the factor 
+\begin{equation}
+  \left( \frac{\hat{f}}{R_f \hat{f} + R_b \hat{b}} \right)^2
+\end{equation}
+becomes small---that is, into background-dominated
+regions---contributes little to reducing the overall uncertainty in the
+foreground rate.  This is consistent with the behavior of the specific
+example in Figure \ref{fig:threshold-rate-dependence}.
 
@@
 Here we discuss a very unattractive feature of the Bayesian loudest
-event estimate of $R$: a small percentage of the time it will yield a
-very large over-estimate.
+event estimate of $R$ \citep{Biswas2009}: a small percentage of the
+time it will yield a very large over-estimate.
 
@@
 previous subsection, and we will begin with a very specific example.
-Let the true rate (times the observation time) be $R_t = 1$, meaning
-(again) that the expected number of actual events with $\rho > \rho_1$
-is one.  Then there is a $1/64$ chance ($\approx 1.6\%$) that the
-loudest event will have $\rho_{LE} > 4 \rho_1$.  Consider this case,
-and let us also assume that there are no events (noise or actual) with
-$\rho_1 < \rho < \rho_{LE}$.  The loudest event estimate basically
-"throws away" the information that there are no events in this
-interval, which (we show below) leads an estimate $R_{LE} \approx 64$,
-which is $64$ times the true rate!  Now, if the true rate really were
-$R_t = 64$, then the expected number of events with $\rho > \rho_1$
-would be $64$.  So in this case, the loudest event estimate ignores
-the fact that there are $\sim 56 -72$ "missing" events.  However a
-Bayesian estimate with $\rho_{th}$ set to $\rho_1$ incorporates this
-information quite naturally, and so (correctly) yields an estimated
-$R$ of order one.
-
-We now essentially repeat the above discussion, but in greater
-generality and with the derivations provided.  First recall that $R_B$
-is given by \be\label{eq:rb1} R_B = \frac{ \int{ R\, p(\{\rho_i\}|R)
-    \, p(R)\, dR}}{ \int{ p(\{\rho_i\}|R)\, p(R)\, dR }}\, , \ee
-\noindent
-where $p(R)$ is the prior on $R$ and (and up to an overall
-normalization) \be\label{eq:rb2} p(\{\rho_i\}|R) = {\rm
-  exp}\bigg[\sum_{i \in list} {\rm ln}\, p(\rho_i|R) -
-  \int_{\rho_{th}}^{\infty} p(\rho)\, d\,\rho \bigg] \, , \ee
-\noindent
-with $p(\rho)$ given by Eq.~(\ref{eq:ptot}).
-
-The loudest event estimate is obtained by setting the lower
-integration limit in the last term on the rhs of Eq.~(\ref{eq:rb2}) to
-$\rho_{th} = \rho_{LE}$, and replacing the sum $\sum_i {\rm ln}\,
-p(\rho_i|R)$ by the single term ${\rm ln}\, p(\rho_{LE}|R)$.  For the
-case of interest here, where $\rho_{LE}$ is (at least) a few times
-larger than $\rho_1$ and $R_t$ is of order $1$, we can neglect the
-noise contribution $p_n(\rho)$. Then the integrals in
-Eq.~(\ref{eq:rb1}) become especially simple, and yield the estimate:
-\be\label{eq:rb3} R_B \approx
-\begin{cases}
-(\rho_{LE}/\rho_1)^3 & \text{if $p(R) \propto 1/R$} \\ 2\,
-  (\rho_{LE}/\rho_1)^3 & \text{if $p(R) =$ const \, .}
-\end{cases}
-\ee
-
-If instead we set the lower integration limit to $\rho_{th} = \rho_1$,
-then the $p_n(\rho)$ contribution to $p(\rho)$ can no longer be
-neglected.  But the integrals are still straightforward, and we find:
-\be\label{eq:rb3} R_B \approx
-\begin{cases}
-1 \ \ \ & \text{if $p(R) \propto 1/R$} \\ 2 \ \ \ & \text{if $p(R) =$
-  const \, .}
-\end{cases}
-\ee
-\noindent Thus (as emphasized above) simply incorporating the
-information that there are no events {\it between} $\rho_1$ and
-$\rho_{LE}$ results in a far beter estimate.
+Let $\Lambda_f = 1$, meaning that the expected number of actual events
+with $x > x_1$ is one.  Then there is a $1/64$ chance ($\approx
+1.6\%$) that the loudest event will have $x_{LE} > 4 x_1$.  Consider
+this case, and let us also assume that there are no events (noise or
+actual) with $x_1 < x < x_{LE}$.  The loudest event estimate basically
+``throws away'' the information that there are no events in this
+interval, which (we show below) leads an estimate
+$\Lambda_f^\mathrm{LE} \approx 64$, which is $64$ times the true rate!
+Now, if the true rate really were $\Lambda_f = 64$, then the expected
+number of events with $x > x_1$ would be $64$.  So in this case, the
+loudest event estimate ignores the fact that there are $\sim 56 -72$
+``missing'' events.  However a Bayesian estimate with $x_\mathrm{th}$
+set to $x_1$ incorporates this information quite naturally, and so
+(correctly) yields an estimated $\Lambda_f$ of order one.
 
@@
 In the following subsection, we will generalize the model to
-overlapping templates.  We rank candidate events by their maxmim SNR
+overlapping templates.  We rank candidate events by their maximum SNR
 over the entire template bank,
@@
 approximations described earlier, the loudest-event statistic and the
-foreground-dominated statistic. The marginalised distribution for the
+foreground-dominated statistic. The marginalized distribution for the
 foreground rate using these alternatives are shown in
@@
 distribution that is essentially equal to the full analysis, for the
-``omniscent'' choice of threshold value that excludes all background
+``omniscient'' choice of threshold value that excludes all background
 data. For lower thresholds, even for a threshold where
@@
 background distribution, to which the foreground-dominated statistic
-is insensitive, and in the loudest event case, after marginalisation
+is insensitive, and in the loudest event case, after marginalization
 over $N_{\rm eff}$ we find $\int_0^{N_{\rm max}} \hat{b}(x_N,N_{\rm
@@
 \label{sec:uncertain-dist}
-The framework outlined above relies on the existence of models for the foreground, $\hat{f}(x,\theta)$, and background, $\hat{b}(x,\theta)$, distributions parameterised by a small number of model parameters, $\theta$.
+The framework outlined above relies on the existence of models for the foreground, $\hat{f}(x,\theta)$, and background, $\hat{b}(x,\theta)$, distributions parameterized by a small number of model parameters, $\theta$.
 While in many situations simple analytic functions such as power laws will provide an adequate description, this will not always be the case.
 In the absence of a good analytic model, the space of the ranking statistic, $x$ could be divided into bins and $\hat{f}(x)$/$\hat{b}(x)$ taken to be flat in each of these bins.
-The number of free parameters characterising each of $\hat{f}$ and $\hat{b}$ is then the number of bins used.
+The number of free parameters characterizing each of $\hat{f}$ and $\hat{b}$ is then the number of bins used.
 While such a framework is model free, the increase in model parameters will mean that more observed events will typically be required to achieve the same precision on the rates and foreground/background distributions.
@@
 The former assumption is probably reasonable, modulo non-stationarities in the detector data, but the latter relies on knowledge of the astrophysical rate of events which is more uncertain.
-These astrophysical uncertainties could be handled with a hybrid approach, in which injections are used to characterise the statistic distribution for sources of a particular type, while additional rate or shape parameters are introduced to characterise the variation in the astrophysical rate of mergers as a function of source type.
+These astrophysical uncertainties could be handled with a hybrid approach, in which injections are used to characterize the statistic distribution for sources of a particular type, while additional rate or shape parameters are introduced to characterize the variation in the astrophysical rate of mergers as a function of source type.
 
@@
     Eq.~\eqref{eq:true-cluster-parameters}.  The points are the
-    realization of this density profile used as sythetic data in
+    realization of this density profile used as synthetic data in
     \S~\ref{sec:star-cluster}; the dashed line encloses one Plummer
@@
 Eq.~\eqref{eq:rate-shape-posterior}.  We did this to take advantage of
-the \textit{emcee} sampler of \citet{ForemanMackey2012}, which
+the \texttt{emcee} sampler of \citet{ForemanMackey2012}, which
 requires all parameters to be in $\mathbb{R}$.  We applied a prior on
@@
 \begin{acknowledgments}
-  We thank Kipp Cannon, Chad Hanna, Drew Keppel, and Richard
-  O'Shaughnessy for discussions and suggestions about this manuscript.
-  IM and WMF thank KITP.
+  We thank Kipp Cannon, Chad Hanna, Drew Keppel, Richard
+  O'Shaughnessy, and Ted von Hippel for discussions and suggestions
+  about this manuscript.  IM and WMF thank KITP.
 \end{acknowledgments}

### 2012-12-16 many.tex
curt made tiny change in title and added acknowledgements

@@
 
-\title{Counting With Confusion: Bayesian Rate Estimation With Multiple
+\title{Counting And Confusion: Bayesian Rate Estimation With Multiple
   Populations}
@@
   O'Shaughnessy, and Ted von Hippel for discussions and suggestions
-  about this manuscript.  IM and WMF thank KITP.
+  about this manuscript.
+IM and WMF thank KITP.
+CC's work was carried out  at the Jet Propulsion Laboratory, California Institute of Technology, under contract to the National Aeronautics and Space Administration.
+CC also gratefully acknowledges support from NSF grant PHY1068881.
+Copyright 2012.
+All rights reserved.
 \end{acknowledgments}

### 2013-01-03 many.tex
Edited through III.A.

@@
-\documentclass[aps,prd,reprint]{revtex4-1}
+\documentclass[aps,prd,reprint,nofootinbib]{revtex4-1}
 
@@
 the second observed event is a fore/background event (if $f_2=1,0$)
-with ranking statistic $x_2$ etc. If the events are ordered by ranking
+with ranking statistic $x_2$, etc. If the events are ordered by ranking
 statistic the corresponding expression is more complicated, since
@@
 
-The second term in Eq.~\eqref{eq:combined-flag-rate-prior} is a
+The last term in Eq.~\eqref{eq:combined-flag-rate-prior} is a
 traditional prior.  Because the rate parameters enter the posterior in
@@
 \end{equation}
-where $\alpha$ is a normalization constant, but of course other
+where $\alpha$ is a normalization constant; but of course other
 choices are possible.  This choice has the advantage that the prior is
@@
 N!\,p(d_{\rm to},N)$, since there are $N!$ ways in which $N$ events
-can be ordered in time and would have the same set of ranking
-statistics.
+with a given set of ranking statistics can be ordered in time.
 
@@
 where $\tilde{b}(x_N,\theta) = \hat{b}(x_N,\theta) /
-(1-\hat{B}(x_N,\theta))$\footnote{The quantity $\Lambda$ from
+(1-\hat{B}(x_N,\theta))$ and similarly for $\tilde{f}(x_N,\theta)$.\footnote{The quantity $\Lambda$ from
   \citet{Biswas2009} is related to $\hat{b}$ and $\tilde{f}$ by
@@
     \Lambda(x, \theta) = \frac{\tilde{f}(x,\theta)}{R_b \hat{b}(x,\theta)}.
-\end{equation}}.  
+\end{equation}}  
 \end{widetext}
@@
 \end{equation}
-and if $\hat{f}(x_N,\theta) \gg R_{\rm max} \hat{b}(x_N.\theta)$ we
+and if $\hat{f}(x_N,\theta) \gg R_{\rm max} \hat{b}(x_N,\theta)$ we
 find the same result as before, $(1- \hat{F}(x_N,\theta))\,R_f \approx

### 2013-01-03 many.tex
Changes in response to Jon's comments.

@@
   or approximated; our method works well even if the foreground and
-  background event distributions overlap significantly and the status
+  background event distributions overlap significantly and the nature
   of any individual event cannot be determined with any certainty.  We
@@
   \erfc\left(\frac{x_1}{\sqrt{2}}\right)\right]^{-1}\exp\left[-\frac{x^2}{2}\right]
-+ \Lambda_f \frac{x_1^3}{x^4}
++ 3 \Lambda_f \frac{x_1^3}{x^4}
 \end{equation}
@@
 While in many situations simple analytic functions such as power laws will provide an adequate description, this will not always be the case.
-In the absence of a good analytic model, the space of the ranking statistic, $x$ could be divided into bins and $\hat{f}(x)$/$\hat{b}(x)$ taken to be flat in each of these bins.
+In the absence of a good analytic model, the space of the ranking statistic $x$ could be divided into bins and $\hat{f}(x)$ and $\hat{b}(x)$ are taken to be flat in each of these bins.
 The number of free parameters characterizing each of $\hat{f}$ and $\hat{b}$ is then the number of bins used.
@@
 We note that this analysis makes the assumption that the background distribution is the same in the time slide and real data and that the foreground distribution is the same between the injection and real data.
-The former assumption is probably reasonable, modulo non-stationarities in the detector data, but the latter relies on knowledge of the astrophysical rate of events which is more uncertain.
+The former assumption is probably reasonable, modulo non-stationarities in the detector data, but the latter relies on knowledge of the relative the astrophysical rates of different events, which is more uncertain.
 These astrophysical uncertainties could be handled with a hybrid approach, in which injections are used to characterize the statistic distribution for sources of a particular type, while additional rate or shape parameters are introduced to characterize the variation in the astrophysical rate of mergers as a function of source type.
@@
   about this manuscript.
-IM and WMF thank KITP.
+IM and WMF acknowledge the hospitality of KITP, supported in part by the National Science Foundation under No.
+NSF Grant PHY11-25915.
 CC's work was carried out  at the Jet Propulsion Laboratory, California Institute of Technology, under contract to the National Aeronautics and Space Administration.
 CC also gratefully acknowledges support from NSF grant PHY1068881.
+JG's work is supported by the Royal Society.
 Copyright 2012.

### 2013-01-07 many.tex
Added more detailed description of connection to Biswas

@@
 posterior (Eq.~\eqref{eq:rate-shape-posterior}) becomes
-\begin{multline}
+\begin{multline}\label{LEeq1}
 p_{\rm LE}\left( R_f, R_b, \theta | d \right) \propto \left(R_f
@@
 \end{multline}
-where $x_N$ denotes the loudness of the loudest event. Marginalizing
-over $R_b$ we obtain
+where $x_N$ denotes the loudness of the loudest event.
+In the loudest event statistic paper~\citep{Biswas2009}, the authors assume the background distribution and rate are known, which corresponds to using a delta function prior on $R_b$.
+They further assume a flat prior on $R_f$ and that the foreground and background distributions do not depend on any unknown free parameters.
+With these assumptions, the posterior on $R_f$, Eq.~\eqref{LEeq1}, is modified to
+\begin{multline}
+p_{\rm LE}(R_f|d) \propto  \left(R_f
+\hat{f}(x_N) + R_b \hat{b}(x_N)\right) \\ \times \exp\left[-\left(
+  R_f (1-\hat{F}(x_N)) + R_b (1-\hat{B}(x_N)) \right)
+  \right] .
+\end{multline}
+Integrating over $R_f$ gives
+\begin{multline}
+\int_0^\infty p_{\rm LE}(R_f|d) {\rm d}R_f =  \frac{R_b \hat{b}(x_N)}{(1-\hat{F}(x_N))} {\rm e}^{-(1-\hat{B}(x_N)) R_b}\\
+%\exp\left[-(1-\hat{B}(x_N))R_b\right] \\ 
+\times \left(\frac{\hat{f}(x_N)}{(1-\hat{F}(x_N)) R_b \hat{b}(x_N)} + 1 \right)
+\end{multline}
+and so the normalised posterior is
+\begin{multline}
+p_{\rm LE}(R_f|d) = \frac{(1-\hat{F}(x_N))}{1+\Lambda} \left( 1+ R_f (1-\hat{F}(x_N))\Lambda \right) \\
+\times \exp\left[-R_f (1-\hat{F}(x_N))\right]
+\end{multline}
+in which we have defined
+\begin{equation}
+\Lambda \equiv \frac{\hat{f}(x_N)}{(1-\hat{F}(x_N)) R_b \hat{b}(x_N)}.
+\end{equation}
+With the further identification $\mu \equiv R_f$ and $\hat{\epsilon} \equiv 1 - \hat{F}(x_N)$, this is Eq.~(18) of~\citep{Biswas2009} and we have shown how their parameter $\Lambda$ is related to the foreground and background distributions used here.
+
+Returning now to Eq.~\eqref{LEeq1} and marginalizing over $R_b$, we obtain
 \begin{multline}
@@
 \end{multline}
-\begin{widetext}
+%\begin{widetext}
+%This posterior has a maximum in $R_f$ at
+%\begin{equation}
+%R_f =
+%\frac{\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta))\tilde{b}(x_N,\theta)  +
+%  \sqrt{\left(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta))\tilde{b}(x_N,\theta)\right)^2  -
+%    4\tilde{b}(x_N,\theta)(1-\hat{F}(x_N,\theta))\hat{f}(x_N,\theta)}}{4
+%  \hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta))}.
+%\end{equation}
+%where $\tilde{b}(x_N,\theta) = \hat{b}(x_N,\theta) /(1-\hat{B}(x_N,\theta))$ and similarly for $\tilde{f}(x_N,\theta)$.
+%\footnote{The quantity $\Lambda$ from  \citet{Biswas2009} is related to $\hat{b}$ and $\tilde{f}$ by  \begin{equation}    \Lambda(x, \theta) = \frac{\tilde{f}(x,\theta)}{R_b \hat{b}(x,\theta)}.\end{equation}}
+%\end{widetext}
 This posterior has a maximum in $R_f$ at
-\begin{equation}
+\begin{multline}
 R_f =
 \frac{\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta))\tilde{b}(x_N,\theta)
-  +
-  \sqrt{\left(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta))\tilde{b}(x_N,\theta)\right)^2
-    -
-    4\tilde{b}(x_N,\theta)(1-\hat{F}(x_N,\theta))\hat{f}(x_N,\theta)}}{4
-  \hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta))}.
-\end{equation}
-where $\tilde{b}(x_N,\theta) = \hat{b}(x_N,\theta) /
-(1-\hat{B}(x_N,\theta))$ and similarly for $\tilde{f}(x_N,\theta)$.\footnote{The quantity $\Lambda$ from
-  \citet{Biswas2009} is related to $\hat{b}$ and $\tilde{f}$ by
-  \begin{equation}
-    \Lambda(x, \theta) = \frac{\tilde{f}(x,\theta)}{R_b \hat{b}(x,\theta)}.
-\end{equation}}  
-\end{widetext}
+  + \sqrt{g(x_N,\theta)}}{4\hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta))} \\
+\mbox{where }  g(x_N,\theta) =\left(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta))\tilde{b}(x_N,\theta)\right)^2
+    \\ - 4\tilde{b}(x_N,\theta)(1-\hat{F}(x_N,\theta))\hat{f}(x_N,\theta)
+\end{multline}
+and $\tilde{b}(x_N,\theta) = \hat{b}(x_N,\theta) /(1-\hat{B}(x_N,\theta))$ and similarly for $\tilde{f}(x_N,\theta)$.
+
 If $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-

### 2013-01-15 many.tex
Edits to V and VI.

@@
 We assume that the signals are sufficiently rare that they do not
-overlap in the data stream.  The signal to noise ratio (SNR) of a
+overlap in the data stream.  The signal-to-noise ratio (SNR) of a
 template, $h(f)$, given data, $d(f)$, is
@@
 
-The SNR of a gravitational wave signal in an interferometric detector
+The SNR of a gravitational-wave signal in an interferometric detector
 scales as $1/d$ \citep{Finn1992}, where $d$ is the distance to the
@@
 In Figure \ref{fig:analytic-rate-recovery}, we show the marginalized
-posterior densities for the foreground and background rates.  (Refer
-to Eq.~\eqref{eq:rate-shape-posterior}.)  Figure
+posterior densities for the foreground and background rates (see
+Eq.~\eqref{eq:rate-shape-posterior}).  Figure
 \ref{fig:analytic-rate-foreground-probs} shows the posterior
@@
 determined because we can examine the synthetic foreground and
-background data samples.  We show results for the first two choices of
+background data samples. The threshold may seem obvious from a 
+visual examination of Figure \ref{fig:analytic-rate-foreground-probs}; 
+however, the construction of this figure relies on the application of the full
+framework in the first place.  We show results for the first two choices of
 $x_{\rm min}$ in Figure~\ref{fig:altmeth}; the omniscient choice
@@
 
-In the context of gravitational wave experiments, additional information on the ranking statistic distributions for the foreground and background can be obtained using injections into and timeslides of the data respectively.
+In the context of gravitational wave experiments, additional information on the ranking statistic distributions for the foreground can be obtained using mock signal injections into the data, while distributions for the background can be estimated by analyzing time slides of data sets from different detectors relative to each other \cite[e.g.,][]{ihope}.
 This information can be readily incorporated in the current framework by assuming there is another set of $N_I$ events with ranking statistics $\{ w_i \}$, known to be drawn from the foreground distribution ($f_i=1$) and a set of $N_T$ events with ranking statistics $\{z_i\}$ known to be drawn from the background distribution ($f_i=0$).
@@
 We note that this analysis makes the assumption that the background distribution is the same in the time slide and real data and that the foreground distribution is the same between the injection and real data.
-The former assumption is probably reasonable, modulo non-stationarities in the detector data, but the latter relies on knowledge of the relative the astrophysical rates of different events, which is more uncertain.
+The former assumption is probably reasonable, modulo correlations of non-gravitational-wave origin between data in different detectors, but the latter relies on knowledge of the relative the astrophysical rates of different events, which is more uncertain.
 These astrophysical uncertainties could be handled with a hybrid approach, in which injections are used to characterize the statistic distribution for sources of a particular type, while additional rate or shape parameters are introduced to characterize the variation in the astrophysical rate of mergers as a function of source type.
@@
 stars.  We showed that this framework is generally superior to both
-the loudest-event statistic and and the foreground-dominated
+the loudest-event statistic and the foreground-dominated
 statistic.
 
-Throughout this paper, we have assumed that the shape of the
+Through most of this paper, we have assumed that the shape of the
 foreground and background distributions is known, or at least can be
@@
 estimate the background at the very tails of the distribution), and
-will require significant future work.
+will require significant future work.  In Section \ref{sec:uncertain-dist}, 
+we discussed some of the possible
+approaches when the shapes of the background and foreground distributions
+cannot be confidently described by models with a few adjustable parameters.
 
@@
 JG's work is supported by the Royal Society.
-Copyright 2012.
+Copyright 2013.
 All rights reserved.

### 2013-01-15 many.tex
Threshold figure converted to B&W, changed x-axis label for pfore.pdf

@@
     rate uncertainty, Eq.~\eqref{eq:delta-R-definition}, as a function
-    of threshold for data sets with $\Lambda_f = 100$ (blue),
-    $\Lambda_f = 10$ (green), and $\Lambda_f = 1$ (red).  Recall that
-    $\Lambda_f$ is the mean number of foreground events above
-    $x_\mathrm{th} = 8$.  The total background rate, $R_b\left(
-    x_\mathrm{th} \right)$, is shown in black; we fix $R_b \left(
-    x_\mathrm{th} = x_1 = 8\right) = 1$, so on average there is one
-    background event above $x = 8$.  For $x_\mathrm{th} \gtrsim x_1$,
-    increasing the threshold tends to increase the foreground rate
-    uncertainty because the rate is foreground-dominated and fewer
-    events are included in the data set.  For $x_\mathrm{th} \lesssim
-    x_1$, the background rate dominates at small loudness, and the
-    foreground rate uncertainty asymptotes to the counting error on
-    the events that stand out from the background, $\Delta R_f / R_f
-    \simeq 1/\sqrt{\Lambda_f}$.  Note that this uncertainty applies
-    even when the total number of background events is orders of
-    magnitude larger than the number of foreground events.  When a
-    threshold must be chosen, it is safest---in the sense of producing
-    the minimal foreground rate uncertainty---to choose the threshold
-    \emph{well into the background-dominated loudness regime}; the
-    extra background events in the data set do not affect the estimate
-    of the foreground rate, and, when the background distribution is
+    of threshold for data sets with $\Lambda_f = 100$ (solid line),
+    $\Lambda_f = 10$ (dashed line), and $\Lambda_f = 1$ (dash-dotted
+    line).  Recall that $\Lambda_f$ is the mean number of foreground
+    events above $x_\mathrm{th} = 8$.  The total background rate,
+    $R_b\left( x_\mathrm{th} \right)$, is shown by the dotted line; we
+    fix $R_b \left( x_\mathrm{th} = x_1 = 8\right) = 1$, so on average
+    there is one background event above $x = 8$.  For $x_\mathrm{th}
+    \gtrsim x_1$, increasing the threshold tends to increase the
+    foreground rate uncertainty because the rate is
+    foreground-dominated and fewer events are included in the data
+    set.  For $x_\mathrm{th} \lesssim x_1$, the background rate
+    dominates at small loudness, and the foreground rate uncertainty
+    asymptotes to the counting error on the events that stand out from
+    the background, $\Delta R_f / R_f \simeq 1/\sqrt{\Lambda_f}$.
+    Note that this uncertainty applies even when the total number of
+    background events is orders of magnitude larger than the number of
+    foreground events.  When a threshold must be chosen, it is
+    safest---in the sense of producing the minimal foreground rate
+    uncertainty---to choose the threshold \emph{well into the
+      background-dominated loudness regime}; the extra background
+    events in the data set do not affect the estimate of the
+    foreground rate, and, when the background distribution is
     parameterized, can help to better determine these parameters (see

### 2013-01-15 many.tex
Added sentence in III.A explaining Rf vs Rf(1-Fhat).

@@
 \end{multline}
-where $x_N$ denotes the loudness of the loudest event.
-In the loudest event statistic paper~\citep{Biswas2009}, the authors assume the background distribution and rate are known, which corresponds to using a delta function prior on $R_b$.
-They further assume a flat prior on $R_f$ and that the foreground and background distributions do not depend on any unknown free parameters.
-With these assumptions, the posterior on $R_f$, Eq.~\eqref{LEeq1}, is modified to
+where $x_N$ denotes the loudness of the loudest event, and $R_f$ and
+$R_b$ are the number of events expected \emph{above our original
+  threshold} (so, for example, $R_f (1 - \hat{F}(x_N,\theta))$ is the
+number of foreground events expected above loudness $x_N$).  In the
+loudest event statistic paper~\citep{Biswas2009}, the authors assume
+the background distribution and rate are known, which corresponds to
+using a delta function prior on $R_b$. They further assume a flat
+prior on $R_f$ and that the foreground and background distributions do
+not depend on any unknown free parameters. With these assumptions, the
+posterior on $R_f$, Eq.~\eqref{LEeq1}, is modified to
 \begin{multline}

### 2013-01-16 many.tex
Edited III.B and IV.A for clarity.

@@
 \times \exp\left[-R_f (1-\hat{F}(x_N))\right]
+\label{loudest-eq}
 \end{multline}
@@
 \hat{b}(x_i,\theta)$ for all ranking statistics $x_i$ in the data
-set. The posterior can then be approximated by
+set.  If we can further be confident that $R_f \hat{f}(x_i,\theta) \gg R_b \hat{b}(x_i,\theta)$ for all events, then the posterior can be approximated by
 \begin{multline}
@@
 \end{multline}
+Note that these are posteriors on the number of events expected \emph{above the
+threshold $x_{\rm min}$}.
+The threshold choice for the foreground-dominated statistic could be different from the threshold choice applied elsewhere.
+If rates $R_{f,1}$ and $R_{f,2}$ are estimated for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$ respectively, these rates should be compared by equating $R_{f,1} (1-\hat{F}(x_2,\theta))$ and $R_{f,2} (1-\hat{F}(x_1,\theta))$.
+
 Normalization over $R_b$ gives a constant factor and the posterior on
@@
 events observed (the $1/2$ comes from our use of the Jeffreys prior on
-the rate).  We note that the rate $R_f$ is defined as the rate of
-events occurring above the specified threshold and therefore if the
-foreground dominated statistic were used to estimate rates $R_{f,1}$
-and $R_{f,2}$ for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$
-respectively, these rates should be compared by equating $R_{f,1}
-(1-\hat{F}(x_2,\theta))$ and $R_{f,2} (1-\hat{F}(x_1,\theta))$.
+the rate).
 
@@
 Let $\Lambda_f = 1$, meaning that the expected number of actual events
-with $x > x_1$ is one.  Then there is a $1/64$ chance ($\approx
+with $x > x_1$ is one.  Then there is a $1/64$ chance ($1-\hat{F}(x_{LE}) \approx
 1.6\%$) that the loudest event will have $x_{LE} > 4 x_1$.  Consider
 this case, and let us also assume that there are no events (noise or
-actual) with $x_1 < x < x_{LE}$.  The loudest event estimate basically
-``throws away'' the information that there are no events in this
-interval, which (we show below) leads an estimate
-$\Lambda_f^\mathrm{LE} \approx 64$, which is $64$ times the true rate!
+actual) with $x_1 < x < x_{LE}$.
+
+The loudest event estimate basically ``throws away'' the information that there are no events in this interval.
+The maximum of the loudest-event-statistic posterior on $R_f$,  Eq.~(\ref{loudest-eq}), is at $R_f = \frac{\Lambda - 1}{ \Lambda (1-\hat{F}(x_{LE}))}$.
+If the value of $\Lambda$ is sufficiently high at $x_1$ (and $\Lambda$ will be even greater at $x_{LE}$), then, for this data set, we would estimate $R_f \approx \frac{1}{1-\hat{F}(x_{LE})} \gtrsim 64$.
+Thus, for our assumed shape of the foreground distribution, we will estimate the rate of events above $x_1$ to be $64$ times the true rate!
+
 Now, if the true rate really were $\Lambda_f = 64$, then the expected

### 2013-01-16 many.tex
Minor edits to IV.

@@
 For each data set in the ensemble, compute the Bayesian estimator 
-$R_{B} = \int{ R p(R|d) dR}$.  Then it is immediate that
+for the mean of the posterior $R_{B} = \int{ R p(R|d) dR}$.  Then it is immediate that
 \begin{equation}
@@
 \begin{equation}
-b(x) = \Lambda_b \exp \left(- \frac{x^2}{2} \right).
+b(x) = \Gamma_b \exp \left(- \frac{x^2}{2} \right).
 \end{equation}
@@
 \begin{equation}
-  \Lambda_b = \left[ \sqrt{\frac{\pi}{2}} \erfc\left(
+  \Gamma_b = \left[ \sqrt{\frac{\pi}{2}} \erfc\left(
     \frac{x_1}{\sqrt{2}} \right) \right]^{-1},
@@
 \label{eq:ps}
-f(x) = 3 \Lambda_f \frac{x_1^3}{x^4},
-\end{equation}
-where $\Lambda_f = R_f - F\left(x_1\right)$ is the mean number of
+f(x) = 3 \Gamma_f \frac{x_1^3}{x^4},
+\end{equation}
+where $\Gamma_f = R_f - F\left(x_1\right)$ is the mean number of
 foreground events with $x > x_1$.  The overall foreground rate is
@@
   \label{eq:foreground-from-lambda}
-  R_f = \Lambda_f \frac{x_1^3}{x_\mathrm{th}^3}.
+  R_f = \Gamma_f \frac{x_1^3}{x_\mathrm{th}^3}.
 \end{equation}
@@
   \erfc\left(\frac{x_1}{\sqrt{2}}\right)\right]^{-1}\exp\left[-\frac{x^2}{2}\right]
-+ 3 \Lambda_f \frac{x_1^3}{x^4}
-\end{equation}
-
-For any pair $(x_1, \Lambda_f)$, it is straightforward to construct
++ 3 \Gamma_f \frac{x_1^3}{x^4}
+\end{equation}
+
+For any pair $(x_1, \Gamma_f)$, it is straightforward to construct
 random event lists drawn from the corresponding $p(x)$, and
 straightforward to apply a threshold by ``throwing away'' all events
-with $x$ less than the threshold value $x_\mathrm{th}$.  If $\Lambda_f
+with $x$ less than the threshold value $x_\mathrm{th}$.  If $\Gamma_f
 \gg 1$, then we are in the foreground-dominated regime at $x = x_1$,
-if $\Lambda_f \ll 1$ we are in the background-dominated regime, and if
-$\Lambda_f \sim 1$ the foreground and background counts above $x_1$
+if $\Gamma_f \ll 1$ we are in the background-dominated regime, and if
+$\Gamma_f \sim 1$ the foreground and background counts above $x_1$
 are about equal.  For any thresholded event list, we use
@@
     rate uncertainty, Eq.~\eqref{eq:delta-R-definition}, as a function
-    of threshold for data sets with $\Lambda_f = 100$ (solid line),
-    $\Lambda_f = 10$ (dashed line), and $\Lambda_f = 1$ (dash-dotted
-    line).  Recall that $\Lambda_f$ is the mean number of foreground
-    events above $x_\mathrm{th} = 8$.  The total background rate,
+    of threshold for data sets with $\Gamma_f = 100$ (solid line),
+    $\Gamma_f = 10$ (dashed line), and $\Gamma_f = 1$ (dash-dotted
+    line).  Recall that $\Gamma_f$ is the mean number of foreground
+    events above $x_\mathrm{th}=x_1= 8$.  The total background rate,
     $R_b\left( x_\mathrm{th} \right)$, is shown by the dotted line; we
@@
     asymptotes to the counting error on the events that stand out from
-    the background, $\Delta R_f / R_f \simeq 1/\sqrt{\Lambda_f}$.
-    Note that this uncertainty applies even when the total number of
-    background events is orders of magnitude larger than the number of
-    foreground events.  When a threshold must be chosen, it is
-    safest---in the sense of producing the minimal foreground rate
-    uncertainty---to choose the threshold \emph{well into the
-      background-dominated loudness regime}; the extra background
-    events in the data set do not affect the estimate of the
-    foreground rate, and, when the background distribution is
-    parameterized, can help to better determine these parameters (see
-    \S~\ref{sec:gw-overlapping-template}). }
+    the background, $\Delta R_f / R_f \simeq 1/\sqrt{\Gamma_f}$.}
 \end{figure}
@@
 \begin{equation}
-  \frac{\Delta R_f}{R_f} \simeq \frac{1}{\sqrt{\Lambda_f}},
+  \frac{\Delta R_f}{R_f} \simeq \frac{1}{\sqrt{\Gamma_f}},
 \end{equation}
@@
 regions---contributes little to reducing the overall uncertainty in the
-foreground rate.  This is consistent with the behavior of the specific
+foreground rate.
+Thus, when the background distribution itself is of no interest and computational costs are high, the threshold does not need to be pushed into background-dominated regions in order to obtain an accurate foreground estimate.
+This is consistent with the behavior of the specific
 example in Figure \ref{fig:threshold-rate-dependence}.
@@
 previous subsection, and we will begin with a very specific example.
-Let $\Lambda_f = 1$, meaning that the expected number of actual events
+Let $\Gamma_f = 1$, meaning that the expected number of actual events
 with $x > x_1$ is one.  Then there is a $1/64$ chance ($1-\hat{F}(x_{LE}) \approx
@@
 
-Now, if the true rate really were $\Lambda_f = 64$, then the expected
+Now, if the true rate really were $\Gamma_f = 64$, then the expected
 number of events with $x > x_1$ would be $64$.  So in this case, the
@@
 set to $x_1$ incorporates this information quite naturally, and so
-(correctly) yields an estimated $\Lambda_f$ of order one.
+(correctly) yields an estimated $\Gamma_f$ of order one.
 

### 2013-01-17 many.tex
Adding small changes based on September e-mail exchange.
Minor acknowledgements changes and address changes prior to P&P submission.

@@
 \author{Curt Cutler} \email{Curt.J.Cutler@jpl.nasa.gov}
-\affiliation{Jet Propulsion Lab, 4800 Oak Grove Dr., Pasadena, CA
-  91109} \affiliation{Theoretical Astrophysics, California Institute
-  of Technology, Pasadena, California 91125}
+\affiliation{Jet Propulsion Lab \\ 4800 Oak Grove Dr \\ Pasadena, CA
+  91109} \affiliation{Theoretical Astrophysics \\ California Institute
+  of Technology \\ Pasadena, CA 91125}
 
@@
 
-If $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-
+If $\tilde{b}(x_N,\theta) \ll \tilde{f}(x_N,\theta)$, we obtain the result $(1-
 \hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be understood as the
@@
 [\ln(1-\hat{B}(x,\theta))]/{\rm d}x$ and $(1-\hat{B}(x,\theta))
-\rightarrow 0$ as $x \rightarrow \infty$, so this term will be
+\rightarrow 0$ as $x \rightarrow \infty$, so this term may be
 divergent and for many reasonable examples, we will find
-$\tilde{b}(x_N,\theta) >> \hat{f}(x_N,\theta)$, in which case the
+$\tilde{b}(x_N,\theta) \gg \tilde{f}(x_N,\theta)$, in which case the
 posterior on $R_f$ is peaked at $0$. This issue highlights the problem
@@
   O'Shaughnessy, and Ted von Hippel for discussions and suggestions
-  about this manuscript.
-IM and WMF acknowledge the hospitality of KITP, supported in part by the National Science Foundation under No.
-NSF Grant PHY11-25915.
-CC's work was carried out  at the Jet Propulsion Laboratory, California Institute of Technology, under contract to the National Aeronautics and Space Administration.
-CC also gratefully acknowledges support from NSF grant PHY1068881.
-JG's work is supported by the Royal Society.
-Copyright 2013.
-All rights reserved.
+  about this manuscript.  IM and WMF acknowledge the hospitality of
+  KITP, supported in part by the National Science Foundation under NSF
+  Grant PHY11-25915.  CC's work was carried out at the Jet Propulsion
+  Laboratory, California Institute of Technology, under contract to
+  the National Aeronautics and Space Administration.  CC also
+  gratefully acknowledges support from NSF grant PHY1068881.  JG's
+  work is supported by the Royal Society.  Copyright 2013.  All rights
+  reserved.
 \end{acknowledgments}

### 2013-02-19 many.tex
Implemented Tom Dent's comments from LVC review.

@@
 in gravitational-wave data analysis
-\cite[e.g.,][]{s5-highmass,s6-lowmass} and in astronomical
-observations of a field of objects of mixed provenance
-\cite{DeGennaro2009}.  In this paper, we introduce a robust formalism
-for estimating event rates from the data when the shape of foreground
-and background distributions are known (or parameterized), but the
-provenance of individual events as either background or foreground is
-unknown.
+\cite[e.g.,][]{s5-lowmass,s5-highmass,s5-highmass-erratum,s5-lowmass-2,s6-lowmass,s6-highmass}
+and in astronomical observations of a field of objects of mixed
+provenance \cite{DeGennaro2009}.  In this paper, we introduce a robust
+formalism for estimating event rates from the data when the shape of
+foreground and background distributions are known (or parameterized),
+but the provenance of individual events as either background or
+foreground is unknown.
 
@@
 are commonly used.  One, known as the \emph{loudest-event statistic}
-\cite{Biswas2009}, uses only the information from the highest-ranked
-event in the data to infer the rate distribution.  This approach has
-been used successfully \cite{s5-highmass,s6-lowmass} when the number
-of loud foreground events is small (typically zero or one) to obtain
-upper limits on foreground rates.  However, the loudest-event
-statistic ignores all events except the loudest one, and so suffers
-from an unnecessary loss of information; therefore, we expect it to
-yield a much larger variance than strictly necessary when multiple
-events are present in the data.
+\cite{Biswas2009,Brady2004,Brady2008}, uses only the information from
+the highest-ranked event in the data to infer the rate distribution.
+This approach has been used successfully
+\cite{s5-lowmass,s5-lowmass-2,s5-highmass,s5-highmass-erratum,s6-lowmass,s6-highmass}
+when the number of loud foreground events is small (typically zero or
+one) to obtain upper limits on foreground rates.  However, the
+loudest-event statistic ignores all events except the loudest one, and
+so suffers from an unnecessary loss of information; therefore, we
+expect it to yield a much larger variance than strictly necessary when
+multiple events are present in the data.  In practice, the
+loudest-event statistic is typically applied repeatedly to multiple
+``chunks'' of data, using the estimated rate posterior from each chunk
+as a rate prior for the next chunk's analysis
+\cite{s5-lowmass-2,s5-highmass,s5-highmass-erratum}.  Even when used
+in this mode, the method discards information, with the amout of
+information loss depending on the (arbitrary) division of the data
+into chunks.
 
@@
 
+Ref.~\cite{Messenger2012} considered the problem of determining an
+intrinsic rate and population parameters in the presence of missing
+data, either due to thresholding, poor sensitivity, or contamination
+from noise events.  The approach is complementary to ours: we consider
+the problem of accurately counting the events of different classes
+present in a dataset, while Ref.~\cite{Messenger2012} deals with
+translating such counts into physical rates by properly accounting for
+the selection effects on the data set.
+
 In order to demonstrate our method, we consider three different
@@
 the background distribution and rate are known, which corresponds to
-using a delta function prior on $R_b$. They further assume a flat
-prior on $R_f$ and that the foreground and background distributions do
-not depend on any unknown free parameters. With these assumptions, the
-posterior on $R_f$, Eq.~\eqref{LEeq1}, is modified to
+using a narrow prior on $R_b$. They further assume a flat prior (in
+the absence of other experimental data) on $R_f$ and that the
+foreground and background distributions do not depend on any unknown
+free parameters. With these assumptions, the posterior on $R_f$,
+Eq.~\eqref{LEeq1}, is modified to
 \begin{multline}
@@
 \end{equation}
-With the further identification $\mu \equiv R_f$ and $\hat{\epsilon} \equiv 1 - \hat{F}(x_N)$, this is Eq.~(18) of~\citep{Biswas2009} and we have shown how their parameter $\Lambda$ is related to the foreground and background distributions used here.
+With the further identification $\mu \equiv R_f$ and $\hat{\epsilon} \equiv 1 - \hat{F}(x_N)$, this is Eq.~(14) of~\citep{Biswas2009} and we have shown how their parameter $\Lambda$ is related to the foreground and background distributions used here.
 
@@
 
-If $\tilde{b}(x_N,\theta) \ll \tilde{f}(x_N,\theta)$, we obtain the result $(1-
-\hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be understood as the
-statement that the rate of foreground events with ranking statistic
-greater than $x_N$, $(1- \hat{F}(x_N,\theta))\,R_f$, is of order $1$,
-as expected. However, $\tilde{b}(x_N,\theta) = -{\rm d}
-[\ln(1-\hat{B}(x,\theta))]/{\rm d}x$ and $(1-\hat{B}(x,\theta))
+If $\tilde{b}(x_N,\theta) \ll \tilde{f}(x_N,\theta)$, we obtain the
+result $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be
+understood as the statement that the rate of foreground events with
+ranking statistic greater than $x_N$, $(1- \hat{F}(x_N,\theta))\,R_f$,
+is of order $1$, as expected. However, $\tilde{b}(x_N,\theta) = -{\rm
+  d} [\ln(1-\hat{B}(x,\theta))]/{\rm d}x$ and $(1-\hat{B}(x,\theta))
 \rightarrow 0$ as $x \rightarrow \infty$, so this term may be
@@
 significantly greater than the total number of observed events are
-strongly disfavored.
+strongly disfavored.  As we will see, the problem can also be avoided
+in the context of the loudest-event framework by even very weak prior
+information on the background rate, $R_b$, of the kind present in
+nearly all experiments.
 
@@
 address the question of how the rate estimate depends on the threshold
-value.
+value.  For a discussion of selection effects, of which thresholding
+is but one, on the estimate of physical rates, see
+Ref.~\cite{Messenger2012}.
 
@@
 stream by matched filtering in the frequency domain against a set of
-$N$ template waveforms \citep[e.g.,][]{findchirppaper,LVC2011}.  In
-our simplistic model, we suppose the data stream consists of
+$N$ template waveforms \citep[e.g.,][]{findchirppaper,s6-lowmass}.  We
+use an extremely simplified model of such a search and the ensuing
+analysis to demonstrate how our framework could be used in practice.
+
+In our simplistic model, we suppose the data stream consists of
 stationary Gaussian noise with a power spectral density $S(f)$
@@
 accurately model both distributions, e.g., through the use of injected
-signals to model the foreground distribution \cite{ihope,CannonHannaKeppel:2012}.
-However, this is a difficult problem (e.g., because of the need to
-estimate the background at the very tails of the distribution), and
-will require significant future work.  In Section \ref{sec:uncertain-dist}, 
-we discussed some of the possible
-approaches when the shapes of the background and foreground distributions
-cannot be confidently described by models with a few adjustable parameters.
+signals \cite{ihope} or other methods \cite{CannonHannaKeppel:2012} to
+model the foreground distribution.  However, this is a difficult
+problem (e.g., because of the need to estimate the background at the
+very tails of the distribution), and will require significant future
+work.  In Section \ref{sec:uncertain-dist}, we discussed some of the
+possible approaches when the shapes of the background and foreground
+distributions cannot be confidently described by models with a few
+adjustable parameters.
 
@@
 data-analysis algorithm and ranking statistic, including any selection
-effects, and would need to be determined on a case-by-case basis.
+effects \cite{Messenger2012}, and would need to be determined on a
+case-by-case basis.  See Ref.~\cite{Brady2008} for an example of such
+conversion when the underlying framework is the loudest-event
+statistic.
 

### 2013-02-20 many.tex
Fixed spelling typo.

@@
 \cite{s5-lowmass-2,s5-highmass,s5-highmass-erratum}.  Even when used
-in this mode, the method discards information, with the amout of
+in this mode, the method discards information, with the amount of
 information loss depending on the (arbitrary) division of the data

### 2013-02-20 many.tex
Added Tom Dent to acknowledgements.

@@
 \begin{acknowledgments}
-  We thank Kipp Cannon, Chad Hanna, Drew Keppel, Richard
+  We thank Kipp Cannon, Thomas Dent, Chad Hanna, Drew Keppel, Richard
   O'Shaughnessy, and Ted von Hippel for discussions and suggestions

### 2013-03-06 many.tex
Added TODO re: citation suggested by Hogg for "detecting the undetectable" in a MW stream.

--- 2013-02-20 many.tex
@@
 
-The task of estimating rates of events when a mixture of foreground
-and background events is present in data is a common one in physical
-and astrophysical applications.  This problem comes up, among others,
-in gravitational-wave data analysis
+\will{Find a way to discuss \cite{Koposov2010}.}  The task of
+estimating rates of events when a mixture of foreground and background
+events is present in data is a common one in physical and
+astrophysical applications.  This problem comes up, among others, in
+gravitational-wave data analysis
 \cite[e.g.,][]{s5-lowmass,s5-highmass,s5-highmass-erratum,s5-lowmass-2,s6-lowmass,s6-highmass}
@@
 \begin{acknowledgments}
-  We thank Kipp Cannon, Chad Hanna, Drew Keppel, Richard
-  O'Shaughnessy, and Ted von Hippel for discussions and suggestions
-  about this manuscript.  IM and WMF acknowledge the hospitality of
-  KITP, supported in part by the National Science Foundation under NSF
-  Grant PHY11-25915.  CC's work was carried out at the Jet Propulsion
-  Laboratory, California Institute of Technology, under contract to
-  the National Aeronautics and Space Administration.  CC also
-  gratefully acknowledges support from NSF grant PHY1068881.  JG's
-  work is supported by the Royal Society.  Copyright 2013.  All rights
-  reserved.
+  We thank Kipp Cannon, Thomas Dent, Chad Hanna, Drew Keppel, Richard
+  O'Shaughnessy, David Hogg, and Ted von Hippel for discussions and
+  suggestions about this manuscript.  IM and WMF acknowledge the
+  hospitality of KITP, supported in part by the National Science
+  Foundation under NSF Grant PHY11-25915.  CC's work was carried out
+  at the Jet Propulsion Laboratory, California Institute of
+  Technology, under contract to the National Aeronautics and Space
+  Administration.  CC also gratefully acknowledges support from NSF
+  grant PHY1068881.  JG's work is supported by the Royal Society.
+  Copyright 2013.  All rights reserved.
 \end{acknowledgments}

### 2014-10-01 response.txt
Partial commit of response to referee; cropped a couple of images, some text changes [through comment 9].

Dear Editor:

This is a response to the referee report on our manuscript DB11163 by Farr et al.

We apologize for the very delayed response.  We are grateful to the referee for a thorough reading of the manuscript and a number of useful comments.  We have incorporated the suggestions of the referee.  The text in bold represents the changes that have been made relative to the initial submission.  We also describe specific changes inline below.

>> ----------------------------------------------------------------------
>> Report of the Referee -- DB11163/Farr
>> ----------------------------------------------------------------------
>>
>> Referee report for manuscript DB11163 Farr
>> ==========================================
>>
>> The manuscript presents an application of Bayesian inference to the
>> problem of estimating the number of events in observed data which
>> belong to two independent populations, an example of a mixture model
>> with latent variables indicating the class of each event. It
>> illustrates the approach with examples from gravitational wave
>> searches and from fitting parameters of a stellar cluster superimposed
>> on a background. Although the authors only use simulated data in their
>> examples, the method employed is interesting and makes an advance on
>> the earlier methods used in the literature for approaching these
>> problems, although I do not think the statistics are novel in
>> themselves.
>>
>> The quality of writing and clarity of explanations is consistently
>> very high, and I believe the article is appropriate for publication in
>> PRD, but I have a few comments noted below that I recommend the
>> authors take into account.
>>
>> 1. There is a little confusion between number and rate in places. The
>> authors estimate rates of Poisson distributions throughout, i.e. the
>> number per unit time or area. The implicit factor of time or area
>> should be noted in the definition of R_f and R_b to avoid confusion,
>> and for dimensional correctness. They also occasionally refer to these
>> as numbers not rates in the text, e.g. after eq 34.

Whatever the units of f and \hat{f} [which depends on the space in which the corresponding rate density is measured], R_f provides the overall scaling amplitude in Eq. 9 and is therefore dimensionless (and similarly for R_b).  We have added a comment to this effect in the text.

>>
>> 2. Physical interpretations. I had difficulty identifying the key
>> result in the paper. The conclusions state the superiority of their
>> "framework", but if I wanted to use the framework where would I start?
>> Equation 18 gives the full posterior for the unknown variables so
>> perhaps that is the place. Or is it eqn 21? I suggest that the authors
>> emphasise the key results of their work with a view to how it might be
>> applied by the reader for his/her own particular problem.

Indeed, (18) and (21) are the two key equations; which one is used depends on whether the marginalized probability distributions associated with individual events are of interest.  We have made this clear in both the introduction and the conclusion.

>
>> 3. Using $f_i$ for the flag variables seems unnecessary when
>> $\hat{f}(x_i)$ and $\tilde{f}$ are also used. Can you choose a
>> different symbol?

We have replaced $f_i$ with $g_i$.

>>
>> 4. Before eqn 32 the start of the paragraph repeats some of the end of
>> the previous paragraph, suggest rewording.
>>
>> 5. I didn't understand the comment about comparing rates for different
>> thresholds just before eqn 35. Could you expand on this paragraph,
>> perhaps with an example?
>>
>> 6. Below eqn 37, "...average of the Bayesian estimate $R_B$" should be
>> an estimatOR. Likewise in the next sentence all threshold values will
>> yield, on average, the same POINT estimate.
>>
>> 7. After eqn 59 the method of time slides is mentioned, and it is
>> stated the background in time slides will be representative of the
>> true background. I don't think this is true when there are real events
>> in the unslid data which can be slid into coincidence with background
>> triggers. This is hardly a key point of the paper but it might be
>> worth mentioning that here.
>>
>> 8. The x-axis in figure 2 is very unclear, having only 1 reference
>> point. I suggest adding another.
>>
>> 9. The axes for figure 9 are illegibly small.
>>
>> Section V.B: As this is the section of most relevance for
>> gravitational wave analyses, it could benefit from some expansion, or
>> perhaps a separate paper if it grows too large. I would also like the
>> authors to clarify these questions below.
>>
>> 10. The authors do not mention the effect of overlapping template
>> banks on the foreground. Won't there also be correlated foreground
>> events as well as background?
>>
>> 11. It is stated that Neff is a property of the background
>> distribution, but shouldn't it be a property of the template bank
>> itself, or a property of the model that is being analysed?
>>
>> 12. It is stated that they used the same dataset as sec. V.A, with the
>> same number of events. Doesn't this mean that there were actually no
>> correlated events in the distributions? A more convincing example
>> would actually simulate the new background.
>>

———————————————————————————————————
Report of the Referee -- DB11163/Farr
----------------------------------------------------------------------

Referee report for manuscript DB11163 Farr
==========================================

The manuscript presents an application of Bayesian inference to the
problem of estimating the number of events in observed data which
belong to two independent populations, an example of a mixture model
with latent variables indicating the class of each event. It
illustrates the approach with examples from gravitational wave
searches and from fitting parameters of a stellar cluster superimposed
on a background. Although the authors only use simulated data in their
examples, the method employed is interesting and makes an advance on
the earlier methods used in the literature for approaching these
problems, although I do not think the statistics are novel in
themselves.

The quality of writing and clarity of explanations is consistently
very high, and I believe the article is appropriate for publication in
PRD, but I have a few comments noted below that I recommend the
authors take into account.

1. There is a little confusion between number and rate in places. The
authors estimate rates of Poisson distributions throughout, i.e. the
number per unit time or area. The implicit factor of time or area
should be noted in the definition of R_f and R_b to avoid confusion,
and for dimensional correctness. They also occasionally refer to these
as numbers not rates in the text, e.g. after eq 34.

2. Physical interpretations. I had difficulty identifying the key
result in the paper. The conclusions state the superiority of their
"framework", but if I wanted to use the framework where would I start?
Equation 18 gives the full posterior for the unknown variables so
perhaps that is the place. Or is it eqn 21? I suggest that the authors
emphasise the key results of their work with a view to how it might be
applied by the reader for his/her own particular problem.

3. Using $f_i$ for the flag variables seems unnecessary when
$\hat{f}(x_i)$ and $\tilde{f}$ are also used. Can you choose a
different symbol?

4. Before eqn 32 the start of the paragraph repeats some of the end of
the previous paragraph, suggest rewording.

5. I didn't understand the comment about comparing rates for different
thresholds just before eqn 35. Could you expand on this paragraph,
perhaps with an example?

6. Below eqn 37, "...average of the Bayesian estimate $R_B$" should be
an estimatOR. Likewise in the next sentence all threshold values will
yield, on average, the same POINT estimate.

7. After eqn 59 the method of time slides is mentioned, and it is
stated the background in time slides will be representative of the
true background. I don't think this is true when there are real events
in the unslid data which can be slid into coincidence with background
triggers. This is hardly a key point of the paper but it might be
worth mentioning that here.

8. The x-axis in figure 2 is very unclear, having only 1 reference
point. I suggest adding another.

9. The axes for figure 9 are illegibly small.

Section V.B: As this is the section of most relevance for
gravitational wave analyses, it could benefit from some expansion, or
perhaps a separate paper if it grows too large. I would also like the
authors to clarify these questions below.

10. The authors do not mention the effect of overlapping template
banks on the foreground. Won't there also be correlated foreground
events as well as background?

11. It is stated that Neff is a property of the background
distribution, but shouldn't it be a property of the template bank
itself, or a property of the model that is being analysed?

12. It is stated that they used the same dataset as sec. V.A, with the
same number of events. Doesn't this mean that there were actually no
correlated events in the distributions? A more convincing example
would actually simulate the new background.

### 2014-10-01 response.txt
Finished edits in response to referee's comments.

@@
 We apologize for the very delayed response.
-We are grateful to the referee for a thorough reading of the manuscript and a number of useful comments.
+We are grateful to the referee for a thorough reading of the manuscript and a number of very useful comments.
 We have incorporated the suggestions of the referee.
@@
 Indeed, (18) and (21) are the two key equations; which one is used depends on whether the marginalized probability distributions associated with individual events are of interest.
-We have made this clear in both the introduction and the conclusion.
+We have added a new paragraph to the introduction (penultimate paragraph) to highlight these equations.
 
->
 >> 3. Using $f_i$ for the flag variables seems unnecessary when
@@
 
->>
 >> 4. Before eqn 32 the start of the paragraph repeats some of the end of
 >> the previous paragraph, suggest rewording.
->>
+
+Reworded.
+
 >> 5. I didn't understand the comment about comparing rates for different
@@
 >> perhaps with an example?
->>
+
+We clarified the text.
+
 >> 6. Below eqn 37, "...average of the Bayesian estimate $R_B$" should be
@@
 >> yield, on average, the same POINT estimate.
->>
+
+Typos fixed.
+
 >> 7. After eqn 59 the method of time slides is mentioned, and it is
@@
 >> worth mentioning that here.
->>
+
+This is indeed a very interesting question that has attracted much debate in the gravitational-wave data analysis community in recent months.
+Whether to remove foreground candidates (those coincident in zero lag, prior to time slides) before performing slides depends on whether a conservative estimate or a more accurate estimate is warranted.
+We decided to leave this discussion out of the present paper.
+
 >> 8. The x-axis in figure 2 is very unclear, having only 1 reference
 >> point. I suggest adding another.
->>
+
+Revised.
+
 >> 9. The axes for figure 9 are illegibly small.
->>
+
+Figure expanded to make the axes more clear.
+
 >> Section V.B: As this is the section of most relevance for
@@
 >> authors to clarify these questions below.
->>
+
+We are indeed planning a follow-up paper that would tackle some of the practical issues raised by the referee, such as background estimates.
+As this paper is focused on the statistical framework, and gravitational-wave data analysis is just one of the examples, we have resisted the urge to go into too many details here.
+
 >> 10. The authors do not mention the effect of overlapping template
@@
 >> events as well as background?
->>
+
+As described around Eq. (54), we only record the loudest event (the one which best matches the template)  at any given time.
+
 >> 11. It is stated that Neff is a property of the background
@@
 >> itself, or a property of the model that is being analysed?
->>
+
+We clarified this to say that it is a property of the distribution of background events identified by filtering with this template bank.
+
 >> 12. It is stated that they used the same dataset as sec. V.A, with the
@@
 >> would actually simulate the new background.
->>
 
-———————————————————————————————————
-Report of the Referee -- DB11163/Farr
-----------------------------------------------------------------------
+There was no filtering as such in either example — the background distribution was constructed by drawing template SNRs as described around Eq.
+(55).
+In the example of section V.B, only N_eff templates were used to construct the background distribution.
 
-Referee report for manuscript DB11163 Farr
-==========================================
+We have also added a new reference, [8] in the present revision.
 
-The manuscript presents an application of Bayesian inference to the
-problem of estimating the number of events in observed data which
-belong to two independent populations, an example of a mixture model
-with latent variables indicating the class of each event. It
-illustrates the approach with examples from gravitational wave
-searches and from fitting parameters of a stellar cluster superimposed
-on a background. Although the authors only use simulated data in their
-examples, the method employed is interesting and makes an advance on
-the earlier methods used in the literature for approaching these
-problems, although I do not think the statistics are novel in
-themselves.
-
-The quality of writing and clarity of explanations is consistently
-very high, and I believe the article is appropriate for publication in
-PRD, but I have a few comments noted below that I recommend the
-authors take into account.
-
-1. There is a little confusion between number and rate in places. The
-authors estimate rates of Poisson distributions throughout, i.e. the
-number per unit time or area. The implicit factor of time or area
-should be noted in the definition of R_f and R_b to avoid confusion,
-and for dimensional correctness. They also occasionally refer to these
-as numbers not rates in the text, e.g. after eq 34.
-
-2. Physical interpretations. I had difficulty identifying the key
-result in the paper. The conclusions state the superiority of their
-"framework", but if I wanted to use the framework where would I start?
-Equation 18 gives the full posterior for the unknown variables so
-perhaps that is the place. Or is it eqn 21? I suggest that the authors
-emphasise the key results of their work with a view to how it might be
-applied by the reader for his/her own particular problem.
-
-3. Using $f_i$ for the flag variables seems unnecessary when
-$\hat{f}(x_i)$ and $\tilde{f}$ are also used. Can you choose a
-different symbol?
-
-4. Before eqn 32 the start of the paragraph repeats some of the end of
-the previous paragraph, suggest rewording.
-
-5. I didn't understand the comment about comparing rates for different
-thresholds just before eqn 35. Could you expand on this paragraph,
-perhaps with an example?
-
-6. Below eqn 37, "...average of the Bayesian estimate $R_B$" should be
-an estimatOR. Likewise in the next sentence all threshold values will
-yield, on average, the same POINT estimate.
-
-7. After eqn 59 the method of time slides is mentioned, and it is
-stated the background in time slides will be representative of the
-true background. I don't think this is true when there are real events
-in the unslid data which can be slid into coincidence with background
-triggers. This is hardly a key point of the paper but it might be
-worth mentioning that here.
-
-8. The x-axis in figure 2 is very unclear, having only 1 reference
-point. I suggest adding another.
-
-9. The axes for figure 9 are illegibly small.
-
-Section V.B: As this is the section of most relevance for
-gravitational wave analyses, it could benefit from some expansion, or
-perhaps a separate paper if it grows too large. I would also like the
-authors to clarify these questions below.
-
-10. The authors do not mention the effect of overlapping template
-banks on the foreground. Won't there also be correlated foreground
-events as well as background?
-
-11. It is stated that Neff is a property of the background
-distribution, but shouldn't it be a property of the template bank
-itself, or a property of the model that is being analysed?
-
-12. It is stated that they used the same dataset as sec. V.A, with the
-same number of events. Doesn't this mean that there were actually no
-correlated events in the distributions? A more convincing example
-would actually simulate the new background.
+Sincerely yours,
+Will Farr
+Jonathan R. Gair
+Ilya Mandel
+Curt Cutler

### 2014-10-01 many.tex
Finished edits in response to referee's comments.

--- 2013-02-20 many.tex
@@
 
-\author{Will M. Farr} \email{w-farr@northwestern.edu}
+\author{Will M. Farr} \email{wfarr@star.sr.bham.ac.uk}
 \homepage{http://faculty.wcas.northwestern.edu/will-farr/}
@@
   University, 2145 Sheridan Road, Evanston, IL 60208}
+\affiliation{School of
+  Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT
+  Birmingham\\United Kingdom}
 
@@
 
-The task of estimating rates of events when a mixture of foreground
-and background events is present in data is a common one in physical
-and astrophysical applications.  This problem comes up, among others,
-in gravitational-wave data analysis
+The task of estimating rates of events when a mixture of foreground and background
+events is present in data is a common one in physical and
+astrophysical applications.  This problem comes up, among others, in
+gravitational-wave data analysis
 \cite[e.g.,][]{s5-lowmass,s5-highmass,s5-highmass-erratum,s5-lowmass-2,s6-lowmass,s6-highmass}
 and in astronomical observations of a field of objects of mixed
-provenance \cite{DeGennaro2009}.  In this paper, we introduce a robust
+provenance \cite{DeGennaro2009,Koposov2010}.  In this paper, we introduce a robust
 formalism for estimating event rates from the data when the shape of
@@
 
+{\bf Our key results appear in Eq.~(\ref{eq:posterior}), which provides the joint posterior probability distribution on the foreground and background rates and shape parameters and the provenance of individual events as either foreground or background.
+Eq.~(\ref{eq:rate-shape-posterior}) is a marginalized version of Eq.~(\ref{eq:posterior}), useful when the provenance of individual events is not relevant.
+In practice, these posteriors are best sampled with stochastic techniques such as Markov chain Monte Carlo.}
+
 In order to demonstrate our method, we consider three different
@@
 probability distributions for these parameters that are implied by the
-data.  At the least, we will want to know the overall amplitude of the
-foreground and background rates.  Let
+data.  At the least, we will want to know the overall {\bf dimensionless} 
+amplitude of the foreground and background rates.  Let
 \begin{equation}
@@
 We do not know a priori which of the events are foreground and which
-are background.  For each event, we introduce a flag, $f_i$, which is
+are background.  For each event, we introduce a flag, $g_i$, which is
 either 0 (background) or 1 (foreground).  These ``state'' flags are
@@
 can marginalize over our uncertainty in the state of any given event
-by summing posteriors over $f_i = \mathset{0,1}$.
+by summing posteriors over $g_i = \mathset{0,1}$.
 
@@
 theorem relates the posterior probability of the state flags, rates,
-and shape parameters, $p\left(\mathset{f_i}, R_f, R_b, \theta | d_{\rm
+and shape parameters, $p\left(\mathset{g_i}, R_f, R_b, \theta | d_{\rm
   to}, N \right)$, the likelihood of the data, $p\left( d_{\rm to} |
-\mathset{f_i}, N, R_f, R_b, \theta\right)$, and the prior probability
+\mathset{g_i}, N, R_f, R_b, \theta\right)$, and the prior probability
 of state flags, rates and shape parameters before any data are
-obtained, $p\left( \mathset{f_i}, N, R_f, R_b, \theta\right)$:
+obtained, $p\left( \mathset{g_i}, N, R_f, R_b, \theta\right)$:
 \begin{multline}
   \label{eq:Bayes}
-  p\left(\mathset{f_i}, R_f, R_b, \theta | d_{\rm to}, N \right) \\ =
-  \frac{p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right)
-    p\left(\mathset{f_i}, N, R_f, R_b, \theta\right)}{p(d_{\rm to},
+  p\left(\mathset{g_i}, R_f, R_b, \theta | d_{\rm to}, N \right) \\ =
+  \frac{p\left( d_{\rm to} | \mathset{g_i}, N, R_f, R_b, \theta\right)
+    p\left(\mathset{g_i}, N, R_f, R_b, \theta\right)}{p(d_{\rm to},
     N)}.
@@
 \begin{multline}
-  p\left( d_{\rm to} | \mathset{f_i}, N, R_f, R_b, \theta\right) \\ =
-  \left[ \prod_{\mathset{i | f_i = 1}} \hat{f}\left(x_i, \theta\right)
-    \right] \left[ \prod_{\mathset{i | f_i = 0}} \hat{b}\left( x_i,
+  p\left( d_{\rm to} | \mathset{g_i}, N, R_f, R_b, \theta\right) \\ =
+  \left[ \prod_{\mathset{i | g_i = 1}} \hat{f}\left(x_i, \theta\right)
+    \right] \left[ \prod_{\mathset{i | g_i = 0}} \hat{b}\left( x_i,
     \theta\right) \right].
@@
 This is the probability that the first observed event is a
-fore/background event (if $f_1=1,0$) with ranking statistic $x_1$ and
-the second observed event is a fore/background event (if $f_2=1,0$)
+fore/background event (if $g_1=1,0$) with ranking statistic $x_1$ and
+the second observed event is a fore/background event (if $g_2=1,0$)
 with ranking statistic $x_2$, etc. If the events are ordered by ranking
@@
   \label{eq:combined-flag-rate-prior}
-  p\left(\mathset{f_i}, N, R_f, R_b, \theta\right) \\ = p\left(
-  \mathset{f_i} | N, R_f, R_b\right) p\left(N |R_f, R_b\right)
-  p\left(R_f, R_b, \theta\right) \\ = p\left( \mathset{f_i},N | R_f,
+  p\left(\mathset{g_i}, N, R_f, R_b, \theta\right) \\ = p\left(
+  \mathset{g_i} | N, R_f, R_b\right) p\left(N |R_f, R_b\right)
+  p\left(R_f, R_b, \theta\right) \\ = p\left( \mathset{g_i},N | R_f,
   R_b\right) p\left(R_f, R_b, \theta\right).
@@
 
-The probability that the $i$'th state flag is $f_i=1$ is given by
+The probability that the $i$'th state flag is $g_i=1$ is given by
 $R_f/(R_f+R_b)$, while the probability that it is zero is
@@
 \begin{multline}
-p\left( \mathset{f_i} | N, R_f, R_b\right) \\ = \prod_{\mathset{i|f_i=1}}
-\left(\frac{R_f}{R_f+R_b}\right) \prod_{\mathset{i|f_i=0}}
+p\left( \mathset{g_i} | N, R_f, R_b\right) \\ = \prod_{\mathset{i|g_i=1}}
+\left(\frac{R_f}{R_f+R_b}\right) \prod_{\mathset{i|g_i=0}}
 \left(\frac{R_b}{R_f+R_b}\right) \\ =
@@
   \label{eq:flag-conditional-prior}
-  p\left(\mathset{f_i},N | R_f, R_b\right) = \frac{R_f^{N_f}
+  p\left(\mathset{g_i},N | R_f, R_b\right) = \frac{R_f^{N_f}
     R_b^{N_b}}{N!} \exp\left[ - \left(R_f + R_b\right) \right].
@@
   \label{eq:posterior}
-  p\left( \mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right) =
-  \frac{\alpha}{p(d_{\rm to},N)\,N!} \\ \times \left[ \prod_{\mathset{i | f_i =
+  p\left( \mathset{g_i}, R_f, R_b, \theta | d_{\rm to},N \right) =
+  \frac{\alpha}{p(d_{\rm to},N)\,N!} \\ \times \left[ \prod_{\mathset{i | g_i =
         1}} R_f \hat{f}\left(x_i, \theta\right) \right] \left[
-    \prod_{\mathset{i | f_i = 0}} R_b \hat{b}\left( x_i, \theta\right)
+    \prod_{\mathset{i | g_i = 0}} R_b \hat{b}\left( x_i, \theta\right)
     \right] \\ \times \exp\left[ - \left(R_f + R_b\right) \right]
@@
 one correspondence between a time-ordered posterior $p\left(
-\mathset{f_i}, R_f, R_b, \theta | d_{\rm to},N \right)$ and the
-corresponding statistic-ordered posterior $ p\left( \mathset{f_i},
-R_f, R_b, \theta | d,N \right)$, which means $p\left( \mathset{f_i},
-R_f, R_b, \theta | d,N \right) = p\left( \mathset{f_i}, R_f, R_b,
+\mathset{g_i}, R_f, R_b, \theta | d_{\rm to},N \right)$ and the
+corresponding statistic-ordered posterior $ p\left( \mathset{g_i},
+R_f, R_b, \theta | d,N \right)$, which means $p\left( \mathset{g_i},
+R_f, R_b, \theta | d,N \right) = p\left( \mathset{g_i}, R_f, R_b,
 \theta | d_{\rm to},N \right)$. However, the evidence $p(d,N) =
@@
 The ranking-statistic ordered posterior can be computed directly by
-assuming that the flags, $\mathset{f_i}$, are un-observed data and
-treating the sets $\mathset{x_i | f_i = 1}$ and $\mathset{x_i | f_i =
+assuming that the flags, $\mathset{g_i}$, are un-observed data and
+treating the sets $\mathset{x_i | g_i = 1}$ and $\mathset{x_i | g_i =
   0}$ as samples from an inhomogeneous Poisson process.  For an
@@
 Applying this once to the foreground samples, once to the background
-samples and taking the product, we obtain $p(d,\mathset{f_i}, N | R_f,
-R_b, \theta)$ and thus $p(\mathset{f_i}, R_f, R_b, \theta | d,N) =
-p(d,\mathset{f_i}, N | R_f, R_b, \theta)\ p(R_f, R_b, \theta)\, /\,
+samples and taking the product, we obtain $p(d,\mathset{g_i}, N | R_f,
+R_b, \theta)$ and thus $p(\mathset{g_i}, R_f, R_b, \theta | d,N) =
+p(d,\mathset{g_i}, N | R_f, R_b, \theta)\ p(R_f, R_b, \theta)\, /\,
 p(d,N)$. With the identification $p(d,N) = N!\,p(d_{\rm to},N)$, as
@@
 
-We can marginalize the posterior over the flags, $f_i$, obtaining
+We can marginalize the posterior over the flags, $g_i$, obtaining
 \begin{multline}
   \label{eq:rate-shape-posterior}
-  p\left( R_f, R_b, \theta | d, N \right) = \sum_{\mathset{f_i} \in
-    \mathset{0,1}^N} p\left( \mathset{f_i}, R_f, R_b, \theta | d, N
+  p\left( R_f, R_b, \theta | d, N \right) = \sum_{\mathset{g_i} \in
+    \mathset{0,1}^N} p\left( \mathset{g_i}, R_f, R_b, \theta | d, N
   \right) \\ \propto \prod_{i} \left[ R_f \hat{f}\left(x_i, \theta\right)
@@
 significantly greater than the total number of observed events are
-strongly disfavored.  As we will see, the problem can also be avoided
+strongly disfavored.
+
+{\bf This problem can also be avoided
 in the context of the loudest-event framework by even very weak prior
 information on the background rate, $R_b$, of the kind present in
-nearly all experiments.
-
-This problem can be avoided in the loudest event framework, by
-including an upper limit on the rate, $R_{\rm max}$, in the prior for
-$R_b$. 
+nearly all experiments.  For example, we can include an upper limit
+on the rate, $R_{\rm max}$, in the prior for $R_b$.}
 \begin{widetext}
@@
 Note that these are posteriors on the number of events expected \emph{above the
-threshold $x_{\rm min}$}.
-The threshold choice for the foreground-dominated statistic could be different from the threshold choice applied elsewhere.
-If rates $R_{f,1}$ and $R_{f,2}$ are estimated for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$ respectively, these rates should be compared by equating $R_{f,1} (1-\hat{F}(x_2,\theta))$ and $R_{f,2} (1-\hat{F}(x_1,\theta))$.
+threshold $x_{\rm min}$}.  The threshold choice for the foreground-dominated statistic could be different from the threshold choice applied elsewhere.  
+%If rates $R_{f,1}$ and $R_{f,2}$ are estimated for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$ respectively, these rates should be compared by equating $R_{f,1} (1-\hat{F}(x_2,\theta))$ and $R_{f,2} (1-\hat{F}(x_1,\theta))$.
+{\bf If the rates are estimated accurately, then a rate estimate $R_{f,1}$ above threshold $x_{\rm min}=x_1$ can be converted into a rate estimate $R_{f,2}$ above threshold $x_{\rm min}=x_2$ via 
+$R_{f,1} (1-\hat{F}(x_2,\theta)) = R_{f,2} (1-\hat{F}(x_1,\theta))$; however, rate point estimates based on thresholding can have significant fluctuations, as discussed in the following section.}
 
@@
 \end{equation}
-i.e.\ the data-weighted average of the Bayesian estimate $R_{B}$
+i.e.\ the data-weighted average of the Bayesian {\bf estimator} $R_{B}$
 equals the prior-weighted average $R$.  Therefore all threshold values
-will yield, on average, the same estimate of the rate.  However this
+will yield, on average, the same {\bf point} estimate of the rate.  However this
 equality of averages does {\it not} imply that all threshold values
@@
 statistic of the synthetic data appears in Figure
-\ref{fig:analytic-data-cumulative}.  We used a Markov Chain Monte
+\ref{fig:analytic-data-cumulative}.  We used a Markov chain Monte
 Carlo simulation to draw samples of state flags and rates from the
@@
 Neither of these alternative methods can inform us about the value of
-$N_{\rm eff}$, a property of the background.  Moreover, the choice of
+$N_{\rm eff}$, a {\bf property of the distribution of background events identified by filtering with this template bank}.  Moreover, the choice of
 threshold value for the foreground-dominated statistic becomes
@@
 In the context of gravitational wave experiments, additional information on the ranking statistic distributions for the foreground can be obtained using mock signal injections into the data, while distributions for the background can be estimated by analyzing time slides of data sets from different detectors relative to each other \cite[e.g.,][]{ihope}.
-This information can be readily incorporated in the current framework by assuming there is another set of $N_I$ events with ranking statistics $\{ w_i \}$, known to be drawn from the foreground distribution ($f_i=1$) and a set of $N_T$ events with ranking statistics $\{z_i\}$ known to be drawn from the background distribution ($f_i=0$).
+This information can be readily incorporated in the current framework by assuming there is another set of $N_I$ events with ranking statistics $\{ w_i \}$, known to be drawn from the foreground distribution ($g_i=1$) and a set of $N_T$ events with ranking statistics $\{z_i\}$ known to be drawn from the background distribution ($g_i=0$).
 These events will typically not be drawn with the correct rate parameters, so they do not contribute to the estimates of $R_f$ and $R_b$, but they do contribute an extra factor
@@
 no single star in the domain that is more likely to be a cluster
-member than a background star (i.e.\ $\left \langle f_i \right \rangle
+member than a background star (i.e.\ $\left \langle g_i \right \rangle
 \lesssim 0.5$ for all stars); nevertheless, we will see that our
@@
 \begin{figure}
-  \includegraphics[width=0.4\columnwidth]{sky_location}\includegraphics[width=0.4\columnwidth]{scale}
+  \includegraphics[width=0.475\columnwidth]{sky_location}\includegraphics[width=0.475\columnwidth]{scale}
   \caption{\label{fig:cluster-params} Posterior density for the

### 2014-12-11 many.tex
Updates to address referee B's comments.

@@
 
-{\bf Our key results appear in Eq.~(\ref{eq:posterior}), which provides the joint posterior probability distribution on the foreground and background rates and shape parameters and the provenance of individual events as either foreground or background.
-Eq.~(\ref{eq:rate-shape-posterior}) is a marginalized version of Eq.~(\ref{eq:posterior}), useful when the provenance of individual events is not relevant.
-In practice, these posteriors are best sampled with stochastic techniques such as Markov chain Monte Carlo.}
+Our key results appear in Eq.~(\ref{eq:posterior}), which provides the
+joint posterior probability distribution on the foreground and
+background rates and shape parameters and the provenance of individual
+events as either foreground or background.
+Eq.~(\ref{eq:rate-shape-posterior}) is a marginalized version of
+Eq.~(\ref{eq:posterior}), useful when the provenance of individual
+events is not relevant.  In practice, these posteriors are best
+sampled with stochastic techniques such as Markov chain Monte Carlo.
 
@@
 probability distributions for these parameters that are implied by the
-data.  At the least, we will want to know the overall {\bf dimensionless} 
+data.  At the least, we will want to know the overall dimensionless
 amplitude of the foreground and background rates.  Let
@@
 
-{\bf This problem can also be avoided
-in the context of the loudest-event framework by even very weak prior
-information on the background rate, $R_b$, of the kind present in
-nearly all experiments.  For example, we can include an upper limit
-on the rate, $R_{\rm max}$, in the prior for $R_b$.}
+This problem can also be avoided in the context of the loudest-event
+framework by even very weak prior information on the background rate,
+$R_b$, of the kind present in nearly all experiments.  For example, we
+can include an upper limit on the rate, $R_{\rm max}$, in the prior
+for $R_b$.
 \begin{widetext}
@@
 %If rates $R_{f,1}$ and $R_{f,2}$ are estimated for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$ respectively, these rates should be compared by equating $R_{f,1} (1-\hat{F}(x_2,\theta))$ and $R_{f,2} (1-\hat{F}(x_1,\theta))$.
-{\bf If the rates are estimated accurately, then a rate estimate $R_{f,1}$ above threshold $x_{\rm min}=x_1$ can be converted into a rate estimate $R_{f,2}$ above threshold $x_{\rm min}=x_2$ via 
-$R_{f,1} (1-\hat{F}(x_2,\theta)) = R_{f,2} (1-\hat{F}(x_1,\theta))$; however, rate point estimates based on thresholding can have significant fluctuations, as discussed in the following section.}
+If the rates are estimated accurately, then a rate estimate $R_{f,1}$
+above threshold $x_{\rm min}=x_1$ can be converted into a rate
+estimate $R_{f,2}$ above threshold $x_{\rm min}=x_2$ via $R_{f,1}
+(1-\hat{F}(x_2,\theta)) = R_{f,2} (1-\hat{F}(x_1,\theta))$; however,
+rate point estimates based on thresholding can have significant
+fluctuations, as discussed in the following section.
 
@@
 \end{equation}
-i.e.\ the data-weighted average of the Bayesian {\bf estimator} $R_{B}$
+i.e.\ the data-weighted average of the Bayesian estimator $R_{B}$
 equals the prior-weighted average $R$.  Therefore all threshold values
-will yield, on average, the same {\bf point} estimate of the rate.  However this
-equality of averages does {\it not} imply that all threshold values
-yield the same information. In general, as the threshold is lowered to
-include more events, the error bar on the estimate shrinks.  In this
-subsection we give quantitative illustrations of how the error bar
-shrinks when the threshold is lowered.
+will yield, on average, the same point estimate of the rate.  However
+this equality of averages does {\it not} imply that all threshold
+values yield the same information. In general, as the threshold is
+lowered to include more events, the error bar on the estimate shrinks.
+In this subsection we give quantitative illustrations of how the error
+bar shrinks when the threshold is lowered.
 
@@
 the Fisher information matrix gives the Cramer-Rao bound on the
-covariance matrix of unbiased estimators of $\theta$.  For the
+covariance matrix of unbiased estimators of $\theta$.  Though our
+Bayesian analysis is not necessarily limited by the Cramer-Rao bound
+(since estimators constructed from it need not be unbiased and can
+also be affected by the prior), the Fisher information is indicative
+of the influence of each measurement on the posterior.  For the
 likelihood that enters Eq.~\eqref{eq:rate-shape-posterior}, the Fisher
@@
 Neither of these alternative methods can inform us about the value of
-$N_{\rm eff}$, a {\bf property of the distribution of background events identified by filtering with this template bank}.  Moreover, the choice of
-threshold value for the foreground-dominated statistic becomes
+$N_{\rm eff}$, a property of the distribution of background events
+identified by filtering with this template bank.  Moreover, the choice
+of threshold value for the foreground-dominated statistic becomes
 significantly more complicated in this case, since $p_\mathrm{thresh}$
@@
 
-Figure \ref{fig:cluster-params} shows the posteriors for the cluster
-parameters.  The center of the cluster, $\vec{x}_0$, is localized to
-within about 5\% of the cluster scale, and the cluster radius with a
-relative error of about 10\%.  In spite of the significant background,
-the cluster parameters are recovered to a relative accuracy consistent
-with the expected uncertainty from $N_\mathrm{eff} \simeq R_f = 1000$
-measurements.  Figure \ref{fig:cluster-number} shows the posteriors
-inferred on the cluster and background numbers, $R_f$ and $R_b$.
+Figures \ref{fig:cluster-loc} and \ref{fig:cluster-scale} shows the
+posteriors for the cluster location and scale parameters.  The center
+of the cluster, $\vec{x}_0$, is localized to within about 5\% of the
+cluster scale, and the cluster radius with a relative error of about
+10\%.  In spite of the significant background, the cluster parameters
+are recovered to a relative accuracy consistent with the expected
+uncertainty from $N_\mathrm{eff} \simeq R_f = 1000$ measurements.
+Figure \ref{fig:cluster-number} shows the posteriors inferred on the
+cluster and background numbers, $R_f$ and $R_b$.
 
 \begin{figure}
-  \includegraphics[width=0.475\columnwidth]{sky_location}\includegraphics[width=0.475\columnwidth]{scale}
-  \caption{\label{fig:cluster-params} Posterior density for the
-    cluster parameters for the example from \S~\ref{sec:star-cluster}.
-    (Left) Contours of the posterior probability distribution for the
-    center of the cluster, $\vec{x}_0$.  The center $(x,y) =
-    \left(x_0, y_0\right)$ is determined to within about 5\% of the
-    structural radius of the cluster, $r_0$ (see
-    Eq.~\eqref{eq:true-cluster-parameters}).  (Right) Posterior
-    density for the scale parameter for the cluster, $r_0$.  The true
-    value is indicated by the vertical line (see
-    Eq.~\ref{eq:true-cluster-parameters}).}
+  \includegraphics[width=\columnwidth]{sky_location}
+  \caption{\label{fig:cluster-loc} Contours of the posterior
+    probability distribution for the center of the cluster,
+    $\vec{x}_0$, for the example from \S~\ref{sec:star-cluster}.  The
+    center $(x,y) = \left(x_0, y_0\right)$ is determined to within
+    about 5\% of the structural radius of the cluster, $r_0$ (see
+    Eq.~\eqref{eq:true-cluster-parameters}).}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{scale}
+  \caption{\label{fig:cluster-scale} Posterior density for the scale
+    parameter for the cluster, $r_0$, for the example from
+    \S~\ref{sec:star-cluster}.  The true value is indicated by the
+    vertical line (see Eq.~\ref{eq:true-cluster-parameters}).}
 \end{figure}
